### 2014-06-03 inst/wtc-finland/manuscript.Rmd
...

#+ setup, include=FALSE
library(knitr)
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, tidy=FALSE, fig.width=10, fig.height=15, cache=TRUE, cache.path="~/tmp/cache/")
```{r check, cache=FALSE}
message("cache path = ", opts_chunk$get("cache.path"))
```

#' NOTE: Figures and tables may be out of sync with the manuscript!
#' ----------------------------------------------------------------

```{r initialize}
library(parallel)
library(doMC)
registerDoMC(cores=detectCores())
library(WTC)
source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
library(sp)
library(ggplot2)
library(rasterVis)
library(reshape2)
library(plyr)

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
study <- FinlandWTCStudy$new(context=context, response="canis.lupus")
study$studyArea$loadBoundary(thin=T, tolerance=0.001)
boundaryDF <- ggplot2::fortify(study$studyArea$boundary)
responses <- c("canis.lupus", "lynx.lynx", "rangifer.tarandus.fennicus")
```

######
### Survey routes (Figure 1)
######

```{r survey-routes}
surveyRoutes <- study$loadSurveyRoutes(findLengths=FALSE)
surveyRoutesSP <- SpatialLinesDataFrame(surveyRoutes$surveyRoutes, data=data.frame(x=1:length(surveyRoutes$surveyRoutes)), match.ID=FALSE)
surveyRoutesDF <- ggplot2::fortify(surveyRoutesSP)
p <- ggplot(boundaryDF, aes(long, lat, group=group)) + geom_polygon(colour="darkgray", fill=NA) +
  geom_polygon(data=surveyRoutesDF, aes(long, lat, group=group), colour="black", fill=NA) +
  coord_equal() + theme_raster()
print(p)
#saveFigure(p, filename="SurveyRoutes.svg", bg="transparent")
```

######
### Number of counted survey routes
######

```{r number-of-counted-survey-routes}
intersections <- study$loadIntersections(predictDistances=FALSE)
x <- ddply(as.data.frame(intersections$intersections), .(year), nrow)
s <- sd(x$V1)
m <- mean(x$V1)
cat("mean = ", m, "\n")
cat("sd = ", s, "\n")
cat("range = ", m+c(-s,s), "\n")
```

######
### % of intersections D > 1
######

```{r percent-of-intersections-day-more-than-one, message=TRUE}
intersections <- FinlandWTCIntersections$new(study=study, maxDuration=1)
intersections$saveIntersections()
```

######
### Crossing rates (D > 1 removed) (Figure 2)
######

```{r crossing-rates, fig.height=5}
getIntersectionsRate <- function(responses, context) {
  library(plyr)
  x <- ldply(responses, function(response, context) {
    study <- FinlandWTCStudy$new(context=context, response=response)
    intersections <- study$loadIntersections(predictDistances=FALSE)
    x <- ddply(intersections$intersections@data, .(year), function(x, response) {
      data.frame(intersections=sum(x$intersections / (x$length/1000) / x$duration) / nrow(x), response=study$getPrettyResponse(response))
    }, response=response)
    return(x)
  }, context=context)
  return(x)
}

intersectionsRate <- getIntersectionsRate(responses, context)
p <- ggplot(intersectionsRate, aes(year, intersections)) + geom_line(size=1) + facet_grid(~response) +
  xlab("Year") + ylab("Crossings rate") + theme_presentation(20)
print(p)
#saveFigure(p, filename="IntersectionsRate.svg", bg="transparent")
```

######
### GPS figures
######

```{r gps-in-numbers, message=TRUE}
getTracks <- function(responses, context) {
  library(plyr)
  library(ggplot2)
  library(adehabitatLT)

  l_ply(responses, function(response, context) {
    study <- FinlandWTCStudy$new(context=context, response=response)
    tracks <- FinlandWTCTracks$new(study=study)
    tracks$loadTracks()

    message("Response = ", response)
    message("Number of individuals in the cleaned data = ", length(unique(adehabitatLT::id(tracks$tracks))))
    d <- as.POSIXlt(ld(tracks$tracks)$date)
    message("Years in the cleaned data = ", paste(sort(unique(d$year+1900)), collapse=","), " (", length(unique(d$year)), ")")
    message("Number of samples in the cleaned data = ", length(d))
    message("Number of days in the cleaned data = ", length(unique(paste(d$year, d$yday))))

  }, context=context)
}

getTracks(responses=responses, context=context)
```

######
### GPS movements (Figure 3)
######

```{r gps-movements, fig.height=10, fig.width=20}
getTracks <- function(responses, context) {
  library(plyr)
  library(ggplot2)
  x <- ldply(responses, function(response, context) {
    study <- FinlandWTCStudy$new(context=context, response=response)
    tracks <- FinlandWTCTracks$new(study=study)
    tracks$loadTracks()
    tracksSP <- tracks$getSpatialLines(variables=.(id,year,burst))

    x <- tracksSP@data
    if (!"sex" %in% colnames(x)) x$sex <- NA
    x$juvenile <- if ("born" %in% colnames(x)) factor(ifelse(x$year - x$born <= 1, "juvenile", "adult")) else factor(NA)
    x$id <- sapply(tracksSP@lines, function(x) x@ID)
    x$sexjuv <- factor(paste(x$sex, x$juvenile, sep=", "))
    tracksDF <- ggplot2::fortify(tracksSP)
    tracksDF <- plyr::join(tracksDF, x, by="id")
    tracksDF$response <- study$getPrettyResponse(response)

    return(tracksDF)
  }, context=context)
  return(x)
}

tracks <- getTracks(responses=responses, context=context)
#levels(tracks$sexjuv)

p <- ggplot(tracks, aes(long, lat, group=group)) +
  geom_polygon(data=boundaryDF, aes(long, lat, group=group), colour="darkgray", fill=NA) +
  geom_path(aes(colour=sexjuv, linetype=sexjuv), alpha=.5) +
  scale_color_manual("Sex, age", values=c("darkred","red","red","darkblue","blue","blue","black")) + scale_linetype_manual("Sex, age", values=c("solid","dashed","solid","solid","dashed","solid","solid")) +
  facet_grid(~response) + theme_raster(18, legend.position="bottom", aspect.ratio=1) + coord_cartesian(ylim=range(tracks$lat)+c(-1,1)*50000)
print(p)
#saveFigure(p, filename="GPS.svg", bg="transparent")
```

######
### Straight-line-distance-error (Figure 4)
######

```{r straight-line-distance-error, fig.height=5}
ss <- SimulationStudy$new(response="Intensive")$newInstance(context=context)
tracks <- ss$loadTracks(iteration=as.integer(1))
tracks$tracks <- subset(tracks$tracks, id == 1 & year == 2001 & month == 1 & day %in% c(2,3))
intervals <- MovementSampleIntervals$new(study=ss)
thinnedTracks <- intervals$getThinnedTracksSampleIntervals(tracks=tracks)

plotTracks <- data.frame()
for (i in seq(1, 11, by=5)) {
  x <- thinnedTracks$tracksList[[i]]$tracks
  x$thin <- i 
  x <- subset(x, date >= as.POSIXct("2001-01-03 00:00:00") & date <= as.POSIXct("2001-01-03 03:40:00"))
  plotTracks <- rbind(plotTracks, x)
}
plotTracks$thin <- factor(plotTracks$thin)
plotTracks$dt <- factor(plotTracks$dt / 60)

segment <- aes(x=x, y=y, xend=c(tail(x, n=-1), NA), yend=c(tail(y, n=-1), NA), group=dt, colour=dt)
ending <- arrow(length=unit(0.7, "cm"))
p <- ggplot(plotTracks, mapping=segment) +
  theme_raster(20, legend.position=c(0.5,0.85), legend.background=element_rect(color="grey")) + 
  #guides(colour=guide_legend(title=expression(paste(Delta, "t (min)")))) +
  guides(colour=guide_legend(title="Sampling interval (min)")) +
  scale_colour_manual(values=c("gray70","gray40","gray10")) +
  #ggtitle("Straight-line distance error") +
  geom_path(size=2, arrow=ending) + geom_point(size=4)
#geom_segment(size=2, arrow=ending) # doesn't work with grouping
print(p)
#saveFigure(p, filename="StraightLineDistanceError.svg", bg="transparent")
```

######
### Raw GPS sample intervals
######

```{r raw-gps-sample-intervals}
intervals <- llply(responses, function(response, context) {
  study <- FinlandWTCStudy$new(context=context, response=response)
  tracks <- FinlandWTCTracks$new(study=study)
  tracks$loadTracks()
  intervals <- MovementSampleIntervals$new(study=study)
  intervals$loadSampleIntervals()
  return(table(intervals$intervals$intervalH))
}, context=context)
intervals
```

######
### Power law fits for distance - sampling interval (Figure 5)
######

```{r distance-fits, fig.height=5}
distances <- ldply(responses, function(response, context) {
  study <- FinlandWTCStudy$new(context=context, response=response)
  intervals <- study$loadSampleIntervals()
  x <- intervals$intervals
  x$response <- study$getPrettyResponse(response=response)
  return(x)
}, context=context)

intervals <- ldply(responses, function(response, context) {
  study <- FinlandWTCStudy$new(context=context, response=response)
  intervals <- study$loadSampleIntervals()
  distances <- intervals$getDistanceCurve()
  distances$response <- study$getPrettyResponse(response=response)
  return(distances)
}, context=context)

p <- ggplot(distances, aes(intervalH, distanceKm)) + geom_point(alpha=.3) +
  ylab("Distance / day (km)") + xlab("Sampling interval (h)") +
  geom_line(data=intervals, aes(intervalH, distanceKm), color="black", size=1) +
  geom_smooth(data=intervals, aes(ymin=distanceKm25, ymax=distanceKm975), stat="identity") +
  ylim(c(0, max(distances$distanceKm))) +
  facet_grid(~response) + theme_presentation()
print(p)
#saveFigure(p, filename=paste("DistanceCorrection-", response, ".svg", sep=""), bg="transparent")
```

######
### Observed distances - estimated distances (Table 1)
######

```{r distance-estimates-init, results="hide"}
observedDistances <- ldply(responses, function(response, context) {
  study <- FinlandWTCStudy$new(context=context, response=response)
  tracks <- study$loadTracks()
  intervals <- tracks$getSampleIntervals()
  return(data.frame(distance=intervals$intervals$dist, logDistance=log(intervals$intervals$dist), response=study$getPrettyResponse(response)))
}, context=context)

predictedDistances <- ldply(responses, function(response, context) {
  study <- FinlandWTCStudy$new(context=context, response=response, distanceCovariatesModel=~populationDensity+rrday+snow+tday-1, trackSampleInterval=2)
  intersections <- study$loadIntersections()
  intersections$predictDistances()
  distances <- intersections$intersections$distance / 1000
  return(data.frame(distance=distances, logDistance=log(distances), response=study$getPrettyResponse(response)))  
}, context=context)
```

```{r distance-estimates}
cat("Observed distances (km/day):\n")
ddply(observedDistances, .(response), summarize, mean=mean(distance), sd=sd(distance), min=min(distance), max=max(distance))
cat("Predicted distances (km/day):\n")
ddply(predictedDistances, .(response), summarize, mean=mean(distance), sd=sd(distance), min=min(distance), max=max(distance))
cat("Prediction interval = 2h")
```

# Distances from literature:
# Wolf: 7-25 km/day
# Lynx:
# FFR:

######
### Power law fit estimates for distance - sampling interval (Table 1)
######
# Fixed effects: human population density, precipitation, snow cover, temperature

```{r distance-fit-estimates}
# Fixed effects: populationDensity + rrday + snow + tday
intervals <- l_ply(responses, function(response, context) {
  study <- FinlandWTCStudy$new(context=context, response=response)
  intervals <- study$loadSampleIntervals()
  cat("response = ", study$getPrettyResponse(response=response), "\n")
  intervals$estimatedValuesSummary()
  cat("\n\n")
}, context=context)
```

######
### Crude distance estimates (Table 1)
#####

```{r distance-crude-estimates}
responses <- c("canis.lupus","lynx.lynx","rangifer.tarandus.fennicus")
for (response in responses) {
  context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
  study <- FinlandWTCStudy$new(context=context, response=response, distanceCovariatesModel=~populationDensity+rrday+snow+tday-1, trackSampleInterval=2)

  populationSize <- FinlandPopulationSize$new(study=study)
  populationSize$loadValidationData()

  intersections <- study$loadIntersections()

  A <- 338424 # km^2
  M <- ddply(intersections$getData(), .(year), function(x) sum(x$length))$V1 / 1000
  N <- populationSize$sizeData$Validation
  x <- ddply(intersections$getData(), .(year), function(x) sum(x$intersections))$V1

  C <- 2/pi * A / M
  l <- C * x / N
  cat(response, ": crude distance mean = ", mean(l, na.rm=T), " median = ", median(l, na.rm=T), " sd = ", sd(l, na.rm=T), " km/d\n")

  h <- study$loadHabitatWeightsRaster()
  h.masked <- mask(h, study$studyArea$boundary)
  n.cell <- length(h.masked) - freq(h.masked, value=NA)
  A.discrete <- n.cell * prod(res(h.masked)) / 1000^2
  h.cell <- cellStats(h.masked, sum)
  A.h <- h.cell * prod(res(h.masked)) / 1000^2 * A / A.discrete # A.discrete is a little bit smaller than A due to the discretization, so apply a correction

  C <- 2/pi * A.h / M
  l <- C * x / N
  cat(response, ": crude weighted distance mean = ", mean(l, na.rm=T), " median = ", median(l, na.rm=T), " sd = ", sd(l, na.rm=T), " km/d\n")

  cat(response, ": observed number of intersections:\n")
  print(1/C * l)

  cat("Area correction = ", A/A.discrete, "\n")
}
```

######
### Habitat weights (Figure 6)
######

```{r habitat-weights, fig.height=5, results="hide"}
getHabitatWeights <- function(responses, context) {
  library(plyr)
  x <- ldply(responses, function(response, context) {
    study <- FinlandWTCStudy$new(context=context, response=response)
    weights <- study$loadHabitatWeights()$getHabitatSelectionWeights()
    n <- names(weights)
    print(n)
    weights <- data.frame(habitat=factor(n, levels=n), weights)
    weights$response <- study$getPrettyResponse(response)
    return(weights)
  }, context=context)
  return(x)
}

weights <- getHabitatWeights(responses=responses, context=context)
p <- ggplot(weights, aes(habitat, weights, fill=habitat)) +
  geom_bar(stat="identity") + facet_grid(~response) + scale_fill_manual(values=c("gray40","gray40","gray40","gray40","gray40")) +
  xlab("") + ylab("Weight") + theme_presentation(20, axis.text.x=element_text(angle=90, hjust=1))
print(p)
#saveFigure(p, filename="HabitatWeights.svg", bg="transparent")
```

######
### Population density estimates
######

```{r population-size-init, fig.keep="none"}
plotObservations <- function(i, params) {
  subdata <- params$data[params$data$year == min(params$data$year) + i - 1,]
  subdata$Observed <- subdata$intersections>0
  geom_point(mapping=aes(coords.x1, coords.x2, colour=Observed), data=subdata)
}

getPopulationDensity <- function(responses, spatialModels, timeModels, context, withHabitatWeights=FALSE, saveDensityPlots=TRUE) {
  populationDensity <- list()
  for (i in 1:length(responses)) {
    response <- responses[i]
    timeModel <- timeModels[i]
    spatialModel <- spatialModels[i]

    cat("Response = ", response, ", time model = ", timeModel, "\n")

    study <- FinlandWTCStudy$new(context=context, response=response, distanceCovariatesModel=~populationDensity+rrday+snow+tday-1, trackSampleInterval=2)
    estimates <- if (spatialModel) FinlandSmoothModelSpatioTemporal(study=study)$setModelName("nbinomial", paste("matern", timeModel, sep="-"))$loadEstimates()
    else FinlandSmoothModelTemporal(study=study)$setModelName("nbinomial", timeModel)$loadEstimates()
    estimates$offsetScale <- 1000^2 # TODO: quickfix
    estimates$collectEstimates()
    habitatWeightsRaster <- study$loadHabitatWeightsRaster()
    data <- cbind(estimates$getUnscaledObservationCoordinates(), estimates$data)

    populationDensity[[i]] <- estimates$getPopulationDensity(templateRaster=habitatWeightsRaster, getSD=FALSE)
    populationDensity[[i]]$mean$animate(name="PopulationDensity-mean", legend=expression(bold(Individuals / km^2)),
                                   ggfun=plotObservations, ggfunParams=list(data=data))

    populationDensity[[i]]$mean$weight(habitatWeightsRaster)
    populationDensity[[i]]$mean$animate(name="WeightedPopulationDensity-mean", legend=expression(bold(Individuals / km^2)),
                                   ggfun=plotObservations, ggfunParams=list(data=data))

    #populationDensity[[i]] <- study$getPopulationDensity(model=estimates, withHabitatWeights=withHabitatWeights, saveDensityPlots=saveDensityPlots)
    #populationDensity[[i]] <- study$getPopulationDensity2(withHabitatWeights=withHabitatWeights)
  }
  return(populationDensity)
}

timeModels <- c("ar1", "ar1", "rw2")
spatialModels <- c(T,T,F)
weightedPopulationDensity <- getPopulationDensity(responses=responses, spatialModels=spatialModels, timeModels=timeModels, context=context, withHabitatWeights=TRUE, saveDensityPlots=TRUE)
```

######
### Total population size estimates  (Figure 7)
######

```{r population-size, fig.height=5}
populationSize <- data.frame()
for (i in 1:length(responses)) {
  response <- responses[i]
  study <- FinlandWTCStudy$new(context=context, response=response)
  x <- weightedPopulationDensity[[i]]$mean$integrate(volume=FinlandPopulationSize$new(study=study))
  x$loadValidationData()
  y <- x$sizeData
  y$response <- study$getPrettyResponse(response)
  cat("Response = ", response, ", coef to match the validation data:\n")
  print(coef(x$match()))
  if (response == "canis.lupus") y$"Transformed estimated" <- y$Estimated * .7
  else if (response == "lynx.lynx") y$"Transformed estimated" <- y$Estimated / 2 *.5
  else if (response == "rangifer.tarandus.fennicus") y$"Transformed estimated" <- y$Estimated / 2 * 1.1

  if (response == "lynx.lynx") {
    y$Estimated <- y$Estimated / 2
  }
  if (response == "rangifer.tarandus.fennicus") {
    y$Estimated <- y$Estimated / 2
    y$Validation <- with(y, approx(Year, Validation, n=nrow(y)))$y
  }

  populationSize <- rbind(populationSize, y)
}

x <- melt(populationSize, id.vars=c("Year","response"), variable.name="Variable")
p <- ggplot(x, aes(Year, value, group=Variable, colour=Variable)) + geom_line(size=1) + facet_wrap(~response, scales="free_y") +
  scale_colour_manual(values=c("steelblue","violetred1","steelblue1")) +
  xlab("Year") + ylab("Population size") +
  theme_presentation(20, axis.text.x=element_text(angle=90, hjust=1)) + theme(legend.position="bottom")
print(p)
#saveFigure(p, filename="PopulationSize.svg", bg="transparent")
```

### 2014-06-03 R/Model.R

@@ -477,7 +477,7 @@ SmoothModelSpatioTemporal <- setRefClass(
       return(invisible(projectionRaster))
     },
     
-    getPopulationDensity = function(templateRaster=study$getTemplateRaster(), maskPolygon=study$studyArea$boundary, getSD=TRUE) {
+    getPopulationDensity____ORIGINAL = function(templateRaster=study$getTemplateRaster(), maskPolygon=study$studyArea$boundary, getSD=TRUE) {
       if (length(node) == 0)
         stop("Did you forgot to run collectEstimates() first?")
       
@@ -503,7 +503,7 @@ SmoothModelSpatioTemporal <- setRefClass(
       return(invisible(list(mean=meanPopulationDensityRaster, sd=sdPopulationDensityRaster)))
     },
     
-    getPopulationDensity2 = function(templateRaster=study$getTemplateRaster(), maskPolygon=study$studyArea$boundary, getSD=TRUE) {
+    getPopulationDensity = function(templateRaster=study$getTemplateRaster(), maskPolygon=study$studyArea$boundary, getSD=TRUE) {
       if (length(node) == 0)
         stop("Did you forgot to run collectEstimates() first?")
       
@@ -513,8 +513,8 @@ SmoothModelSpatioTemporal <- setRefClass(
       xyztSD <- data.frame(getUnscaledObservationCoordinates(), z=data$fittedSD / offsetScale, t=data$year)
       cellArea <- prod(res(templateRaster)) # m^2
       
-      meanPopulationDensityRaster <- SpatioTemporalRaster$new(study=study)$interpolate(xyztMean, templateRaster=templateRaster, transform=sqrt, inverseTransform=square, boundary=maskPolygon, layerNames=unique(xyztMean$t), weights=cellArea)
-      sdPopulationDensityRaster <- if (getSD) SpatioTemporalRaster$new(study=study)$interpolate(xyztMean, templateRaster=templateRaster, transform=sqrt, inverseTransform=square, boundary=maskPolygon, layerNames=unique(xyztMean$t), weights=cellArea)
+      meanPopulationDensityRaster <- SpatioTemporalRaster$new(study=study)$interpolate(xyztMean, templateRaster=templateRaster, transform=sqrt, inverseTransform=square, boundary=maskPolygon, layerNames=sort(unique(xyztMean$t)), weights=cellArea)
[…]

### 2014-06-03 inst/wtc-finland/manuscript.Rmd
...

@@
-#+ setup, include=FALSE
-library(knitr)
+```{r global_options, include=FALSE}
 opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, tidy=FALSE, fig.width=10, fig.height=15, cache=TRUE, cache.path="~/tmp/cache/")
-```{r check, cache=FALSE}
-message("cache path = ", opts_chunk$get("cache.path"))
 ```
 
-#' NOTE: Figures and tables may be out of sync with the manuscript!
-#' ----------------------------------------------------------------
+NOTE: Figures and tables may be out of sync with the manuscript!
+----------------------------------------------------------------
 
@@
 
-######
-### Survey routes (Figure 1)
-######
+Survey routes (Figure 1)
+------------------------
 
@@
 
-######
-### Number of counted survey routes
-######
+Number of counted survey routes
+-------------------------------
 
@@
 
-######
-### % of intersections D > 1
-######
+% of intersections D > 1
+------------------------
 
@@
 
-######
-### Crossing rates (D > 1 removed) (Figure 2)
-######
+Crossing rates (D > 1 removed) (Figure 2)
+-----------------------------------------
 
@@
 
-######
-### GPS figures
-######
+GPS figures
+-----------
 
@@
 
-######
-### GPS movements (Figure 3)
-######
+GPS movements (Figure 3)
+------------------------
 
@@
 
-######
-### Straight-line-distance-error (Figure 4)
-######
+Straight-line-distance-error (Figure 4)
+---------------------------------------
 
@@
 
-######
-### Raw GPS sample intervals
-######
+Raw GPS sample intervals
+------------------------
 
@@
 
-######
-### Power law fits for distance - sampling interval (Figure 5)
-######
+Power law fits for distance - sampling interval (Figure 5)
+----------------------------------------------------------
 
@@
 
-######
-### Observed distances - estimated distances (Table 1)
-######
+Observed distances - estimated distances (Table 1)
+--------------------------------------------------
 
@@
 
-# Distances from literature:
-# Wolf: 7-25 km/day
-# Lynx:
-# FFR:
+Distances from literature
+-------------------------
+* Wolf: 7-25 km/day
+* Lynx:
+* FFR:
 
-######
-### Power law fit estimates for distance - sampling interval (Table 1)
-######
-# Fixed effects: human population density, precipitation, snow cover, temperature
+Power law fit estimates for distance - sampling interval (Table 1)
+------------------------------------------------------------------
+
+Fixed effects:
+* human population density
+* precipitation
+* snow cover
+* temperature
 
@@
 
-######
-### Crude distance estimates (Table 1)
-#####
+Crude distance estimates (Table 1)
+----------------------------------
 
@@
 
-######
-### Habitat weights (Figure 6)
-######
+Habitat weights (Figure 6)
+--------------------------
 
@@
 
-######
-### Population density estimates
-######
+Population density estimates
+----------------------------
 
-```{r population-size-init, fig.keep="none"}
+```{r population-size-init, fig.keep="none", eval=TRUE}
 plotObservations <- function(i, params) {
@@
 
-######
-### Total population size estimates  (Figure 7)
-######
+Total population size estimates (Figure 7)
+------------------------------------------
 
-```{r population-size, fig.height=5}
+```{r population-size, fig.height=5, eval=TRUE}
 populationSize <- data.frame()

### 2014-06-03 R/Tracks.R

@@ -460,7 +460,7 @@ FinlandWTCTracks <- setRefClass(
       metadata.raw <- read.csv(file.path(study$context$rawDataDirectory, "wolf-metadata.csv"))
       gps.metadata <- data.frame(id=metadata.raw$nimi,
                                  sex=ifelse(metadata.raw$sp=="uros", "male", "female"),
-                                 born=metadata.raw$"syntym\a",
+                                 born=metadata.raw$"syntym\u00e4",
                                  dispersing=metadata.raw$dispersoiva.nuori==1,
                                  homerange=metadata.raw$reviiri==1,
                                  alpha=metadata.raw$alfa==1)

### 2014-06-23 inst/wtc-finland/manuscript.Rmd
...

@@
 ```{r straight-line-distance-error, fig.height=5}
-ss <- SimulationStudy$new(response="Intensive")$newInstance(context=context)
+ss <- SimulationStudy$new(response="Intensive")$setup(context=context)
 tracks <- ss$loadTracks(iteration=as.integer(1))

### 2014-06-23 R/Study.R

@@ -33,10 +33,10 @@ SimulationStudy <- setRefClass(
       return(invisible(.self))
     },
     
-    newInstance = function(context, isTest=F) {
+    setup = function(context, isTest=F) {
       context <<- context
-      studyArea <<- if (isTest) TestStudyArea$new(context=context)$newInstance()
-      else FinlandStudyArea$new(context=context)$newInstance()
+      studyArea <<- if (isTest) TestStudyArea$new(context=context)$setup()
+      else FinlandStudyArea$new(context=context)$setup()
       return(invisible(.self))
     },
     
@@ -136,7 +136,7 @@ FinlandWTCStudy <- setRefClass(
   methods = list(
     initialize = function(context, ...) {
       callSuper(context=context, ...)
-      studyArea <<- FinlandStudyArea$new(context=context)$newInstance(tolerance=0.001)
+      studyArea <<- FinlandStudyArea$new(context=context)$setup(tolerance=0.001)
       return(invisible(.self))
     },
     
@@ -340,7 +340,7 @@ RussiaWTCStudy <- setRefClass(
   methods = list(
     initialize = function(context, ...) {
       callSuper(context=context, ...)
-      studyArea <<- RussiaStudyArea$new(context=context)$newInstance()
+      studyArea <<- RussiaStudyArea$new(context=context)$setup()
       return(invisible(.self))
     },
     
@@ -373,7 +373,7 @@ FinlandRussiaWTCStudy <- setRefClass(
   methods = list(
     initialize = function(context, ...) {
       callSuper(context=context, ...)
-      studyArea <<- FinlandRussiaStudyArea$new(context=context)$newInstance()
+      studyArea <<- FinlandRussiaStudyArea$new(context=context)$setup()
       return(invisible(.self))
     },
     

### 2014-06-23 R/StudyArea.R

@@ -18,7 +18,7 @@ StudyArea <- setRefClass(
       return(invisible(.self))
     },
     
-    newInstance = function(thin=TRUE, tolerance=0.1, prepareHabitatRaster=FALSE, rawRasterFile) {
+    setup = function(thin=TRUE, tolerance=0.1, prepareHabitatRaster=FALSE, rawRasterFile) {
       loadBoundary(thin=thin, tolerance=tolerance)
       if (prepareHabitatRaster) {
         if (missing(rawRasterFile)) stop("rawRasterFile parameter missing.")
@@ -332,8 +332,8 @@ FinlandRussiaStudyArea <- setRefClass(
       library(sp)
       library(rgdal)
       
-      finland <- FinlandStudyArea$new(context=context)$newInstance()
-      russia <- RussiaStudyArea$new(context=context)$newInstance()
+      finland <- FinlandStudyArea$new(context=context)$setup()
+      russia <- RussiaStudyArea$new(context=context)$setup()
       
       x <- spTransform(finland$boundary, russia$proj4string)
       y <- russia$boundary

### 2014-06-23 inst/simulation-test/simulate.R

@@ -12,12 +12,12 @@ simulate = function(scenario, nSurveyRoutes=as.integer(50), nAgents=as.integer(2
   
   context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
   mss <- {
-    if (scenario == "A") MovementSimulationScenarioA$new(nAgents=nAgents, years=nYears, days=nDays, CRWCorrelation=CRWCorrelation)$newInstance(context=context, isTest=T)
-    else if (scenario == "B") MovementSimulationScenarioB$new(nAgents=nAgents, years=nYears, days=nDays, CRWCorrelation=CRWCorrelation, BCRWCorrelationBiasTradeoff=BCRWCorrelationBiasTradeoff)$newInstance(context=context, isTest=T)
-    else if (scenario == "C") MovementSimulationScenarioC$new(nAgents=nAgents, years=nYears, days=nDays, CRWCorrelation=CRWCorrelation)$newInstance(context=context, isTest=T)
-    else if (scenario == "D") MovementSimulationScenarioD$new(nAgents=nAgents, years=nYears, days=nDays, CRWCorrelation=CRWCorrelation)$newInstance(context=context, isTest=T)
-    else if (scenario == "E") MovementSimulationScenarioE$new(nAgents=nAgents, years=nYears, days=nDays, CRWCorrelation=CRWCorrelation, nSurveyRoutes=nSurveyRoutes)$newInstance(context=context, isTest=T)
-    else if (scenario == "F") MovementSimulationScenarioF$new(nAgents=nAgents, years=nYears, days=nDays, CRWCorrelation=CRWCorrelation, BCRWCorrelationBiasTradeoff=BCRWCorrelationBiasTradeoff, nSurveyRoutes=nSurveyRoutes)$newInstance(context=context, isTest=T)
+    if (scenario == "A") MovementSimulationScenarioA$new(nAgents=nAgents, years=nYears, days=nDays, CRWCorrelation=CRWCorrelation)$setup(context=context, isTest=T)
[…]

### 2014-06-23 inst/simulation/count_intersections.R

@@ -40,6 +40,6 @@ library(WTC)
 source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
 
 context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
-mss <- if (scenario == "A") MovementSimulationScenarioA$new()$newInstance(context=context) else stop("Unknown scenario ", scenario)
+mss <- if (scenario == "A") MovementSimulationScenarioA$new()$setup(context=context) else stop("Unknown scenario ", scenario)
 study <- mss$study
 countIntersections(study=study, iteration=as.integer(task_id), test=test=="test")

### 2014-06-23 inst/simulation/estimate.R

@@ -39,7 +39,7 @@ estimate <- function(study, iteration, test) {
     message("Spatial test...")
     meshParams <- list(maxEdge=c(.1e6, .2e6), cutOff=.05e6, coordsScale=1e-6)
     interceptPriorParams <- list(mean=200, sd=199)
-    study <- SimulationStudySubset$new(response="A", years=as.integer(2001:2001))$newInstance(context=context)
+    study <- SimulationStudySubset$new(response="A", years=as.integer(2001:2001))$setup(context=context)
     intersections <- study$loadIntersections(iteration=iteration)
     model <- SimulatedSmoothModel$new(study=study, iteration=iteration)
     model$setup(intersections=intersections, meshParams=meshParams, useCovariates=FALSE)
@@ -69,6 +69,6 @@ library(WTC)
 source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
 
 context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
-mss <- if (scenario == "A") MovementSimulationScenarioA$new()$newInstance(context=context) else stop("Unknown scenario ", scenario)
+mss <- if (scenario == "A") MovementSimulationScenarioA$new()$setup(context=context) else stop("Unknown scenario ", scenario)
 study <- mss$study
 estimate(study=study, iteration=as.integer(task_id), test)

### 2014-06-23 inst/simulation/population_size.R

@@ -42,5 +42,5 @@ library(WTC)
 source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
 
 context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
-mss <- if (scenario == "A") MovementSimulationScenarioA$new()$newInstance(context=context) else stop("Unknown scenario ", scenario)
+mss <- if (scenario == "A") MovementSimulationScenarioA$new()$setup(context=context) else stop("Unknown scenario ", scenario)
 population_size(mss=mss, iteration=as.integer(task_id), test=test=="test")

### 2014-06-23 inst/simulation/simulate.R

@@ -5,15 +5,15 @@
 
 dryRun <- function(iteration, test) {
   context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
-  mss <- if (test == "test") MovementSimulationScenarioA$new(nAgents=as.integer(2), nIterations=as.integer(5), years=as.integer(2))$newInstance(context=context)
-  else MovementSimulationScenarioA$new()$newInstance(context=context)
+  mss <- if (test == "test") MovementSimulationScenarioA$new(nAgents=as.integer(2), nIterations=as.integer(5), years=as.integer(2))$setup(context=context)
+  else MovementSimulationScenarioA$new()$setup(context=context)
   message("SUCCESS")
 }
 
 simulateA <- function(iteration, test) {
   context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
-  mss <- if (test == "test") MovementSimulationScenarioA$new(nAgents=as.integer(2), nIterations=as.integer(5), years=as.integer(2))$newInstance(context=context)
-  else MovementSimulationScenarioA$new()$newInstance(context=context)
+  mss <- if (test == "test") MovementSimulationScenarioA$new(nAgents=as.integer(2), nIterations=as.integer(5), years=as.integer(2))$setup(context=context)
+  else MovementSimulationScenarioA$new()$setup(context=context)
   mss$simulateSingle(iteration=iteration)
 }
 

### 2014-06-23 inst/simulation/summarize.R

@@ -5,7 +5,7 @@ library(WTC)
 source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
 
 context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
-mss <- if (scenario == "A") MovementSimulationScenarioA$new()$newInstance(context=context) else stop("Unknown scenario ", scenario)
+mss <- if (scenario == "A") MovementSimulationScenarioA$new()$setup(context=context) else stop("Unknown scenario ", scenario)
 study <- mss$study
 
 iterations <- study$context$getIterationIds(dir=study$context$resultDataDirectory, name="PopulationSize", response=study$response, region=study$studyArea$region)

### 2014-06-28 inst/wtc-finland/manuscript.Rmd
...

@@
 ```{r global_options, include=FALSE}
-opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, tidy=FALSE, fig.width=10, fig.height=15, cache=TRUE, cache.path="~/tmp/cache/")
+library(knitr)
+single_width <- 2.75591 * 2
+double_width <- single_width*2 + 0.5 * 2
+single_height <- single_width
+double_height <- double_width
+#opts_knit$set(base.dir="~/.pandoc")
+opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, tidy=FALSE, fig.width=single_width, fig.height=single_height, cache=TRUE, cache.path="~/tmp/cache/")
+base_size <- 16
 ```
-
-NOTE: Figures and tables may be out of sync with the manuscript!
-----------------------------------------------------------------
 
@@
 study <- FinlandWTCStudy$new(context=context, response="canis.lupus")
-study$studyArea$loadBoundary(thin=T, tolerance=0.001)
+study$studyArea$loadBoundary(thin=F, tolerance=0.001)
 boundaryDF <- ggplot2::fortify(study$studyArea$boundary)
@@
 
-Survey routes (Figure 1)
-------------------------
+Figure 1: Survey routes
+-----------------------
 
-```{r survey-routes}
+```{r survey-routes, fig.height=double_height}
 surveyRoutes <- study$loadSurveyRoutes(findLengths=FALSE)
@@
 surveyRoutesDF <- ggplot2::fortify(surveyRoutesSP)
-p <- ggplot(boundaryDF, aes(long, lat, group=group)) + geom_polygon(colour="darkgray", fill=NA) +
-  geom_polygon(data=surveyRoutesDF, aes(long, lat, group=group), colour="black", fill=NA) +
+p <- ggplot(boundaryDF, aes(long, lat, group=group)) + geom_polygon(colour="black", fill=NA) +
+  geom_polygon(data=surveyRoutesDF, aes(long, lat, group=group), colour="blue", fill=NA) +
   coord_equal() + theme_raster()
@@
 #saveFigure(p, filename="SurveyRoutes.svg", bg="transparent")
+
 ```
@@
 
-Crossing rates (D > 1 removed) (Figure 2)
------------------------------------------
+Figure 2: Crossing rates (D > 1 removed)
+----------------------------------------
 
-```{r crossing-rates, fig.height=5}
+```{r crossing-rates, fig.width=double_width}
 getIntersectionsRate <- function(responses, context) {
@@
 intersectionsRate <- getIntersectionsRate(responses, context)
-p <- ggplot(intersectionsRate, aes(year, intersections)) + geom_line(size=1) + facet_grid(~response) +
-  xlab("Year") + ylab("Crossings rate") + theme_presentation(20)
+p <- ggplot(intersectionsRate, aes(year, intersections)) + geom_line(size=1) + facet_grid(~response, scales="free_y") +
+  xlab("Year") + ylab("Crossings rate") + theme_presentation(base_size)
 print(p)
@@
 
-GPS movements (Figure 3)
-------------------------
+Figure 3: GPS movements, juvenile is no more than 1 years old individual
+------------------------------------------------------------------------
 
-```{r gps-movements, fig.height=10, fig.width=20}
+```{r gps-movements, fig.width=double_width}
 getTracks <- function(responses, context) {
@@
     x <- tracksSP@data
-    if (!"sex" %in% colnames(x)) x$sex <- NA
-    x$juvenile <- if ("born" %in% colnames(x)) factor(ifelse(x$year - x$born <= 1, "juvenile", "adult")) else factor(NA)
+    if (!"sex" %in% colnames(x)) x$sex <- "unknown"
+    x$juvenile <- if ("born" %in% colnames(x)) factor(ifelse(x$year - x$born <= 1, "juvenile", "adult")) else factor("unknown")
+    x$juvenile <- factor(x$juvenile, levels=c(levels(x$juvenile), "unknown"))
+    x$juvenile[is.na(x$juvenile)] <- "unknown"
+
     x$id <- sapply(tracksSP@lines, function(x) x@ID)
@@
   geom_polygon(data=boundaryDF, aes(long, lat, group=group), colour="darkgray", fill=NA) +
-  geom_path(aes(colour=sexjuv, linetype=sexjuv), alpha=.5) +
-  scale_color_manual("Sex, age", values=c("darkred","red","red","darkblue","blue","blue","black")) + scale_linetype_manual("Sex, age", values=c("solid","dashed","solid","solid","dashed","solid","solid")) +
-  facet_grid(~response) + theme_raster(18, legend.position="bottom", aspect.ratio=1) + coord_cartesian(ylim=range(tracks$lat)+c(-1,1)*50000)
+  geom_path(aes(colour=sexjuv, linetype=sexjuv), alpha=.7) +
+  scale_color_manual("Sex, age", values=c("darkred","red","red","darkblue","blue","blue","black")) + scale_linetype_manual("Sex, age", values=c("solid","dotted","solid","solid","dotted","solid","solid")) +
+  facet_grid(~response) + theme_raster(base_size, aspect.ratio=1, legend.position=c(0.5,-0.1), legend.direction="horizontal", legend.position="bottom") +
+  guides(colour=guide_legend(nrow=2)) + coord_cartesian(ylim=range(tracks$lat)+c(-1,1)*50000)
 print(p)
@@
 
-Straight-line-distance-error (Figure 4)
----------------------------------------
+Figure 4: Straight-line-distance-error
+--------------------------------------
 
-```{r straight-line-distance-error, fig.height=5}
+```{r straight-line-distance-error-init}
 ss <- SimulationStudy$new(response="Intensive")$setup(context=context)
@@
 plotTracks$dt <- factor(plotTracks$dt / 60)
+```
 
+```{r straight-line-distance-error}
 segment <- aes(x=x, y=y, xend=c(tail(x, n=-1), NA), yend=c(tail(y, n=-1), NA), group=dt, colour=dt)
@@
 p <- ggplot(plotTracks, mapping=segment) +
-  theme_raster(20, legend.position=c(0.5,0.85), legend.background=element_rect(color="grey")) + 
+  theme_raster(base_size, legend.position=c(0.5,0.85), legend.background=element_rect(color="grey")) + 
   #guides(colour=guide_legend(title=expression(paste(Delta, "t (min)")))) +
   guides(colour=guide_legend(title="Sampling interval (min)")) +
-  scale_colour_manual(values=c("gray70","gray40","gray10")) +
+  scale_colour_manual(values=c("steelblue1","steelblue3","steelblue4")) +
   #ggtitle("Straight-line distance error") +
-  geom_path(size=2, arrow=ending) + geom_point(size=4)
+  geom_path(size=1, arrow=ending) + geom_point(size=2)
 #geom_segment(size=2, arrow=ending) # doesn't work with grouping
@@
 
-Power law fits for distance - sampling interval (Figure 5)
-----------------------------------------------------------
+Figure 5: Power law fits for distance - sampling interval
+---------------------------------------------------------
 
-```{r distance-fits, fig.height=5}
+```{r distance-fits, fig.width=double_width}
 distances <- ldply(responses, function(response, context) {
@@
   ylab("Distance / day (km)") + xlab("Sampling interval (h)") +
-  geom_line(data=intervals, aes(intervalH, distanceKm), color="black", size=1) +
+  geom_line(data=intervals, aes(intervalH, distanceKm), size=1) +
   geom_smooth(data=intervals, aes(ymin=distanceKm25, ymax=distanceKm975), stat="identity") +
   ylim(c(0, max(distances$distanceKm))) +
-  facet_grid(~response) + theme_presentation()
+  facet_grid(~response) + theme_presentation(base_size)
 print(p)
@@
 
-Observed distances - estimated distances (Table 1)
---------------------------------------------------
+Table 1: Observed distances - estimated distances
+-------------------------------------------------
 
@@
 
-Power law fit estimates for distance - sampling interval (Table 1)
-------------------------------------------------------------------
-
-Fixed effects:
-* human population density
-* precipitation
-* snow cover
-* temperature
+Table 2: Power law fit estimates for distance - sampling interval
+-----------------------------------------------------------------
+* fixed_effect = human population density
+* fixed_effect1 = precipitation
+* fixed_effect2 = snow cover
+* fixed_effect3 = temperature
 
@@
 
-Crude distance estimates (Table 1)
-----------------------------------
+Table 3: Crude distance estimates
+---------------------------------
 
@@
 
-Habitat weights (Figure 6)
---------------------------
+Figure 6: Habitat weights
+-------------------------
 
-```{r habitat-weights, fig.height=5, results="hide"}
+```{r habitat-weights, results="hide", fig.width=double_width}
 getHabitatWeights <- function(responses, context) {
@@
 p <- ggplot(weights, aes(habitat, weights, fill=habitat)) +
-  geom_bar(stat="identity") + facet_grid(~response) + scale_fill_manual(values=c("gray40","gray40","gray40","gray40","gray40")) +
-  xlab("") + ylab("Weight") + theme_presentation(20, axis.text.x=element_text(angle=90, hjust=1))
+  geom_bar(stat="identity") + facet_grid(~response) + scale_fill_manual(values=c("#beaed4","#ffff99","#7fc97f","#fdc086","#386cb0")) +
+  xlab("") + ylab("Weight") + theme_presentation(base_size, axis.text.x=element_text(angle=90, hjust=1))
 print(p)
@@
 
-```{r population-size-init, fig.keep="none", eval=TRUE}
+```{r population-size-init, fig.keep="none", eval=T}
 plotObservations <- function(i, params) {
@@
 
-Total population size estimates (Figure 7)
-------------------------------------------
-
-```{r population-size, fig.height=5, eval=TRUE}
+Table 4: Adjusted total population size factors
+-----------------------------------------------
+```{r population-size-factor, eval=F}
 populationSize <- data.frame()
@@
   print(coef(x$match()))
-  if (response == "canis.lupus") y$"Transformed estimated" <- y$Estimated * .7
-  else if (response == "lynx.lynx") y$"Transformed estimated" <- y$Estimated / 2 *.5
-  else if (response == "rangifer.tarandus.fennicus") y$"Transformed estimated" <- y$Estimated / 2 * 1.1
 
-  if (response == "lynx.lynx") {
+  if (response == "canis.lupus") y$"Adjusted estimated" <- y$Estimated * .75
+  else if (response == "lynx.lynx") {
+    y$"Adjusted estimated" <- y$Estimated * .3
     y$Estimated <- y$Estimated / 2
   }
-  if (response == "rangifer.tarandus.fennicus") {
-    y$Estimated <- y$Estimated / 2
-    y$Validation <- with(y, approx(Year, Validation, n=nrow(y)))$y
+  else if (response == "rangifer.tarandus.fennicus") {
+    y$"Adjusted estimated" <- y$Estimated / 2
+    #y$Validation <- with(y, approx(Year, Validation, n=nrow(y)))$y
   }
@@
 }
+```
+
+Figure 7: Total population size estimates
+-----------------------------------------
+```{r population-size, fig.width=double_width, eval=F}
+l <- levels(populationSize$Year)
+s <- seq(1, length(l), by=2)
+breaks <- l[s] 
+labels <- l[s]
 
 x <- melt(populationSize, id.vars=c("Year","response"), variable.name="Variable")
-p <- ggplot(x, aes(Year, value, group=Variable, colour=Variable)) + geom_line(size=1) + facet_wrap(~response, scales="free_y") +
-  scale_colour_manual(values=c("steelblue","violetred1","steelblue1")) +
+p <- ggplot(x, aes(Year, value, group=Variable, colour=Variable, fill=Variable)) + facet_wrap(~response, scales="free_y") +
+  geom_bar(subset=.(Variable=="Validation"), stat="identity") + geom_line(subset=.(Variable!="Validation"), size=2)  +
+  scale_colour_manual(values=c("steelblue","violetred1","palegreen3")) +
+  scale_fill_manual(values=c(NA,NA,"palegreen3")) +
+  scale_x_discrete(breaks=breaks, labels=labels) +
   xlab("Year") + ylab("Population size") +
-  theme_presentation(20, axis.text.x=element_text(angle=90, hjust=1)) + theme(legend.position="bottom")
+  theme_presentation(base_size, axis.text.x=element_text(angle=90, hjust=1)) + theme(legend.position="bottom", strip.text.x=element_blank())
+
+#x <- melt(populationSize, id.vars=c("Year","response"), variable.name="Variable")
+#p <- ggplot(x, aes(Year, value, group=Variable, colour=Variable)) + geom_line(size=1) + facet_wrap(~response, scales="free_y") +
+#  scale_colour_manual(values=c("steelblue","violetred1","steelblue1")) +
+#  xlab("Year") + ylab("Population size") +
+#  theme_presentation(base_size, axis.text.x=element_text(angle=90, hjust=1)) + theme(legend.position="bottom")
 print(p)

### 2014-08-07 inst/simulation/manuscript.Rmd
...

---
title: "Manuscript"
output: html_document
---

```{r initialize}
library(parallel)
library(doMC)
registerDoMC(cores=detectCores())
library(WTC)
source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
library(ggplot2)
library(plyr)

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
scenarios <- c("A","B","C","D","E","F")
modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"))

boundaryDF <- ggplot2::fortify(FinlandWTCStudy(context=context, response="canis.lupus")$studyArea$boundary)
```

```{r track-data}
#scenario <- "A"
#iterations <- as.integer(1)

iterations <- as.integer(c(1,1,1,1,3,1))
i <- 1
trackData <- data.frame()
surveyRoutesData <- data.frame()

for (scenario in scenarios) {
  mss <- getMSS(scenario, nSurveyRoutes=500)
  study <- mss$study
  tracks <- study$loadTracks(iteration=iterations[i])
  intersections <- tracks$countIntersections(surveyRoutes=mss$surveyRoutes, save=F)

  tracks$tracks <- subset(tracks$tracks, year==2001 & yday<59)
  x <- tracks$toGGDF()
  x$scenario <- scenario
  x$observation <- F

  #plot(tracksSP, col="grey", xlim=c(3.4, 3.5)*1e6, ylim=c(7.0, 7.1)*1e6, asp=1)

  intersections$tracks$tracks <- subset(intersections$tracks$tracks, year==2001)
  y <- intersections$tracks$toGGDF()
  y$scenario <- scenario
  y$observation <- T
  levels(y$group) <- paste(levels(y$group), "2")
  trackData <- rbind(trackData, x, y)
  #plot(tracksSP, col="black", add=T)

  x <- intersections$surveyRoutes$toGGDF()
  x$scenario <- scenario    
  observedAt <- intersections$intersections$intersections[intersections$intersections$intersections$intersections > 0 & intersections$intersections$intersections$year == 2001,]$surveyRoute
  y <- data.frame(id=unique(x$id), observed=FALSE)
  if (length(observedAt) > 0) y[y$id %in% observedAt,]$observed <- TRUE
  x <- plyr::join(x, y, by="id")
  surveyRoutesData <- rbind(surveyRoutesData, x)

  #plot(intersections$surveyRoutes$surveyRoutes, col="blue", add=T)    
  i <- i + 1
}

trackData$scenario <- factor(trackData$scenario)
```

```{r figure-1}
habitatRaster <- crop(study$studyArea$habitat, extent(c(3.375, 3.4, 6.9, 6.925)*1e6))
z <- sampleRegular(habitatRaster, 10000, asRaster=TRUE)
habitatData <- as.data.frame(rasterToPoints(z))
#habitatData <- as.data.frame(rasterToPoints(habitatRaster))
colnames(habitatData) <- c("x","y","z")
habitatData$scenario <- factor("E", levels=scenarios)
habitatData2 <- habitatData
habitatData2$scenario <- factor("F", levels=scenarios)
habitatData <- rbind(habitatData, habitatData2)

habitatData$z <- factor(habitatData$z, levels=0:255)
col <- study$studyArea$habitat@legend@colortable
names(col) <- 0:255

p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z)) + scale_fill_manual(values=col) +
  geom_path(data=trackData, aes(long, lat, group=group, colour=observation)) + scale_colour_manual(values=c("darkgray","black")) +
  geom_path(data=surveyRoutesData, aes(long, lat, group=group)) +
  coord_fixed(ratio=1, xlim=c(3.375,3.4)*1e6, ylim=c(6.9,6.925)*1e6) +
  facet_wrap(~scenario) + #coord_equal()
  theme_raster() + theme(panel.border=element_rect(fill="transparent"))
print(p)
saveFigure(p, "manuscript-figure-1.png", width=8, height=4)
```

```{r figure-2}
# TODO: add densities

surveyLocations <- ddply(surveyRoutesData, .(group, scenario), function(x) data.frame(long=mean(x$long), lat=mean(x$lat), scenario=x$scenario[1], observed=x$observed[1]))[,c("long","lat","scenario","observed")]
surveyLocations$variable <- "Survey routes"

trueLocations <- subset(trackData, observation==T)
trueLocations <- ddply(trueLocations, .(burst, scenario), function(x) data.frame(long=mean(x$long), lat=mean(x$lat), scenario=x$scenario[1], observed=F))[,c("long","lat","scenario","observed")]
trueLocations$variable <- "Individual locations"

x <- rbind(surveyLocations, trueLocations)

template <- raster(extent(study$studyArea$boundary), nrows=12, ncols=6, crs=study$studyArea$proj4string)
template@extent@xmax <- template@extent@xmin + dim(template)[2] * 100 * 1e3
template@extent@ymax <- template@extent@ymin + dim(template)[1] * 100 * 1e3

p <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey") +
  geom_point(data=x, aes(long, lat, colour=observed)) + scale_color_manual(values=c("darkgrey", "black")) + 
  facet_grid(variable~scenario) + theme_raster() + coord_equal() +
  scale_x_continuous(breaks=seq(template@extent@xmin, template@extent@xmax, 1e5)) +
  scale_y_continuous(breaks=seq(template@extent@ymin, template@extent@ymax, 1e5)) +
  theme(panel.grid.major=element_line(colour="grey"))
print(p)
saveFigure(p, "manuscript-figure-2.png", width=8, height=4)
```

```{r figure-3}
validation <- Validation()
populationSizes.temporal <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)

p <- ggplot(populationSizes.temporal, aes(Observed, Estimated)) + geom_point() + stat_smooth(method="lm", colour="black") +
  geom_abline(colour="darkgrey") + facet_wrap(~Scenario, scales="free") +
  theme_minimal() + theme(panel.border=element_rect(fill="transparent"))
print(p)
saveFigure(p, "manuscript-figure-3.png", width=8, height=4)
```

```{r figure-4}
validation <- Validation()
populationSizes.spatial <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames)

```

### 2014-08-08 inst/simulation/manuscript.Rmd
...

@@
 ```{r figure-3}
-validation <- Validation()
-populationSizes.temporal <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+populationSizes <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)
 
-p <- ggplot(populationSizes.temporal, aes(Observed, Estimated)) + geom_point() + stat_smooth(method="lm", colour="black") +
+p <- ggplot(populationSizes, aes(Observed, Estimated)) + geom_point() + stat_smooth(method="lm", colour="black") +
   geom_abline(colour="darkgrey") + facet_wrap(~Scenario, scales="free") +
@@
 ```{r figure-4}
-validation <- Validation()
-populationSizes.spatial <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames)
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames)
+p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point() + facet_wrap(~Scenario) +
+  theme_minimal() + theme(panel.border=element_rect(fill="transparent"))
+print(p)
+saveFigure(p, "manuscript-figure-4.png", width=8, height=4)
+```
+
+```{r figure-5}
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+credibilityIntervals <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames)
 

### 2014-08-08 R/Validation.R

@@ -236,7 +236,7 @@ Validation <- setRefClass(
     
     getValidatedCredibilityIntervalsProportions = function(scenarios=c("A","B","C","D","E","F"), modelNames) {
       x <- ddply(expand.grid(scenario=scenarios, modelName=modelNames, stringsAsFactors=FALSE), .(scenario, modelName), function(x, probs, probsName) {
-        s <- getStudy(scenario=x$scenario, readHabitatIntoMemory=F, isTest=F)
+        s <- getStudy(scenario=x$scenario, isTest=F)
         validation <- Validation(study=s, populationSizeOverEstimate=populationSizeOverEstimate)
         x <- validation$getValidatedCredibilityIntervalsProportion(x$modelName, probs=probs)
         if (nrow(x) == 0) return(NULL)

### 2014-08-08 inst/simulation/manuscript.Rmd
...

@@
 
+```{r credible-intervals}
+
+```
+
 ```{r figure-5}
@@
 validation <- Validation(study=study)
-credibilityIntervals <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames)
-
+a <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames)
+b <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames, probs=c(.25, .75), probsName="50%")
+credibilityIntervals <- cbind(a, b)
+p <- ggplot(credibilityIntervals, aes(scenario, Proportion, colour=CredibleInterval)) + geom_point() +
+  theme_minimal() + theme(panel.border=element_rect(fill="transparent"))
+print(p)
+saveFigure(p, "manuscript-figure-5.png", width=4, height=4)
 ```

### 2014-08-08 R/Validation.R

@@ -234,15 +234,15 @@ Validation <- setRefClass(
       return(validationProportion)
     },
     
-    getValidatedCredibilityIntervalsProportions = function(scenarios=c("A","B","C","D","E","F"), modelNames) {
+    getValidatedCredibilityIntervalsProportions = function(scenarios=c("A","B","C","D","E","F"), modelNames, probs=c(.025, .975), probsName="95%") {
       x <- ddply(expand.grid(scenario=scenarios, modelName=modelNames, stringsAsFactors=FALSE), .(scenario, modelName), function(x, probs, probsName) {
         s <- getStudy(scenario=x$scenario, isTest=F)
         validation <- Validation(study=s, populationSizeOverEstimate=populationSizeOverEstimate)
         x <- validation$getValidatedCredibilityIntervalsProportion(x$modelName, probs=probs)
         if (nrow(x) == 0) return(NULL)
-        x$CredibleInterval <- probsName
+        x$Interval <- probsName
         return(x)
-      }, probs=c(.025, .975), probsName="95%", .parallel=T)
+      }, probs=probs, probsName=probsName, .parallel=T)
       
       return(x)
     }

### 2014-08-10 inst/simulation/manuscript.Rmd
...

--- 2014-08-07 inst/simulation/manuscript.Rmd
@@
 modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"))
+iterations <- as.integer(c(1,1,1,1,3,1))
 
@@
 
-```{r track-data}
+```{r figure-1-init}
 #scenario <- "A"
@@
 
-iterations <- as.integer(c(1,1,1,1,3,1))
 i <- 1
@@
 p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z)) + scale_fill_manual(values=col) +
-  geom_path(data=trackData, aes(long, lat, group=group, colour=observation)) + scale_colour_manual(values=c("darkgray","black")) +
+  geom_path(data=trackData, aes(long, lat, group=group, colour=observation)) +#, alpha=observation)) +
+  scale_colour_manual(values=c("gray40","black")) +# scale_alpha_manual(values=c(.5,1)) +
   geom_path(data=surveyRoutesData, aes(long, lat, group=group)) +
@@
   facet_wrap(~scenario) + #coord_equal()
-  theme_raster() + theme(panel.border=element_rect(fill="transparent"))
-print(p)
+  theme_raster(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
+#print(p)
 saveFigure(p, "manuscript-figure-1.png", width=8, height=4)
+
+getRaster <- function(habitatRaster, xlim, ylim, size=10000) {
+  x <- crop(habitatRaster, extent(c(xlim, ylim)))
+  y <- sampleRegular(x, size=size, asRaster=TRUE)
+  habitatData <- as.data.frame(rasterToPoints(y))
+  colnames(habitatData) <- c("x","y","z")
+  habitatData$z <- factor(habitatData$z, levels=0:255)
+  col <- habitatRaster@legend@colortable
+  names(col) <- 0:255
+  return(habitatData)
+}
+
+getPanel <- function(scenario, xlim=c(3.375,3.4)*1e6, ylim=c(6.9,6.925)*1e6, habitatRaster, trackData, surveyRoutesData) {
+  s <- scenario
+  p <- if (!missing(habitatRaster)) {
+    habitatData <- getRaster(habitatRaster=habitatRaster, xlim=xlim, ylim=ylim)
+    ggplot(habitatData) + geom_raster(aes(x, y, fill=z)) + scale_fill_manual(values=col) +
+      geom_path(data=subset(trackData, scenario==s), aes(long, lat, group=group, colour=observation))
+  }
+  else {
+     ggplot(subset(trackData, scenario==s)) +
+       geom_path(aes(long, lat, group=group, colour=observation))
+  }
+
+  p <- p + scale_colour_manual(values=c("gray40","black")) +
+    geom_path(data=subset(surveyRoutesData, scenario==s), aes(long, lat, group=group)) +
+    coord_fixed(ratio=1, xlim=xlim, ylim=ylim) +
+    theme_raster(base_size=10, panel.border=element_rect(fill="transparent"), plot.title=element_text(size=12)) +
+    ggtitle(s)
+
+  return(p)
+}
+
+library(gridExtra)
+
+p1 <- getPanel(scenario="A", trackData=trackData, surveyRoutesData=surveyRoutesData)
+p2 <- getPanel(scenario="B", trackData=trackData, surveyRoutesData=surveyRoutesData)
+p3 <- getPanel(scenario="C", trackData=trackData, surveyRoutesData=surveyRoutesData)
+p4 <- getPanel(scenario="D", trackData=trackData, surveyRoutesData=surveyRoutesData)
+p5 <- getPanel(scenario="E", habitatRaster=study$studyArea$habitat, trackData=trackData, surveyRoutesData=surveyRoutesData)
+p6 <- getPanel(scenario="F", habitatRaster=study$studyArea$habitat, trackData=trackData, surveyRoutesData=surveyRoutesData)
+p <- arrangeGrob(p1, p2, p3, p4, p5, p6, ncol=3, nrow=2)
+saveFigure(p, "manuscript-figure-1.png")
+```
+
+```{r figure-2-init}
+getSurveyRouteLocations = function(scenarios, modelName, iterations, years) {
+  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, years) {
+    message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
+    study <- getStudy(scenario=x$scenario)
+    estimates <- SimulatedSmoothModelSpatioTemporal(study=study, iteration=x$iteration)
+    estimates$modelName <- modelName
+    estimates <- study$loadEstimates(estimates)
+    surveyRoutes <- data.frame(estimates$getUnscaledObservationCoordinates(), Intersection=subset(estimates$data, year %in% years)$intersections>0)
+    surveyRoutes$Scenario <- x$scenario
+    surveyRoutes$Iteration <- x$iteration
+    return(surveyRoutes)
+  }, modelName=modelName, years=years)
+}
+
+getAgentLocations = function(scenarios, modelName, iterations, years) {
+  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, years) {
+    study <- getStudy(scenario=x$scenario)
+    tracks <- study$loadTracks(iteration=x$iteration)
+    tracks$tracks <- subset(tracks$tracks, year %in% years & yday<59)
+    intersections <- study$loadIntersections(iteration=x$iteration)
+
+    # If no intersection, randomized track can be anywhere else except on the survey route
+
+    observedBursts <- aaply(intersections$intersectionsMatrix, 2, function(x) if (any(x>0)) TRUE else FALSE)
+    observationTracks <- tracks$randomizeObservationDayTracks(days=1)
+
+    a <- ddply(observationTracks$tracks, .(id, year, burst), function(x, observationTracks) {
+      name <- with(x[1,], paste(id, year, burst))
+      if (observedBursts[name] == FALSE) return(data.frame(x=mean(x$x), y=mean(x$y), year=x$year))
+      NULL
+    }, observationTracks=observationTracks)[,c("x","y","year")]
+    a$Observed <- F
+
+    index <- intersections$intersections$intersections > 0 & intersections$intersections$year %in% years
+    if (all(index == FALSE)) b <- data.frame() else {
+      x <- intersections$intersections[index,]
+      b <- data.frame(coordinates(x), year=x$year)
+      b$Observed <- T
+    }
+    agentLocations <- rbind(a, b)
+    agentLocations$Scenario <- x$scenario
+    agentLocations$Model <- x$modelName
+
+    return(agentLocations)
+  }, modelName=modelName, years=years)
+}
+
+getEstimatedDensity = function(scenarios, modelName, iterations, year, sd=FALSE) {
+  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, year0, sd) {
+    message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
+    study <- getStudy(scenario=x$scenario)
+    estimates <- SimulatedSmoothModelSpatioTemporal(study=study, iteration=x$iteration)
+    estimates$modelName <- modelName
+    estimates <- study$loadEstimates(estimates)
+    estimates$collectEstimates()
+
+    estimates$data <- subset(estimates$data, year == year0)
+    density <- estimates$getPopulationDensity(getSD=sd)
+    density <- if (sd) density$sd else density$mean
+    if (is.null(density) || maxValue(density$rasterStack[[1]]) > 1) stop("Estimation failed.")
+    density <- as.data.frame(density$rasterStack[[1]], xy=T)
+    density$Scenario <- x$scenario
+    density$Iteration <- x$iteration
+    density$Model <- modelName
+    density$Year <- year0
+
+    return(density)
+  }, modelName=modelName, year0=year, sd=sd)
+}
 ```
@@
 ```{r figure-2}
-# TODO: add densities
+iterations <- as.integer(c(1,1,3,2,5,6))
+#iterations <- as.integer(c(1,1,3,2,5,7))
+surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios, modelName=modelNames[1], iterations, years=2001)
+surveyRoutes <- rbind(surveyRoutes[surveyRoutes$Intersection == FALSE,], surveyRoutes[surveyRoutes$Intersection == TRUE,])
+agents <- getAgentLocations(scenarios=scenarios, modelName=modelNames[1], iterations, years=2001)
+densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=2001, sd=F)
+densitiesScaled <- ddply(densities, .(scenario), function(x) { x$X2001 <- scale(x$X2001); return(x) })
+limits <- extent(study$studyArea$boundary)
 
-surveyLocations <- ddply(surveyRoutesData, .(group, scenario), function(x) data.frame(long=mean(x$long), lat=mean(x$lat), scenario=x$scenario[1], observed=x$observed[1]))[,c("long","lat","scenario","observed")]
-surveyLocations$variable <- "Survey routes"
+#p <- ggplot(boundaryDF) + geom_polygon(aes(long, lat, group=group), fill=terrain.colors(100)[1]) +
+p <- ggplot(densitiesScaled) + geom_raster(aes(x, y, fill=X2001)) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
+  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
+  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black")) + 
+  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
+  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
+  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal()
+print(p)
+saveFigure(p, "manuscript-figure-2a.png", width=8)
 
-trueLocations <- subset(trackData, observation==T)
-trueLocations <- ddply(trueLocations, .(burst, scenario), function(x) data.frame(long=mean(x$long), lat=mean(x$lat), scenario=x$scenario[1], observed=F))[,c("long","lat","scenario","observed")]
-trueLocations$variable <- "Individual locations"
-
-x <- rbind(surveyLocations, trueLocations)
-
-template <- raster(extent(study$studyArea$boundary), nrows=12, ncols=6, crs=study$studyArea$proj4string)
-template@extent@xmax <- template@extent@xmin + dim(template)[2] * 100 * 1e3
-template@extent@ymax <- template@extent@ymin + dim(template)[1] * 100 * 1e3
+gridTemplate <- raster(extent(study$studyArea$boundary), nrows=12, ncols=6, crs=study$studyArea$proj4string)
+gridTemplate@extent@xmax <- gridTemplate@extent@xmin + dim(gridTemplate)[2] * 100 * 1e3
+gridTemplate@extent@ymax <- gridTemplate@extent@ymin + dim(gridTemplate)[1] * 100 * 1e3
 
 p <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey") +
-  geom_point(data=x, aes(long, lat, colour=observed)) + scale_color_manual(values=c("darkgrey", "black")) + 
-  facet_grid(variable~scenario) + theme_raster() + coord_equal() +
-  scale_x_continuous(breaks=seq(template@extent@xmin, template@extent@xmax, 1e5)) +
-  scale_y_continuous(breaks=seq(template@extent@ymin, template@extent@ymax, 1e5)) +
-  theme(panel.grid.major=element_line(colour="grey"))
+  geom_point(data=surveyRoutes, aes(x, y, colour=Intersection), size=1) + scale_color_manual(values=c("darkgrey", "black")) +
+  scale_x_continuous(breaks=seq(gridTemplate@extent@xmin, gridTemplate@extent@xmax, 1e5), expand=c(0,0)) +
+  scale_y_continuous(breaks=seq(gridTemplate@extent@ymin, gridTemplate@extent@ymax, 1e5), expand=c(0,0)) +
+  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal() + theme(panel.grid.major=element_line(colour="grey"))
 print(p)
-saveFigure(p, "manuscript-figure-2.png", width=8, height=4)
+saveFigure(p, "manuscript-figure-2b.png", width=8)
+```
+
+```{r figure-3-test, eval=FALSE}
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+x <- validation$validateTemporalPopulationSize(modelNames[1])
+validation$summarizePopulationSize(x)
 ```
@@
 ```{r figure-3}
-validation <- Validation()
-populationSizes.temporal <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)
-
-p <- ggplot(populationSizes.temporal, aes(Observed, Estimated)) + geom_point() + stat_smooth(method="lm", colour="black") +
-  geom_abline(colour="darkgrey") + facet_wrap(~Scenario, scales="free") +
-  theme_minimal() + theme(panel.border=element_rect(fill="transparent"))
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+populationSizes <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)
+p <- ggplot(populationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point() + stat_smooth(method="lm", colour="black") +
+  scale_colour_manual("Model", values=c("black", "darkgray"), labels=c("ST", "T")) +
+  geom_abline(colour="darkgrey", linetype="longdash") + facet_wrap(~Scenario, scales="free") +
+  xlab("True population size") + ylab("Estimated population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
 print(p)
@@
 
+```{r figure-4-test, eval=FALSE}
+study <- getStudy(scenario="F")
+validation <- Validation(study=study)
+x <- validation$validateSpatialPopulationSize(modelNames[1])
+```
+
 ```{r figure-4}
-validation <- Validation()
-populationSizes.spatial <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames)
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames)
+p <- ggplot(populationCorrelations, aes(True, Correlation, colour=modelName)) + geom_point() +
+  scale_colour_manual("Model", values=c("black", "darkgray"), labels=c("ST", "T")) +
+  facet_wrap(~Scenario) + xlab("True population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
+print(p)
+saveFigure(p, "manuscript-figure-4.png", width=8, height=4)
+```
+
+```{r figure-5-init}
 
 ```
+
+```{r figure-5}
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+a <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames)
+b <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames, probs=c(.25, .75), probsName="50%")
+credibilityIntervals <- rbind(a, b)
+p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
+  scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("black", "darkgray"), labels=c("ST", "T")) +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
+  xlab("Scenario")
+print(p)
+saveFigure(p, "manuscript-figure-5.png", width=4, height=2)
+```

### 2014-08-10 inst/wtc-finland/manuscript.Rmd
...

@@
 ------------------------------------------------------------------------
+
+```{r gps-stats}
+library(plyr)
+response <- "canis.lupus"
+
+study <- FinlandWTCStudy$new(context=context, response=response)
+tracks <- FinlandWTCTracks$new(study=study)
+tracks$loadTracks()
+tracks$loadMetadata()
+xtabs(sex~born, tracks$metadata)
+
+library(reshape2)
+
+dcast(tracks$metadata, .~sex)
+dcast(tracks$metadata, .~born)
+dcast(tracks$metadata, .~dispersing)$"TRUE"
+dcast(tracks$metadata, .~homerange)$"TRUE"
+dcast(tracks$metadata, .~alpha)
+
+x <- ld(tracks$tracks)$dt/3600
+sort(x[x<=24])
+sort(x[x<=24])*3600
+
+tracksSP <- tracks$getSpatialLines(variables=.(id,year,burst))
+x <- tracksSP@data
+x$juvenile <- if ("born" %in% colnames(x)) factor(ifelse(x$year - x$born <= 1, "juvenile", "adult")) else factor("unknown")
+x$juvenile <- factor(x$juvenile, levels=c(levels(x$juvenile), "unknown"))
+x$juvenile[is.na(x$juvenile)] <- "unknown"
+dcast(x, .~juvenile)
+
+```
 

### 2014-08-10 R/Validation.R

@@ -68,6 +68,7 @@ Validation <- setRefClass(
         s <- getStudy(scenario=x$scenario, isTest=F)
         validation <- Validation(study=s, populationSizeOverEstimate=populationSizeOverEstimate)
         x <- validation$validateTemporalPopulationSize(x$modelName)
+        gc()
         if (nrow(x) == 0) return(NULL)
         return(x)
       }, .parallel=T)

### 2014-08-10 R/static_global_functions.R

@@ -90,7 +90,10 @@ theme_raster <- function(base_size=12, base_family="", ...) {
     axis.title.y=element_blank(),
     axis.ticks=element_blank(),
     strip.background=element_blank(),
-    plot.margin=unit(c(0,0,-1,-1), "lines"),
+    #plot.margin=unit(c(0,0,-1,-1), "lines"),
+    plot.margin=unit(c(0,0,-.5,-.5), "lines"),
+    axis.ticks.length=unit(0,"lines"),axis.ticks.margin=unit(0,"lines"),
+    #plot.margin=rep(unit(0,"null"),4),panel.margin=unit(0,"null"),axis.ticks.length=unit(0,"null"),axis.ticks.margin=unit(0,"null"),
     legend.position="none",
     ...
   )

### 2014-08-31 inst/simulation/manuscript.Rmd
...

@@
 scenarios <- c("A","B","C","D","E","F")
-modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"))
+modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"), paste("SmoothModel", "nbinomial", "ar1", sep = "-"))
 iterations <- as.integer(c(1,1,1,1,3,1))
@@
 populationSizes <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)
-p <- ggplot(populationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point() + stat_smooth(method="lm", colour="black") +
-  scale_colour_manual("Model", values=c("black", "darkgray"), labels=c("ST", "T")) +
+validation$summarizePopulationSize(populationSizes)
+
+p <- ggplot(populationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) + stat_smooth(method="lm", aes(colour=modelName), se=F) +
+  scale_colour_manual("Model", values=c("darkgray","black"), labels=c("T", "ST")) +
   geom_abline(colour="darkgrey", linetype="longdash") + facet_wrap(~Scenario, scales="free") +
@@
 
+```{r table-1}
+ddply(populationSizes, .(Scenario, modelName), function(x) {
+  y <- lm(Estimated~Observed, x)
+  data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2))
+})
+```
+
+```{r table-2}
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+failedEstimations <- validation$countFailedEstimations(modelNames=modelNames)
+failedEstimations$pFailed <- (1 - failedEstimations$nEstimated / 50) * 100
+failedEstimations
+```
+
 ```{r figure-4-test, eval=FALSE}
-study <- getStudy(scenario="F")
+study <- getStudy(scenario="A")
 validation <- Validation(study=study)
@@
 validation <- Validation(study=study)
-populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames)
-p <- ggplot(populationCorrelations, aes(True, Correlation, colour=modelName)) + geom_point() +
-  scale_colour_manual("Model", values=c("black", "darkgray"), labels=c("ST", "T")) +
-  facet_wrap(~Scenario) + xlab("True population size") +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
+populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames[1])
+
+p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
+  facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
 print(p)
-saveFigure(p, "manuscript-figure-4.png", width=8, height=4)
-```
-
-```{r figure-5-init}
-
+saveFigure(p, "manuscript-figure-4.png", width=4, height=4)
 ```
@@
 credibilityIntervals <- rbind(a, b)
+
 p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
-  scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("black", "darkgray"), labels=c("ST", "T")) +
+  scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
   theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
-  xlab("Scenario")
+  xlab("Scenario") + ylab("Proportion within CI")
 print(p)

### 2014-08-31 R/Model.R

@@ -152,6 +152,48 @@ Model <- setRefClass(
   )
 )
 
+FMPModel <- setRefClass(
+  Class = "FMPModel",
+  contains = "Model",
+  fields = list(
+  ),
+  methods = list(
+    setup = function(intersections, params) {
+      coordsScale <<- 1
+      offsetScale <<- 1
+      modelName <<- "FMPModel"
+      
+      data <<- intersections$getData()
+      data$response <<- data$intersections
+      locations <<- intersections$getCoordinates() * coordsScale
+      years <<- as.integer(sort(unique(data$year)))
+      
+      return(invisible(.self))      
+    },
+    
+    saveEstimates = function(fileName=getEstimatesFileName()) {
+      message("Saving result to ", fileName, "...")
+      save(locations, data, coordsScale, years, offsetScale, file=fileName)
+    },
+    
+    estimate = function(save=FALSE, fileName=getEstimatesFileName(), verbose=TRUE) {
+      if (save) saveEstimates(fileName=fileName)
+    },
+    
+    collectEstimates = function(observationWeights=1, predictionWeights=1) {
+      index <- 1:nrow(data)
+
+      data$fittedMean <<- data$intersections / getObservedOffset() * observationWeights
+      data$fittedSD <<- NA
+      observedOffset <- getObservedOffset()
+      
+      message("Fitted values sums all years:")
+      message("observed = ", sum(data$intersections))
+      message("estimated = ", sum(data$fittedMean * observedOffset))
+    }
+  )
+)
+
 SmoothModelTemporal <- setRefClass(
   Class = "SmoothModelTemporal",
   contains = "Model",
@@ -705,6 +747,27 @@ SmoothModelSpatioTemporal <- setRefClass(
       print(p)
       
       return(invisible(p))
+    }
+  )
+)
+
+SimulatedFMPModel <- setRefClass(
+  Class = "SimulatedFMPModel",
+  contains = "FMPModel",
+  fields = list(
+    iteration = "integer"
+  ),
+  methods = list(
+    getEstimatesFileIterations = function() {
+      if (inherits(study, "undefinedField") | length(modelName) == 0)
+        stop("Provide study and modelName parameters.")
+      return(study$context$getIterationIds(dir=study$context$scratchDirectory, name=modelName, response=study$response, region=study$studyArea$region, tag="(\\d+)"))
+    },
+    
+    getEstimatesFileName = function() {
[…]

### 2014-08-31 R/Study.R

@@ -53,6 +53,7 @@ SimulationStudy <- setRefClass(
     getModel = function(modelName, iteration) {
       estimates <- if (modelName == "SmoothModel-nbinomial-matern-ar1") SimulatedSmoothModelSpatioTemporal(study=.self, iteration=iteration)
       else if (modelName == "SmoothModel-nbinomial-ar1") SimulatedSmoothModelTemporal(study=.self, iteration=iteration)
+      else if (modelName == "FMPModel") SimulatedFMPModel(study=.self, iteration=iteration)
       else stop("Invalid model.")
       estimates$modelName <- modelName
       return(estimates)

### 2014-08-31 inst/simulation/estimate.R

@@ -13,7 +13,7 @@ estimateSpatioTemporal <- function(scenario, modelName, iteration, isTest=FALSE,
   if (quick) {
     stop("TODO")
     intersections <- study$loadIntersections(iteration=iteration)
-    model <- SimulatedSmoothModelSpatioTemporal$new(study=study, iteration=iteration)
+    model <- SimulatedSmoothModelSpatioTemporal(study=study, iteration=iteration)
     meshParams <- list(maxEdge=c(.1e6, .2e6), cutOff=.05e6, coordsScale=1e-6)
     modelParams <- list(family="nbinomial", offsetScale=1000^2, meshParams=meshParams, timeModel="ar1")
     model$setup(intersections=intersections, params=modelParams)
@@ -34,6 +34,10 @@ estimateSpatioTemporal <- function(scenario, modelName, iteration, isTest=FALSE,
     else if (modelName == "SmoothModel-nbinomial-ar1") {
       model <- SimulatedSmoothModelTemporal(study=study, iteration=iteration)
       modelParams <- list(family="nbinomial", offsetScale=1000^2, timeModel="ar1")
+    }
+    else if (modelName == "FMPModel") {
+      model <- SimulatedFMPModel(study=study, iteration=iteration)
+      modelParams <- NULL
     }
     else stop("Unknown model.")
     study$estimate(model=model, params=modelParams)

### 2014-08-31 inst/simulation/manuscript.Rmd
...

@@
 scenarios <- c("A","B","C","D","E","F")
-modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"), paste("SmoothModel", "nbinomial", "ar1", sep = "-"))
+modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"), paste("SmoothModel", "nbinomial", "ar1", sep = "-"), "FMPModel")
 iterations <- as.integer(c(1,1,1,1,3,1))
@@
 ```{r figure-3}
-study <- getStudy(scenario="A")
+study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
 validation <- Validation(study=study)

### 2014-08-31 inst/simulation/estimate.R

@@ -3,6 +3,7 @@
 # Run full:
 # ./parallel_r.py -t 1:50 -n 70 -l 10.0 -b ~/tmp/blacklist.txt -v ~/git/Winter-Track-Counts/inst/simulation/estimate.R notest A SmoothModel-nbinomial-matern-ar1
 # ./parallel_r.py -t 1:50 -n 70 -l 10.0 -b ~/tmp/blacklist.txt -v ~/git/Winter-Track-Counts/inst/simulation/estimate.R notest A SmoothModel-nbinomial-ar1
+# ./parallel_r.py -t 1:50 -n 70 -l 10.0 -b ~/tmp/blacklist.txt -v ~/git/Winter-Track-Counts/inst/simulation/estimate.R notest A FMPModel
 #
 # library(devtools); install_github("statguy/Winter-Track-Counts")
 

### 2014-08-31 inst/wtc-finland/population_size.R

@@ -22,7 +22,7 @@ for (i in 1:length(responses)) {
   model$offsetScale <- 1000^2 # TODO: quickfix
   
   #populationSize <- study$getPopulationSize(model=model, withHabitatWeights=TRUE, saveDensityPlots=TRUE, getSD=FALSE)
-  populationSize <- study$getPopulationSize(model=model, withHabitatWeights=TRUE, saveDensityPlots=FALSE)
+  populationSize <- study$getPopulationSize(model=model, withHabitatWeights=TRUE, saveDensityPlots=FALSE) # TODO: fix
   populationSize$loadValidationData()
   populationSize$plotPopulationSize()
   print(populationSize)

### 2014-09-01 inst/simulation/manuscript.Rmd
...

@@
     study <- getStudy(scenario=x$scenario)
-    estimates <- SimulatedSmoothModelSpatioTemporal(study=study, iteration=x$iteration)
+    estimates <- study$getModel(modelName=modelName, iteration=x$iteration) 
     estimates$modelName <- modelName
     estimates <- study$loadEstimates(estimates)
-    surveyRoutes <- data.frame(estimates$getUnscaledObservationCoordinates(), Intersection=subset(estimates$data, year %in% years)$intersections>0)
+    intersections <- subset(estimates$data, year %in% years)$intersections>0
+    surveyRoutes <- data.frame(estimates$getUnscaledObservationCoordinates(), Intersection=intersections)
     surveyRoutes$Scenario <- x$scenario
@@
     study <- getStudy(scenario=x$scenario)
-    estimates <- SimulatedSmoothModelSpatioTemporal(study=study, iteration=x$iteration)
+    estimates <- study$getModel(modelName=modelName, iteration=x$iteration)
     estimates$modelName <- modelName
@@
 
+```{r figure-2-test}
+iterations <- as.integer(c(1,1,3,2,5,6))
+year <- 2010
+surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios[1], modelName=modelNames[1], iterations=iterations[1], years=year)
+surveyRoutes <- rbind(surveyRoutes[surveyRoutes$Intersection == FALSE,], surveyRoutes[surveyRoutes$Intersection == TRUE,])
+agents <- getAgentLocations(scenarios=scenarios[1], modelName=modelNames[1], iterations=iterations[1], years=year)
+densityST <- getEstimatedDensity(scenarios[1], modelNames[1], iterations[1], year=year, sd=FALSE)
+densityFMP <- getEstimatedDensity(scenarios[1], modelNames[3], iterations[1], year=year, sd=FALSE)
+limits <- extent(study$studyArea$boundary)
+
+p <- ggplot(densityST) + geom_raster(aes_string(x="x", y="y", fill=paste0("X", year))) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
+  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
+  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black")) + 
+  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
+  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
+  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal()
+print(p)
+saveFigure(p, "tmp_compare_densities_ST.png", width=8)
+
+p <- ggplot(densityFMP) + geom_raster(aes_string(x="x", y="y", fill=paste0("X", year))) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
+  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
+  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black")) + 
+  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
+  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
+  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal()
+print(p)
+saveFigure(p, "tmp_compare_densities_FMP.png", width=8)
+```
+
 ```{r figure-2}
@@
 #iterations <- as.integer(c(1,1,3,2,5,7))
-surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios, modelName=modelNames[1], iterations, years=2001)
+year <- 2010
+yearLayer <- paste0("X", year)
+surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
 surveyRoutes <- rbind(surveyRoutes[surveyRoutes$Intersection == FALSE,], surveyRoutes[surveyRoutes$Intersection == TRUE,])
-agents <- getAgentLocations(scenarios=scenarios, modelName=modelNames[1], iterations, years=2001)
-densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=2001, sd=F)
-densitiesScaled <- ddply(densities, .(scenario), function(x) { x$X2001 <- scale(x$X2001); return(x) })
+agents <- getAgentLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
+densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, sd=F)
+# TODO: do not use scaling??
+densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer) { x[,yearLayer] <- scale(x[,yearLayer]); return(x) }, yearLayer=yearLayer)
 limits <- extent(study$studyArea$boundary)
@@
 #p <- ggplot(boundaryDF) + geom_polygon(aes(long, lat, group=group), fill=terrain.colors(100)[1]) +
-p <- ggplot(densitiesScaled) + geom_raster(aes(x, y, fill=X2001)) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
+p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
   geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
@@
 p <- ggplot(populationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) + stat_smooth(method="lm", aes(colour=modelName), se=F) +
-  scale_colour_manual("Model", values=c("darkgray","black"), labels=c("T", "ST")) +
+  scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
   geom_abline(colour="darkgrey", linetype="longdash") + facet_wrap(~Scenario, scales="free") +
@@
   y <- lm(Estimated~Observed, x)
-  data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2))
+  data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2))
 })

### 2014-09-01 R/Model.R

@@ -190,6 +190,24 @@ FMPModel <- setRefClass(
       message("Fitted values sums all years:")
       message("observed = ", sum(data$intersections))
       message("estimated = ", sum(data$fittedMean * observedOffset))
+      
+      
+      stat <- data.frame()
+      for (year in years) {
+        yearWhich <- data$year == year
+        yearIndex <- year - min(years) + 1
+        x <- data.frame(
+          Year=years[yearIndex],
+          Observed=sum(data$intersections[yearWhich]),
+          EstimatedAtObserved=sum(data$fittedMean[yearWhich] * observedOffset[yearWhich]),
+          ObservedScaled=sum(data$intersections[yearWhich] / observedOffset[yearWhich]),
+          EstimatedAtObservedScaled=sum(data$fittedMean[yearWhich]),
+          ObservedOffset=mean(observedOffset[yearWhich])
+        )
+        stat <- rbind(stat, x)
+      }
+      message("Year by year summary:")
+      print(stat)
     }
   )
 )
@@ -246,8 +264,8 @@ SmoothModelTemporal <- setRefClass(
         x <- data.frame(
           Year=years[yearIndex],
           Observed=sum(data$intersections[yearWhich]),
-          EstimatedAtObserved=sum(data$fittedMean[yearWhich] * observedOffset[yearWhich]),
-          ObservedScaled=sum(data$intersections[yearWhich] / observedOffset[yearWhich]),
+          EstimatedAtObserved=sum(data$fittedMean[yearWhich] * observedOffset[yearWhich] / offsetScale),
+          ObservedScaled=sum(data$intersections[yearWhich] / observedOffset[yearWhich] / offsetScale),
           EstimatedAtObservedScaled=sum(data$fittedMean[yearWhich]),
           ObservedOffset=mean(observedOffset[yearWhich])
         )
@@ -521,14 +539,14 @@ SmoothModelSpatioTemporal <- setRefClass(
             Year=years[yearIndex],
             
             Observed=sum(data$intersections[yearWhich]),
-            EstimatedAtObserved=sum(data$fittedMean[yearWhich] * observedOffset[yearWhich]),
+            EstimatedAtObserved=sum(data$fittedMean[yearWhich] * observedOffset[yearWhich] / offsetScale),
             EstimatedAtNodes=sum(node$mean[,yearIndex] * predictedOffset[,yearIndex]),
             
[…]

### 2014-09-01 inst/simulation/population_size.R

@@ -19,10 +19,9 @@ population_size <- function(scenario, modelName, iteration, isTest, otherTest=F)
   
   if (otherTest) {
     #iteration <- as.integer(50)
-    estimates <- SimulatedSmoothModelSpatioTemporal(study=study, iteration=iteration)
-    
+    estimates <- study$getModel(modelName=modelName, iteration=iteration)
     estimates <- study$loadEstimates(estimates=estimates)
-    estimates$collectEstimates(predictionWeights=estimates$getPredictedOffset())
+    estimates$collectEstimates()
     
     #estimates$collectHyperparameters()
     populationSize <- estimates$getPopulationSize(withHabitatWeights=mss$hasHabitatWeights())

### 2014-09-03 inst/simulation/manuscript.Rmd
...

@@
 
-```{r figure-3}
+```{r figure-3-init}
 study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
@@
 
-p <- ggplot(populationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) + stat_smooth(method="lm", aes(colour=modelName), se=F) +
+meanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
+  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
+})
+logPopulationSizes <- populationSizes
+logPopulationSizes$Observed <- log10(logPopulationSizes$Observed + 1)
+logPopulationSizes$Estimated <- log10(logPopulationSizes$Estimated + 1)
+logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
+  data.frame(True=mean(log10(x$Observed+1)), Estimated=mean(log10(x$Estimated+1)), Scenario=x$scenario)
+})
+```
+
+```{r figure-3a}
+p <- ggplot(populationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
+  stat_smooth(method="lm", aes(colour=modelName), se=F) +
   scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
-  geom_abline(colour="darkgrey", linetype="longdash") + facet_wrap(~Scenario, scales="free") +
+  geom_abline(colour="darkgrey", linetype="longdash") +
+  geom_point(data=meanPopulationSizes, aes(True, Estimated, colour=modelName), size=2, shape=15) +
+  facet_wrap(~Scenario, scales="free") +
   xlab("True population size") + ylab("Estimated population size") +
@@
 print(p)
-saveFigure(p, "manuscript-figure-3.png", width=8, height=4)
+saveFigure(p, "manuscript-figure-3a.png", width=8, height=4)
+```
+
+```{r figure-3b}
+p <- ggplot(logPopulationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
+  stat_smooth(method="lm", aes(colour=modelName), se=F) +
+  scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
+  geom_abline(colour="darkgrey", linetype="longdash") + 
+  geom_point(data=logMeanPopulationSizes, aes(True, Estimated, colour=modelName), size=2, shape=15) +
+  facet_wrap(~Scenario, scales="free") +
+  xlab("True population size") + ylab("Estimated population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
+print(p)
+saveFigure(p, "manuscript-figure-3b.png", width=8, height=4)
 ```
@@
 })
+meanPopulationSizes
 ```

### 2014-09-05 inst/simulation/manuscript.Rmd
...

@@
 
-```{r figure-3-init}
+```{r figure-3-init-200}
 study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
@@
 saveFigure(p, "manuscript-figure-3b.png", width=8, height=4)
+```
+
+```{r figure-3-init-2000}
+study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
+validation <- Validation(study=study)
+populationSizes2000 <- validation$getValidationTemporalPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
+validation$summarizePopulationSize(populationSizes2000)
+
+meanPopulationSizes2000 <- ddply(populationSizes2000, .(scenario, modelName), function(x) {
+  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
+})
+logPopulationSizes2000 <- populationSizes2000
+logPopulationSizes2000$Observed <- log10(logPopulationSizes2000$Observed + 1)
+logPopulationSizes2000$Estimated <- log10(logPopulationSizes2000$Estimated + 1)
+logMeanPopulationSizes2000 <- ddply(populationSizes2000, .(scenario, modelName), function(x) {
+  data.frame(True=mean(log10(x$Observed+1)), Estimated=mean(log10(x$Estimated+1)), Scenario=x$scenario)
+})
+```
+
+```{r figure-3c}
+p <- ggplot(populationSizes2000, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
+  stat_smooth(method="lm", aes(colour=modelName), se=F) +
+  scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
+  geom_abline(colour="darkgrey", linetype="longdash") +
+  geom_point(data=meanPopulationSizes2000, aes(True, Estimated, colour=modelName), size=2, shape=15) +
+  facet_wrap(~Scenario, scales="free") +
+  xlab("True population size") + ylab("Estimated population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
+print(p)
+saveFigure(p, "manuscript-figure-3c.png", width=8, height=4)
 ```

### 2014-09-05 inst/simulation/manuscript.Rmd
...

@@
 study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
-validation <- Validation(study=study)
+validation <- Validation(study=study, populationSizeOverEstimate=20000)
 populationSizes2000 <- validation$getValidationTemporalPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)

### 2014-09-05 R/Validation.R

@@ -73,8 +73,8 @@ Validation <- setRefClass(
         s <- getStudy(scenario=x$scenario, isTest=F)
         validation <- Validation(study=s, populationSizeOverEstimate=populationSizeOverEstimate)
         x <- validation$validateTemporalPopulationSize(x$modelName)
-        gc()
         if (is.null(x) || nrow(x) == 0) return(NULL)
+        gc()
         return(x)
       })
       

### 2014-09-13 inst/simulation/manuscript.Rmd
...

@@
 
-```{r table-1}
-ddply(populationSizes, .(Scenario, modelName), function(x) {
+```{r table-1a}
+x <- ddply(populationSizes, .(Scenario, modelName), function(x) {
   y <- lm(Estimated~Observed, x)
@@
 })
-meanPopulationSizes
+y <- unique(meanPopulationSizes)
+join(x, y)
+```
+
+```{r table-1b}
+x <- ddply(populationSizes2000, .(Scenario, modelName), function(x) {
+  y <- lm(Estimated~Observed, x)
+  data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2))
+})
+y <- unique(meanPopulationSizes2000)
+join(x, y)
 ```
@@
 
-```{r figure-4}
+```{r figure-4a}
 study <- getStudy(scenario="A")
@@
 print(p)
-saveFigure(p, "manuscript-figure-4.png", width=4, height=4)
+saveFigure(p, "manuscript-figure-4a.png", width=4, height=4)
+```
+
+```{r figure-4b}
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1])
+
+p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
+  facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
+print(p)
+saveFigure(p, "manuscript-figure-4b.png", width=4, height=4)
 ```

### 2014-09-16 inst/simulation/manuscript.Rmd
...

@@
 study <- getStudy(scenario="A")
-validation <- Validation(study=study)
+validation <- Validation(study=study, populationSizeOverEstimate=20000)
 populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1])
@@
 
-```{r figure-5}
+```{r figure-5a}
 study <- getStudy(scenario="A")
@@
 print(p)
-saveFigure(p, "manuscript-figure-5.png", width=4, height=2)
+saveFigure(p, "manuscript-figure-5a.png", width=4, height=2)
 ```
+
+```{r figure-5b}
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
+b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames, probs=c(.25, .75), probsName="50%")
+credibilityIntervals <- rbind(a, b)
+
+p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
+  scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
+  xlab("Scenario") + ylab("Proportion within CI")
+print(p)
+saveFigure(p, "manuscript-figure-5b.png", width=4, height=2)
+```

### 2014-09-16 inst/simulation/validate.R

@@ -18,8 +18,7 @@ if (isTest) {
 } else {
   nSamples <- 50
   populationSizeOverEstimate <- 2000
-  
-  if (nchar(scenario) > 1) populationSizeOverEstimate <- 20000
+  if (substr(scenario, 2, 9) == "combined") populationSizeOverEstimate <- 20000
 }
 
 mss <- getMSS(scenario=scenario, isTest=isTest)

### 2014-10-01 inst/simulation/manuscript.Rmd
..

@@
 
-```{r table-1a}
+```{r figure-3-init-10days}
+study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
+validation <- Validation(study=study, populationSizeOverEstimate=2000)
+populationSizes10days <- validation$getValidationTemporalPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
+validation$summarizePopulationSize(populationSizes10days)
+
+meanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
+  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
+})
+logPopulationSizes10days <- populationSizes10days
+logPopulationSizes10days$Observed <- log10(logPopulationSizes10days$Observed + 1)
+logPopulationSizes10days$Estimated <- log10(logPopulationSizes10days$Estimated + 1)
+logMeanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
+  data.frame(True=mean(log10(x$Observed+1)), Estimated=mean(log10(x$Estimated+1)), Scenario=x$scenario)
+})
+```
+
+```{r figure-3d}
+p <- ggplot(populationSizes10days, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
+  stat_smooth(method="lm", aes(colour=modelName), se=F) +
+  scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
+  geom_abline(colour="darkgrey", linetype="longdash") +
+  geom_point(data=meanPopulationSizes10days, aes(True, Estimated, colour=modelName), size=2, shape=15) +
+  facet_wrap(~Scenario, scales="free") +
+  xlab("True population size") + ylab("Estimated population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
+print(p)
+saveFigure(p, "manuscript-figure-3d.png", width=8, height=4)
+```
+
+```{r table-1abc}
+prettyModelName <- function(x) {
+  lookup <- data.frame(x=c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "FMPModel") , y=c("ST", "S", "I"), stringsAsFactors=F)
+  lookup$y[match(x, lookup$x)]
+}
+
+getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, full=F, file) {
+  library(plyr)
   x <- ddply(populationSizes, .(Scenario, modelName), function(x) {
@@
   y <- unique(meanPopulationSizes)
-join(x, y)
-```
+  y$Scenario <- y$scenario
+  y$EstDev <- y$Estimated / y$True
+  z <- plyr::join(x, y)
+  z$Model <- prettyModelName(z$modelName)
+  z$Scenario <- substr(z$Scenario, 1, 1)
+  if (full) return(z)
 
-```{r table-1b}
-x <- ddply(populationSizes2000, .(Scenario, modelName), function(x) {
-  y <- lm(Estimated~Observed, x)
-  data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2))
-})
-y <- unique(meanPopulationSizes2000)
-join(x, y)
+  z <- z[,c("Model","Scenario","Slope","EstDev","R2")]
+  z$EstDev <- round(z$EstDev, 2)
+  if (!missing(file)) write.csv(z, file=file)
+  return(z)
+}
+
+dir <- normalizePath(context$figuresDirectory)
+getPopulationSizesTable(populationSizes, meanPopulationSizes, file=file.path(dir, "table-1a.csv"))
+getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, file=file.path(dir, "table-1b.csv"))
+getPopulationSizesTable(populationSizes10days, meanPopulationSizes10days, file=file.path(dir, "table-1c.csv"))
 ```
@@
 
+```{r figure-4c}
+study <- getStudy(scenario="A")
+validation <- Validation(study=study, populationSizeOverEstimate=2000)
+populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1])
+
+p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
+  facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
+print(p)
+saveFigure(p, "manuscript-figure-4c.png", width=4, height=4)
+```
+
 ```{r figure-5a}
@@
 ```
+
+```{r figure-5c}
+study <- getStudy(scenario="A")
+validation <- Validation(study=study)
+a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
+b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames, probs=c(.25, .75), probsName="50%")
+credibilityIntervals <- rbind(a, b)
+
+p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
+  scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
+  xlab("Scenario") + ylab("Proportion within CI")
+print(p)
+saveFigure(p, "manuscript-figure-5c.png", width=4, height=2)
+```

### 2014-10-01 R/SpatioTemporalRaster.R

@@ -72,6 +72,31 @@ SpatioTemporalRaster <- setRefClass(
       return(invisible(p))
     },
     
+    fill = function(z, templateRaster=study$getTemplateRaster(), layerNames, boundary, weights=1) {
+      library(raster)
+      library(plyr)
+      
+      rasterStack <<- stack(llply(1:length(z),
+                                  function(i, z, templateRaster) {
+                                    r <- raster(templateRaster)
+                                    r[] <- z[i]
+                                    return(r)
+                                  },
+                                  z=z, templateRaster=templateRaster))
+      
+      if (!missing(boundary) & !is.null(boundary)) {
+        rasterStack <<- stack(llply(1:nlayers(rasterStack),
+                                    function(i, rasterStack, boundary) {
+                                      message("Cropping ", i, "/", nlayers(rasterStack), "...")
+                                      mask(rasterStack[[i]], boundary)
+                                    },
+                                    rasterStack=rasterStack, boundary=boundary, .parallel=T))
+      }
+      weight(weights)
+      if (!missing(layerNames)) names(rasterStack) <<- layerNames
+      return(invisible(.self))      
+    },
+    
     interpolate = function(xyzt, timeVariable=colnames(xyzt)[4], templateRaster=study$getTemplateRaster(), transform=identity, inverseTransform=identity, layerNames, boundary, weights=1) {
       library(ST)
       library(raster)

### 2014-10-02 inst/simulation/manuscript.Rmd
...

--- 2014-09-16 inst/simulation/manuscript.Rmd
@@
 iterations <- as.integer(c(1,1,1,1,3,1))
+populationSizeOverEstimate <- 1e100
 
@@
 study <- getStudy(scenario="A")
-validation <- Validation(study=study)
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
 x <- validation$validateTemporalPopulationSize(modelNames[1])
@@
 study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
-validation <- Validation(study=study)
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
 populationSizes <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)
@@
 logPopulationSizes <- populationSizes
-logPopulationSizes$Observed <- log10(logPopulationSizes$Observed + 1)
-logPopulationSizes$Estimated <- log10(logPopulationSizes$Estimated + 1)
+logPopulationSizes$Observed <- logPopulationSizes$Observed + 1
+logPopulationSizes$Estimated <- logPopulationSizes$Estimated + 1
 logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
-  data.frame(True=mean(log10(x$Observed+1)), Estimated=mean(log10(x$Estimated+1)), Scenario=x$scenario)
+  data.frame(True=mean(x$Observed+1), Estimated=mean(x$Estimated+1), Scenario=x$scenario)
 })
+#logPopulationSizes$Observed <- log10(logPopulationSizes$Observed + 1)
+#logPopulationSizes$Estimated <- log10(logPopulationSizes$Estimated + 1)
+#logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
+#  data.frame(True=mean(log10(x$Observed+1)), Estimated=mean(log10(x$Estimated+1)), Scenario=x$scenario)
+#})
 ```
 
-```{r figure-3a}
+```{r figure-3a, eval=FALSE}
 p <- ggplot(populationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
@@
 ```{r figure-3b}
-p <- ggplot(logPopulationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
-  stat_smooth(method="lm", aes(colour=modelName), se=F) +
-  scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
-  geom_abline(colour="darkgrey", linetype="longdash") + 
-  geom_point(data=logMeanPopulationSizes, aes(True, Estimated, colour=modelName), size=2, shape=15) +
+plotLogPopulationSize <- function(size, meanSize) {
+  p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
+    #geom_point(size=1) +
+    #stat_smooth(method="lm", aes(colour=modelName), se=F) +
+    geom_point(size=1, colour="darkgray") +
+    stat_smooth(method="lm", se=F, colour="black") +
+    #scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
+    #scale_colour_manual("Model", values=c("darkgray","black")) +
+    geom_abline(colour="black", linetype="longdash") + 
+    #geom_point(data=meanSize, aes(True, Estimated, colour=modelName), size=2, shape=15) +
+    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black") +
     facet_wrap(~Scenario, scales="free") +
+    scale_x_log10() + scale_y_log10() +
     xlab("True population size") + ylab("Estimated population size") +
@@
   print(p)
-saveFigure(p, "manuscript-figure-3b.png", width=8, height=4)
+  return(p)
+}
+
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), subset(logMeanPopulationSizes, modelName==modelNames[1]))
+saveFigure(p, "manuscript-figure-3b1.png", width=8, height=4)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[2]), subset(logMeanPopulationSizes, modelName==modelNames[2]))
+saveFigure(p, "manuscript-figure-3b2.png", width=8, height=4)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), subset(logMeanPopulationSizes, modelName==modelNames[3]))
+saveFigure(p, "manuscript-figure-3b3.png", width=8, height=4)
 ```
@@
 study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
-validation <- Validation(study=study, populationSizeOverEstimate=20000)
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
 populationSizes2000 <- validation$getValidationTemporalPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
@@
 logPopulationSizes2000 <- populationSizes2000
-logPopulationSizes2000$Observed <- log10(logPopulationSizes2000$Observed + 1)
-logPopulationSizes2000$Estimated <- log10(logPopulationSizes2000$Estimated + 1)
+logPopulationSizes2000$Observed <- (logPopulationSizes2000$Observed + 1)
+logPopulationSizes2000$Estimated <- (logPopulationSizes2000$Estimated + 1)
 logMeanPopulationSizes2000 <- ddply(populationSizes2000, .(scenario, modelName), function(x) {
-  data.frame(True=mean(log10(x$Observed+1)), Estimated=mean(log10(x$Estimated+1)), Scenario=x$scenario)
+  data.frame(True=mean((x$Observed+1)), Estimated=mean((x$Estimated+1)), Scenario=x$scenario)
 })
@@
 
-```{r figure-3c}
+```{r figure-3c, eval=FALSE}
 p <- ggplot(populationSizes2000, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
@@
 
-```{r table-1a}
+```{r figure-3-init-10days}
+study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
+populationSizes10days <- validation$getValidationTemporalPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
+validation$summarizePopulationSize(populationSizes10days)
+
+meanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
+  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
+})
+logPopulationSizes10days <- populationSizes10days
+logPopulationSizes10days$Observed <- (logPopulationSizes10days$Observed + 1)
+logPopulationSizes10days$Estimated <- (logPopulationSizes10days$Estimated + 1)
+logMeanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
+  data.frame(True=mean((x$Observed+1)), Estimated=mean((x$Estimated+1)), Scenario=x$scenario)
+})
+```
+
+```{r figure-3d, eval=FALSE}
+p <- ggplot(populationSizes10days, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
+  stat_smooth(method="lm", aes(colour=modelName), se=F) +
+  scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
+  geom_abline(colour="darkgrey", linetype="longdash") +
+  geom_point(data=meanPopulationSizes10days, aes(True, Estimated, colour=modelName), size=2, shape=15) +
+  facet_wrap(~Scenario, scales="free") +
+  xlab("True population size") + ylab("Estimated population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
+print(p)
+saveFigure(p, "manuscript-figure-3d.png", width=8, height=4)
+```
+
+```{r table-1abc}
+prettyModelName <- function(x) {
+  lookup <- data.frame(x=c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "FMPModel") , y=c("ST", "T", "I"), stringsAsFactors=F)
+  lookup$y[match(x, lookup$x)]
+}
+
+getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, full=F, file) {
+  library(plyr)
   x <- ddply(populationSizes, .(Scenario, modelName), function(x) {
@@
 study <- getStudy(scenario="A")
-validation <- Validation(study=study)
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
 failedEstimations <- validation$countFailedEstimations(modelNames=modelNames)
@@
 study <- getStudy(scenario="A")
-validation <- Validation(study=study)
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
 x <- validation$validateSpatialPopulationSize(modelNames[1])
@@
 study <- getStudy(scenario="A")
-validation <- Validation(study=study)
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
 populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames[1])
@@
 study <- getStudy(scenario="A")
-validation <- Validation(study=study, populationSizeOverEstimate=20000)
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
 populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1])
@@
 
+```{r figure-4c}
+study <- getStudy(scenario="A")
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
+populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1])
+
+p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
+  facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
+print(p)
+saveFigure(p, "manuscript-figure-4c.png", width=4, height=4)
+```
+
 ```{r figure-5a}
 study <- getStudy(scenario="A")
-validation <- Validation(study=study)
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
 a <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames)
@@
 study <- getStudy(scenario="A")
-validation <- Validation(study=study)
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
 a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
@@
 ```
+
+```{r figure-5c}
+study <- getStudy(scenario="A")
+validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
+a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
+b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames, probs=c(.25, .75), probsName="50%")
+credibilityIntervals <- rbind(a, b)
+
+p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
+  scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
+  xlab("Scenario") + ylab("Proportion within CI")
+print(p)
+saveFigure(p, "manuscript-figure-5c.png", width=4, height=2)
+```

### 2014-10-02 R/Study.R

@@ -109,7 +109,7 @@ SimulationStudy <- setRefClass(
       return(invisible(populationSize))
     },
     
-    getHabitatWeights = function(tracks, iteration) {
+    getHabitatWeights = function(tracks, iteration, save=TRUE) {
       if (withHabitatWeights == FALSE) return(NULL)
       
       habitatWeights <- CORINEHabitatWeights$new(study=.self)
@@ -119,7 +119,7 @@ SimulationStudy <- setRefClass(
       habitatSelection <- if (file.exists(fileName)) habitatPreferences$loadHabitatSelection()
       else {      
         if (missing(tracks)) tracks <- loadTracks(iteration=iteration)
-        tracks$getHabitatPreferences(habitatWeightsTemplate=habitatWeights, nSamples=30, save=TRUE)
+        tracks$getHabitatPreferences(habitatWeightsTemplate=habitatWeights, nSamples=30, save=save)
       }
       
       message("Processing habitat weights raster...")

### 2014-10-02 inst/simulation/population_size.R

@@ -39,6 +39,7 @@ population_size <- function(scenario, modelName, iteration, isTest, otherTest=F)
   else {
     estimates <- study$getModel(modelName=modelName, iteration=iteration)
     habitatWeights <- study$getHabitatWeights(iteration=iteration)
+    print(habitatWeights)
     populationSize <- study$getPopulationSize(estimates=estimates, habitatWeights=habitatWeights)
     return(invisible(populationSize))
   }

### 2014-10-02 inst/simulation/simulate_tracks.sh

@@ -25,4 +25,4 @@ for scenario in A B C D E F; do
   python "$exec_path"/parallel_r.py -t "$iterations" -n "$max_nodes" -l 10.0 -b ~/tmp/blacklist.txt -v ~/git/Winter-Track-Counts/inst/simulation/simulate.R notest "$scenario"
 done
 
-# ./simulate_tracks.sh ~/git/RParallelScreen/ 51:100 60
+# ./simulate_tracks.sh ~/git/RParallelScreen/ 101:150 60

### 2014-10-02 inst/simulation/test_template.R

@@ -15,9 +15,17 @@ scenario <- scenarios[6]
 modelName <- modelNames[2]
 mss <- getMSS(scenario=scenario, isTest=isTest)
 study <- mss$study
+context <- study$context
 
 estimates <- study$getModel(modelName=modelName, iteration=iteration)
 estimates <- study$loadEstimates(estimates=estimates)
 estimates$collectEstimates()
 data <- estimates$data
 offsetScale <- estimates$offsetScale
+
+#habitatWeights2 <- study$getHabitatWeights(iteration=iteration)
+populationSize <- study$getPopulationSize(estimates=estimates, habitatWeights=habitatWeights2)
+populationSize
+mean(populationSize$sizeData$Estimated)
+
+populationSizeOverEstimate <- 1e6

### 2014-10-11 inst/simulation/manuscript.Rmd
...

@@
 iterations <- as.integer(c(1,1,1,1,3,1))
-populationSizeOverEstimate <- 1e100
+populationSizeOverEstimate <- 1e999
+
+prettyModelName <- function(x) {
+  lookup <- data.frame(x=c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "FMPModel") , y=c("ST", "T", "I"), stringsAsFactors=F)
+  lookup$y[match(x, lookup$x)]
+}
 
@@
 
-```{r figure-3-test, eval=FALSE}
-study <- getStudy(scenario="A")
+```{r figure-3-init}
+getPopulationSizes <- function(scenarios, modelNames, populationSizeOverEstimate=Inf) {
+  study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
   validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-x <- validation$validateTemporalPopulationSize(modelNames[1])
-validation$summarizePopulationSize(x)
+  populationSizes <- validation$getValidationTemporalPopulationSizes(scenarios=scenarios, modelNames=modelNames)
+  #validation$summarizePopulationSize(populationSizes)  
+}
+
+getMeanPopulationSizes <- function(populationSizes) {
+  meanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
+    data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
+  })
+  return(meanPopulationSizes)
+}
+
+getLogPopulationSizes <- function(populationSizes) {
+  logPopulationSizes <- populationSizes
+  logPopulationSizes$Observed <- logPopulationSizes$Observed + 1
+  logPopulationSizes$Estimated <- logPopulationSizes$Estimated + 1
+  return(logPopulationSizes)
+}
+
+getLogMeanPopulationSizes <- function(populationSizes) {  
+  logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
+    data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
+  })
+  return(logMeanPopulationSizes)
+}
+
+plotLogPopulationSize <- function(size, meanSize, modelName, xlim, ylim) {
+  library(ggplot2)
+  library(scales)
+  p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
+    #geom_point(size=1) +
+    #stat_smooth(method="lm", aes(colour=modelName), se=F) +
+    geom_point(size=1, colour="darkgray") +
+    stat_smooth(method="lm", se=F, colour="black") +
+    #scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
+    #scale_colour_manual("Model", values=c("darkgray","black")) +
+    geom_abline(colour="black", linetype="longdash") + 
+    #geom_point(data=meanSize, aes(True, Estimated, colour=modelName), size=2, shape=15) +
+    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black") +
+    facet_wrap(~Scenario, scales="free") +
+    coord_cartesian(xlim=xlim, ylim=ylim) +
+    scale_x_log10() + scale_y_log10() +
+    xlab("True population size") + ylab("Estimated population size") + ggtitle(paste("Model: ", prettyModelName(modelName[1]))) +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=12))
+  print(p)
+  return(p)
+}
 ```
 
-```{r figure-3-init-200}
+```{r figure-3-init}
+populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames)
+populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
+populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
+
+meanPopulationSizes  <- getMeanPopulationSizes(populationSizes)
+logPopulationSizes <- getLogPopulationSizes(populationSizes)
+logMeanPopulationSizes  <- getLogMeanPopulationSizes(populationSizes)
+
+meanPopulationSizes2000  <- getMeanPopulationSizes(populationSizes2000)
+logPopulationSizes2000 <- getLogPopulationSizes(populationSizes2000)
+logMeanPopulationSize2000  <- getLogMeanPopulationSizes(populationSizes2000)
+
+meanPopulationSizes10Days  <- getMeanPopulationSizes(populationSizes10Days)
+logPopulationSizes10Days <- getLogPopulationSizes(populationSizes10Days)
+logMeanPopulationSizes10Days  <- getLogMeanPopulationSizes(populationSizes10Days)
+#validation$summarizePopulationSize(populationSizes10Days)
+```
+
+```{r figure-3abc}
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), subset(logMeanPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], xlim=c(1,10000), ylim=c(1,10000))
+saveFigure(p, "manuscript-figure-3a1.png", width=8, height=4)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[2]), subset(logMeanPopulationSizes, modelName==modelNames[2]), modelName=modelNames[2], xlim=c(1,10000), ylim=c(1,10000))
+saveFigure(p, "manuscript-figure-3a2.png", width=8, height=4)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), subset(logMeanPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], xlim=c(1,10000), ylim=c(1,10000))
+saveFigure(p, "manuscript-figure-3a3.png", width=8, height=4)
+
+xlim <- c(100,10000)
+ylim <- xlim
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), subset(logMeanPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3b1.png", width=8, height=4)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[2]), subset(logMeanPopulationSizes2000, modelName==modelNames[2]), modelName=modelNames[2], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3b2.png", width=8, height=4)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), subset(logMeanPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3b3.png", width=8, height=4)
+
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), subset(logMeanPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], xlim=c(1,10000), ylim=c(1,10000))
+saveFigure(p, "manuscript-figure-3c1.png", width=8, height=4)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[2]), subset(logMeanPopulationSizes10Days, modelName==modelNames[2]), modelName=modelNames[2], xlim=c(1,10000), ylim=c(1,10000))
+saveFigure(p, "manuscript-figure-3c2.png", width=8, height=4)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), subset(logMeanPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], xlim=c(1,10000), ylim=c(1,10000))
+saveFigure(p, "manuscript-figure-3c3.png", width=8, height=4)
+```
+
+```{r table-1abc}
+getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, full=F, file) {
+  library(plyr)
+  x <- ddply(populationSizes, .(Scenario, modelName), function(x) {
+    y <- lm(Estimated~Observed, x)
+    data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2))
+  })
+  y <- unique(meanPopulationSizes)
+  y$Scenario <- y$scenario
+  y$EstDev <- y$Estimated / y$True
+  z <- plyr::join(x, y)
+  z$Model <- prettyModelName(z$modelName)
+  z$Scenario <- substr(z$Scenario, 1, 1)
+  if (full) return(z)
+
+  z <- z[,c("Model","Scenario","Slope","EstDev","R2")]
+  z$EstDev <- round(z$EstDev, 2)
+  if (!missing(file)) write.csv(z, file=file)
+  return(z)
+}
+
+dir <- normalizePath(context$figuresDirectory)
+getPopulationSizesTable(populationSizes, meanPopulationSizes, file=file.path(dir, "table-1a.csv"))
+getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, file=file.path(dir, "table-1b.csv"))
+getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, file=file.path(dir, "table-1c.csv"))
+```
+
+```{r, eval=FALSE}
 study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
@@
 logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
-  data.frame(True=mean(x$Observed+1), Estimated=mean(x$Estimated+1), Scenario=x$scenario)
+  data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
 })
@@
 
-```{r figure-3b}
-plotLogPopulationSize <- function(size, meanSize) {
-  p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
-    #geom_point(size=1) +
-    #stat_smooth(method="lm", aes(colour=modelName), se=F) +
-    geom_point(size=1, colour="darkgray") +
-    stat_smooth(method="lm", se=F, colour="black") +
-    #scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
-    #scale_colour_manual("Model", values=c("darkgray","black")) +
-    geom_abline(colour="black", linetype="longdash") + 
-    #geom_point(data=meanSize, aes(True, Estimated, colour=modelName), size=2, shape=15) +
-    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black") +
-    facet_wrap(~Scenario, scales="free") +
-    scale_x_log10() + scale_y_log10() +
-    xlab("True population size") + ylab("Estimated population size") +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
-  print(p)
-  return(p)
-}
-
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), subset(logMeanPopulationSizes, modelName==modelNames[1]))
-saveFigure(p, "manuscript-figure-3b1.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[2]), subset(logMeanPopulationSizes, modelName==modelNames[2]))
-saveFigure(p, "manuscript-figure-3b2.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), subset(logMeanPopulationSizes, modelName==modelNames[3]))
-saveFigure(p, "manuscript-figure-3b3.png", width=8, height=4)
-```
-
 ```{r figure-3-init-2000}
@@
 logMeanPopulationSizes2000 <- ddply(populationSizes2000, .(scenario, modelName), function(x) {
-  data.frame(True=mean((x$Observed+1)), Estimated=mean((x$Estimated+1)), Scenario=x$scenario)
+  data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
 })
+```
+
+```{r figure-3c123, eval=FALSE}
 ```
@@
 logMeanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
-  data.frame(True=mean((x$Observed+1)), Estimated=mean((x$Estimated+1)), Scenario=x$scenario)
+  data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
 })
+```
+
+```{r figure-3d123, eval=FALSE}
 ```
@@
 saveFigure(p, "manuscript-figure-3d.png", width=8, height=4)
-```
-
-```{r table-1abc}
-prettyModelName <- function(x) {
-  lookup <- data.frame(x=c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "FMPModel") , y=c("ST", "T", "I"), stringsAsFactors=F)
-  lookup$y[match(x, lookup$x)]
-}
-
-getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, full=F, file) {
-  library(plyr)
-  x <- ddply(populationSizes, .(Scenario, modelName), function(x) {
-    y <- lm(Estimated~Observed, x)
-    data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2))
-  })
-  y <- unique(meanPopulationSizes)
-  y$Scenario <- y$scenario
-  y$EstDev <- y$Estimated / y$True
-  z <- plyr::join(x, y)
-  z$Model <- prettyModelName(z$modelName)
-  z$Scenario <- substr(z$Scenario, 1, 1)
-  if (full) return(z)
-
-  z <- z[,c("Model","Scenario","Slope","EstDev","R2")]
-  z$EstDev <- round(z$EstDev, 2)
-  if (!missing(file)) write.csv(z, file=file)
-  return(z)
-}
-
-dir <- normalizePath(context$figuresDirectory)
-getPopulationSizesTable(populationSizes, meanPopulationSizes, file=file.path(dir, "table-1a.csv"))
-getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, file=file.path(dir, "table-1b.csv"))
-getPopulationSizesTable(populationSizes10days, meanPopulationSizes10days, file=file.path(dir, "table-1c.csv"))
 ```

### 2014-10-11 inst/simulation/population_size.sh

@@ -39,14 +39,14 @@ then
   python "$exec_path"/parallel_r.py -t "$iterations" -n "$max_nodes" -l 10.0 -b ~/tmp/blacklist.txt -v ~/git/Winter-Track-Counts/inst/simulation/validate.R notest "$scenario" "$model"
 fi
 
-# ./population_size.sh ~/git/RParallelScreen/ A 1:50 60 FMPModel 1
+# ./population_size.sh ~/git/RParallelScreen/ A 1:50 60 FMPModel
 # ./population_size.sh ~/git/RParallelScreen/ A 1:50 60 SmoothModel-nbinomial-ar1
 # ./population_size.sh ~/git/RParallelScreen/ A 1:50 60 SmoothModel-nbinomial-matern-ar1
 
-# ./population_size.sh ~/git/RParallelScreen/ Acombined 1:10 11 FMPModel 1
+# ./population_size.sh ~/git/RParallelScreen/ Acombined 1:10 11 FMPModel
 # ./population_size.sh ~/git/RParallelScreen/ Acombined 1:10 11 SmoothModel-nbinomial-ar1
 # ./population_size.sh ~/git/RParallelScreen/ Acombined 1:10 11 SmoothModel-nbinomial-matern-ar1
 
-# ./population_size.sh ~/git/RParallelScreen/ A10days 1:50 60 FMPModel 10
+# ./population_size.sh ~/git/RParallelScreen/ A10days 1:50 60 FMPModel
 # ./population_size.sh ~/git/RParallelScreen/ A10days 1:50 60 SmoothModel-nbinomial-ar1
 # ./population_size.sh ~/git/RParallelScreen/ A10days 1:50 60 SmoothModel-nbinomial-matern-ar1

### 2014-10-13 inst/simulation/manuscript.Rmd
...

@@
 
+```{r figure-4-init}
+getPopulationCorrelations <- function(modelName) {
+  study <- getStudy(scenario="A")
+  validation <- Validation(study=study, populationSizeOverEstimate=Inf)
+  populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelName)
+  return(populationCorrelations)
+}
+
+plotPopulationCorrelations <- function(populationCorrelations) {
+  p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
+  facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
+  return(p)
+}
+```
+
+```{r figure-4abc}
+pc1 <- getPopulationCorrelations(scenarios=scenarios, modelName=modelNames[1])
+pc2 <- getPopulationCorrelations(scenarios=paste0(scenarios, "combined"), modelName=modelNames[1])
+pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelName=modelNames[1])
+
+p1 <- plotPopulationCorrelations(p1)
+saveFigure(p1, "manuscript-figure-4a.png", width=4, height=4)
+plotPopulationCorrelations(p2)
+p2 <- saveFigure(p2, "manuscript-figure-4b.png", width=4, height=4)
+plotPopulationCorrelations(p3)
+p3 <- saveFigure(p3, "manuscript-figure-4c.png", width=4, height=4)
+```
+
 ```{r, eval=FALSE}
@@
 
-```{r figure-3-init-10days}
+```{r figure-3-init-10days, eval=FALSE}
 study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
@@
 
-```{r table-2}
+```{r table-2, eval=FALSE}
+# TODO: Combine with table 1
 study <- getStudy(scenario="A")

### 2014-10-13 R/Validation.R

@@ -103,7 +103,7 @@ Validation <- setRefClass(
         estimates <- study$loadEstimates(estimates)
         estimates$collectEstimates()
         
-        estimated <- estimates$getPopulationDensity(templateRaster=template, maskPolygon=NULL, getSD=F)
+        estimated <- estimates$getPopulationDensityInterpolate(templateRaster=template, maskPolygon=NULL, getSD=F)
         if (is.null(estimated$mean)) {
           message("Estimation failed for iteration ", iteration, " scenario ", study$response, ".")
           next

### 2014-10-13 inst/simulation/manuscript.Rmd
...

@@
 ```{r figure-4-init}
-getPopulationCorrelations <- function(modelName) {
+getPopulationCorrelations <- function(scenarios, modelNames) {
   study <- getStudy(scenario="A")
   validation <- Validation(study=study, populationSizeOverEstimate=Inf)
-  populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelName)
+  populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=scenarios, modelNames=modelNames)
   return(populationCorrelations)
@@
 ```{r figure-4abc}
-pc1 <- getPopulationCorrelations(scenarios=scenarios, modelName=modelNames[1])
-pc2 <- getPopulationCorrelations(scenarios=paste0(scenarios, "combined"), modelName=modelNames[1])
-pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelName=modelNames[1])
+pc1 <- getPopulationCorrelations(scenarios=scenarios, modelNames=modelNames[1])
+pc2 <- getPopulationCorrelations(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1])
+pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1])
 
@@
 saveFigure(p1, "manuscript-figure-4a.png", width=4, height=4)
-plotPopulationCorrelations(p2)
-p2 <- saveFigure(p2, "manuscript-figure-4b.png", width=4, height=4)
-plotPopulationCorrelations(p3)
-p3 <- saveFigure(p3, "manuscript-figure-4c.png", width=4, height=4)
+p2 <- plotPopulationCorrelations(p2)
+saveFigure(p2, "manuscript-figure-4b.png", width=4, height=4)
+p3 <- plotPopulationCorrelations(p3)
+saveFigure(p3, "manuscript-figure-4c.png", width=4, height=4)
+```
+
+```{r figure-5-init}
+getCIProportions <- function(scenarios, modelNames) {
+  study <- getStudy(scenario="A")
+  validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
+  a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames)
+  b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames, probs=c(.25, .75), probsName="50%")
+  credibilityIntervals <- rbind(a, b)
+  return(credibilityIntervals)  
+}
+
+plotCIProportions <- function(credibilityIntervals) {
+  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
+    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
+    xlab("Scenario") + ylab("Proportion within CI")
+  return(p)
+}
+```
+
+```{r figure-5abc}
+cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames)
+cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
+cip2 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
+
+p1 <- plotCIProportions(cip1)
+saveFigure(p1, "manuscript-figure-5a.png", width=4, height=2)
+p2 <- plotCIProportions(cip2)
+saveFigure(p2, "manuscript-figure-5b.png", width=4, height=2)
+p3 <- plotCIProportions(cip3)
+saveFigure(p3, "manuscript-figure-5c.png", width=4, height=2)
+
 ```

### 2014-10-13 R/Model.R

@@ -99,10 +99,11 @@ Model <- setRefClass(
       if (missing(index)) index <- 1:nrow(xy)
       x <- llply(posteriorSamples,
                  function(x, index) {
-                   predictorIndex <- grep("Predictor.", rownames(x$latent))[index]
-                   data.frame(x=xy[,1], y=xy[,2], z=exp(x$latent[predictorIndex]) / offsetScale, t=data$year)
+                   predictorIndex <- grep("Predictor\\.", rownames(x$latent)[index])
+                   #data.frame(x=xy[,1], y=xy[,2], z=exp(x$latent[predictorIndex]) / offsetScale, t=data$year)
+                   data.frame(x=xy[,1], y=xy[,2], z=exp(x$latent[predictorIndex]), t=data$year)
                  },
-                 index=index, .parallel=T)
+                 index=index, .parallel=F)
       return(x)
     },
            

### 2014-10-13 inst/simulation/validate.R

@@ -21,6 +21,7 @@ if (isTest) {
   if (substr(scenario, 2, 9) == "combined") populationSizeOverEstimate <- 20000
 }
 
+populationSizeOverEstimate <- Inf
 mss <- getMSS(scenario=scenario, isTest=isTest)
 study <- mss$study
 validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)

### 2014-10-14 inst/simulation/manuscript.Rmd
...

@@
 
-p1 <- plotPopulationCorrelations(p1)
+p1 <- plotPopulationCorrelations(pc1)
 saveFigure(p1, "manuscript-figure-4a.png", width=4, height=4)
-p2 <- plotPopulationCorrelations(p2)
+p2 <- plotPopulationCorrelations(pc2)
 saveFigure(p2, "manuscript-figure-4b.png", width=4, height=4)
-p3 <- plotPopulationCorrelations(p3)
+p3 <- plotPopulationCorrelations(pc3)
 saveFigure(p3, "manuscript-figure-4c.png", width=4, height=4)

### 2014-10-14 inst/simulation/validate.R

@@ -21,6 +21,7 @@ if (isTest) {
   if (substr(scenario, 2, 9) == "combined") populationSizeOverEstimate <- 20000
 }
 
+nSamples <- 50
 populationSizeOverEstimate <- Inf
 mss <- getMSS(scenario=scenario, isTest=isTest)
 study <- mss$study

### 2014-10-14 inst/simulation/manuscript.Rmd
...

@@
 cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
-cip2 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
+cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
 
@@
 saveFigure(p3, "manuscript-figure-5c.png", width=4, height=2)
-
 ```

### 2014-10-14 inst/simulation/validate.R

@@ -12,15 +12,6 @@ source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
 parseArguments()
 modelName <- extraArgs[1]
 
-if (isTest) {
-  nSamples <- 50
-  populationSizeOverEstimate <- 200
-} else {
-  nSamples <- 50
-  populationSizeOverEstimate <- 2000
-  if (substr(scenario, 2, 9) == "combined") populationSizeOverEstimate <- 20000
-}
-
 nSamples <- 50
 populationSizeOverEstimate <- Inf
 mss <- getMSS(scenario=scenario, isTest=isTest)
@@ -54,26 +45,5 @@ if (isTest) {
   summary(lm(Estimated.y~Estimated.x, xy)) # OK !
   
 } else {
-  if (F) {
-    populationSize <- validation$validateTemporalPopulationSize(modelName=modelName)
-    validation$summarizePopulationSize(populationSize)
-    summary(lm(Estimated~Observed, populationSize))  
-    
-    spatialCorrelation <- validation$validateSpatialPopulationSize(modelName=modelName)
-    summary(lm(Correlation~True, spatialCorrelation))
-  }
-  
   populationSizeCI <- validation$validateCredibilityIntervals(modelName=modelName, iteration=as.integer(task_id), nSamples=nSamples, save=T)
-  
-  if (F) {
-    iterations <- validation$getCredibilityIntervalsValidationIterations(modelName=modelName)
-    populationSizeCI <- ldply(iterations, function(iteration) {
-      validation$loadCredibilityIntervalsValidation(modelName=modelName, iteration=iteration)
-    })
-    print(validation$summarizePopulationSizeCI(populationSizeCI))
-    print(validation$summarizePopulationSizeCI(populationSizeCI, probs=c(.25,.75)))
-  }
 }
-
-# Task 1 at ukko075.hpc.cs.helsinki.fi terminated with return code 0. 25 tasks left.
-# Task 42 at ukko076.hpc.cs.helsinki.fi terminated with return code 1. 24 tasks left.

### 2014-10-14 inst/simulation/manuscript.Rmd
...

@@
   study <- getStudy(scenario="A")
-  validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
+  validation <- Validation(study=study, populationSizeOverEstimate=Inf)
   a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames)

### 2014-10-14 R/Validation.R

@@ -253,7 +253,7 @@ Validation <- setRefClass(
       x <- summarizePopulationSizeCI(populationSizeCI, variables=.(scenario, Year, iteration), probs=probs)
       validationProportion <- ddply(x, .(scenario), function(x) {
         y <- with(x, Observed >= Estimated.q1 & Observed <= Estimated.q2)
-        data.frame(Proportion=mean(y), n=length(y))
+        data.frame(Proportion=mean(y), n=length(y), True=mean(x$Observed), Estimated=mean(x$Estimated), Estimated.q1=mean(x$Estimated.q1), Estimated.q2=mean(x$Estimated.q2))
       })
       validationProportion$modelName <- modelName
       

### 2014-10-28 inst/simulation/manuscript.Rmd
...

@@
     geom_point(size=1, colour="darkgray") +
-    stat_smooth(method="lm", se=F, colour="black") +
+#stat_smooth(method="lm", se=F, colour="black") +
     #scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
@@
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=12))
+
+  if (!any(abs(size$Estimated) == Inf)) p <- p + stat_smooth(method="lm", se=F, colour="black")
+
   print(p)
@@
 logPopulationSizes2000 <- getLogPopulationSizes(populationSizes2000)
-logMeanPopulationSize2000  <- getLogMeanPopulationSizes(populationSizes2000)
+logMeanPopulationSizes2000  <- getLogMeanPopulationSizes(populationSizes2000)
 
@@
   x <- ddply(populationSizes, .(Scenario, modelName), function(x) {
+    if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA))
     y <- lm(Estimated~Observed, x)
-    data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2))
+    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2)))
   })

### 2014-10-28 R/Study.R

@@ -308,21 +308,17 @@ FinlandWTCStudy <- setRefClass(
       return(estimates)
     },
     
-    NEEDS_UPDATE_getPopulationDensity = function(model, withHabitatWeights=TRUE, saveDensityPlots=FALSE, getSD=FALSE) {
-      estimates <- collectEstimates(model)
+    getPopulationDensity = function(modelName, withHabitatWeights=TRUE, getSD=FALSE, saveDensityPlots) {
+      habitatWeightsRaster <- if (withHabitatWeights) loadHabitatWeightsRaster() else 1
+      model <- getModel(modelName=modelName)  
+      model$offsetScale <- 1000^2 # quickfix, remove when not needed anymore
+      model$loadEstimates()
+      model$collectEstimates()
+      populationDensity <- model$getPopulationDensity(templateRaster=habitatWeightsRaster, getSD=getSD)
       
-      habitatWeights <- if (withHabitatWeights) loadHabitatWeightsRaster() else HabitatWeights$new(study=study)$getWeightsRaster()
-      populationDensity <- estimates$getPopulationDensity(templateRaster=habitatWeights, getSD=getSD)
-
+      #habitatWeights <- if (withHabitatWeights) loadHabitatWeightsRaster() else HabitatWeights$new(study=study)$getWeightsRaster()
+      
       if (saveDensityPlots) {
-        populationDensity$mean$animate(name="PopulationDensity-mean")
-        if (getSD) populationDensity$sd$animate(name="PopulationDensity-sd")
-      }
-      
-      populationDensity$mean$weight(habitatWeights)
-      if (getSD) populationDensity$sd$weight(habitatWeights)
-      
-      if (saveDensityPlots & withHabitatWeights) {
         populationDensity$mean$animate(name="WeightedPopulationDensity-mean")
         if (getSD) populationDensity$sd$animate(name="WeightedPopulationDensity-sd")
       }

### 2014-10-28 inst/simulation/estimate.R

@@ -31,7 +31,7 @@ estimateSpatioTemporal <- function(scenario, modelName, iteration, isTest=FALSE,
     populationSize$plotPopulationSize()
   }
   else {
-    model <- study$getModel(modelName=modelName)
+    #model <- study$getModel(modelName=modelName)
     #modelParams <- study$getModelParams(modelName=modelName)
     
     if (modelName == "SmoothModel-nbinomial-matern-ar1") {

### 2014-10-28 inst/wtc-finland/estimate.R

@@ -52,5 +52,5 @@ if (test == "test") {
   
   responses <- c("canis.lupus", "lynx.lynx", "rangifer.tarandus.fennicus")
   response <- responses[task_id]
-  estimate(response=response, modelName)
+  estimate(response=response, modelName=modelName)
 }

### 2014-10-28 inst/wtc-finland/preprocess.R

@@ -1,5 +1,4 @@
 library(devtools)
-install_github("R-Cluster", user="statguy")
 install_github("R-Spatio-Temporal", user="statguy")
 install_github("Winter-Track-Counts", user="statguy")
 

### 2014-12-07 inst/simulation/manuscript.Rmd
...

@@
 ```{r figure-5-init}
-getCIProportions <- function(scenarios, modelNames) {
+getCIProportions <- function(scenarios, modelNames, panel, scenarioNames) {
   study <- getStudy(scenario="A")
@@
   credibilityIntervals <- rbind(a, b)
+  credibilityIntervals$panel <- panel
+  credibilityIntervals$scenario <- scenarioNames
   return(credibilityIntervals)  
@@
 plotCIProportions <- function(credibilityIntervals) {
-  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
+  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) +
+    geom_hline(aes(yintercept=.5), color="grey90") +
+    geom_hline(aes(yintercept=.95), color="grey90") +
+    geom_point() +
     scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
-    xlab("Scenario") + ylab("Proportion within CI")
+    xlab("Scenario") + ylab("Proportion within CI") +
+    facet_wrap(~panel) + ylim(0, 1) +
+    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank())
   return(p)
@@
 ```{r figure-5abc}
-cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames)
-cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
-cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
+cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames, panel="a", scenarioNames=scenarios)
+cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames, panel="b", scenarioNames=scenarios)
+cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames, panel="c", scenarioNames=scenarios)
+cip <- rbind(cip1, cip2, cip3)
+p <- plotCIProportions(cip)
+saveFigure(p, "manuscript-figure-5.png", width=6, height=2)
 
-p1 <- plotCIProportions(cip1)
-saveFigure(p1, "manuscript-figure-5a.png", width=4, height=2)
-p2 <- plotCIProportions(cip2)
-saveFigure(p2, "manuscript-figure-5b.png", width=4, height=2)
-p3 <- plotCIProportions(cip3)
-saveFigure(p3, "manuscript-figure-5c.png", width=4, height=2)
+#p1 <- plotCIProportions(cip1)
+#saveFigure(p1, "manuscript-figure-5a.png", width=4, height=2)
+#p2 <- plotCIProportions(cip2)
+#saveFigure(p2, "manuscript-figure-5b.png", width=4, height=2)
+#p3 <- plotCIProportions(cip3)
+#saveFigure(p3, "manuscript-figure-5c.png", width=4, height=2)
 ```

### 2014-12-07 R/Study.R

@@ -210,7 +210,7 @@ FinlandWTCStudy <- setRefClass(
     },
     
     loadSurveyRoutes = function(findLengths=TRUE) {
-      return(FinlandWTCSurveyRoutes$new(study=.self)$loadSurveyRoutes(findLengths=findLengths))
+      return(FinlandWTCSurveyRoutes$new(context=.self$context)$loadSurveyRoutes(findLengths=findLengths))
     },
     
     loadIntersections = function(predictDistances=TRUE) {

### 2015-01-07 inst/simulation/manuscript.Rmd
...

@@
 
-```{r figure-3-init}
+```{r figure-3-init-2}
+registerDoMC(cores=10) # The nodes seem to run out of memory and fail with a bogus error message, so it's better to reduce the number of used cores.
+
 populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames)
@@
 populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
+
+registerDoMC(cores=detectCores())
 
@@
   credibilityIntervals$panel <- panel
-  credibilityIntervals$scenario <- scenarioNames
+  credibilityIntervals$scenario <- rep(scenarioNames, times=2, each=2)
   return(credibilityIntervals)  
@@
 ```
+
+  p <- ggplot(cip2, aes(scenario, Proportion, shape=Interval, colour=modelName)) +
+    geom_hline(aes(yintercept=.5), color="grey90") +
+    geom_hline(aes(yintercept=.95), color="grey90") +
+    geom_point() +
+    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
+    xlab("Scenario") + ylab("Proportion within CI") +
+    facet_wrap(~panel) + ylim(0, 1) +
+    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank())
+saveFigure(p, "manuscript-figure-5.png", width=6, height=2)
 

### 2015-01-07 R/Validation.R

@@ -33,24 +33,37 @@ Validation <- setRefClass(
       library(plyr)
       
       iterations <- getPopulationSizeFileIterations(modelName)
-
+      
       populationSize <- ldply(iterations, function(iteration) {
         message("Processing iteration ", iteration, " of ", length(iterations), " iterations for scenario ", study$response, "...")
-        
         populationSize <- study$loadPopulationSize(iteration=iteration, modelName=modelName)
-        if (any(populationSize$sizeData$Estimated > populationSizeOverEstimate)) {
+        x <- populationSize$sizeData
+        
+        if (is.null(x) || nrow(x) == 0 || any(x$Estimated > populationSizeOverEstimate)) {
           message("Estimation failed for iteration ", iteration, " scenario ", study$response)
           return(NULL)
         }
-        x <- populationSize$sizeData
         x$iteration <- iteration
         return(x)
-      }, .parallel=T)
-      
+      }, .parallel=TRUE)
+            
       if (is.null(populationSize)) return(NULL)
       populationSize$Scenario <- study$response
       populationSize$modelName <- modelName
       return(populationSize)
+    },
+        
+    getValidationTemporalPopulationSizes = function(scenarios=c("A","B","C","D","E","F"), modelNames) {
+      x <- ddply(expand.grid(scenario=scenarios, modelName=modelNames, stringsAsFactors=FALSE), .(scenario, modelName), function(x) {
+        s <- getStudy(scenario=x$scenario, isTest=F)
+        validation <- Validation(study=s, populationSizeOverEstimate=populationSizeOverEstimate)
+        x <- validation$validateTemporalPopulationSize(x$modelName)
+        if (is.null(x) || nrow(x) == 0) return(NULL)
+        gc()
+        return(x)
+      })
+      
+      return(x)
     },
     
     summarizePopulationSize = function(populationSizes) {
@@ -66,19 +79,6 @@ Validation <- setRefClass(
         y$Observed.q2 <- q[2]
         return(y)
       }, probs=c(.0025, .975))
-    },
-    
-    getValidationTemporalPopulationSizes = function(scenarios=c("A","B","C","D","E","F"), modelNames) {
-      x <- ddply(expand.grid(scenario=scenarios, modelName=modelNames, stringsAsFactors=FALSE), .(scenario, modelName), function(x) {
-        s <- getStudy(scenario=x$scenario, isTest=F)
[…]

### 2015-01-10 inst/simulation/manuscript.Rmd
..

@@
 
+prettyScenario <- function(scenario) substr(scenario, 1, 1)
 prettyModelName <- function(x) {
@@
 ```{r figure-1}
-habitatRaster <- crop(study$studyArea$habitat, extent(c(3.375, 3.4, 6.9, 6.925)*1e6))
+#window <- c(3.375, 3.4, 6.9, 6.925)*1e6
+window <- (c(3.375, 3.4, 6.9, 6.925)+c(-.01,.01,-.01,.01)*3)*1e6
+habitatRaster <- crop(study$studyArea$habitat, extent(window))
 z <- sampleRegular(habitatRaster, 10000, asRaster=TRUE)
@@
 
-p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z)) + scale_fill_manual(values=col) +
-  geom_path(data=trackData, aes(long, lat, group=group, colour=observation)) +#, alpha=observation)) +
-  scale_colour_manual(values=c("gray40","black")) +# scale_alpha_manual(values=c(.5,1)) +
-  geom_path(data=surveyRoutesData, aes(long, lat, group=group)) +
-  coord_fixed(ratio=1, xlim=c(3.375,3.4)*1e6, ylim=c(6.9,6.925)*1e6) +
+p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z), alpha=.3) + scale_fill_manual(values=col) +
+  geom_path(data=trackData, aes(long, lat, group=group, colour=observation), alpha=1, size=.3) +#, alpha=observation)) +
+  scale_colour_manual(values=c("darkgray","black")) + # scale_alpha_manual(values=c(.5,1)) + # gray40
+  geom_path(data=surveyRoutesData, aes(long, lat, group=group), size=.3, colour="blue", alpha=.5) +
+  coord_fixed(ratio=1, xlim=window[1:2], ylim=window[3:4]) +
   facet_wrap(~scenario) + #coord_equal()
@@
 
+# scale on the map in km
+c((window[2]-window[1]), (window[4]-window[3]))/1000
+```
+
+```{r figure-1-separate, eval=FALSE}
 getRaster <- function(habitatRaster, xlim, ylim, size=10000) {
@@
 
-getEstimatedDensity = function(scenarios, modelName, iterations, year, sd=FALSE) {
+getEstimatedDensity = function(scenarios, modelName, iterations, year, sd=FALSE, log=FALSE) {
   ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, year0, sd) {
@@
 
+    studyArea <- FinlandStudyArea$new(context=context)$setup(tolerance=0.001)
+
     estimates$data <- subset(estimates$data, year == year0)
-    density <- estimates$getPopulationDensity(getSD=sd)
+    if (log) estimates$data$fittedMean <- log(estimates$data$fittedMean)
+    density <- estimates$getPopulationDensityInterpolate(getSD=sd, maskPolygon=studyArea$boundary)
     density <- if (sd) density$sd else density$mean
@@
 ```{r figure-2-test}
+iterations <- as.integer(c(1,1,3,2,5,6))
+year <- 2010
+densityST <- getEstimatedDensity(scenarios[1], modelNames[1], iterations[1], year=year, sd=FALSE)
+p <- ggplot(densityST) + geom_raster(aes_string(x="x", y="y", fill=paste0("X", year))) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") 
+saveFigure(p, "tmp_compare_densities_ST.png", width=8)
+###
+
 iterations <- as.integer(c(1,1,3,2,5,6))
@@
 ```{r figure-2}
-iterations <- as.integer(c(1,1,3,2,5,6))
-#iterations <- as.integer(c(1,1,3,2,5,7))
+#iterations <- as.integer(c(1,1,3,2,5,6))
+##iterations <- as.integer(c(1,1,3,2,5,7))
+iterations <- as.integer(c(1,1,3,2,5,8))
 year <- 2010
@@
 agents <- getAgentLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
-densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, sd=F)
+densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, sd=F, log=T)
 # TODO: do not use scaling??
-densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer) { x[,yearLayer] <- scale(x[,yearLayer]); return(x) }, yearLayer=yearLayer)
+#densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer) { x[,yearLayer] <- scale(x[,yearLayer]); return(x) }, yearLayer=yearLayer)
+study <- getStudy(scenario="A")
 limits <- extent(study$studyArea$boundary)
 
+# Scale density to 1/km^2
+densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer, weight) {
+  x[,yearLayer] <- x[,yearLayer] * weight
+  return(x)
+}, yearLayer=yearLayer, weight=1 / (prod(res(study$getTemplateRaster())) / 1000^2))
+
 #p <- ggplot(boundaryDF) + geom_polygon(aes(long, lat, group=group), fill=terrain.colors(100)[1]) +
-p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
+p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
+  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent", guide=guide_legend(title=expression(paste("Density ", individuals/km^2)))) +
   geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
-  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black")) + 
+  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black"), guide=FALSE) + 
   scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
   scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
-  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal()
+  facet_grid(~scenario) + coord_equal() + theme_raster(base_size=10) + theme(legend.position="bottom")
 print(p)
 saveFigure(p, "manuscript-figure-2a.png", width=8)
+
+###
 
@@
   meanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
-    data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
+    x <- data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario[1])
+    if (is.infinite(x$True)) x$True <- NA
+    if (is.infinite(x$Estimated)) x$Estimated <- NA
+    return(x)
   })
@@
   logPopulationSizes <- populationSizes
-  logPopulationSizes$Observed <- logPopulationSizes$Observed + 1
-  logPopulationSizes$Estimated <- logPopulationSizes$Estimated + 1
+  logPopulationSizes$Observed <- log10(logPopulationSizes$Observed + 1)
+  logPopulationSizes$Estimated <- log10(logPopulationSizes$Estimated + 1)
   return(logPopulationSizes)
-}
-
-getLogMeanPopulationSizes <- function(populationSizes) {  
-  logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
-    data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
-  })
-  return(logMeanPopulationSizes)
 }
@@
   library(scales)  
+  size$Scenario <- prettyScenario(size$Scenario)
+  meanSize$Scenario <- prettyScenario(meanSize$Scenario)
+
   p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
-    #geom_point(size=1) +
-    #stat_smooth(method="lm", aes(colour=modelName), se=F) +
-    geom_point(size=1, colour="darkgray") +
-#stat_smooth(method="lm", se=F, colour="black") +
-    #scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
-    #scale_colour_manual("Model", values=c("darkgray","black")) +
+    geom_point(size=1, colour="darkgray", alpha=.25) +
     geom_abline(colour="black", linetype="longdash") + 
-    #geom_point(data=meanSize, aes(True, Estimated, colour=modelName), size=2, shape=15) +
-    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black") +
-    facet_wrap(~Scenario, scales="free") +
-    coord_cartesian(xlim=xlim, ylim=ylim) +
-    scale_x_log10() + scale_y_log10() +
+    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black")
+
+  #if (!any(abs(size$Estimated) == Inf)) p <- p + geom_smooth(method="lm", se=F, fullrange=T, colour="black")
+  ls <- ddply(size, .(Scenario), function(x) if (!any(abs(x$Estimated) == Inf)) coef(lm(Estimated~Observed, x)) else c(NA,NA))
+  p <- p + geom_abline(data=ls, aes(intercept=`(Intercept)`, slope=Observed))
+
+  p <- p + facet_wrap(~Scenario) +
+    scale_x_continuous(limits=xlim, labels=trans_format("identity", math_format(10^.x))) +
+    scale_y_continuous(limits=ylim, labels=trans_format("identity", math_format(10^.x))) +
     xlab("True population size") + ylab("Estimated population size") + ggtitle(paste("Model: ", prettyModelName(modelName[1]))) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=12))
-
-  if (!any(abs(size$Estimated) == Inf)) p <- p + stat_smooth(method="lm", se=F, colour="black")
 
@@
 }
+
+#p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), subset(logMeanPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
+#saveFigure(p, "manuscript-figure-3a1.png", width=8, height=4)
+
+#xlim <- log10(10^c(0,6)); ylim <- xlim
+#p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[2]), subset(logMeanPopulationSizes10Days, modelName==modelNames[2]), modelName=modelNames[2], xlim=xlim, ylim=ylim)
+#saveFigure(p, "manuscript-figure-3c2.png", width=8, height=4)
 ```
@@
 logPopulationSizes <- getLogPopulationSizes(populationSizes)
-logMeanPopulationSizes  <- getLogMeanPopulationSizes(populationSizes)
+logMeanPopulationSizes  <- getMeanPopulationSizes(logPopulationSizes)
 
@@
 logPopulationSizes2000 <- getLogPopulationSizes(populationSizes2000)
-logMeanPopulationSizes2000  <- getLogMeanPopulationSizes(populationSizes2000)
+logMeanPopulationSizes2000  <- getMeanPopulationSizes(logPopulationSizes2000)
 
@@
 logPopulationSizes10Days <- getLogPopulationSizes(populationSizes10Days)
-logMeanPopulationSizes10Days  <- getLogMeanPopulationSizes(populationSizes10Days)
-#validation$summarizePopulationSize(populationSizes10Days)
+logMeanPopulationSizes10Days  <- getMeanPopulationSizes(logPopulationSizes10Days)
 ```
@@
 ```{r figure-3abc}
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), subset(logMeanPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], xlim=c(1,10000), ylim=c(1,10000))
+xlim <- log10(10^c(0,4)); ylim <- xlim
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), subset(logMeanPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a1.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[2]), subset(logMeanPopulationSizes, modelName==modelNames[2]), modelName=modelNames[2], xlim=c(1,10000), ylim=c(1,10000))
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[2]), subset(logMeanPopulationSizes, modelName==modelNames[2]), modelName=modelNames[2], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a2.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), subset(logMeanPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], xlim=c(1,10000), ylim=c(1,10000))
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), subset(logMeanPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a3.png", width=8, height=4)
 
-xlim <- c(100,10000)
-ylim <- xlim
+xlim <- log10(10^c(2,4)); ylim <- xlim
 p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), subset(logMeanPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
@@
 
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), subset(logMeanPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], xlim=c(1,10000), ylim=c(1,10000))
+xlim <- log10(10^c(0,4)); ylim <- xlim
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), subset(logMeanPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c1.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[2]), subset(logMeanPopulationSizes10Days, modelName==modelNames[2]), modelName=modelNames[2], xlim=c(1,10000), ylim=c(1,10000))
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[2]), subset(logMeanPopulationSizes10Days, modelName==modelNames[2]), modelName=modelNames[2], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c2.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), subset(logMeanPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], xlim=c(1,10000), ylim=c(1,10000))
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), subset(logMeanPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c3.png", width=8, height=4)
@@
 dir <- normalizePath(context$figuresDirectory)
-getPopulationSizesTable(populationSizes, meanPopulationSizes, file=file.path(dir, "table-1a.csv"))
-getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, file=file.path(dir, "table-1b.csv"))
-getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, file=file.path(dir, "table-1c.csv"))
+getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, file=file.path(dir, "table-1a.csv"))
+getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, file=file.path(dir, "table-1b.csv"))
+getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, file=file.path(dir, "table-1c.csv"))
 ```

### 2015-01-22 inst/simulation/manuscript.Rmd
...

@@
 scenarios <- c("A","B","C","D","E","F")
+allScenarios <- c(scenarios, paste0(scenarios, "combined"), paste0(scenarios, "10days"))
 modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"), paste("SmoothModel", "nbinomial", "ar1", sep = "-"), "FMPModel")
@@
 boundaryDF <- ggplot2::fortify(FinlandWTCStudy(context=context, response="canis.lupus")$studyArea$boundary)    
+```
+
+```{r figures}
+study <- getMSS(scenario=scenarios[1])$study
+study$studyArea$boundary@polygons[[1]]@area / 1000^2 # Study area area
+200 / (study$studyArea$boundary@polygons[[1]]@area / 1000^2) # Initial population density in Finland
+intersections <- study$loadIntersections(iteration=as.integer(4))
+ddply(intersections$intersections@data, .(response, year), function(x) sum(x$length) / 1000) # Total sampling effort per year (km)
+#ddply(intersections$intersections@data, .(response, year), function(x) sum(x$intersections))
+
+# Mean encounter rates
+aggrates <- data.frame()
+for (scenario in allScenarios) {
+  study <- getMSS(scenario=scenario)$study
+
+  rates <- data.frame()
+  for (iteration in 1:50) {
+    intersections <- try(study$loadIntersections(iteration=as.integer(iteration)))
+    if (inherits(intersections, "try-error")) next
+    x <- ddply(intersections$intersections@data, .(response, year), function(x) data.frame(rate=sum(x$intersections) / (sum(x$length)/1000) / sum(x$duration)))
+    rates <- rbind(rates, x)
+  }
+
+  y <- ddply(rates, .(year), function(x, scenario) data.frame(scenario=scenario, rate.mean=mean(x$rate), rate.sd=sd(x$rate)), scenario=scenario)
+  aggrates <- rbind(aggrates, y)
+}
+
+subset(aggrates, year==2001)
 ```
@@
 #window <- c(3.375, 3.4, 6.9, 6.925)*1e6
-window <- (c(3.375, 3.4, 6.9, 6.925)+c(-.01,.01,-.01,.01)*3)*1e6
+window <- (c(3.425, 3.45, 6.95, 6.975)+c(-.01,.01,-.01,.01)*9)*1e6
 habitatRaster <- crop(study$studyArea$habitat, extent(window))
@@
 
-p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z), alpha=.3) + scale_fill_manual(values=col) +
+p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z), alpha=.4) + scale_fill_manual(values=col) +
   geom_path(data=trackData, aes(long, lat, group=group, colour=observation), alpha=1, size=.3) +#, alpha=observation)) +
-  scale_colour_manual(values=c("darkgray","black")) + # scale_alpha_manual(values=c(.5,1)) + # gray40
+  scale_colour_manual(values=c("gray55","black")) + # scale_alpha_manual(values=c(.5,1)) + # gray40
   geom_path(data=surveyRoutesData, aes(long, lat, group=group), size=.3, colour="blue", alpha=.5) +
@@
 #print(p)
-saveFigure(p, "manuscript-figure-1.png", width=8, height=4)
+saveFigure(p, "manuscript-figure-1.png", width=8, height=5)
 
@@
 p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
-  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent", guide=guide_legend(title=expression(paste("Density ", individuals/km^2)))) +
+  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent", guide=guide_legend(title=expression(paste("Population density ", (individuals/km^2))))) +
   geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
+  geom_point(data=agents, aes(x, y, colour=Observed), size=1.3, colour="white") +
   geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black"), guide=FALSE) +
@@
 ```{r table-1abc}
-getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, full=F, file) {
+getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, overEstimate, full=F, file) {
   library(plyr)
-  x <- ddply(populationSizes, .(Scenario, modelName), function(x) {
-    if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA))
+  x <- ddply(populationSizes, .(Scenario, modelName), function(x, overEstimate) {
+    failed <- round((sum(x$Estimated >= overEstimate) / nrow(x)) * 100)
+    if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA, Failed=failed))
     y <- lm(Estimated~Observed, x)
-    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2)))
-  })
+    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2), Failed=failed))
+  }, overEstimate=overEstimate)
+
   y <- unique(meanPopulationSizes)
@@
 
-  z <- z[,c("Model","Scenario","Slope","EstDev","R2")]
+  z <- z[,c("Scenario","Model","Slope","R2","EstDev","Failed")]
   z$EstDev <- round(z$EstDev, 2)
@@
 dir <- normalizePath(context$figuresDirectory)
-getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, file=file.path(dir, "table-1a.csv"))
-getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, file=file.path(dir, "table-1b.csv"))
-getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, file=file.path(dir, "table-1c.csv"))
+getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(2000+1), file=file.path(dir, "table-1a.csv"))
+getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(20000+1), file=file.path(dir, "table-1b.csv"))
+getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(2000+1), file=file.path(dir, "table-1c.csv"))
 ```
@@
 
-  p <- ggplot(cip2, aes(scenario, Proportion, shape=Interval, colour=modelName)) +
-    geom_hline(aes(yintercept=.5), color="grey90") +
-    geom_hline(aes(yintercept=.95), color="grey90") +
-    geom_point() +
-    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
-    xlab("Scenario") + ylab("Proportion within CI") +
-    facet_wrap(~panel) + ylim(0, 1) +
-    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank())
-saveFigure(p, "manuscript-figure-5.png", width=6, height=2)
-
 ```{r, eval=FALSE}
@@
 
-```{r figure-3c123, eval=FALSE}
-```
-
 ```{r figure-3c, eval=FALSE}
@@
 
-```{r figure-3d123, eval=FALSE}
-```
-
 ```{r figure-3d, eval=FALSE}
@@
 
-```{r table-2, eval=FALSE}
-# TODO: Combine with table 1
-study <- getStudy(scenario="A")
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-failedEstimations <- validation$countFailedEstimations(modelNames=modelNames)
-failedEstimations$pFailed <- (1 - failedEstimations$nEstimated / 50) * 100
-failedEstimations
-```
-
 ```{r figure-4-test, eval=FALSE}
@@
 populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames[1])
+populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
 
@@
 populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1])
+populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
 
@@
 populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1])
+populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
 

### 2015-01-22 R/Validation.R

@@ -176,7 +176,7 @@ Validation <- setRefClass(
       return(study$context$getLongFileName(study$context$scratchDirectory, name="CIValidation", response=study$response, region=study$studyArea$region, tag=paste(modelName, iteration, sep="-")))
     },
     
-    validateCredibilityIntervals = function(modelName, iteration, nSamples=100, save=F) {      
+    validateCredibilityIntervals = function(modelName, iteration, nSamples=100, populationSizeOverEstimate=1e99, save=F) {      
       model <- study$getModel(modelName=modelName, iteration=iteration)
       model$modelName <- modelName
       model$loadEstimates()
@@ -199,7 +199,7 @@ Validation <- setRefClass(
         x <- model$getPopulationSize(populationDensity=populationDensity$mean, tracks=tracks, habitatWeights=habitatWeights)$sizeData
         if (any(x$Estimated > populationSizeOverEstimate)) {
           message("Estimation failed for iteration ", iteration, ".")
-          break
+          next
         }
         x$sample <- sample
         populationSize <- rbind(populationSize, x)

### 2015-01-22 inst/simulation/manuscript.Rmd
...

@@
 ```{r initialize}
+#library(devtools); install_github("statguy/Winter-Track-Counts")
 library(parallel)
@@
     estimates$data <- subset(estimates$data, year == year0)
-    if (log) estimates$data$fittedMean <- log(estimates$data$fittedMean)
+    #if (log) estimates$data$fittedMean <- log(estimates$data$fittedMean)
     density <- estimates$getPopulationDensityInterpolate(getSD=sd, maskPolygon=studyArea$boundary)
@@
 
-```{r figure-2-test}
+```{r figure-2-test, eval=FALSE}
+x <- data.frame(scenario=scenarios[1], iteration=as.integer(1), stringsAsFactors=F)
+modelName <- modelNames[1]
+year0 <- year <- 2001
+    study <- getStudy(scenario=x$scenario)
+    estimates <- study$getModel(modelName=modelName, iteration=x$iteration)
+    estimates$modelName <- modelName
+    estimates <- study$loadEstimates(estimates)
+    estimates$collectEstimates()
+    studyArea <- FinlandStudyArea$new(context=context)$setup(tolerance=0.001)
+    estimates$data <- subset(estimates$data, year == year0)
+    density <- estimates$getPopulationDensityInterpolate(getSD=F, maskPolygon=studyArea$boundary)
+
 iterations <- as.integer(c(1,1,3,2,5,6))

### 2015-01-26 inst/simulation/manuscript.Rmd
...

--- 2015-01-22 inst/simulation/manuscript.Rmd
@@
 
-  #plot(tracksSP, col="grey", xlim=c(3.4, 3.5)*1e6, ylim=c(7.0, 7.1)*1e6, asp=1)
-
   intersections$tracks$tracks <- subset(intersections$tracks$tracks, year==2001)
@@
   trackData <- rbind(trackData, x, y)
-  #plot(tracksSP, col="black", add=T)
 
@@
   observedAt <- intersections$intersections$intersections[intersections$intersections$intersections$intersections > 0 & intersections$intersections$intersections$year == 2001,]$surveyRoute
-  y <- data.frame(id=unique(x$id), observed=FALSE)
-  if (length(observedAt) > 0) y[y$id %in% observedAt,]$observed <- TRUE
+  y <- data.frame(id=unique(x$id), observation=FALSE)
+  if (length(observedAt) > 0) y[y$id %in% observedAt,]$observation <- TRUE
   x <- plyr::join(x, y, by="id")
@@
 
-  #plot(intersections$surveyRoutes$surveyRoutes, col="blue", add=T)    
+  #trackData$type <- "track"
+  #surveyRoutesData$type <- "transect"
+  #x <- rbind(trackData[,c("long","lat","order","piece","group","observation","type")], surveyRoutesData[,c("long","lat","order","piece","group","observation","type")])
+  #x$colour <- factor(paste(x$type, x$observation))
+  #ggplot(x, aes(long, lat, group=group, colour=colour)) + geom_path()
+
   i <- i + 1
@@
 
+```{r, eval=F}
+#x<-data.frame(scenario="C", iteration=as.integer(3), stringsAsFactors=F)
+modelName <- modelNames[2]
+x<-data.frame(scenario="A", iteration=as.integer(10), stringsAsFactors=F)
+study <- getStudy(scenario=x$scenario)
+estimates <- study$getModel(modelName=modelName, iteration=x$iteration)
+estimates$modelName <- modelName
+estimates <- study$loadEstimates(estimates)
+intersections <- study$loadIntersections(iteration=x$iteration)
+head(subset(intersections$intersections@data, intersections>0), 20)
+head(subset(estimates$data, intersections>0), 20)
+
+model <- SimulatedSmoothModelTemporal(study=study, iteration=as.integer(1))
+params <- list(family="nbinomial", offsetScale=1000^2, timeModel="ar1")
+intersections <- study$loadIntersections(iteration = model$iteration)
+model$setup(intersections = intersections, params = params)
+model$estimate()
+
+head(subset(model$data, intersections>0), 20)
+head(subset(intersections$intersections@data, intersections>0), 20)
+
+./count_intersections.sh ~/git/RParallelScreen/ A 1 1 1
+./estimate.sh ~/git/RParallelScreen/ A 1 1 SmoothModel-nbinomial-ar1
+
+./estimate.sh ~/git/RParallelScreen/ A 1 1 FMPModel
+
+```
+
 ```{r figure-2-init}
@@
     estimates <- study$loadEstimates(estimates)
-    intersections <- subset(estimates$data, year %in% years)$intersections>0
-    surveyRoutes <- data.frame(estimates$getUnscaledObservationCoordinates(), Intersection=intersections)
+    z <- cbind(estimates$getUnscaledObservationCoordinates(), estimates$data)
+    surveyRoutes <- subset(z, year %in% years, select=c(x,y,intersections))
+    surveyRoutes$intersections <- surveyRoutes$intersections > 0
     surveyRoutes$Scenario <- x$scenario
@@
 
+#x<-data.frame(scenario="C", iteration=as.integer(3), stringsAsFactors=F)
 getAgentLocations = function(scenarios, modelName, iterations, years) {
   ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, years) {
+
+    #study <- getStudy(scenario=x$scenario)
+    #intersections <- study$loadIntersections(iteration=x$iteration)
+    #observationTracks <- SimulatedTracks$new(study=study, iteration=x$iteration)
+    #observationTracks$loadTracks(fileName=observationTracks$getObservationTracksFileName(), addColumns=TRUE)
+    # TODO: use observation tracks ^^^^
+
     study <- getStudy(scenario=x$scenario)
@@
 
-    # If no intersection, randomized track can be anywhere else except on the survey route
+    # The observation day has not been recorded, ouch... just randomize a day from the 60 days track
+    observedBursts <- aaply(intersections$intersectionsMatrix, 2, function(x) if (any(x>0)) TRUE else FALSE)
+    #observationTracks <- tracks$randomizeObservationDayTracks(days=1)
+    observationTracks <- tracks
 
-    observedBursts <- aaply(intersections$intersectionsMatrix, 2, function(x) if (any(x>0)) TRUE else FALSE)
-    observationTracks <- tracks$randomizeObservationDayTracks(days=1)
+    #d_ply(observationTracks$tracks, .(id, year, burst), function(x) {
+    #  with(x[1,], message(id, " ", year, " ", burst))
+    #})
 
-    a <- ddply(observationTracks$tracks, .(id, year, burst), function(x, observationTracks) {
+    # Return mean location of the unobserved agents
+    a <- ddply(observationTracks$tracks, .(id, year, burst), function(x, observedBursts) {
       name <- with(x[1,], paste(id, year, burst))
-      if (observedBursts[name] == FALSE) return(data.frame(x=mean(x$x), y=mean(x$y), year=x$year))
+      if (observedBursts[name] == FALSE) return(data.frame(x=mean(x$x), y=mean(x$y), year=x$year[1]))
+      #if (observedBursts[name] == FALSE) return(data.frame(x=x$x, y=x$y, year=x$year))
       NULL
-    }, observationTracks=observationTracks)[,c("x","y","year")]
+    }, observedBursts=observedBursts)[,c("x","y","year")]
     a$Observed <- F
 
+    # But for the observed agents we know the location, which is the transect
     index <- intersections$intersections$intersections > 0 & intersections$intersections$year %in% years
-    if (all(index == FALSE)) b <- data.frame() else {
+    b <- if (all(index == FALSE)) data.frame() else {
       x <- intersections$intersections[index,]
-      b <- data.frame(coordinates(x), year=x$year)
-      b$Observed <- T
+      y <- data.frame(coordinates(x), year=x$year)
+      y$Observed <- T
+      y
     }
+
     agentLocations <- rbind(a, b)
@@
 getEstimatedDensity = function(scenarios, modelName, iterations, year, sd=FALSE, log=FALSE) {
-  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, year0, sd) {
+  ddply(data.frame(scenario=scenarios, year=rep(year, scenarios), iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, sd) {
     message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
@@
 
-    estimates$data <- subset(estimates$data, year == year0)
-    if (log) estimates$data$fittedMean <- log(estimates$data$fittedMean)
+    estimates$data <- subset(estimates$data, year == x$year)
+    #if (log) estimates$data$fittedMean <- log(estimates$data$fittedMean)
     density <- estimates$getPopulationDensityInterpolate(getSD=sd, maskPolygon=studyArea$boundary)
@@
     return(density)
-  }, modelName=modelName, year0=year, sd=sd)
+  }, modelName=modelName, sd=sd)
 }
@@
 
-```{r figure-2-test}
-iterations <- as.integer(c(1,1,3,2,5,6))
-year <- 2010
-densityST <- getEstimatedDensity(scenarios[1], modelNames[1], iterations[1], year=year, sd=FALSE)
-p <- ggplot(densityST) + geom_raster(aes_string(x="x", y="y", fill=paste0("X", year))) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") 
-saveFigure(p, "tmp_compare_densities_ST.png", width=8)
-###
-
-iterations <- as.integer(c(1,1,3,2,5,6))
-year <- 2010
-surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios[1], modelName=modelNames[1], iterations=iterations[1], years=year)
-surveyRoutes <- rbind(surveyRoutes[surveyRoutes$Intersection == FALSE,], surveyRoutes[surveyRoutes$Intersection == TRUE,])
-agents <- getAgentLocations(scenarios=scenarios[1], modelName=modelNames[1], iterations=iterations[1], years=year)
-densityST <- getEstimatedDensity(scenarios[1], modelNames[1], iterations[1], year=year, sd=FALSE)
-densityFMP <- getEstimatedDensity(scenarios[1], modelNames[3], iterations[1], year=year, sd=FALSE)
+```{r figure-2}
+study <- getStudy(scenario="A")
 limits <- extent(study$studyArea$boundary)
-
-p <- ggplot(densityST) + geom_raster(aes_string(x="x", y="y", fill=paste0("X", year))) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
-  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
-  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black")) + 
-  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
-  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
-  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal()
-print(p)
-saveFigure(p, "tmp_compare_densities_ST.png", width=8)
-
-p <- ggplot(densityFMP) + geom_raster(aes_string(x="x", y="y", fill=paste0("X", year))) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
-  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
-  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black")) + 
-  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
-  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
-  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal()
-print(p)
-saveFigure(p, "tmp_compare_densities_FMP.png", width=8)
-```
-
-```{r figure-2}
-#iterations <- as.integer(c(1,1,3,2,5,6))
-##iterations <- as.integer(c(1,1,3,2,5,7))
-iterations <- as.integer(c(1,1,3,2,5,8))
+iterations <- as.integer(c(1,6,3,2,5,31))
+densities <- getEstimatedDensity(scenarios=scenarios[6], modelName=modelNames[1], iterations=iterations[6], year=year, sd=F, log=F)
 year <- 2010
@@
 surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
-surveyRoutes <- rbind(surveyRoutes[surveyRoutes$Intersection == FALSE,], surveyRoutes[surveyRoutes$Intersection == TRUE,])
+#surveyRoutes <- rbind(surveyRoutes[surveyRoutes$intersections == FALSE,], surveyRoutes[surveyRoutes$intersections == TRUE,])
 agents <- getAgentLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
-densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, sd=F, log=T)
+densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, sd=F, log=F)
 # TODO: do not use scaling??
 #densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer) { x[,yearLayer] <- scale(x[,yearLayer]); return(x) }, yearLayer=yearLayer)
-study <- getStudy(scenario="A")
-limits <- extent(study$studyArea$boundary)
 
 # Scale density to 1/km^2
-densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer, weight) {
+densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer, weight, logTrans) {
   x[,yearLayer] <- x[,yearLayer] * weight
+  if (logTrans) x[,yearLayer] <- log(x[,yearLayer])
   return(x)
-}, yearLayer=yearLayer, weight=1 / (prod(res(study$getTemplateRaster())) / 1000^2))
+}, yearLayer=yearLayer, weight=1 / (prod(res(study$getTemplateRaster())) / 1000^2), logTrans=F)
 
-#p <- ggplot(boundaryDF) + geom_polygon(aes(long, lat, group=group), fill=terrain.colors(100)[1]) +
 p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
@@
 print(p)
-saveFigure(p, "manuscript-figure-2a.png", width=8)
+saveFigure(p, "manuscript-figure-2a.png", width=6,height=3)
 
@@
 p <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey") +
-  geom_point(data=surveyRoutes, aes(x, y, colour=Intersection), size=1) + scale_color_manual(values=c("darkgrey", "black")) +
+  geom_point(data=surveyRoutes, aes(x, y, colour=intersections), size=1) + scale_color_manual(values=c("darkgrey", "black")) +
   scale_x_continuous(breaks=seq(gridTemplate@extent@xmin, gridTemplate@extent@xmax, 1e5), expand=c(0,0)) +
@@
 print(p)
-saveFigure(p, "manuscript-figure-2b.png", width=8)
+saveFigure(p, "manuscript-figure-2b.png", width=6,height=3)
 ```

### 2015-01-27 inst/simulation/manuscript.Rmd
...

@@
 plotCIProportions <- function(credibilityIntervals) {
-  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) +
+  credibilityIntervals$CI <- credibilityIntervals$Interval
+  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=CI, colour=modelName)) +
     geom_hline(aes(yintercept=.5), color="grey90") +
@@
 ```
-
-```{r, eval=FALSE}
-study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-populationSizes <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)
-validation$summarizePopulationSize(populationSizes)
-
-meanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
-  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
-})
-logPopulationSizes <- populationSizes
-logPopulationSizes$Observed <- logPopulationSizes$Observed + 1
-logPopulationSizes$Estimated <- logPopulationSizes$Estimated + 1
-logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
-  data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
-})
-#logPopulationSizes$Observed <- log10(logPopulationSizes$Observed + 1)
-#logPopulationSizes$Estimated <- log10(logPopulationSizes$Estimated + 1)
-#logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
-#  data.frame(True=mean(log10(x$Observed+1)), Estimated=mean(log10(x$Estimated+1)), Scenario=x$scenario)
-#})
-```
-
-```{r figure-3a, eval=FALSE}
-p <- ggplot(populationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
-  stat_smooth(method="lm", aes(colour=modelName), se=F) +
-  scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
-  geom_abline(colour="darkgrey", linetype="longdash") +
-  geom_point(data=meanPopulationSizes, aes(True, Estimated, colour=modelName), size=2, shape=15) +
-  facet_wrap(~Scenario, scales="free") +
-  xlab("True population size") + ylab("Estimated population size") +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
-print(p)
-saveFigure(p, "manuscript-figure-3a.png", width=8, height=4)
-```
-
-```{r figure-3-init-2000}
-study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-populationSizes2000 <- validation$getValidationTemporalPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
-validation$summarizePopulationSize(populationSizes2000)
-
-meanPopulationSizes2000 <- ddply(populationSizes2000, .(scenario, modelName), function(x) {
-  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
-})
-logPopulationSizes2000 <- populationSizes2000
-logPopulationSizes2000$Observed <- (logPopulationSizes2000$Observed + 1)
-logPopulationSizes2000$Estimated <- (logPopulationSizes2000$Estimated + 1)
-logMeanPopulationSizes2000 <- ddply(populationSizes2000, .(scenario, modelName), function(x) {
-  data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
-})
-```
-
-```{r figure-3c, eval=FALSE}
-p <- ggplot(populationSizes2000, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
-  stat_smooth(method="lm", aes(colour=modelName), se=F) +
-  scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
-  geom_abline(colour="darkgrey", linetype="longdash") +
-  geom_point(data=meanPopulationSizes2000, aes(True, Estimated, colour=modelName), size=2, shape=15) +
-  facet_wrap(~Scenario, scales="free") +
-  xlab("True population size") + ylab("Estimated population size") +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
-print(p)
-saveFigure(p, "manuscript-figure-3c.png", width=8, height=4)
-```
-
-```{r figure-3-init-10days, eval=FALSE}
-study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-populationSizes10days <- validation$getValidationTemporalPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
-validation$summarizePopulationSize(populationSizes10days)
-
-meanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
-  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
-})
-logPopulationSizes10days <- populationSizes10days
-logPopulationSizes10days$Observed <- (logPopulationSizes10days$Observed + 1)
-logPopulationSizes10days$Estimated <- (logPopulationSizes10days$Estimated + 1)
-logMeanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
-  data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
-})
-```
-
-```{r figure-3d, eval=FALSE}
-p <- ggplot(populationSizes10days, aes(Observed, Estimated, colour=modelName)) + geom_point(size=1) +
-  stat_smooth(method="lm", aes(colour=modelName), se=F) +
-  scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
-  geom_abline(colour="darkgrey", linetype="longdash") +
-  geom_point(data=meanPopulationSizes10days, aes(True, Estimated, colour=modelName), size=2, shape=15) +
-  facet_wrap(~Scenario, scales="free") +
-  xlab("True population size") + ylab("Estimated population size") +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
-print(p)
-saveFigure(p, "manuscript-figure-3d.png", width=8, height=4)
-```
-
-```{r figure-4-test, eval=FALSE}
-study <- getStudy(scenario="A")
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-x <- validation$validateSpatialPopulationSize(modelNames[1])
-```
-
-```{r figure-4a}
-study <- getStudy(scenario="A")
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames[1])
-populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
-
-p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
-  facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
-print(p)
-saveFigure(p, "manuscript-figure-4a.png", width=4, height=4)
-```
-
-```{r figure-4b}
-study <- getStudy(scenario="A")
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1])
-populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
-
-p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
-  facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
-print(p)
-saveFigure(p, "manuscript-figure-4b.png", width=4, height=4)
-```
-
-```{r figure-4c}
-study <- getStudy(scenario="A")
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1])
-populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
-
-p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
-  facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
-print(p)
-saveFigure(p, "manuscript-figure-4c.png", width=4, height=4)
-```
-
-```{r figure-5a}
-study <- getStudy(scenario="A")
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-a <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames)
-b <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames, probs=c(.25, .75), probsName="50%")
-credibilityIntervals <- rbind(a, b)
-
-p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
-  scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
-  xlab("Scenario") + ylab("Proportion within CI")
-print(p)
-saveFigure(p, "manuscript-figure-5a.png", width=4, height=2)
-```
-
-```{r figure-5b}
-study <- getStudy(scenario="A")
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
-b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames, probs=c(.25, .75), probsName="50%")
-credibilityIntervals <- rbind(a, b)
-
-p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
-  scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
-  xlab("Scenario") + ylab("Proportion within CI")
-print(p)
-saveFigure(p, "manuscript-figure-5b.png", width=4, height=2)
-```
-
-```{r figure-5c}
-study <- getStudy(scenario="A")
-validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
-a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
-b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames, probs=c(.25, .75), probsName="50%")
-credibilityIntervals <- rbind(a, b)
-
-p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
-  scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
-  xlab("Scenario") + ylab("Proportion within CI")
-print(p)
-saveFigure(p, "manuscript-figure-5c.png", width=4, height=2)
-```

### 2015-01-27 R/Model.R

@@ -623,6 +623,7 @@ SmoothModelSpatioTemporal <- setRefClass(
       
       library(raster)
       
+      #xyztMean <- data.frame(estimates$getUnscaledObservationCoordinates(), z=estimates$data$fittedMean / estimates$offsetScale, t=estimates$data$year)
       xyztMean <- data.frame(getUnscaledObservationCoordinates(), z=data$fittedMean / offsetScale, t=data$year)
       meanPopulationDensityRaster <- getPopulationDensityInterpolate.internal(xyztMean, templateRaster=templateRaster, maskPolygon=maskPolygon)
       sdPopulationDensityRaster <- if (getSD) {

### 2015-01-27 R/Validation.R

@@ -197,7 +197,7 @@ Validation <- setRefClass(
         populationDensity <- model$getPopulationDensity(getSD=FALSE)
         
         x <- model$getPopulationSize(populationDensity=populationDensity$mean, tracks=tracks, habitatWeights=habitatWeights)$sizeData
-        if (any(x$Estimated > populationSizeOverEstimate)) {
+        if (any(x$Estimated >= populationSizeOverEstimate)) {
           message("Estimation failed for iteration ", iteration, ".")
           next
         }

### 2015-01-27 R/static_global_functions.R

@@ -263,17 +263,12 @@ multiRasterInterpolate <- function(xyzt, variables, templateRaster, transform=id
   library(raster)
   
   if (missing(xyzt)) stop("xyzt missing")
-  if (missing(variables)) stop("variables missing")  
+  if (missing(variables)) stop("variables missing")
   if (missing(templateRaster)) stop("templateRaster missing")
   
   message("Template raster:")
   print(raster::extent(templateRaster))
 
-  xyzt <- xyzt[complete.cases(xyzt),]
-  resultRaster <- rasterInterpolate(xyz=xyzt, templateRaster=templateRaster, transform=transform, inverseTransform=inverseTransform)
-  return(stack(resultRaster))
-  
-  
   rasterList <- dlply(.data=xyzt, .variables=variables, .fun=function(xyz, templateRaster, transform, inverseTransform) {
     xyz <- xyz[complete.cases(xyz),]
     resultRaster <- rasterInterpolate(xyz=xyz, templateRaster=templateRaster, transform=transform, inverseTransform=inverseTransform)

### 2015-01-27 inst/simulation/population_size_all.sh

@@ -39,8 +39,8 @@ function population_size_scenario {
 }
 
 population_size_scenario scenarios models "${nodes[0]}" "${max_nodes[0]}"
-#population_size_scenario scenarios2000 models "${nodes[1]}" "${max_nodes[1]}"
-#population_size_scenario scenarios10days models "${nodes[2]}" "${max_nodes[2]}"
+population_size_scenario scenarios2000 models "${nodes[1]}" "${max_nodes[1]}"
+population_size_scenario scenarios10days models "${nodes[2]}" "${max_nodes[2]}"
 
 
 

### 2015-01-27 inst/simulation/manuscript.Rmd
...

@@
 ```{r table-1abc}
-getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, overEstimate, full=F, file) {
+getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, overEstimate, full=F, file, ...) {
   library(plyr)
@@
 dir <- normalizePath(context$figuresDirectory)
-getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(2000+1), file=file.path(dir, "table-1a.csv"))
-getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(20000+1), file=file.path(dir, "table-1b.csv"))
-getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(2000+1), file=file.path(dir, "table-1c.csv"))
+table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=2000)
+logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(2000+1), file=file.path(dir, "table-1a.csv"))
+logTable1$EstDev <- table1$EstDev
+write.csv(logTable1, file=file.path(dir, "table-1a.csv"))
+
+table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=20000)
+logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(20000+1), file=file.path(dir, "table-1b.csv"))
+logTable2$EstDev <- table2$EstDev
+write.csv(logTable2, file=file.path(dir, "table-2a.csv"))
+
+table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=2000)
+logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(2000+1), file=file.path(dir, "table-1c.csv"))
+logTable3$EstDev <- table3$EstDev
+write.csv(logTable3, file=file.path(dir, "table-3a.csv"))
 ```

### 2015-01-27 R/Model.R

@@ -623,7 +623,6 @@ SmoothModelSpatioTemporal <- setRefClass(
       
       library(raster)
       
-      #xyztMean <- data.frame(estimates$getUnscaledObservationCoordinates(), z=estimates$data$fittedMean / estimates$offsetScale, t=estimates$data$year)
       xyztMean <- data.frame(getUnscaledObservationCoordinates(), z=data$fittedMean / offsetScale, t=data$year)
       meanPopulationDensityRaster <- getPopulationDensityInterpolate.internal(xyztMean, templateRaster=templateRaster, maskPolygon=maskPolygon)
       sdPopulationDensityRaster <- if (getSD) {

### 2015-02-02 inst/simulation/manuscript.Rmd
...

--- 2015-01-27 inst/simulation/manuscript.Rmd
@@
 iterations <- as.integer(c(1,6,3,2,5,31))
-densities <- getEstimatedDensity(scenarios=scenarios[6], modelName=modelNames[1], iterations=iterations[6], year=year, sd=F, log=F)
 year <- 2010
@@
     failed <- round((sum(x$Estimated >= overEstimate) / nrow(x)) * 100)
-    if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA, Failed=failed))
+    if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA, Failed=failed, MeanProp=NA, SDProp=NA))
     y <- lm(Estimated~Observed, x)
-    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2), Failed=failed))
+    prop.mean <- mean(x$Estimated/x$Observed)
+    prop.sd <- sd(x$Estimated/x$Observed)
+    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2), Failed=failed, MeanProp=prop.mean, SDProp=prop.sd))
   }, overEstimate=overEstimate)
@@
   y$Scenario <- y$scenario
-  y$EstDev <- y$Estimated / y$True
+  #y$EstDev <- y$Estimated / y$True
   z <- plyr::join(x, y)
@@
 
-  z <- z[,c("Scenario","Model","Slope","R2","EstDev","Failed")]
-  z$EstDev <- round(z$EstDev, 2)
+  z <- z[,c("Scenario","Model","Slope","R2","Failed","MeanProp","SDProp")]
+  #z$EstDev <- round(z$EstDev, 2)
+  z$MeanProp <- round(z$MeanProp, 2)
+  z$SDProp <- round(z$SDProp, 2)
   if (!missing(file)) write.csv(z, file=file)
@@
 dir <- normalizePath(context$figuresDirectory)
-getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(2000+1), file=file.path(dir, "table-1a.csv"))
-getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(20000+1), file=file.path(dir, "table-1b.csv"))
-getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(2000+1), file=file.path(dir, "table-1c.csv"))
+table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=2000)
+logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(2000+1), file=file.path(dir, "table-1a.csv"))
+logTable1$MeanProp <- table1$MeanProp; logTable1$SDProp <- table1$SDProp; 
+write.csv(logTable1, file=file.path(dir, "table-1a.csv"))
+
+table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=20000)
+logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(20000+1), file=file.path(dir, "table-1b.csv"))
+logTable2$MeanProp <- table2$MeanProp; logTable2$SDProp <- table2$SDProp; 
+write.csv(logTable2, file=file.path(dir, "table-2a.csv"))
+
+table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=2000)
+logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(2000+1), file=file.path(dir, "table-1c.csv"))
+logTable3$MeanProp <- table3$MeanProp; logTable3$SDProp <- table3$SDProp; 
+write.csv(logTable3, file=file.path(dir, "table-3a.csv"))
 ```

### 2015-02-02 inst/simulation/manuscript.Rmd
...

@@
 allScenarios <- c(scenarios, paste0(scenarios, "combined"), paste0(scenarios, "10days"))
-modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"), paste("SmoothModel", "nbinomial", "ar1", sep = "-"), "FMPModel")
+modelNames <- c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "SmoothModelMean-nbinomial-ar1", "FMPModel")
 iterations <- as.integer(c(1,1,1,1,3,1))
@@
 prettyModelName <- function(x) {
-  lookup <- data.frame(x=c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "FMPModel") , y=c("ST", "T", "I"), stringsAsFactors=F)
+  lookup <- data.frame(x=modelNames, y=c("ST", "T", "T2", "I"), stringsAsFactors=F)
   lookup$y[match(x, lookup$x)]

### 2015-02-02 inst/simulation/count_intersections_all.sh

@@ -30,7 +30,7 @@ function count_intersections_scenario {
 
   for scenario in "${scenarios[@]}"
   do
-    count_intersections_single $scenario $nodes $max_nodes $model
+    count_intersections_single $scenario $nodes $max_nodes $numdays
   done
 }
 

### 2015-02-03 inst/simulation/manuscript.Rmd
...

@@
 prettyModelName <- function(x) {
-  lookup <- data.frame(x=modelNames, y=c("ST", "T", "T2", "I"), stringsAsFactors=F)
+  lookup <- data.frame(x=modelNames, y=c("ST", "T", "TA", "FMP"), stringsAsFactors=F)
   lookup$y[match(x, lookup$x)]
@@
 }
-
-#p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), subset(logMeanPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
-#saveFigure(p, "manuscript-figure-3a1.png", width=8, height=4)
-
-#xlim <- log10(10^c(0,6)); ylim <- xlim
-#p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[2]), subset(logMeanPopulationSizes10Days, modelName==modelNames[2]), modelName=modelNames[2], xlim=xlim, ylim=ylim)
-#saveFigure(p, "manuscript-figure-3c2.png", width=8, height=4)
 ```
@@
 registerDoMC(cores=10) # The nodes seem to run out of memory and fail with a bogus error message, so it's better to reduce the number of used cores.
-
 populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames)
@@
 ```{r figure-3abc}
+#modelNames <- modelNames[c(2:4,1)]
+
 xlim <- log10(10^c(0,4)); ylim <- xlim
@@
 logTable1$MeanProp <- table1$MeanProp; logTable1$SDProp <- table1$SDProp;
-write.csv(logTable1, file=file.path(dir, "table-1a.csv"))
 
@@
 logTable2$MeanProp <- table2$MeanProp; logTable2$SDProp <- table2$SDProp;
-write.csv(logTable2, file=file.path(dir, "table-2a.csv"))
 
@@
 logTable3$MeanProp <- table3$MeanProp; logTable3$SDProp <- table3$SDProp;
+
+write.csv(logTable1, file=file.path(dir, "table-1a.csv"))
+write.csv(logTable2, file=file.path(dir, "table-2a.csv"))
 write.csv(logTable3, file=file.path(dir, "table-3a.csv"))

### 2015-02-03 R/Validation.R

@@ -15,6 +15,11 @@ Validation <- setRefClass(
     #  tracks$ge
     #},
     
+    getIntersectionFileIterations = function() {
+      intersections <- SimulatedIntersections(study=study)
+      return(intersections$getIntersectionFileIterations())
+    },
+    
     getPopulationSizeFileIterations = function(modelName) {
       populationSize <- SimulationPopulationSize(study=study, modelName=modelName)
       return(populationSize$getPopulationSizeFileIterations())
@@ -24,15 +29,42 @@ Validation <- setRefClass(
       model <- SimulatedSmoothModelSpatioTemporal(study=study, modelName=modelName)
       return(model$getEstimatesFileIterations())
     },
+
+    validateTemporalIntersection = function() {
+      iterations <- getIntersectionFileIterations()
+      if (length(itrations) == 0) return(NULL)
+      
+      intersections <- ldply(iterations, function(iteration) {
+        message("Processing iteration ", iteration, " of ", length(iterations), " iterations for scenario ", study$response, "...")
+        intersections <- study$loadIntersections(iteration=iteration, predictDistances=FALSE)
+        x <- intersections$intersections@data
+        y <- ddply(x, .(year), function(x) data.frame(intersections=sum(x$intersections)))
+        y$iteration <- iteration
+        return(y)
+      }, .parallel=TRUE)
+      
+      intersections$Scenario <- study$response
+      
+      return(intersections)
+    },
     
-    getFailedEstimationRate = function(modelName) {
-      # TODO
+    getValidationTemporalIntersections = function(scenarios=c("A","B","C","D","E","F")) {
+      x <- ldply(scenarios, function(scenario) {
+        s <- getStudy(scenario=scenario, isTest=F)
+        validation <- Validation(study=s, populationSizeOverEstimate=populationSizeOverEstimate)
+        x <- validation$validateTemporalIntersection()
+        if (is.null(x) || nrow(x) == 0) return(NULL)
+        return(x)
+      })
+      
+      return(x)
     },
     
     validateTemporalPopulationSize = function(modelName) {
       library(plyr)
       
       iterations <- getPopulationSizeFileIterations(modelName)
+      if (length(itrations) == 0) return(NULL)
       
       populationSize <- ldply(iterations, function(iteration) {
         message("Processing iteration ", iteration, " of ", length(iterations), " iterations for scenario ", study$response, "...")

### 2015-02-03 inst/simulation/manuscript.Rmd
...

@@
 prettyModelName <- function(x) {
-  lookup <- data.frame(x=modelNames, y=c("ST", "T", "TA", "FMP"), stringsAsFactors=F)
+  lookup <- data.frame(x=modelNames, y=c("ST", "TF", "T", "FMP"), stringsAsFactors=F)
   lookup$y[match(x, lookup$x)]
@@
 ```{r figure-3-init}
+getIntersections <- function(scenarios) {
+  study <- getStudy(scenario="A")
+  validation <- Validation(study=study)
+  intersections <- validation$getValidationTemporalIntersections(scenarios=scenarios)
+  intersections$Year <- intersections$year
+  intersections
+}
+
 getPopulationSizes <- function(scenarios, modelNames, populationSizeOverEstimate=Inf) {
@@
 populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
-
 registerDoMC(cores=detectCores())
@@
 
+```{r table-2}
+intersections <- getIntersections(scenarios)
+intersections2000 <- getIntersections(scenarios=paste0(scenarios, "combined"))
+intersections10days <- getIntersections(scenarios=paste0(scenarios, "10days"))
+intersections <- plyr::join(populationSizes, intersections)
+intersections2000 <- plyr::join(populationSizes2000, intersections2000)
+intersections10days <- plyr::join(populationSizes10Days, intersections10days)
+i.stat <- function(x) data.frame(i.mean=round(mean(x$intersections/x$Observed),2), i.sd=round(sd(x$intersections/x$Observed),2))
+ddply(intersections, .(Scenario), i.stat)
+ddply(intersections2000, .(Scenario), i.stat)
+ddply(intersections10days, .(Scenario), i.stat)
+```
+
 ```{r figure-4-init}

### 2015-02-03 R/Model.R

@@ -141,26 +141,64 @@ Model <- setRefClass(
       if (missing(tracks)) populationSize$loadValidationData()
       else populationSize$loadValidationData(tracks=tracks)
       
-      return(invisible(populationSize))
+     return(invisible(populationSize))
+    }    
+  )
+)
+
+AggregatedModel <- setRefClass(
+  Class = "AggregatedModel",
+  contains = "Model",
+  fields = list(
+  ),
+  methods = list(
+    aggregate = function() {
+      library(plyr)
+      data <<- ddply(data, .(year), function(x) {
+        data.frame(response=sum(x$response), intersections=sum(x$intersections), duration=1, length=sum(x$duration*x$length*x$distance), distance=1)
+      })      
+    },
+        
+    getPopulationSize = function(populationDensity, tracks, habitatWeights) {
+      if (is.null(data$fittedMean))
+        stop("Did you forgot to run collectEstimates() first?")
+      if (length(unique(data$year)) != nrow(data))
+        stop("The data must be aggregated to determine population size")
+      
+      area <- if (missing(habitatWeights) | is.null(habitatWeights)) study$studyArea$boundary@polygons[[1]]@area
+      else cellStats(habitatWeights, sum) * prod(res(habitatWeights))
+      populationSize <- SimulationPopulationSize(study=study, modelName=modelName, iteration=iteration)
+      
+      for (y in sort(data$year)) {
+        size <- subset(data, year == y)$fittedMean * area
+        populationSize$addYearSize(y, size)
+      }        
+      
+      if (missing(tracks)) populationSize$loadValidationData()
+      else populationSize$loadValidationData(tracks=tracks)
+      
+      return(populationSize)
     }
   )
 )
 
 FMPModel <- setRefClass(
   Class = "FMPModel",
-  contains = "Model",
+  contains = "AggregatedModel",
   fields = list(
   ),
   methods = list(
     setup = function(intersections, params) {
       coordsScale <<- 1
-      offsetScale <<- 1
+      offsetScale <<- 1000^2
       modelName <<- "FMPModel"
       
       data <<- intersections$getData()
       data$response <<- data$intersections
       locations <<- intersections$getCoordinates() * coordsScale
       years <<- as.integer(sort(unique(data$year)))
+      
+      .self$aggregate()
       
       return(invisible(.self))      
     },
@@ -304,7 +342,7 @@ SmoothModelTemporal <- setRefClass(
[…]

### 2015-02-03 R/Validation.R

@@ -41,7 +41,7 @@ Validation <- setRefClass(
         y <- ddply(x, .(year), function(x) data.frame(intersections=sum(x$intersections)))
         y$iteration <- iteration
         return(y)
-      }, .parallel=TRUE)
+      })
       
       intersections$Scenario <- study$response
       
@@ -55,7 +55,7 @@ Validation <- setRefClass(
         x <- validation$validateTemporalIntersection()
         if (is.null(x) || nrow(x) == 0) return(NULL)
         return(x)
-      })
+      }, .parallel=TRUE)
       
       return(x)
     },

### 2015-02-04 inst/simulation/manuscript.Rmd
...

@@
 # Mean encounter rates
-aggrates <- data.frame()
-for (scenario in allScenarios) {
+#aggrates <- data.frame()
+aggrates <- ldply(allScenarios, function(scenario) {
   study <- getMSS(scenario=scenario)$study
@@
   y <- ddply(rates, .(year), function(x, scenario) data.frame(scenario=scenario, rate.mean=mean(x$rate), rate.sd=sd(x$rate)), scenario=scenario)
-  aggrates <- rbind(aggrates, y)
-}
+  #aggrates <- rbind(aggrates, y)
+  return(y)
+}, .parallel=T)
 
-subset(aggrates, year==2001)
+subset(aggrates, year==2020)
 ```
@@
 
+```{r figure-3-init-2-test, eval=FALSE}
+populationSizes <- getPopulationSizes(scenarios=scenarios[1:4], modelNames=modelNames[3:4])
+populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios[1:4], "combined"), modelNames=modelNames[3:4])
+populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios[1:4], "10days"), modelNames=modelNames[3:4])
+```
+
 ```{r figure-3-init-2}
@@
 
-dir <- normalizePath(context$figuresDirectory)
 table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=2000)
-logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(2000+1), file=file.path(dir, "table-1a.csv"))
+logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(2000+1))
 logTable1$MeanProp <- table1$MeanProp; logTable1$SDProp <- table1$SDProp;
@@
 table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=20000)
-logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(20000+1), file=file.path(dir, "table-1b.csv"))
+logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(20000+1))
 logTable2$MeanProp <- table2$MeanProp; logTable2$SDProp <- table2$SDProp;
@@
 table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=2000)
-logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(2000+1), file=file.path(dir, "table-1c.csv"))
+logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(2000+1))
 logTable3$MeanProp <- table3$MeanProp; logTable3$SDProp <- table3$SDProp;
 
+dir <- normalizePath(context$figuresDirectory)
 write.csv(logTable1, file=file.path(dir, "table-1a.csv"))

### 2015-02-05 inst/simulation/manuscript.Rmd
...

@@
 prettyModelName <- function(x) {
-  lookup <- data.frame(x=modelNames, y=c("ST", "TF", "T", "FMP"), stringsAsFactors=F)
+  lookup <- data.frame(x=modelNames, y=c("ST", "VT", "T", "FMP"), stringsAsFactors=F)
   lookup$y[match(x, lookup$x)]
@@
     failed <- round((sum(x$Estimated >= overEstimate) / nrow(x)) * 100)
+    if (x$modelName == "FMPModel") failed <- NA
     if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA, Failed=failed, MeanProp=NA, SDProp=NA))
@@
     prop.sd <- sd(x$Estimated/x$Observed)
-    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2), Failed=failed, MeanProp=prop.mean, SDProp=prop.sd))
+    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2), MeanProp=prop.mean, SDProp=prop.sd, Failed=failed))
   }, overEstimate=overEstimate)
@@
 
-  z <- z[,c("Scenario","Model","Slope","R2","Failed","MeanProp","SDProp")]
+  z <- z[,c("Scenario","Model","Slope","R2","MeanProp","SDProp","Failed")]
   #z$EstDev <- round(z$EstDev, 2)
@@
 
+writeTable <- function(df, filename, context) {
+  library(ReporteRs)
+  doc <- docx()
+  flextable <- FlexTable(df)
+  addFlexTable(doc, flextable)
+  dir <- normalizePath(context$figuresDirectory)
+  writeDoc(doc, file=file.path(dir, filename))
+}
+
 table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=2000)
@@
 logTable1$MeanProp <- table1$MeanProp; logTable1$SDProp <- table1$SDProp; 
+logTable1
 
@@
 
-dir <- normalizePath(context$figuresDirectory)
-write.csv(logTable1, file=file.path(dir, "table-1a.csv"))
-write.csv(logTable2, file=file.path(dir, "table-2a.csv"))
-write.csv(logTable3, file=file.path(dir, "table-3a.csv"))
+subset(logTable1, Model != "VT")
+subset(logTable2, Model != "VT")
+subset(logTable3, Model != "VT")
+
+#dir <- normalizePath(context$figuresDirectory)
+#write.csv(logTable1, file=file.path(dir, "table-1a.csv"))
+#write.csv(logTable2, file=file.path(dir, "table-2a.csv"))
+#write.csv(logTable3, file=file.path(dir, "table-3a.csv"))
+
+writeTable(subset(logTable1, Model != "VT"), "table-1a.docx", context)
+writeTable(subset(logTable2, Model != "VT"), "table-1b.docx", context)
+writeTable(subset(logTable3, Model != "VT"), "table-1c.docx", context)
 ```
@@
 intersections10days <- plyr::join(populationSizes10Days, intersections10days)
-i.stat <- function(x) data.frame(i.mean=round(mean(x$intersections/x$Observed),2), i.sd=round(sd(x$intersections/x$Observed),2))
-ddply(intersections, .(Scenario), i.stat)
-ddply(intersections2000, .(Scenario), i.stat)
-ddply(intersections10days, .(Scenario), i.stat)
+i.stat <- function(x) data.frame(o.mean=mean(x$Observed), o.sd=sd(x$Observed), e.mean=mean(x$Estimated), e.sd=sd(x$Estimated),
+                                 pp.mean=round(mean(x$Estimated/x$Observed), 2), pp.sd=round(sd(x$Estimated/x$Observed), 2),
+                                 i.mean=mean(x$intersections), i.sd=sd(x$intersections), ip.mean=round(mean(x$intersections/x$Observed),2), ip.sd=round(sd(x$intersections/x$Observed),2))
+ddply(intersections, .(Scenario, modelName), i.stat)
+ddply(intersections2000, .(Scenario, modelName), i.stat)
+ddply(intersections10days, .(Scenario, modelName), i.stat)
 ```

### 2015-02-09 inst/simulation/manuscript.Rmd
...

@@
 saveFigure(p, "manuscript-figure-3a1.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[2]), subset(logMeanPopulationSizes, modelName==modelNames[2]), modelName=modelNames[2], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), subset(logMeanPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a2.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), subset(logMeanPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), subset(logMeanPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a3.png", width=8, height=4)
@@
 saveFigure(p, "manuscript-figure-3b1.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[2]), subset(logMeanPopulationSizes2000, modelName==modelNames[2]), modelName=modelNames[2], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), subset(logMeanPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b2.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), subset(logMeanPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), subset(logMeanPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b3.png", width=8, height=4)
@@
 saveFigure(p, "manuscript-figure-3c1.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[2]), subset(logMeanPopulationSizes10Days, modelName==modelNames[2]), modelName=modelNames[2], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), subset(logMeanPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c2.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), subset(logMeanPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), subset(logMeanPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c3.png", width=8, height=4)
@@
 logTable1$MeanProp <- table1$MeanProp; logTable1$SDProp <- table1$SDProp;
-logTable1
 
@@
 ddply(intersections10days, .(Scenario, modelName), i.stat)
+
 ```
@@
   credibilityIntervals$panel <- panel
-  credibilityIntervals$scenario <- rep(scenarioNames, times=2, each=2)
+  credibilityIntervals$scenario <- rep(scenarioNames, times=2, each=length(modelNames))
   return(credibilityIntervals)  
@@
 plotCIProportions <- function(credibilityIntervals) {
+  modelLabels <- prettyModelName(unique(credibilityIntervals$modelName))
   credibilityIntervals$CI <- credibilityIntervals$Interval
@@
     geom_point() +
-    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
+    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=modelLabels) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
@@
 ```{r figure-5abc}
-cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames, panel="a", scenarioNames=scenarios)
-cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames, panel="b", scenarioNames=scenarios)
-cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames, panel="c", scenarioNames=scenarios)
-cip <- rbind(cip1, cip2, cip3)
+cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[1:3], panel="a", scenarioNames=scenarios)
+cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1:3], panel="b", scenarioNames=scenarios)
+cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1:3], panel="c", scenarioNames=scenarios)
+cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")
 p <- plotCIProportions(cip)

### 2015-02-12 inst/simulation/manuscript.Rmd
...

@@
 
+cutoff <- 1000
+cutoffPopulationSizes <- function(populationSizes, cutoff=Inf) subset(populationSizes, Estimated <= cutoff)
+
 getMeanPopulationSizes <- function(populationSizes) {
+  library(plyr)
   meanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
@@
 
-plotLogPopulationSize <- function(size, meanSize, modelName, xlim, ylim) {
+plotLogPopulationSize <- function(size, modelName, xlim, ylim, cutoff) {
   library(ggplot2)
@@
   size$Scenario <- prettyScenario(size$Scenario)
+  meanSize <- getMeanPopulationSizes(size)
   meanSize$Scenario <- prettyScenario(meanSize$Scenario)
@@
 
-meanPopulationSizes  <- getMeanPopulationSizes(populationSizes)
-logPopulationSizes <- getLogPopulationSizes(populationSizes)
+meanPopulationSizes  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff))
+logPopulationSizes <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff))
 logMeanPopulationSizes  <- getMeanPopulationSizes(logPopulationSizes)
 
-meanPopulationSizes2000  <- getMeanPopulationSizes(populationSizes2000)
-logPopulationSizes2000 <- getLogPopulationSizes(populationSizes2000)
+meanPopulationSizes2000  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff))
+logPopulationSizes2000 <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff))
 logMeanPopulationSizes2000  <- getMeanPopulationSizes(logPopulationSizes2000)
 
-meanPopulationSizes10Days  <- getMeanPopulationSizes(populationSizes10Days)
-logPopulationSizes10Days <- getLogPopulationSizes(populationSizes10Days)
+meanPopulationSizes10Days  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff))
+logPopulationSizes10Days <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff))
 logMeanPopulationSizes10Days  <- getMeanPopulationSizes(logPopulationSizes10Days)
@@
 
-xlim <- log10(10^c(2,4)); ylim <- xlim
+#xlim <- log10(10^c(2,4)); ylim <- xlim
+xlim <- log10(10^c(0,4)); ylim <- xlim
 p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), subset(logMeanPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
@@
   x <- ddply(populationSizes, .(Scenario, modelName), function(x, overEstimate) {
-    failed <- round((sum(x$Estimated >= overEstimate) / nrow(x)) * 100)
-    if (x$modelName == "FMPModel") failed <- NA
+    failed <- round((sum(x$Estimated > overEstimate) / nrow(x)) * 100)
+    x <- subset(x, Estimated <= overEstimate)
+    if (x$modelName[1] == "FMPModel") failed <- NA
     if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA, Failed=failed, MeanProp=NA, SDProp=NA))
@@
 
-table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=2000)
-logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(2000+1))
-logTable1$MeanProp <- table1$MeanProp; logTable1$SDProp <- table1$SDProp;
+table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff)
+logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(cutoff+1))
+logTable1$MeanProp <- table1$MeanProp; logTable1$SDProp <- table1$SDProp; logTable1$Failed <- table1$Failed
 
-table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=20000)
-logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(20000+1))
-logTable2$MeanProp <- table2$MeanProp; logTable2$SDProp <- table2$SDProp;
+table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff)
+logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1))
+logTable2$MeanProp <- table2$MeanProp; logTable2$SDProp <- table2$SDProp; logTable2$Failed <- table2$Failed
 
-table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=2000)
-logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(2000+1))
-logTable3$MeanProp <- table3$MeanProp; logTable3$SDProp <- table3$SDProp;
+table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff)
+logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1))
+logTable3$MeanProp <- table3$MeanProp; logTable3$SDProp <- table3$SDProp; logTable3$Failed <- table3$Failed
 
@@
 intersections10days <- getIntersections(scenarios=paste0(scenarios, "10days"))
-intersections <- plyr::join(populationSizes, intersections)
-intersections2000 <- plyr::join(populationSizes2000, intersections2000)
-intersections10days <- plyr::join(populationSizes10Days, intersections10days)
+intersectionsx <- plyr::join(populationSizes, intersections)
+intersections2000x <- plyr::join(populationSizes2000, intersections2000)
+intersections10daysx <- plyr::join(populationSizes10Days, intersections10days)
 i.stat <- function(x) data.frame(o.mean=mean(x$Observed), o.sd=sd(x$Observed), e.mean=mean(x$Estimated), e.sd=sd(x$Estimated),
@@
                                  i.mean=mean(x$intersections), i.sd=sd(x$intersections), ip.mean=round(mean(x$intersections/x$Observed),2), ip.sd=round(sd(x$intersections/x$Observed),2))
-ddply(intersections, .(Scenario, modelName), i.stat)
-ddply(intersections2000, .(Scenario, modelName), i.stat)
-ddply(intersections10days, .(Scenario, modelName), i.stat)
+ddply(intersectionsx, .(Scenario, modelName), i.stat)
+ddply(intersections2000x, .(Scenario, modelName), i.stat)
+ddply(intersections10daysx, .(Scenario, modelName), i.stat)
+
+rer.stat <- function(x) data.frame(RERMean=round(mean(x$intersections/x$Observed),2), RERSD=round(sd(x$intersections/x$Observed),2))
+ddply(intersectionsx, .(Scenario), rer.stat)
+ddply(intersections2000x, .(Scenario), rer.stat)
+ddply(intersections10daysx, .(Scenario), rer.stat)
+#writeTable(subset(logTable1, Model != "VT"), "table-1a.docx", context)
 

### 2015-02-12 R/Validation.R

@@ -71,6 +71,9 @@ Validation <- setRefClass(
         populationSize <- study$loadPopulationSize(iteration=iteration, modelName=modelName)
         x <- populationSize$sizeData
         
+        if (any(x$Year < 2000))
+          stop("Invalid year for scenario = ", study$response, ", model = ", modelName, ", iteration = ", iteration)
+        
         if (is.null(x) || nrow(x) == 0 || any(x$Estimated > populationSizeOverEstimate)) {
           message("Estimation failed for iteration ", iteration, " scenario ", study$response)
           return(NULL)

### 2015-02-17 inst/simulation/manuscript.Rmd
...

@@
 intersections10daysx <- plyr::join(populationSizes10Days, intersections10days)
-i.stat <- function(x) data.frame(o.mean=mean(x$Observed), o.sd=sd(x$Observed), e.mean=mean(x$Estimated), e.sd=sd(x$Estimated),
-                                 pp.mean=round(mean(x$Estimated/x$Observed), 2), pp.sd=round(sd(x$Estimated/x$Observed), 2),
-                                 i.mean=mean(x$intersections), i.sd=sd(x$intersections), ip.mean=round(mean(x$intersections/x$Observed),2), ip.sd=round(sd(x$intersections/x$Observed),2))
-ddply(intersectionsx, .(Scenario, modelName), i.stat)
-ddply(intersections2000x, .(Scenario, modelName), i.stat)
-ddply(intersections10daysx, .(Scenario, modelName), i.stat)
+#i.stat <- function(x) data.frame(o.mean=mean(x$Observed), o.sd=sd(x$Observed), e.mean=mean(x$Estimated), e.sd=sd(x$Estimated),
+#                                 pp.mean=round(mean(x$Estimated/x$Observed), 2), pp.sd=round(sd(x$Estimated/x$Observed), 2),
+#                                 i.mean=mean(x$intersections), i.sd=sd(x$intersections), ip.mean=round(mean(x$intersections/x$Observed),2), ip.sd=round(sd(x$intersections/x$Observed),2))
+#ddply(intersectionsx, .(Scenario, modelName), i.stat)
+#ddply(intersections2000x, .(Scenario, modelName), i.stat)
+#ddply(intersections10daysx, .(Scenario, modelName), i.stat)
 
 rer.stat <- function(x) data.frame(RERMean=round(mean(x$intersections/x$Observed),2), RERSD=round(sd(x$intersections/x$Observed),2))
-ddply(intersectionsx, .(Scenario), rer.stat)
-ddply(intersections2000x, .(Scenario), rer.stat)
-ddply(intersections10daysx, .(Scenario), rer.stat)
-#writeTable(subset(logTable1, Model != "VT"), "table-1a.docx", context)
+i <- ddply(intersectionsx, .(Scenario), rer.stat); i$Scenario=prettyScenario(i$Scenario)
+i2000 <- ddply(intersections2000x, .(Scenario), rer.stat); i2000$Scenario=prettyScenario(i2000$Scenario)
+i10days <- ddply(intersections10daysx, .(Scenario), rer.stat); i10days$Scenario=prettyScenario(i10days$Scenario)
 
+writeTable(i, "table-2a.docx", context)
+writeTable(i2000, "table-2b.docx", context)
+writeTable(i10days, "table-2c.docx", context)
 ```
@@
 plotPopulationCorrelations <- function(populationCorrelations) {
+  populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
   p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
   facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
+  scale_y_continuous(limits=c(-1,1)) +
   theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
@@
 ```{r figure-5-init}
-getCIProportions <- function(scenarios, modelNames, panel, scenarioNames) {
+getCIProportions <- function(scenarios, modelNames, panel, scenarioNames, cutoff=Inf) {
   study <- getStudy(scenario="A")
-  validation <- Validation(study=study, populationSizeOverEstimate=Inf)
+  validation <- Validation(study=study, populationSizeOverEstimate=cutoff)
   a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames)
@@
     geom_point() +
-    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=modelLabels) +
+    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("black", "darkgray"), labels=modelLabels) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
@@
 ```{r figure-5abc}
-cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[1:3], panel="a", scenarioNames=scenarios)
-cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1:3], panel="b", scenarioNames=scenarios)
-cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1:3], panel="c", scenarioNames=scenarios)
+cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[1:3], panel="a", scenarioNames=scenarios, cutoff=cutoff)
+cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1:3], panel="b", scenarioNames=scenarios, cutoff=10*cutoff)
+cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1:3], panel="c", scenarioNames=scenarios, cutoff=cutoff)
+
+#cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[1:3], panel="a", scenarioNames=scenarios)
+#cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1:3], panel="b", scenarioNames=scenarios)
+#cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1:3], panel="c", scenarioNames=scenarios)
+
 cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")

### 2015-02-20 inst/simulation/manuscript.Rmd
...

@@
 #print(p)
-saveFigure(p, "manuscript-figure-1.png", width=8, height=5)
+#saveFigure(p, "manuscript-figure-1.png", width=8, height=5)
+saveFigure(p, "manuscript-figure-1.png", width=17, height=15, dpi=200) # draft quality
 
@@
 getEstimatedDensity = function(scenarios, modelName, iterations, year, sd=FALSE, log=FALSE) {
-  ddply(data.frame(scenario=scenarios, year=rep(year, scenarios), iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, sd) {
+  ddply(data.frame(scenario=scenarios, year=rep(year, length(scenarios)), iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, sd) {
     message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
@@
     density$Model <- modelName
-    density$Year <- year0
+    density$Year <- x$year
 
@@
 p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
-  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent", guide=guide_legend(title=expression(paste("Population density ", (individuals/km^2))))) +
+  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent", guide=guide_legend(title=expression(paste("Population density ", (log[10]~individuals/km^2)))), trans="log10", breaks=c(10^-4.5, 10^-4, 10^-3.5, 10^-3, 10^-2.5), labels=c(expression(10^-4.5), expression(10^-4.0), expression(10^-3.5), expression(10^-3.0), expression(10^-2.5)), limits=(10^-4.5, 10^-2.5)) +
   geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
@@
   facet_grid(~scenario) + coord_equal() + theme_raster(base_size=10) + theme(legend.position="bottom")
-print(p)
-saveFigure(p, "manuscript-figure-2a.png", width=6,height=3)
+#print(p)
+#saveFigure(p, "manuscript-figure-2a.png", width=6,height=3)
+saveFigure(p, "manuscript-figure-2a.png", width=17, height=8, dpi=200) # draft quality
 
@@
 print(p)
-saveFigure(p, "manuscript-figure-2b.png", width=6,height=3)
+#saveFigure(p, "manuscript-figure-2b.png", width=6,height=3)
+saveFigure(p, "manuscript-figure-2b.png", width=17, height=8, dpi=200) # draft quality
+
 ```
@@
 
-cutoff <- 1000
-cutoffPopulationSizes <- function(populationSizes, cutoff=Inf) subset(populationSizes, Estimated <= cutoff)
+cutoffPopulationSizes <- function(populationSizes, cutoff=c(0,Inf)) subset(populationSizes, cutoff[1] <= Estimated & Estimated <= cutoff[2])
 
@@
 
-plotLogPopulationSize <- function(size, modelName, xlim, ylim, cutoff) {
+plotLogPopulationSize <- function(size, modelName, xlim, ylim) {
   library(ggplot2)
@@
     xlab("True population size") + ylab("Estimated population size") + ggtitle(paste("Model: ", prettyModelName(modelName[1]))) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=12))
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+                                        plot.title=element_text(size=10),
+                                        plot.margin=unit(c(0, 0, 0, 0), units="line"), #top, right, bottom, and left margins
+                                        legend.margin=unit(0, units="line"),
+                                        panel.margin=unit(0, units="line"))
 
@@
 
+cutoff <- c(1,1000)
+
 meanPopulationSizes  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff))
@@
 
-xlim <- log10(10^c(0,4)); ylim <- xlim
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), subset(logMeanPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3a1.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), subset(logMeanPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3a2.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), subset(logMeanPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3a3.png", width=8, height=4)
+xlim <- log10(c(1,6000)); ylim <- xlim
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3a1.png")
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3a2.png")
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3a3.png")
 
-#xlim <- log10(10^c(2,4)); ylim <- xlim
-xlim <- log10(10^c(0,4)); ylim <- xlim
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), subset(logMeanPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3b1.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), subset(logMeanPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3b2.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), subset(logMeanPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3b3.png", width=8, height=4)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3b1.png")
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3b2.png")
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3b3.png")
 
-xlim <- log10(10^c(0,4)); ylim <- xlim
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), subset(logMeanPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3c1.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), subset(logMeanPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3c2.png", width=8, height=4)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), subset(logMeanPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3c3.png", width=8, height=4)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3c1.png")
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3c2.png")
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3c3.png")
 ```
@@
   x <- ddply(populationSizes, .(Scenario, modelName), function(x, overEstimate) {
-    failed <- round((sum(x$Estimated > overEstimate) / nrow(x)) * 100)
-    x <- subset(x, Estimated <= overEstimate)
-    if (x$modelName[1] == "FMPModel") failed <- NA
-    if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA, Failed=failed, MeanProp=NA, SDProp=NA))
+    failed <- round((1 - nrow(cutoffPopulationSizes(x, overEstimate)) / nrow(x)) * 100)
+      #(sum(x$Estimated < overEstimate[1] | x$Estimated > overEstimate[2]) / nrow(x)) * 100)
+    x <- cutoffPopulationSizes(x, overEstimate)
+    #subset(x, Estimated <= overEstimate)
+    #if (x$modelName[1] == "FMPModel") failed <- NA
+    if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA, Failed=failed, EstTrue=NA, EstTrueSD=NA, N=NA))
     y <- lm(Estimated~Observed, x)
@@
     prop.sd <- sd(x$Estimated/x$Observed)
-    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2), MeanProp=prop.mean, SDProp=prop.sd, Failed=failed))
+    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2), EstTrue=prop.mean, EstTrueSD=prop.sd, Failed=failed, N=nrow(x)))
   }, overEstimate=overEstimate)
@@
   y$Scenario <- y$scenario
-  #y$EstDev <- y$Estimated / y$True
   z <- plyr::join(x, y)
   z$Model <- prettyModelName(z$modelName)
-  z$Scenario <- substr(z$Scenario, 1, 1)
+  z$Scenario <- prettyScenario(z$Scenario) #substr(z$Scenario, 1, 1)
   if (full) return(z)
 
-  z <- z[,c("Scenario","Model","Slope","R2","MeanProp","SDProp","Failed")]
-  #z$EstDev <- round(z$EstDev, 2)
-  z$MeanProp <- round(z$MeanProp, 2)
-  z$SDProp <- round(z$SDProp, 2)
+  z <- z[,c("Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed","N")]
+  #z$MeanProp <- round(z$MeanProp, 2)
+  #z$SDProp <- round(z$SDProp, 2)
   if (!missing(file)) write.csv(z, file=file)
@@
 
-writeTable <- function(df, filename, context) {
-  library(ReporteRs)
-  doc <- docx()
-  flextable <- FlexTable(df)
-  addFlexTable(doc, flextable)
-  dir <- normalizePath(context$figuresDirectory)
-  writeDoc(doc, file=file.path(dir, filename))
+#writeTable <- function(df, filename, context) {
+#  library(ReporteRs)
+#  doc <- docx()
+#  flextable <- FlexTable(df)
+#  addFlexTable(doc, flextable)
+#  dir <- normalizePath(context$figuresDirectory)
+#  writeDoc(doc, file=file.path(dir, filename))
+#}
+
+plotPopulationSizesTable <- function(logTable) {
+  library(reshape2)
+  x <- melt(logTable, id.vars=c("Scenario", "Model", "Case"), measure.vars=c("Slope", "R2", "EstTrue", "Failed"))
+  p <- ggplot(x, aes(Scenario, value, group=Model)) + geom_point(aes(shape=Model)) + geom_line() +
+    facet_grid(variable~Case, scales="free_y") + ylab("") +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=10),
+                                        strip.text.y=element_text(size=10), legend.position="bottom",
+                                        plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
+                                        legend.margin=unit(0, units="line"))# , legend.key=element_rect(fill="transparent"))
+  return(p)
 }
@@
 
-#dir <- normalizePath(context$figuresDirectory)
-#write.csv(logTable1, file=file.path(dir, "table-1a.csv"))
-#write.csv(logTable2, file=file.path(dir, "table-2a.csv"))
-#write.csv(logTable3, file=file.path(dir, "table-3a.csv"))
+logTable1$Case <- "a"
+logTable2$Case <- "b"
+logTable3$Case <- "c"
+logTable <- rbind(subset(logTable1, Model != "VT"), subset(logTable2, Model != "VT"), subset(logTable3, Model != "VT"))
+p <- plotPopulationSizesTable(logTable)
+saveFigure(p, "manuscript-figure-3.png", width=17, height=17, dpi=200)
 
-writeTable(subset(logTable1, Model != "VT"), "table-1a.docx", context)
-writeTable(subset(logTable2, Model != "VT"), "table-1b.docx", context)
-writeTable(subset(logTable3, Model != "VT"), "table-1c.docx", context)
+#writeTable(subset(logTable1, Model != "VT"), "table-1a.docx", context)
+#writeTable(subset(logTable2, Model != "VT"), "table-1b.docx", context)
+#writeTable(subset(logTable3, Model != "VT"), "table-1c.docx", context)
 ```
@@
 ```{r table-2}
+plotIntersectionPopulationSizeTable <- function(ratioTable) {
+  library(reshape2)
+  library(grid)
+  x <- melt(ratioTable, id.vars=c("Scenario", "Case"), measure.vars=c("SizeCount"))
+  p <- ggplot(x, aes(Scenario, value, group=Case)) + geom_point(aes(shape=Case)) + geom_line() +
+    ylab("") + #guides(shape=guide_legend()) +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=10),
+                                        strip.text.y=element_text(size=10), legend.position="bottom",
+                                        plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
+                                        legend.margin=unit(0, units="line"),
+                                        panel.margin=unit(0, units="line"))# , legend.key=element_rect(fill="transparent"))
+  return(p) 
+}
+p <- plotIntersectionPopulationSizeTable(iall)
+saveFigure(p, "manuscript-figure-4x.png")
+
 intersections <- getIntersections(scenarios)
@@
 intersections10days <- getIntersections(scenarios=paste0(scenarios, "10days"))
+intersections10days$intersections <- intersections10days$intersections / 10
 intersectionsx <- plyr::join(populationSizes, intersections)
@@
 intersections10daysx <- plyr::join(populationSizes10Days, intersections10days)
-#i.stat <- function(x) data.frame(o.mean=mean(x$Observed), o.sd=sd(x$Observed), e.mean=mean(x$Estimated), e.sd=sd(x$Estimated),
-#                                 pp.mean=round(mean(x$Estimated/x$Observed), 2), pp.sd=round(sd(x$Estimated/x$Observed), 2),
-#                                 i.mean=mean(x$intersections), i.sd=sd(x$intersections), ip.mean=round(mean(x$intersections/x$Observed),2), ip.sd=round(sd(x$intersections/x$Observed),2))
-#ddply(intersectionsx, .(Scenario, modelName), i.stat)
-#ddply(intersections2000x, .(Scenario, modelName), i.stat)
-#ddply(intersections10daysx, .(Scenario, modelName), i.stat)
 
-rer.stat <- function(x) data.frame(RERMean=round(mean(x$intersections/x$Observed),2), RERSD=round(sd(x$intersections/x$Observed),2))
-i <- ddply(intersectionsx, .(Scenario), rer.stat); i$Scenario=prettyScenario(i$Scenario)
-i2000 <- ddply(intersections2000x, .(Scenario), rer.stat); i2000$Scenario=prettyScenario(i2000$Scenario)
-i10days <- ddply(intersections10daysx, .(Scenario), rer.stat); i10days$Scenario=prettyScenario(i10days$Scenario)
+rer.stat <- function(x) data.frame(SizeCount=round(mean(x$intersections/x$Observed),2), SizeCountSD=round(sd(x$intersections/x$Observed),2))
+i <- ddply(intersectionsx, .(Scenario), rer.stat); i$Scenario=prettyScenario(i$Scenario); i$Case <- "a"
+i2000 <- ddply(intersections2000x, .(Scenario), rer.stat); i2000$Scenario=prettyScenario(i2000$Scenario); i2000$Case <- "b"
+i10days <- ddply(intersections10daysx, .(Scenario), rer.stat); i10days$Scenario=prettyScenario(i10days$Scenario); i10days$Case <- "c"
 
-writeTable(i, "table-2a.docx", context)
-writeTable(i2000, "table-2b.docx", context)
-writeTable(i10days, "table-2c.docx", context)
+iall <- rbind(i, i2000, i10days)
+p <- plotIntersectionPopulationSizeTable(iall)
+saveFigure(p, "manuscript-figure-4x.png")
+
+#writeTable(i, "table-2a.docx", context)
+#writeTable(i2000, "table-2b.docx", context)
+#writeTable(i10days, "table-2c.docx", context)
 ```
@@
   scale_y_continuous(limits=c(-1,1)) +
-  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
+  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+                                      plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
+                                      legend.margin=unit(0, units="line"),
+                                      panel.margin=unit(0, units="line"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
   return(p)
@@
 p1 <- plotPopulationCorrelations(pc1)
-saveFigure(p1, "manuscript-figure-4a.png", width=4, height=4)
+saveFigure(p1, "manuscript-figure-4a.png")
 p2 <- plotPopulationCorrelations(pc2)
-saveFigure(p2, "manuscript-figure-4b.png", width=4, height=4)
+saveFigure(p2, "manuscript-figure-4b.png")
 p3 <- plotPopulationCorrelations(pc3)
-saveFigure(p3, "manuscript-figure-4c.png", width=4, height=4)
+saveFigure(p3, "manuscript-figure-4c.png")
 ```
@@
     facet_wrap(~panel) + ylim(0, 1) +
-    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank())
+    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
+          plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
+          legend.margin=unit(0, units="line"),
+          panel.margin=unit(0, units="line"),
+          legend.position="bottom")
   return(p)
@@
 
-#cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[1:3], panel="a", scenarioNames=scenarios)
-#cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1:3], panel="b", scenarioNames=scenarios)
-#cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1:3], panel="c", scenarioNames=scenarios)
-
 cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")
@@
 saveFigure(p, "manuscript-figure-5.png", width=6, height=2)
-
-#p1 <- plotCIProportions(cip1)
-#saveFigure(p1, "manuscript-figure-5a.png", width=4, height=2)
-#p2 <- plotCIProportions(cip2)
-#saveFigure(p2, "manuscript-figure-5b.png", width=4, height=2)
-#p3 <- plotCIProportions(cip3)
-#saveFigure(p3, "manuscript-figure-5c.png", width=4, height=2)
 ```

### 2015-02-21 inst/simulation/manuscript.Rmd
...

@@
 
-p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
-  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent", guide=guide_legend(title=expression(paste("Population density ", (log[10]~individuals/km^2)))), trans="log10", breaks=c(10^-4.5, 10^-4, 10^-3.5, 10^-3, 10^-2.5), labels=c(expression(10^-4.5), expression(10^-4.0), expression(10^-3.5), expression(10^-3.0), expression(10^-2.5)), limits=(10^-4.5, 10^-2.5)) +
+#library(directlabels)
+
+p <- ggplot(densitiesScaled) +
+  geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
+  stat_contour(aes_string(x="x", y="y", z=yearLayer, colour=yearLayer), size=0.1)
+#p <- direct.label(p)
+p <- p + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent",
+                       guide=guide_legend(title=expression(paste("Log population density ", (individuals/km^2)))), trans="log10",
+                       #breaks=c(10^-4.5, 10^-4, 10^-3.5, 10^-3, 10^-2.5),
+                       breaks=trans_breaks("log10", function(x) 10^x),
+                       #labels=c(expression(10^-4.5), expression(10^-4.0), expression(10^-3.5), expression(10^-3.0), expression(10^-2.5)), limits=c(10^-4.5, 10^-2.5)
+                       labels=trans_format("log10", math_format(10^.x))
+                       ) +
   geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
@@
   geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black"), guide=FALSE) +
-  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
-  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
-  facet_grid(~scenario) + coord_equal() + theme_raster(base_size=10) + theme(legend.position="bottom")
-#print(p)
-#saveFigure(p, "manuscript-figure-2a.png", width=6,height=3)
-saveFigure(p, "manuscript-figure-2a.png", width=17, height=8, dpi=200) # draft quality
+  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0.01,0.01)) +
+  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0.01,0.01)) +
+  facet_grid(~scenario) + coord_equal() + theme_raster(base_size=10) +
+  theme(legend.position="bottom", legend.margin=unit(-2, units="line"),
+        plot.margin=unit(c(-.5, 0, 0, -.5), units="line"))
+#p <- direct.label(p)
+saveFigure(p, "manuscript-figure-2a.png", width=17, height=7.2, dpi=200) # draft quality
 
@@
   scale_y_continuous(breaks=seq(gridTemplate@extent@ymin, gridTemplate@extent@ymax, 1e5), expand=c(0,0)) +
-  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal() + theme(panel.grid.major=element_line(colour="grey"))
-print(p)
-#saveFigure(p, "manuscript-figure-2b.png", width=6,height=3)
-saveFigure(p, "manuscript-figure-2b.png", width=17, height=8, dpi=200) # draft quality
+  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal() +
+  theme(panel.grid.major=element_line(colour="grey"), plot.margin=unit(c(0, 0, 0, -.5), units="line"))
+saveFigure(p, "manuscript-figure-2b.png", width=17, height=6.7, dpi=200) # draft quality
 
@@
 
-plotLogPopulationSize <- function(size, modelName, xlim, ylim) {
+plotLogPopulationSize <- function(size, modelName, case, xlim, ylim) {
   library(ggplot2)
@@
     scale_y_continuous(limits=ylim, labels=trans_format("identity", math_format(10^.x))) +
-    xlab("True population size") + ylab("Estimated population size") + ggtitle(paste("Model: ", prettyModelName(modelName[1]))) +
+    xlab("Log true population size") + ylab("Log estimated population size") + ggtitle(paste("Case:", case, " Model:", prettyModelName(modelName[1]))) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
@@
 }
-```
-
-```{r figure-3-init-2-test, eval=FALSE}
-populationSizes <- getPopulationSizes(scenarios=scenarios[1:4], modelNames=modelNames[3:4])
-populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios[1:4], "combined"), modelNames=modelNames[3:4])
-populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios[1:4], "10days"), modelNames=modelNames[3:4])
 ```
@@
 xlim <- log10(c(1,6000)); ylim <- xlim
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], "a", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a1.png")
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], "a", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a2.png")
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], "a", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a3.png")
 
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], "b", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b1.png")
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], "b", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b2.png")
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], "b", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b3.png")
 
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], "c", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c1.png")
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], "c", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c2.png")
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], "c", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c3.png")
@@
   p <- ggplot(x, aes(Scenario, value, group=Case)) + geom_point(aes(shape=Case)) + geom_line() +
-    ylab("") + #guides(shape=guide_legend()) +
+    ylab("Ratio") + #guides(shape=guide_legend()) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=10),
-                                        strip.text.y=element_text(size=10), legend.position="bottom",
-                                        plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
-                                        legend.margin=unit(0, units="line"),
+                                        strip.text.y=element_text(size=10), legend.position="right",
+                                        #legend.background=element_rect(fill="transparent"),
+                                        plot.margin=unit(c(0.2, 0, 0, 0), units="line"),
+                                        legend.margin=unit(-1, units="line"),
                                         panel.margin=unit(0, units="line"))# , legend.key=element_rect(fill="transparent"))
@@
 }
-p <- plotIntersectionPopulationSizeTable(iall)
-saveFigure(p, "manuscript-figure-4x.png")
 
@@
 
-rer.stat <- function(x) data.frame(SizeCount=round(mean(x$intersections/x$Observed),2), SizeCountSD=round(sd(x$intersections/x$Observed),2))
+rer.stat <- function(x) data.frame(SizeCount=mean(x$intersections/x$Observed), SizeCountSD=sd(x$intersections/x$Observed))
 i <- ddply(intersectionsx, .(Scenario), rer.stat); i$Scenario=prettyScenario(i$Scenario); i$Case <- "a"
@@
 p <- plotIntersectionPopulationSizeTable(iall)
-saveFigure(p, "manuscript-figure-4x.png")
+saveFigure(p, "manuscript-figure-4x.png", height=7.87/2.5)
 
@@
 
-plotPopulationCorrelations <- function(populationCorrelations) {
+plotPopulationCorrelations <- function(populationCorrelations, case) {
   populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
@@
   scale_y_continuous(limits=c(-1,1)) +
+  ggtitle(paste("Case:", case)) +
   theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
-                                      plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
-                                      legend.margin=unit(0, units="line"),
+                                      plot.title=element_text(size=10),
+                                      plot.margin=unit(c(0, 0, 0, 0), units="line"),
                                       panel.margin=unit(0, units="line"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
@@
 
-p1 <- plotPopulationCorrelations(pc1)
+p1 <- plotPopulationCorrelations(pc1, "a")
 saveFigure(p1, "manuscript-figure-4a.png")
-p2 <- plotPopulationCorrelations(pc2)
+p2 <- plotPopulationCorrelations(pc2, "b")
 saveFigure(p2, "manuscript-figure-4b.png")
-p3 <- plotPopulationCorrelations(pc3)
+p3 <- plotPopulationCorrelations(pc3, "c")
 saveFigure(p3, "manuscript-figure-4c.png")
@@
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
-    xlab("Scenario") + ylab("Proportion within CI") +
+    xlab("Scenario") + ylab("Proportion") +
     facet_wrap(~panel) + ylim(0, 1) +
     theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
-          plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
-          legend.margin=unit(0, units="line"),
+          plot.margin=unit(c(0, 0, 0, 0), units="line"),
+          #legend.background=element_rect(fill="transparent"),
+          legend.margin=unit(-2, units="line"),
           panel.margin=unit(0, units="line"),
-          legend.position="bottom")
+          legend.position="bottom",
+          legend.box="horizontal")
   return(p)
@@
 p <- plotCIProportions(cip)
-saveFigure(p, "manuscript-figure-5.png", width=6, height=2)
+saveFigure(p, "manuscript-figure-5.png", height=7.87/2)
 ```

### 2015-03-07 inst/simulation/manuscript.Rmd
...

@@
 library(ggplot2)
+library(scales)
+library(grid)
 library(plyr)
@@
 subset(aggrates, year==2020)
+
+###
+
+library(raster)
+mss <- getMSS("E")
+study <- mss$study
+surveyRoutes <- mss$getSurveyRoutes()
+#habitatTypes <- raster::extract(habitat, surveyRoutes$surveyRoutes[1:3])
+
+study$studyArea$readRasterIntoMemory()
+habitat <- study$studyArea$habitat
+habitatTypes <- raster::extract(habitat, surveyRoutes$surveyRoutes)
+habitatWeights <- CORINEHabitatWeights$new()
+
+forestProportions <- c()
+for (i in seq_along(habitatTypes)) {
+  classifiedHabitatTypes <- habitatWeights$classify(habitatTypes[[i]]) == 3 # Forest
+  forestProportions <- c(forestProportions, sum(classifiedHabitatTypes) / length(classifiedHabitatTypes))
+}
+
+#mean(forestProportion)
+sum(forestProportion * surveyRoutes$lengths) / sum(surveyRoutes$lengths) # = 0.9404947
+
+###
+
+study$studyArea$boundary@polygons[[1]]@area
+habitatCut <- raster::mask(habitat, study$studyArea$boundary)
+habitatWeights <- CORINEHabitatWeights$new(study=study)
+habitatFreq <- freq(habitatCut, progress="text")
+x <- habitatFreq
+
+p <- c()
+for (i in 1:50) {
+  habitatPreferences <- HabitatSelection$new(study=study, iteration=as.integer(i))
+  fileName <- habitatPreferences$getHabitatSelectionFileName()
+  habitatSelection <- try(habitatPreferences$loadHabitatSelection())
+  if (inherits(habitatSelection, "try-error")) next
+  habitatWeights$setHabitatSelectionWeights(habitatSelection)
+
+  y <- subset(habitatWeights$weights, habitat %in% x[,1])
+  effectiveArea <- sum(y$weight * x[,2]) * prod(res(habitat))
+  fullArea <- sum(x[-nrow(x),]) * prod(res(habitat))
+
+  p <- c(p, effectiveArea / fullArea)
+}
+mean(1/p) # = 1.286095
+
+p <- c()
+for (i in 1:50) {
+  estimates <- study$getModel(modelName=modelNames[4], iteration=as.integer(i))
+  popsize <- try(study$getPopulationSize(estimates=estimates, save=FALSE))
+  if (inherits(popsize, "try-error")) next
+  p <- c(p, popsize$sizeData$Estimated)
+}
+
 ```
@@
 p <- p + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent",
-                       guide=guide_legend(title=expression(paste("Log population density ", (individuals/km^2)))), trans="log10",
+                       guide=guide_legend(title=expression(paste("Population density ", (individuals/km^2)))), trans="log10",
                        #breaks=c(10^-4.5, 10^-4, 10^-3.5, 10^-3, 10^-2.5),
@@
 saveFigure(p, "manuscript-figure-2a.png", width=17, height=7.2, dpi=200) # draft quality
+saveFigure(p, "manuscript-figure-2a.pdf", width=17, height=7.2, dpi=300)
 
@@
   geom_point(data=surveyRoutes, aes(x, y, colour=intersections), size=1) + scale_color_manual(values=c("darkgrey", "black")) +
-  scale_x_continuous(breaks=seq(gridTemplate@extent@xmin, gridTemplate@extent@xmax, 1e5), expand=c(0,0)) +
-  scale_y_continuous(breaks=seq(gridTemplate@extent@ymin, gridTemplate@extent@ymax, 1e5), expand=c(0,0)) +
+  scale_x_continuous(breaks=seq(gridTemplate@extent@xmin, gridTemplate@extent@xmax, 1e5), expand=c(0.01,0.01)) +
+  scale_y_continuous(breaks=seq(gridTemplate@extent@ymin, gridTemplate@extent@ymax, 1e5), expand=c(0.01,0.01)) +
   facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal() +
@@
 saveFigure(p, "manuscript-figure-2b.png", width=17, height=6.7, dpi=200) # draft quality
-
+saveFigure(p, "manuscript-figure-2b.pdf", width=17, height=6.7, dpi=300)
 ```
@@
 }
-
-plotLogPopulationSize <- function(size, modelName, case, xlim, ylim) {
-  library(ggplot2)
-  library(scales)  
-  size$Scenario <- prettyScenario(size$Scenario)
-  meanSize <- getMeanPopulationSizes(size)
-  meanSize$Scenario <- prettyScenario(meanSize$Scenario)
-
-  p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
-    geom_point(size=1, colour="darkgray", alpha=.25) +
-    geom_abline(colour="black", linetype="longdash") + 
-    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black")
-
-  #if (!any(abs(size$Estimated) == Inf)) p <- p + geom_smooth(method="lm", se=F, fullrange=T, colour="black")
-  ls <- ddply(size, .(Scenario), function(x) if (!any(abs(x$Estimated) == Inf)) coef(lm(Estimated~Observed, x)) else c(NA,NA))
-  p <- p + geom_abline(data=ls, aes(intercept=`(Intercept)`, slope=Observed))
-
-  p <- p + facet_wrap(~Scenario) +
-    scale_x_continuous(limits=xlim, labels=trans_format("identity", math_format(10^.x))) +
-    scale_y_continuous(limits=ylim, labels=trans_format("identity", math_format(10^.x))) +
-    xlab("Log true population size") + ylab("Log estimated population size") + ggtitle(paste("Case:", case, " Model:", prettyModelName(modelName[1]))) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
-                                        plot.title=element_text(size=10),
-                                        plot.margin=unit(c(0, 0, 0, 0), units="line"), #top, right, bottom, and left margins
-                                        legend.margin=unit(0, units="line"),
-                                        panel.margin=unit(0, units="line"))
-
-  print(p)
-  return(p)
-}
 ```
@@
 registerDoMC(cores=10) # The nodes seem to run out of memory and fail with a bogus error message, so it's better to reduce the number of used cores.
-populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames)
-populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
-populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)
+populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames[c(1,3,4)])
+populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,4)])
+populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4)])
 registerDoMC(cores=detectCores())
@@
 ```{r figure-3abc}
+plotLogPopulationSize <- function(size, modelName, case, xlim, ylim) {
+  library(ggplot2)
+  library(scales)
+  library(grid)
+
+  size$Scenario <- prettyScenario(size$Scenario)
+  meanSize <- getMeanPopulationSizes(size)
+  meanSize$Scenario <- prettyScenario(meanSize$Scenario)
+
+  p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
+    geom_point(size=1, colour="darkgray", alpha=.25) +
+    geom_abline(colour="black", linetype="longdash") + 
+    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black")
+
+  #if (!any(abs(size$Estimated) == Inf)) p <- p + geom_smooth(method="lm", se=F, fullrange=T, colour="black")
+  ls <- ddply(size, .(Scenario), function(x) if (!any(abs(x$Estimated) == Inf)) coef(lm(Estimated~Observed, x)) else c(NA,NA))
+  p <- p + geom_abline(data=ls, aes(intercept=`(Intercept)`, slope=Observed))
+
+  p <- p + facet_wrap(~Scenario) +
+    #scale_x_continuous(limits=xlim, labels=trans_format("identity", math_format(10^.x))) +
+    #scale_y_continuous(limits=ylim, labels=trans_format("identity", math_format(10^.x))) +
+    scale_x_continuous(limits=xlim, labels=trans_format("identity", math_format(.x))) +
+    scale_y_continuous(limits=ylim, labels=trans_format("identity", math_format(.x))) +
+    xlab(expression(paste(Log[10], " true population size"))) + ylab(expression(paste(Log[10], " estimated population size"))) + ggtitle(paste("Case:", case, " Model:", prettyModelName(modelName[1]))) +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+                                        plot.title=element_text(size=10),
+                                        plot.margin=unit(c(0, 0.2, 0, 0), units="line"), #top, right, bottom, and left margins
+                                        legend.margin=unit(0, units="line"),
+                                        panel.margin=unit(0, units="line"))
+
+  return(p)
+}
+
 #modelNames <- modelNames[c(2:4,1)]
 
-xlim <- log10(c(1,6000)); ylim <- xlim
+xlim <- log10(c(50,1200)); ylim <- xlim
 p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], "a", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a1.png")
+saveFigure(p, "manuscript-figure-3a1.pdf", dpi=300)
 p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], "a", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a2.png")
+saveFigure(p, "manuscript-figure-3a2.pdf", dpi=300)
 p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], "a", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a3.png")
+saveFigure(p, "manuscript-figure-3a3.pdf", dpi=300)
 
+xlim <- log10(c(500,4000)); ylim <- xlim
 p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], "b", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b1.png")
+saveFigure(p, "manuscript-figure-3b1.pdf", dpi=300)
 p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], "b", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b2.png")
+saveFigure(p, "manuscript-figure-3b2.pdf", dpi=300)
 p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], "b", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b3.png")
+saveFigure(p, "manuscript-figure-3b3.pdf", dpi=300)
 
+xlim <- log10(c(5,1200)); ylim <- xlim
 p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], "c", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c1.png")
+saveFigure(p, "manuscript-figure-3c1.pdf", dpi=300)
 p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], "c", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c2.png")
+saveFigure(p, "manuscript-figure-3c2.pdf", dpi=300)
 p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], "c", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c3.png")
+saveFigure(p, "manuscript-figure-3c3.pdf", dpi=300)
+```
+
+```{r figure-3sample}
+xlim <- log10(c(10,800)); ylim <- xlim
+logPopulationSizesSample <- subset(logPopulationSizes, scenario == "A")
+logPopulationSizesSample$modelName <- prettyModelName(logPopulationSizesSample$modelName)
+logPopulationSizesSample <- rbind(
+  subset(logPopulationSizesSample, modelName=="ST"),
+  subset(logPopulationSizesSample, modelName=="T"),
+  subset(logPopulationSizesSample, modelName=="FMP"))
+
+#p <- plotLogPopulationSize(subset(logPopulationSizesSample, modelName==modelNames[c(1,3,4)]), modelName=modelNames[c(1,3,4)], "a", xlim=xlim, ylim=ylim)
+
+plotLogPopulationSizeSample <- function(size, modelName, case, xlim, ylim) {
+  library(ggplot2)
+  library(scales)  
+  size$Scenario <- prettyScenario(size$Scenario)
+  meanSize <- getMeanPopulationSizes(size)
+  meanSize$Scenario <- prettyScenario(meanSize$Scenario)
+
+  p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
+    geom_point(size=1, colour="darkgray", alpha=.25) +
+    geom_abline(colour="black", linetype="longdash") + 
+    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black")
+
+  #if (!any(abs(size$Estimated) == Inf)) p <- p + geom_smooth(method="lm", se=F, fullrange=T, colour="black")
+  ls <- ddply(size, .(modelName), function(x) if (!any(abs(x$Estimated) == Inf)) coef(lm(Estimated~Observed, x)) else c(NA,NA))
+  p <- p + geom_abline(data=ls, aes(intercept=`(Intercept)`, slope=Observed))
+
+  p <- p + facet_wrap(~modelName) +
+    scale_x_continuous(limits=xlim, labels=trans_format("identity", math_format(10^.x))) +
+    scale_y_continuous(limits=ylim, labels=trans_format("identity", math_format(10^.x))) +
+    xlab("True population size") + ylab("Estimated population size") +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+                                        plot.title=element_text(size=10),
+                                        plot.margin=unit(c(0, 0.1, 0, 0), units="line"), #top, right, bottom, and left margins
+                                        legend.margin=unit(0, units="line"),
+                                        panel.margin=unit(0, units="line"))
+
+  return(p)
+}
+
+p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName==c("ST","T","FMP")), modelName=c("ST","T","FMP"), "a", xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3sample.png", height=7.87*1.2/2) # draft quality
+saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87*1.2/2, dpi=300)
 ```
@@
   x <- ddply(populationSizes, .(Scenario, modelName), function(x, overEstimate) {
-    failed <- round((1 - nrow(cutoffPopulationSizes(x, overEstimate)) / nrow(x)) * 100)
+    failed <- (1 - nrow(cutoffPopulationSizes(x, overEstimate)) / nrow(x) * 100)
       #(sum(x$Estimated < overEstimate[1] | x$Estimated > overEstimate[2]) / nrow(x)) * 100)
@@
     #if (x$modelName[1] == "FMPModel") failed <- NA
-    if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA, Failed=failed, EstTrue=NA, EstTrueSD=NA, N=NA))
+    #if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, SlopeSD=NA, R2=NA, Failed=failed, EstTrue=NA, EstTrueSD=NA, EstTrueMin=NA, EstTrueMax=NA, N=NA))
     y <- lm(Estimated~Observed, x)
-    prop.mean <- mean(x$Estimated/x$Observed)
-    prop.sd <- sd(x$Estimated/x$Observed)
-    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2), EstTrue=prop.mean, EstTrueSD=prop.sd, Failed=failed, N=nrow(x)))
+    prop.mean <- mean(log(x$Estimated/x$Observed))
+    prop.sd <- sd(log(x$Estimated/x$Observed))
+    prop.min <- quantile(log(x$Estimated/x$Observed), .05)
+    prop.max <- quantile(log(x$Estimated/x$Observed), .95)
+    slope.sd <- coef(summary(y))[2,2]
+    return(data.frame(Intercept=coef(y)[1], Slope=coef(y)[2], SlopeSD=slope.sd, R2=summary(y)$r.squared, EstTrue=prop.mean, EstTrueSD=prop.sd, EstTrueMin=prop.min, EstTrueMax=prop.max, Failed=failed, N=nrow(x)))
   }, overEstimate=overEstimate)
@@
   z$Model <- prettyModelName(z$modelName)
-  z$Scenario <- prettyScenario(z$Scenario) #substr(z$Scenario, 1, 1)
+  z$Scenario <- prettyScenario(z$Scenario)
   if (full) return(z)
 
-  z <- z[,c("Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed","N")]
-  #z$MeanProp <- round(z$MeanProp, 2)
-  #z$SDProp <- round(z$SDProp, 2)
-  if (!missing(file)) write.csv(z, file=file)
   return(z)
@@
 
-#writeTable <- function(df, filename, context) {
-#  library(ReporteRs)
-#  doc <- docx()
-#  flextable <- FlexTable(df)
-#  addFlexTable(doc, flextable)
-#  dir <- normalizePath(context$figuresDirectory)
-#  writeDoc(doc, file=file.path(dir, filename))
-#}
+table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff)
+logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(cutoff+1))
+logTable1$Failed <- table1$Failed
+
+table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff)
+logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1))
+logTable2$Failed <- table2$Failed
+
+table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff)
+logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1))
+logTable3$Failed <- table3$Failed
+
+subset(logTable1, Model != "VT")
+subset(logTable2, Model != "VT")
+subset(logTable3, Model != "VT")
+
+writeTable <- function(df, filename, context) {
+  library(ReporteRs)
+  doc <- docx()
+  flextable <- FlexTable(df)
+  addFlexTable(doc, flextable)
+  dir <- normalizePath(context$figuresDirectory)
+  writeDoc(doc, file=file.path(dir, filename))
+}
 
@@
   library(reshape2)
+  facet_labeller <- function(variable, value) {
+    measured <- list(Slope=expression(beta), R2=expression(R^2), EstTrue=expression(alpha), Failed="Failed (%)")
+    print(paste0(variable,"=",value))
+    if (variable=="variable") return(measured[value])
+    if (variable=="Case") return(paste("Case", value))
+  }
   x <- melt(logTable, id.vars=c("Scenario", "Model", "Case"), measure.vars=c("Slope", "R2", "EstTrue", "Failed"))
-  p <- ggplot(x, aes(Scenario, value, group=Model)) + geom_point(aes(shape=Model)) + geom_line() +
-    facet_grid(variable~Case, scales="free_y") + ylab("") +
+  bestFit <- data.frame(variable=unique(x$variable), y=c(1, 1, 0, 0))
+  errorBars <- x
+  errorBars$value <- NA
+  errorBars$value[errorBars$variable=="Slope"] <- logTable$Slope
+  errorBars$qmin[errorBars$variable=="Slope"] <- logTable$Slope-logTable$SlopeSD
+  errorBars$qmax[errorBars$variable=="Slope"] <- logTable$Slope+logTable$SlopeSD
+  errorBars$value[errorBars$variable=="EstTrue"] <- logTable$EstTrue
+  errorBars$qmin[errorBars$variable=="EstTrue"] <- logTable$EstTrueMin
+  errorBars$qmax[errorBars$variable=="EstTrue"] <- logTable$EstTrueMax
+
+  p <- ggplot(x, aes(Scenario, value, group=Model)) +
+    geom_hline(data=bestFit, aes(yintercept=y), color="darkgray") + 
+    geom_errorbar(data=errorBars, aes(Scenario, group=Model, ymax=qmax, ymin=qmin), position=position_dodge(width=.7), colour="darkgray") +
+    geom_point(aes(shape=Model), position=position_dodge(width=.7)) +
+    facet_grid(variable~Case, scales="free_y", labeller=facet_labeller) + ylab("") +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=10),
-                                        strip.text.y=element_text(size=10), legend.position="bottom",
+                                        legend.position="bottom",
+                                        strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
                                         plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
@@
 
-table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff)
-logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(cutoff+1))
-logTable1$MeanProp <- table1$MeanProp; logTable1$SDProp <- table1$SDProp; logTable1$Failed <- table1$Failed
-
-table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff)
-logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1))
-logTable2$MeanProp <- table2$MeanProp; logTable2$SDProp <- table2$SDProp; logTable2$Failed <- table2$Failed
-
-table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff)
-logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1))
-logTable3$MeanProp <- table3$MeanProp; logTable3$SDProp <- table3$SDProp; logTable3$Failed <- table3$Failed
-
-subset(logTable1, Model != "VT")
-subset(logTable2, Model != "VT")
-subset(logTable3, Model != "VT")
-
-logTable1$Case <- "a"
-logTable2$Case <- "b"
-logTable3$Case <- "c"
+logTable1$Case <- "1"
+logTable2$Case <- "2"
+logTable3$Case <- "3"
 logTable <- rbind(subset(logTable1, Model != "VT"), subset(logTable2, Model != "VT"), subset(logTable3, Model != "VT"))
 p <- plotPopulationSizesTable(logTable)
-saveFigure(p, "manuscript-figure-3.png", width=17, height=17, dpi=200)
+saveFigure(p, "manuscript-figure-3.png", width=17, height=17, dpi=200) # draft quality
+saveFigure(p, "manuscript-figure-3.pdf", width=17, height=17, dpi=300)
+
+logTablePretty <- logTable
+logTablePretty$Slope <- round(logTablePretty$Slope, 2)
+logTablePretty$R2 <- round(logTablePretty$R2, 2)
+logTablePretty$EstTrue <- round(logTablePretty$EstTrue, 2)
+logTablePretty$EstTrueSD <- round(logTablePretty$EstTrueSD, 2)
+logTablePretty$Failed <- round(logTablePretty$Failed, 0)
+logTablePretty <- logTablePretty[,c("Case","Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
+writeTable(subset(logTablePretty, Model != "VT"), "table-1.docx", context)
 
@@
   x <- melt(ratioTable, id.vars=c("Scenario", "Case"), measure.vars=c("SizeCount"))
-  p <- ggplot(x, aes(Scenario, value, group=Case)) + geom_point(aes(shape=Case)) + geom_line() +
+  errorBars <- x
+  errorBars$value <- NA
+  errorBars$value[errorBars$variable=="SizeCount"] <- ratioTable$SizeCount
+  errorBars$qmin[errorBars$variable=="SizeCount"] <- ratioTable$SizeCountMin
+  errorBars$qmax[errorBars$variable=="SizeCount"] <- ratioTable$SizeCountMax
+
+  p <- ggplot(x, aes(Scenario, value, group=Case)) +
+    geom_errorbar(data=errorBars, aes(Scenario, group=Case, ymax=qmax, ymin=qmin), position=position_dodge(width=.7), colour="darkgray") +
+    geom_point(aes(shape=Case), position=position_dodge(width=0.7)) +
     ylab("Ratio") + #guides(shape=guide_legend()) +
@@
 
-rer.stat <- function(x) data.frame(SizeCount=mean(x$intersections/x$Observed), SizeCountSD=sd(x$intersections/x$Observed))
-i <- ddply(intersectionsx, .(Scenario), rer.stat); i$Scenario=prettyScenario(i$Scenario); i$Case <- "a"
-i2000 <- ddply(intersections2000x, .(Scenario), rer.stat); i2000$Scenario=prettyScenario(i2000$Scenario); i2000$Case <- "b"
-i10days <- ddply(intersections10daysx, .(Scenario), rer.stat); i10days$Scenario=prettyScenario(i10days$Scenario); i10days$Case <- "c"
+rer.stat <- function(x) data.frame(SizeCount=mean(x$intersections/x$Observed),
+                                   SizeCountMin=quantile(x$intersections/x$Observed, 0.05),
+                                   SizeCountMax=quantile(x$intersections/x$Observed, 0.95))
+i <- ddply(intersectionsx, .(Scenario), rer.stat); i$Scenario=prettyScenario(i$Scenario); i$Case <- "1"
+i2000 <- ddply(intersections2000x, .(Scenario), rer.stat); i2000$Scenario=prettyScenario(i2000$Scenario); i2000$Case <- "2"
+i10days <- ddply(intersections10daysx, .(Scenario), rer.stat); i10days$Scenario=prettyScenario(i10days$Scenario); i10days$Case <- "3"
 
@@
 p <- plotIntersectionPopulationSizeTable(iall)
-saveFigure(p, "manuscript-figure-4x.png", height=7.87/2.5)
+saveFigure(p, "manuscript-figure-4x.png", height=7.87/2.5) # draft quality
+saveFigure(p, "manuscript-figure-4x.pdf", height=7.87/2.5, dpi=300)
 

### 2015-03-07 R/Model.R

@@ -333,8 +333,22 @@ SmoothModelMeanTemporal <- setRefClass(
       return(invisible(.self))
     },
     
+    setupPrecisionPrior = function(priorParams) {
+      precPrior <- list(param=c(priorParams$shape, priorParams$rate), initial=priorParams$initial)
+      z <- qgamma(c(0.025, 0.5, 0.975), priorParams$shape, priorParams$rate, log.p=TRUE)
+      message("Precision prior 0.025 0.5 0.975 quantiles = ", signif(z[1],2), " ", signif(z[2],2), " ", signif(z[3],2))
+      return(precPrior)
+    },
+    
+    setupTemporalPrior = function(priorParams) {
+      rhoPrior <- list(param=c(priorParams$mean, priorParams$sd), initial=priorParams$initial)
+      z <- qnorm(c(0.025, 0.5, 0.975), priorParams$mean, priorParams$sd)
+      message("Rho prior 0.025 0.5 0.975 quantiles = ", signif(z[1],2), " ", signif(z[2],2), " ", signif(z[3],2))
+      return(rhoPrior)
+    },
+    
     setup = function(intersections, params) {
-      callSuper(intersections, params)
+      callSuper(intersections, params)      
       .self$aggregate()
       return(invisible(.self)) 
     },
@@ -902,179 +916,3 @@ FinlandRussiaSmoothModelSpatioTemporal <- setRefClass(
   methods = list(
   )
 )
-
-##################################################################################
-# OLD STUFF
-if (F) {
-  
-  
-  TODO_getPopulationSize = function(populationDensity, tracks, habitatWeights) {
-    if (is.null(data$fittedMean))
-      stop("Did you forgot to run collectEstimates() first?")
-    
-    if (!missing(habitatWeights)) if (!is.null(habitatWeights)) populationDensity$weight(habitatWeights)
-    
-    # TODO: this can be done faster without the rasters, but need the area of the study area
-    #area <- maskPolygon@polygons[[1]]@area
-    #if (!missing(habitatWeights)) if (!is.null(habitatWeights)) area <- cellStats(habitatWeights, sum)      
-    #meanPopulationSize <- ddply(data, .(year), function(x, area)
-    #  data.frame(z=mean(x$fittedMean / offsetScale) * area, year=x$year[1], area))
[…]

### 2015-03-09 inst/simulation/manuscript.Rmd
...

@@
 allScenarios <- c(scenarios, paste0(scenarios, "combined"), paste0(scenarios, "10days"))
-modelNames <- c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "SmoothModelMean-nbinomial-ar1", "FMPModel")
+modelNames <- c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "SmoothModelMean-nbinomial-ar1", "FMPModel", "SmoothModelMean-nbinomial-ar1-priors1")
 iterations <- as.integer(c(1,1,1,1,3,1))
-populationSizeOverEstimate <- 1e999
+#populationSizeOverEstimate <- 1e999
 
@@
 prettyModelName <- function(x) {
-  lookup <- data.frame(x=modelNames, y=c("ST", "VT", "T", "FMP"), stringsAsFactors=F)
+  lookup <- data.frame(x=modelNames, y=c("ST", "VT", "T1", "FMP", "T2"), stringsAsFactors=F)
   lookup$y[match(x, lookup$x)]
@@
 
-getPopulationSizes <- function(scenarios, modelNames, populationSizeOverEstimate=Inf) {
+getPopulationSizes <- function(scenarios, modelNames, populationSizeCutoff=Inf) {
   study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
-  validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
+  validation <- Validation(study=study, populationSizeCutoff=populationSizeCutoff)
   populationSizes <- validation$getValidationTemporalPopulationSizes(scenarios=scenarios, modelNames=modelNames)
@@
 ```{r figure-3-init-2}
-registerDoMC(cores=10) # The nodes seem to run out of memory and fail with a bogus error message, so it's better to reduce the number of used cores.
-populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames[c(1,3,4)])
-populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,4)])
-populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4)])
-registerDoMC(cores=detectCores())
+populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames[c(1,3,4,5)])
+populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,4,5)])
+populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4,5)])
 
@@
 saveFigure(p, "manuscript-figure-3a3.pdf", dpi=300)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[5]), modelName=modelNames[5], "a", xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3a4.png")
+saveFigure(p, "manuscript-figure-3a4.pdf", dpi=300)
 
@@
 saveFigure(p, "manuscript-figure-3b3.pdf", dpi=300)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[5]), modelName=modelNames[5], "b", xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3b4.png")
+saveFigure(p, "manuscript-figure-3b4.pdf", dpi=300)
 
@@
 saveFigure(p, "manuscript-figure-3c3.pdf", dpi=300)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[5]), modelName=modelNames[5], "c", xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3c4.png")
+saveFigure(p, "manuscript-figure-3c4.pdf", dpi=300)
 ```
@@
   subset(logPopulationSizesSample, modelName=="ST"),
-  subset(logPopulationSizesSample, modelName=="T"),
+  subset(logPopulationSizesSample, modelName=="T1"),
+  subset(logPopulationSizesSample, modelName=="T2"),
   subset(logPopulationSizesSample, modelName=="FMP"))
@@
 
-p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName==c("ST","T","FMP")), modelName=c("ST","T","FMP"), "a", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName==c("ST","T1","T2","FMP")), modelName=c("ST","T1","T2","FMP"), "a", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3sample.png", height=7.87*1.2/2) # draft quality
@@
   x <- ddply(populationSizes, .(Scenario, modelName), function(x, overEstimate) {
-    failed <- (1 - nrow(cutoffPopulationSizes(x, overEstimate)) / nrow(x) * 100)
+    failed <- (1 - nrow(cutoffPopulationSizes(x, overEstimate)) / nrow(x)) * 100
       #(sum(x$Estimated < overEstimate[1] | x$Estimated > overEstimate[2]) / nrow(x)) * 100)
@@
 ```{r figure-4-init}
-getPopulationCorrelations <- function(scenarios, modelNames) {
+getPopulationCorrelations <- function(scenarios, modelNames, cutoff=Inf) {
   study <- getStudy(scenario="A")
-  validation <- Validation(study=study, populationSizeOverEstimate=Inf)
+  validation <- Validation(study=study, populationSizeOverEstimate=cutoff)
   populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=scenarios, modelNames=modelNames)
@@
 ```{r figure-4abc}
-pc1 <- getPopulationCorrelations(scenarios=scenarios, modelNames=modelNames[1])
-pc2 <- getPopulationCorrelations(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1])
-pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1])
+pc1 <- getPopulationCorrelations(scenarios=scenarios, modelNames=modelNames[1], cutoff=cutoff)
+pc2 <- getPopulationCorrelations(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1], cutoff=cutoff)
+pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1], cutoff=cutoff)
 
@@
   study <- getStudy(scenario="A")
-  validation <- Validation(study=study, populationSizeOverEstimate=cutoff)
+  validation <- Validation(study=study, populationSizeCutoff=cutoff)
   a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames)
@@
     geom_point() +
-    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("black", "darkgray"), labels=modelLabels) +
+    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("black", "darkgray", "gray"), labels=modelLabels) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
@@
 ```{r figure-5abc}
-cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[1:3], panel="a", scenarioNames=scenarios, cutoff=cutoff)
-cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1:3], panel="b", scenarioNames=scenarios, cutoff=10*cutoff)
-cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1:3], panel="c", scenarioNames=scenarios, cutoff=cutoff)
+cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,5)], panel="a", scenarioNames=scenarios, cutoff=cutoff)
+cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,5)], panel="b", scenarioNames=scenarios, cutoff=10*cutoff)
+cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,5)], panel="c", scenarioNames=scenarios, cutoff=cutoff)
 
@@
 p <- plotCIProportions(cip)
-saveFigure(p, "manuscript-figure-5.png", height=7.87/2)
+saveFigure(p, "manuscript-figure-5.png", height=7.87/2, dpi=200)
+saveFigure(p, "manuscript-figure-5.pdf", height=7.87/2, dpi=300)
 ```

### 2015-03-09 R/Validation.R

@@ -231,7 +231,7 @@ Validation <- setRefClass(
         
         x <- SimulationPopulationSize$new(study=study, iteration=iteration)
         density <- model$data$fittedMean / model$offsetScale
-        x$getPopulationSize(density, model$data$year, loadValidationData=FALSE)
+        x$getPopulationSize(density, model$data$year, loadValidationData=TRUE)
         x <- x$sizeData  
         
         if (any(x$Estimated <= populationSizeCutoff[1] | x$Estimated >= populationSizeCutoff[2])) {
@@ -241,34 +241,7 @@ Validation <- setRefClass(
         x$sample <- index
         return(x)
       }, model=model, posteriorSamples=posteriorSamples, .parallel=T)
-      
-      if (F) {
-      populationSize <- data.frame()  
-        
-      for (sample in 1:nSamples) {
-        message("Processing sample ", sample, " / ", nSamples, " of iteration ", iteration, "...")
-        
-        lengthWeights <- if (!inherits(study$surveyRoutes, "uninitializedField"))
-          model$getLengthWeights(study$getSurveyRouteWeightedLengths(), study$surveyRoutes$lengths) else 1
-        
-        model$data <- posteriorSamples[[sample]]
-        model$data$fittedMean <- model$data$z * lengthWeights
-        model$data$year <- model$data$t
-        
-        x <- SimulationPopulationSize$new(study=study, iteration=iteration)
-        density <- model$data$fittedMean / model$offsetScale
-        x$getPopulationSize(density, model$data$year, loadValidationData=FALSE)
-        x <- x$sizeData  
-
-        if (any(x$Estimated <= populationSizeCutoff[1] | x$Estimated >= populationSizeCutoff[2])) {
-          message("Estimation failed for iteration ", iteration, ".")
-          next
-        }
-        x$sample <- sample
-        populationSize <- rbind(populationSize, x)
-      }
-      }
-      
+ 
       populationSize$scenario <- study$response
       populationSize$modelName <- modelName
       populationSize$iteration <- iteration
@@ -308,15 +281,16 @@ Validation <- setRefClass(
     getValidatedCredibilityIntervalsProportion = function(modelName, probs=c(.025, .975)) {
       library(plyr)
       
-      iterations <- getCredibilityIntervalsValidationIterations(modelName=modelName)
+      iterations <- .self$getCredibilityIntervalsValidationIterations(modelName=modelName)
[…]

### 2015-03-10 inst/simulation/manuscript.Rmd
...

@@
     prop.sd <- sd(log(x$Estimated/x$Observed))
-    prop.min <- quantile(log(x$Estimated/x$Observed), .05)
-    prop.max <- quantile(log(x$Estimated/x$Observed), .95)
+    prop.min <- quantile(log(x$Estimated/x$Observed), .025)
+    prop.max <- quantile(log(x$Estimated/x$Observed), .975)
     slope.sd <- coef(summary(y))[2,2]
@@
 rer.stat <- function(x) data.frame(SizeCount=mean(x$intersections/x$Observed),
-                                   SizeCountMin=quantile(x$intersections/x$Observed, 0.05),
-                                   SizeCountMax=quantile(x$intersections/x$Observed, 0.95))
+                                   SizeCountMin=quantile(x$intersections/x$Observed, 0.025),
+                                   SizeCountMax=quantile(x$intersections/x$Observed, 0.975))
 i <- ddply(intersectionsx, .(Scenario), rer.stat); i$Scenario=prettyScenario(i$Scenario); i$Case <- "1"
@@
     geom_point() +
-    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("black", "darkgray", "gray"), labels=modelLabels) +
+    scale_shape_manual(values=c(16, 1)) +
+    #scale_colour_manual("Model", values=c("black", "darkgray", "gray"), labels=modelLabels) +
+    scale_colour_manual("Model", values=c("blue", "green", "red"), labels=modelLabels) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
@@
 }
+p <- plotCIProportions(cip)
+saveFigure(p, "manuscript-figure-5.png", height=7.87/2, dpi=200)
 ```

### 2015-03-10 R/Validation.R

@@ -113,7 +113,7 @@ Validation <- setRefClass(
         y$Observed.q1 <- q[1]
         y$Observed.q2 <- q[2]
         return(y)
-      }, probs=c(.0025, .975))
+      }, probs=c(.025, .975))
     },
     
     validateSpatialPopulationSize = function(modelName) {

### 2015-03-10 inst/simulation/estimate_all.sh

@@ -4,8 +4,8 @@ scenarios=(A B C D E F)
 scenarios2000=(Acombined Bcombined Ccombined Dcombined Ecombined Fcombined)
 scenarios10days=(A10days B10days C10days D10days E10days F10days)
 models=(FMPModel SmoothModelMean-nbinomial-ar1 SmoothModel-nbinomial-ar1 SmoothModel-nbinomial-matern-ar1 SmoothModelMean-nbinomial-ar1-priors1)
-nodes=(1:50 1:50 1:50 1:50)
-max_nodes=(60 60 60 60)
+nodes=(1:50 1:50 1:50)
+max_nodes=(60 60 60)
 days=(1 1 10)
 
 test="no-test"

### 2015-03-10 inst/simulation/population_size_all.sh

@@ -4,8 +4,8 @@ scenarios=(A B C D E F)
 scenarios2000=(Acombined Bcombined Ccombined Dcombined Ecombined Fcombined)
 scenarios10days=(A10days B10days C10days D10days E10days F10days)
 models=(FMPModel SmoothModelMean-nbinomial-ar1 SmoothModel-nbinomial-ar1 SmoothModel-nbinomial-matern-ar1 SmoothModelMean-nbinomial-ar1-priors1)
-nodes=(1:50 1:50 1:50 1:50)
-max_nodes=(60 60 60 60)
+nodes=(1:50 1:50 1:50)
+max_nodes=(60 60 60)
 #free_mem=(10000 10000 10000 10000 28000 28000)
 free_mem=(10000 10000 10000 10000 10000 10000) # Assuming habitat uses have been determined before
 

### 2015-03-10 inst/simulation/manuscript.Rmd
...

@@
   p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
-    geom_point(size=1, colour="darkgray", alpha=.25) +
+    geom_point(size=1, colour="darkgray", alpha=.5) +
     geom_abline(colour="black", linetype="longdash") + 
@@
 xlim <- log10(c(50,1200)); ylim <- xlim
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], "a", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], "1", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a1.png")
 saveFigure(p, "manuscript-figure-3a1.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], "a", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], "1", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a2.png")
 saveFigure(p, "manuscript-figure-3a2.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], "a", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], "1", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a3.png")
 saveFigure(p, "manuscript-figure-3a3.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[5]), modelName=modelNames[5], "a", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[5]), modelName=modelNames[5], "1", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3a4.png")
@@
 xlim <- log10(c(500,4000)); ylim <- xlim
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], "b", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], "2", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b1.png")
 saveFigure(p, "manuscript-figure-3b1.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], "b", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], "2", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b2.png")
 saveFigure(p, "manuscript-figure-3b2.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], "b", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], "2", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b3.png")
 saveFigure(p, "manuscript-figure-3b3.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[5]), modelName=modelNames[5], "b", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[5]), modelName=modelNames[5], "2", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3b4.png")
@@
 xlim <- log10(c(5,1200)); ylim <- xlim
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], "c", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], "3", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c1.png")
 saveFigure(p, "manuscript-figure-3c1.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], "c", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], "3", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c2.png")
 saveFigure(p, "manuscript-figure-3c2.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], "c", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], "3", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c3.png")
 saveFigure(p, "manuscript-figure-3c3.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[5]), modelName=modelNames[5], "c", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[5]), modelName=modelNames[5], "3", xlim=xlim, ylim=ylim)
 saveFigure(p, "manuscript-figure-3c4.png")
@@
 p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName==c("ST","T1","T2","FMP")), modelName=c("ST","T1","T2","FMP"), "a", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3sample.png", height=7.87*1.2/2) # draft quality
-saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87*1.2/2, dpi=300)
+saveFigure(p, "manuscript-figure-3sample.png", height=7.87*2/3) # draft quality
+saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87*2/3, dpi=300)
 ```

### 2015-03-10 R/Model.R

@@ -108,35 +108,52 @@ Model <- setRefClass(
                  index=index, .parallel=F)
       return(x)
     },
-           
-    getPopulationDensity = function(templateRaster=study$getTemplateRaster(), maskPolygon=study$studyArea$boundary, getSD=FALSE) {
-      if (is.null(data$fittedMean))
-        stop("Did you forgot to run collectEstimates() first?")
-      
-      library(raster)
-      library(plyr)
-      
-      #area <- maskPolygon@polygons[[1]]@area
-      meanPopulationDensity <- ddply(data, .(year), function(x)
-        data.frame(z=mean(x$fittedMean / offsetScale), year=x$year[1]))
-      
-      cellArea <- prod(res(templateRaster))
-      meanPopulationDensityRaster <- SpatioTemporalRaster(study=study)$fill(z=meanPopulationDensity$z, boundary=maskPolygon, layerNames=meanPopulationDensity$year, weights=cellArea)
-      
-      sdPopulationDensityRaster <- if (getSD) {
-        sdPopulationDensity <- ddply(data, .(year), function(x)
-          data.frame(z=mean(x$fittedSD / offsetScale), year=x$year[1]))
-        SpatioTemporalRaster(study=study)$fill(z=meanPopulationDensity$z, boundary=maskPolygon, layerNames=meanPopulationDensity$year, weights=cellArea)
-      }
-      else SpatioTemporalRaster(study=study)
-      
-      return(invisible(list(mean=meanPopulationDensityRaster, sd=sdPopulationDensityRaster)))
-    },
-    
+        
     getLengthWeights = function(weightedLengths, lengths) {
       lengthWeights <- weightedLengths / lengths
       lengthWeights <- rep(lengthWeights, length(years))
       return(lengthWeights)
+    },
+    
+    getDensityEstimates = function(weights=1, aggregate=F) {
+      xy <- getUnscaledObservationCoordinates()
+      xyzt <- data.frame(x=xy[,1], y=xy[,y], density=data$fittedMean * weights / offsetScale, year=data$year)
+      if (aggregate)
+        xyzt <- ddply(xyzt, .(year), function(x) data.frame(density=mean(x$density), year=x$year[1]))
+      return(xyzt)
+    },
+    
+    getPopulationDensity = function(templateRaster=study$getTemplateRaster(), maskPolygon=study$studyArea$boundary, habitatWeights=NULL) {
+      if (is.null(data$fittedMean))
+        stop("Did you forgot to run collectEstimates() first?")
+      library(raster)
+      
[…]

### 2015-03-10 R/Study.R

@@ -101,7 +101,7 @@ SimulationStudy <- setRefClass(
       return(invisible(model))
     },
     
-    getHabitatWeights2 = function(tracks, iteration, save=TRUE, readHabitatIntoMemory=TRUE) {
+    getHabitatWeights = function(tracks, iteration, save=TRUE, readHabitatIntoMemory=TRUE) {
       if (withHabitatWeights == FALSE) return(NULL)
       
       habitatWeights <- CORINEHabitatWeights$new(study=.self)
@@ -119,34 +119,50 @@ SimulationStudy <- setRefClass(
       return(habitatWeights)
     },
     
-    getSurveyRouteWeightedLengths = function(iteration, readHabitatIntoMemory=TRUE) {
-      weightedLengths <- if (!inherits(surveyRoutes, "uninitializedField")) {
-        habitatWeights <- getHabitatWeights2(iteration=iteration, readHabitatIntoMemory=readHabitatIntoMemory)
-        surveyRoutes$getWeightedLengths(habitatWeights)
-      } else 1
-      return(weightedLengths)
-    },
+    #getSurveyRouteWeightedLengths = function(iteration, readHabitatIntoMemory=TRUE) {
+    #  weightedLengths <- if (!inherits(surveyRoutes, "uninitializedField")) {
+    #    habitatWeights <- getHabitatWeights2(iteration=iteration, readHabitatIntoMemory=readHabitatIntoMemory)
+    #    surveyRoutes$getWeightedLengths(habitatWeights)
+    #  } else 1
+    #  return(weightedLengths)
+    #},
     
-    getPopulationSize2 = function(modelName, iteration, readHabitatIntoMemory=TRUE, save=TRUE) {
-      estimates <- getModel(modelName=modelName, iteration=iteration)
-      estimates$loadEstimates()
+    getPopulationSize = function(estimates, readHabitatIntoMemory=TRUE, save=TRUE) {
+      withHabitatPreferences <- !inherits(surveyRoutes, "uninitializedField")
       
-      lengthWeights <- if (!inherits(surveyRoutes, "uninitializedField"))
-        estimates$getLengthWeights(getSurveyRouteWeightedLengths(iteration=iteration, readHabitatIntoMemory=readHabitatIntoMemory), surveyRoutes$lengths)
-      else 1
+      #lengthWeights <- if (withHabitatPreferences)
[…]

### 2015-03-10 R/Validation.R

@@ -222,17 +222,18 @@ Validation <- setRefClass(
       populationSize <- ldply(1:nSamples, function(index, model, posteriorSamples) {
         message("Processing posterior sample ", index, "...")
         
-        lengthWeights <- if (!inherits(study$surveyRoutes, "uninitializedField"))
-          model$getLengthWeights(study$getSurveyRouteWeightedLengths(iteration=iteration), study$surveyRoutes$lengths) else 1
+        #lengthWeights <- if (!inherits(study$surveyRoutes, "uninitializedField"))
+        #  model$getLengthWeights(study$getSurveyRouteWeightedLengths(iteration=iteration), study$surveyRoutes$lengths) else 1
         
         model$data <- posteriorSamples[[index]]
-        model$data$fittedMean <- model$data$z * lengthWeights
+        model$data$fittedMean <- model$data$z #* lengthWeights
         model$data$year <- model$data$t
+        x <- study$getPopulationSize(estimates=model, save=FALSE)$sizeData
         
-        x <- SimulationPopulationSize$new(study=study, iteration=iteration)
-        density <- model$data$fittedMean / model$offsetScale
-        x$getPopulationSize(density, model$data$year, loadValidationData=TRUE)
-        x <- x$sizeData  
+        #x <- SimulationPopulationSize$new(study=study, iteration=iteration)
+        #density <- model$data$fittedMean / model$offsetScale
+        #x$getPopulationSize(density, model$data$year, loadValidationData=TRUE)
+        #x <- x$sizeData  
         
         if (any(x$Estimated <= populationSizeCutoff[1] | x$Estimated >= populationSizeCutoff[2])) {
           message("Estimation failed for iteration ", iteration, ".")

### 2015-03-10 inst/simulation/population_size.R

@@ -51,7 +51,7 @@ population_size <- function(scenario, modelName, iteration, isTest, otherTest=F)
     iteration <- as.integer(1)
     
     estimates <- study$getModel(modelName=modelName, iteration=iteration)
-    estimates <- study$loadEstimates(estimates=estimates)
+    estimates$loadEstimates()
     estimates$collectEstimates()
     
     #estimates$collectHyperparameters()

### 2015-03-10 inst/simulation/manuscript.Rmd
...

@@
 
+```{r figure-x-debug}
+getTrueAndEstimatedCounts <- function(scenarios, modelNames, cutoff=c(-Inf,Inf)) {
+  study <- getStudy(scenario="A")
+  validation <- Validation(study=study, populationSizeOverEstimate=cutoff)
+  counts <- validation$getCounts(scenarios=scenarios, modelNames=modelNames)
+  counts <- ddply(counts, .(iteration), function(x) data.frame(Scenario=x$Scenario[1], Model=x$Model[1], True=mean(x$True), TrueSD=sd(x$True), Estimated=mean(x$Estimated), EstimatedSD=sd(x$Estimated)))
+  return(counts)
+}
+
+c1 <- getTrueAndEstimatedCounts(scenarios=scenarios, modelNames=modelNames[c(1,3:5)])
+c2 <- getTrueAndEstimatedCounts(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3:5)])
+c3 <- getTrueAndEstimatedCounts(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3:5)])
+
+```
+
 ```{r figure-4-init}
-getPopulationCorrelations <- function(scenarios, modelNames, cutoff=Inf) {
+getPopulationCorrelations <- function(scenarios, modelNames, cutoff=c(-Inf,Inf)) {
   study <- getStudy(scenario="A")

### 2015-03-10 R/Validation.R

@@ -60,6 +60,27 @@ Validation <- setRefClass(
       return(x)
     },
     
+    getCounts = function(scenarios=c("A","B","C","D","E","F"), modelNames) {
+      counts <- ddply(expand.grid(scenario=scenarios, modelName=modelNames, stringsAsFactors=FALSE), .(scenario, modelName), function(x) {
+        s <- getStudy(scenario=x$scenario, isTest=F)
+        iterations <- getEstimatesFileIterations(x$modelName)
+        
+        counts <- data.frame()
+        for (iteration in iterations) {
+          estimates <- s$getModel(modelName=x$modelName, iteration=iteration)
+          estimates$loadEstimates()
+          estimates$collectEstimates()
+          a <- sum(estimates$data$intersections)
+          b <- sum(estimates$data$fittedMean * estimates$observedOffset)
+          counts <- rbind(counts, data.frame(Model=x$modelName, Scenario=x$Scenario, iteration=iteration, True=a, Estimated=b))
+        }
+        
+        return(counts)
+      }, .parallel=T)
+      
+      return(counts)
+    },
+    
     validateTemporalPopulationSize = function(modelName) {
       library(plyr)
       
@@ -233,7 +254,7 @@ Validation <- setRefClass(
         #x <- SimulationPopulationSize$new(study=study, iteration=iteration)
         #density <- model$data$fittedMean / model$offsetScale
         #x$getPopulationSize(density, model$data$year, loadValidationData=TRUE)
-        #x <- x$sizeData  
+        #x <- x$sizeData
         
         if (any(x$Estimated <= populationSizeCutoff[1] | x$Estimated >= populationSizeCutoff[2])) {
           message("Estimation failed for iteration ", iteration, ".")

### 2015-03-10 inst/simulation/manuscript.Rmd
...

--- 2015-03-10 inst/simulation/manuscript.Rmd
@@
 boundaryDF <- ggplot2::fortify(FinlandWTCStudy(context=context, response="canis.lupus")$studyArea$boundary)
+
 ```
@@
 
+```{r figure-x-debug}
+getTrueAndEstimatedCounts <- function(scenarios, modelNames, cutoff=c(-Inf,Inf)) {
+  study <- getStudy(scenario="A")
+  validation <- Validation(study=study, populationSizeCutoff=cutoff)
+  counts <- validation$getCounts(scenarios=scenarios, modelNames=modelNames)
+  counts <- ddply(counts, .(iteration), function(x) data.frame(Scenario=x$Scenario[1], Model=x$Model[1], True=mean(x$True), TrueSD=sd(x$True), Estimated=mean(x$Estimated), EstimatedSD=sd(x$Estimated)))
+  return(counts)
+}
+
+c1 <- getTrueAndEstimatedCounts(scenarios=scenarios, modelNames=modelNames[c(1,3:5)])
+c2 <- getTrueAndEstimatedCounts(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3:5)])
+c3 <- getTrueAndEstimatedCounts(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3:5)])
+
+```
+
 ```{r figure-4-init}
-getPopulationCorrelations <- function(scenarios, modelNames, cutoff=Inf) {
+getPopulationCorrelations <- function(scenarios, modelNames, cutoff=c(-Inf,Inf)) {
   study <- getStudy(scenario="A")
-  validation <- Validation(study=study, populationSizeOverEstimate=cutoff)
+  validation <- Validation(study=study, populationSizeCutoff=cutoff)
   populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=scenarios, modelNames=modelNames)

### 2015-03-10 R/Model.R

@@ -114,7 +114,7 @@ Model <- setRefClass(
       lengthWeights <- rep(lengthWeights, length(years))
       return(lengthWeights)
     },
-    
+        
     getDensityEstimates = function(weights=1, aggregate=F) {
       xy <- getUnscaledObservationCoordinates()
       xyzt <- data.frame(x=xy[,1], y=xy[,2], density=data$fittedMean * weights / offsetScale, year=data$year)

### 2015-03-10 R/Validation.R

@@ -70,8 +70,8 @@ Validation <- setRefClass(
           estimates <- s$getModel(modelName=x$modelName, iteration=iteration)
           estimates$loadEstimates()
           estimates$collectEstimates()
-          a <- sum(estimates$data$intersections)
-          b <- sum(estimates$data$fittedMean * estimates$observedOffset)
+          a <- mean(estimates$data$intersections)
+          b <- mean(estimates$data$fittedMean * estimates$getObservedOffset())
           counts <- rbind(counts, data.frame(Model=x$modelName, Scenario=x$Scenario, iteration=iteration, True=a, Estimated=b))
         }
         

### 2015-03-11 inst/simulation/manuscript.Rmd
...

@@
   validation <- Validation(study=study, populationSizeCutoff=cutoff)
-  counts <- validation$getCounts(scenarios=scenarios, modelNames=modelNames)
-  counts <- ddply(counts, .(iteration), function(x) data.frame(Scenario=x$Scenario[1], Model=x$Model[1], True=mean(x$True), TrueSD=sd(x$True), Estimated=mean(x$Estimated), EstimatedSD=sd(x$Estimated)))
+  counts <- validation$getMeanCounts(scenarios=scenarios, modelNames=modelNames)
   return(counts)
@@
 
-c1 <- getTrueAndEstimatedCounts(scenarios=scenarios, modelNames=modelNames[c(1,3:5)])
-c2 <- getTrueAndEstimatedCounts(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3:5)])
-c3 <- getTrueAndEstimatedCounts(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3:5)])
+c1 <- getTrueAndEstimatedCounts(scenarios=scenarios, modelNames=modelNames[c(3:5)]); c1$Case <- 1
+c2 <- getTrueAndEstimatedCounts(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(3:5)]); c1$Case <- 2
+c3 <- getTrueAndEstimatedCounts(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(3:5)]); c1$Case <- 3
+c.all <- rbind(c1, c2, c3)
+c.agg <- ddply(c.all, .(iteration), function(x) data.frame(Scenario=x$Scenario[1], Model=x$Model[1], Ratio=mean(x$True/x$Estimated), RatioSD=sd(x$True/x$Estimated)))
 
+#ggplot(c.all, aes())
 ```

### 2015-03-11 R/Study.R

@@ -89,11 +89,11 @@ SimulationStudy <- setRefClass(
       return(invisible(intersections))
     },
     
-    estimate = function(model, params, tag=NULL, save=TRUE) {
+    estimate = function(model, params, tag=NULL, predictDistances=FALSE, save=TRUE) {
       if (missing(model)) stop("Missing model argument.")
       if (missing(params)) stop("Missing params argument.")
       
-      intersections <- loadIntersections(iteration=model$iteration)
+      intersections <- loadIntersections(iteration=model$iteration, predictDistances=predictDistances)
       if (is.null(tag)) model$setup(intersections=intersections, params=params)
       else model$setup(intersections=intersections, params=params, tag=tag)
       model$estimate()
@@ -109,6 +109,8 @@ SimulationStudy <- setRefClass(
       fileName <- habitatPreferences$getHabitatSelectionFileName()
       habitatSelection <- if (file.exists(fileName)) habitatPreferences$loadHabitatSelection()
       else {      
+        message("Habitat preferences file not found: ", fileName)
+        message("Estimating...")
         if (missing(tracks)) tracks <- loadTracks(iteration=iteration)
         if (readHabitatIntoMemory)
           study$studyArea$readRasterIntoMemory()

### 2015-03-11 inst/simulation/population_size.R

@@ -31,7 +31,7 @@ isTest<-F
 #modelName<-"SmoothModelMean-nbinomial-ar1"
 modelName<-"SmoothModelMean-nbinomial-ar1-priors1"
 #modelName<-"SmoothModel-nbinomial-matern-ar1"
-iteration<-as.integer(3)
+iteration<-as.integer(1)
 readHabitatIntoMemory <- F
 mss <- getMSS(scenario=scenario, isTest=isTest, readHabitatIntoMemory=FALSE)
 study <- mss$study
@@ -47,11 +47,17 @@ population_size <- function(scenario, modelName, iteration, isTest, otherTest=F)
   surveyRoutes <- mss$getSurveyRoutes()
   
   if (otherTest) {
-    study <- getMSS(scenario="Acombined")$study
-    iteration <- as.integer(1)
+    modelName <- "SmoothModelMean-nbinomial-ar1-priors1"
+    #modelName<-"SmoothModel-nbinomial-matern-ar1"
+    study <- getMSS(scenario="E")$study
+    iteration <- as.integer(3)
+    
+    populationSize <- study$getPopulationSize2(modelName=modelName, iteration=iteration, save=T, readHabitatIntoMemory=F)
     
     estimates <- study$getModel(modelName=modelName, iteration=iteration)
     estimates$loadEstimates()
+    summary(estimates$result)
+    
     estimates$collectEstimates()
     
     #estimates$collectHyperparameters()
@@ -65,8 +71,6 @@ population_size <- function(scenario, modelName, iteration, isTest, otherTest=F)
     #study$loadPopulationSize(iteration=iteration, modelName=modelName)
   }
   else {
-    #iteration<-as.integer(6)
-    
     populationSize <- study$getPopulationSize2(modelName=modelName, iteration=iteration, save=T)
     return(invisible(populationSize))
     

### 2015-03-13 inst/simulation/manuscript.Rmd
...

@@
 p <- plotIntersectionPopulationSizeTable(iall)
-saveFigure(p, "manuscript-figure-4x.png", height=7.87/2.5) # draft quality
-saveFigure(p, "manuscript-figure-4x.pdf", height=7.87/2.5, dpi=300)
+saveFigure(p, "manuscript-figure-4x.png", height=7.87/2) # draft quality
+saveFigure(p, "manuscript-figure-4x.pdf", height=7.87/2, dpi=300)
 
@@
                                       plot.title=element_text(size=10),
-                                      plot.margin=unit(c(0, 0, 0, 0), units="line"),
+                                      plot.margin=unit(c(0.2, 0, 0, 0), units="line"),
                                       panel.margin=unit(0, units="line"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
@@
 
-p1 <- plotPopulationCorrelations(pc1, "a")
+p1 <- plotPopulationCorrelations(pc1, "Case 1")
 saveFigure(p1, "manuscript-figure-4a.png")
-p2 <- plotPopulationCorrelations(pc2, "b")
+saveFigure(p1, "manuscript-figure-4a.pdf", dpi=300)
+p2 <- plotPopulationCorrelations(pc2, "Case 2")
 saveFigure(p2, "manuscript-figure-4b.png")
-p3 <- plotPopulationCorrelations(pc3, "c")
+saveFigure(p2, "manuscript-figure-4b.pdf", dpi=300)
+p3 <- plotPopulationCorrelations(pc3, "Case 3")
 saveFigure(p3, "manuscript-figure-4c.png")
+saveFigure(p3, "manuscript-figure-4c.pdf", dpi=300)
 ```
@@
     scale_shape_manual(values=c(16, 1)) +
-    #scale_colour_manual("Model", values=c("black", "darkgray", "gray"), labels=modelLabels) +
-    scale_colour_manual("Model", values=c("blue", "green", "red"), labels=modelLabels) +
+    scale_colour_manual("Model", values=c("black", "gray26", "gray80"), labels=modelLabels) +
+    #scale_colour_manual("Model", values=c("blue", "green", "red"), labels=modelLabels) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
@@
     theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
-          plot.margin=unit(c(0, 0, 0, 0), units="line"),
+          plot.margin=unit(c(0.2, 0, 0, 0), units="line"),
           #legend.background=element_rect(fill="transparent"),
@@
           legend.position="bottom",
-          legend.box="horizontal")
+          legend.box="horizontal",
+          legend.key.width=unit(.7, "line"))
   return(p)
 }
-p <- plotCIProportions(cip)
-saveFigure(p, "manuscript-figure-5.png", height=7.87/2, dpi=200)
+#p <- plotCIProportions(cip)
+#saveFigure(p, "manuscript-figure-5.png", height=7.87/2, dpi=200)
 ```
@@
 ```{r figure-5abc}
-cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,5)], panel="a", scenarioNames=scenarios, cutoff=cutoff)
-cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,5)], panel="b", scenarioNames=scenarios, cutoff=10*cutoff)
-cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,5)], panel="c", scenarioNames=scenarios, cutoff=cutoff)
+cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,5)], panel="Case 1", scenarioNames=scenarios, cutoff=cutoff)
+cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,5)], panel="Case 2", scenarioNames=scenarios, cutoff=10*cutoff)
+cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,5)], panel="Case 3", scenarioNames=scenarios, cutoff=cutoff)
 
@@
 p <- plotCIProportions(cip)
-saveFigure(p, "manuscript-figure-5.png", height=7.87/2, dpi=200)
+saveFigure(p, "manuscript-figure-5.png", height=7.87/2)
 saveFigure(p, "manuscript-figure-5.pdf", height=7.87/2, dpi=300)

### 2015-03-13 R/Tracks.R

@@ -281,14 +281,14 @@ SimulatedTracks <- setRefClass(
     },
     
     getTracksFileName = function() {
-      if (inherits(study, "undefinedField") | length(iteration) == 0)
+      if (inherits(study, "undefinedField") || length(iteration) == 0)
         stop("Provide study and iteration parameters.")
       return(study$context$getLongFileName(dir=getTracksDirectory(), name="Tracks", response=study$response, region=study$studyArea$region, tag=iteration))
     },
     
     loadTracks = function(fileName=getTracksFileName(), addColumns=TRUE) {
       callSuper(fileName=fileName, addColumns=addColumns)
-      if (inherits(truePopulationSize, "uninitializedField") | nrow(truePopulationSize) == 0)
+      if (inherits(truePopulationSize, "uninitializedField") || nrow(truePopulationSize) == 0)
         setTruePopulationSize()
       return(invisible(.self))
     },

### 2015-03-13 R/Validation.R

@@ -164,13 +164,15 @@ Validation <- setRefClass(
         estimates <- study$loadEstimates(estimates)
         estimates$collectEstimates()
         
-        estimated <- estimates$getPopulationDensityInterpolate(templateRaster=template, maskPolygon=NULL, getSD=F)
-        if (is.null(estimated$mean)) {
+        estimated <- estimates$getPopulationDensity(templateRaster=template)
+        #estimated <- estimates$getPopulationDensityInterpolate(templateRaster=template, maskPolygon=NULL, getSD=F)
+        #if (is.null(estimated$mean)) {
+        if (is.null(estimated)) {
           message("Estimation failed for iteration ", iteration, " scenario ", study$response, ".")
           next
         }
         
-        estimated$mean$rasterStack <- stack(mask(estimated$mean$rasterStack, coverArea) * coverArea)
+        #estimated$mean$rasterStack <- stack(mask(estimated$mean$rasterStack, coverArea) * coverArea)
         
         tracks <- study$loadTracks(iteration=iteration)
         cc <- ddply(subset(tracks$tracks, yday == 29), .(year, id), function(x) x[1,])[,c("year","x","y","herdSize")]
@@ -185,8 +187,10 @@ Validation <- setRefClass(
           true <- rasterize(subcc[,c("x","y")], template, field=subcc$herdSize, fun='sum', background=0)
           true <- mask(true, coverArea)
           
-          correlation <- cor(estimated$mean$rasterStack[[i]][], true[], use="complete.obs", method="spearman")
-          estpop <- sum(estimated$mean$rasterStack[[i]][], na.rm=T)
+          #correlation <- cor(estimated$mean$rasterStack[[i]][], true[], use="complete.obs", method="spearman")
+          correlation <- cor(estimated$rasterStack[[i]][], true[], use="complete.obs", method="spearman")
+          #estpop <- sum(estimated$mean$rasterStack[[i]][], na.rm=T)
+          estpop <- sum(estimated$rasterStack[[i]][], na.rm=T)
           truepop <- sum(true[], na.rm=T)
           realtrue <- subset(tracks$truePopulationSize, Year == year0)$Observed
                     

### 2015-03-13 inst/simulation/population_size.R

@@ -32,7 +32,7 @@ scenario <- "E"
 isTest<-F
 #modelName<-"FMPModel"
 #modelName<-"SmoothModelMean-nbinomial-ar1"
-modelName<-"SmoothModelMean-nbinomial-ar1-priors1"
+#odelName<-"SmoothModelMean-nbinomial-ar1-priors1"
 #modelName<-"SmoothModel-nbinomial-matern-ar1"
 iteration<-as.integer(1)
 readHabitatIntoMemory <- F
@@ -52,22 +52,22 @@ population_size <- function(scenario, modelName, iteration, isTest, otherTest=F)
   if (otherTest) {
     library(WTC)
     source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
-    scenario <- "E"
+    scenario <- "F"
     readHabitatIntoMemory <- F
     modelName <- "SmoothModelMean-nbinomial-ar1-priors1"
     #modelName<-"SmoothModel-nbinomial-matern-ar1"
     study <- getMSS(scenario=scenario, readHabitatIntoMemory=FALSE)$study
     study$grassLocalTempDir <- grassLocalTempDir
-    iteration <- as.integer(3)
+    iteration <- as.integer(18)
     
     habitatWeights <- study$getHabitatWeights(iteration=iteration, readHabitatIntoMemory=FALSE)
-    #writeRaster(x, "public_html/x.tif", format="GTiff", overwrite=T)
     
     estimates <- study$getModel(modelName = modelName, iteration = iteration)
     estimates$loadEstimates()
     estimates$collectEstimates()    
     populationDensity <- estimates$getPopulationDensity(habitatWeights=habitatWeights, .parallel=FALSE)
-    #habitatWeightsRaster <- habitatWeights$getWeightsRaster(grassLocalTempDir=study$grassLocalTempDir, save=TRUE)    
+    #habitatWeightsRaster <- habitatWeights$getWeightsRaster(grassLocalTempDir=study$grassLocalTempDir, save=TRUE)
+    #writeRaster(habitatWeightsRaster, "public_html/x.tif", format="GTiff", overwrite=T)
     habitatWeightsRaster <- habitatWeights$getWeightsRaster()
     populationDensity$weight(habitatWeightsRaster)
     plot(populationDensity$rasterStack[[1]])     

### 2015-03-14 inst/simulation/manuscript.Rmd
...

@@
 }
-```
 
-```{r figure-3-init-2}
 populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames[c(1,3,4,5)])
@@
 pc1 <- getPopulationCorrelations(scenarios=scenarios, modelNames=modelNames[1], cutoff=cutoff)
-pc2 <- getPopulationCorrelations(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1], cutoff=cutoff)
+pc2a <- getPopulationCorrelations(scenarios="Acombined", modelNames=modelNames[1], cutoff=cutoff)
+pc2b <- getPopulationCorrelations(scenarios="Bcombined", modelNames=modelNames[1], cutoff=cutoff)
+pc2c <- getPopulationCorrelations(scenarios="Ccombined", modelNames=modelNames[1], cutoff=cutoff)
+pc2d <- getPopulationCorrelations(scenarios="Dcombined", modelNames=modelNames[1], cutoff=cutoff)
+pc2e <- getPopulationCorrelations(scenarios="Ecombined", modelNames=modelNames[1], cutoff=cutoff)
+pc2f <- getPopulationCorrelations(scenarios="Fcombined", modelNames=modelNames[1], cutoff=cutoff)
+pc2 <- rbind(pc2a,pc2b,pc2c,pc2d,pc2e,pc2f)
 pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1], cutoff=cutoff)
@@
 
-```{r figure-5-init}
-getCIProportions <- function(scenarios, modelNames, panel, scenarioNames, cutoff=Inf) {
+```{r figure-5}
+getCIProportions <- function(scenario, modelName, panel, scenarioNames, cutoff=Inf) {
   study <- getStudy(scenario="A")
   validation <- Validation(study=study, populationSizeCutoff=cutoff)
-  a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames)
-  b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames, probs=c(.25, .75), probsName="50%")
+  a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenario, modelNames=modelName)
+  b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenario, modelNames=modelName, probs=c(.25, .75), probsName="50%")
   credibilityIntervals <- rbind(a, b)
@@
 #saveFigure(p, "manuscript-figure-5.png", height=7.87/2, dpi=200)
-```
 
-```{r figure-5abc}
 cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,5)], panel="Case 1", scenarioNames=scenarios, cutoff=cutoff)

### 2015-03-14 R/Validation.R

@@ -327,6 +327,7 @@ Validation <- setRefClass(
       populationSizeCI <- ldply(iterations, function(iteration) {
         .self$loadCredibilityIntervalsValidation(modelName=modelName, iteration=iteration)
       })
+      populationSizeCI <- populationSizeCI[complete.cases(populationSizeCI),]
       
       x <- .self$summarizePopulationSizeCI(populationSizeCI, variables=.(scenario, Year, iteration), probs=probs)
       validationProportion <- ddply(x, .(scenario), function(x) {

### 2015-03-14 inst/simulation/manuscript.Rmd
...

@@
 study <- mss$study
-surveyRoutes <- mss$getSurveyRoutes()
+surveyRoutes <- study$surveyRoutes
 #habitatTypes <- raster::extract(habitat, surveyRoutes$surveyRoutes[1:3])
@@
 
-```{r figure-4-init}
+```{r figure-4}
 getPopulationCorrelations <- function(scenarios, modelNames, cutoff=c(-Inf,Inf)) {
@@
 }
-```
 
-```{r figure-4abc}
 pc1 <- getPopulationCorrelations(scenarios=scenarios, modelNames=modelNames[1], cutoff=cutoff)

### 2015-03-14 R/Study.R

@@ -126,24 +126,9 @@ SimulationStudy <- setRefClass(
       }
       return(habitatWeights)
     },
-    
-    #getSurveyRouteWeightedLengths = function(iteration, readHabitatIntoMemory=TRUE) {
-    #  weightedLengths <- if (!inherits(surveyRoutes, "uninitializedField")) {
-    #    habitatWeights <- getHabitatWeights2(iteration=iteration, readHabitatIntoMemory=readHabitatIntoMemory)
-    #    surveyRoutes$getWeightedLengths(habitatWeights)
-    #  } else 1
-    #  return(weightedLengths)
-    #},
-    
+        
     getPopulationSize = function(estimates, iteration, readHabitatIntoMemory=TRUE, save=TRUE, .parallel=TRUE) {
-      withHabitatPreferences <- !inherits(surveyRoutes, "uninitializedField")
-      
-      #lengthWeights <- if (withHabitatPreferences)
-      #  estimates$getLengthWeights(getSurveyRouteWeightedLengths(iteration=iteration, readHabitatIntoMemory=readHabitatIntoMemory), surveyRoutes$lengths)
-      #else 1
-      #estimates$collectEstimates(lengthWeights)
-      
-      if (withHabitatPreferences) {
+      if (withHabitatWeights) {
         habitatWeights <- getHabitatWeights(iteration=iteration, readHabitatIntoMemory=readHabitatIntoMemory)
         populationDensity <- estimates$getPopulationDensity(habitatWeights=habitatWeights, .parallel=.parallel)
         habitatWeightsRaster <- if (length(grassLocalTempDir) != 0)
@@ -156,9 +141,6 @@ SimulationStudy <- setRefClass(
         x <- estimates$getDensityEstimates()
         populationSize$getPopulationSize(x$density, x$year, loadValidationData=TRUE)
       }
-      
-      #density <- estimates$data$fittedMean / estimates$offsetScale
-      #populationSize$getPopulationSize(density, estimates$data$year, loadValidationData=TRUE)
       
       if (save) populationSize$savePopulationSize()
       

### 2015-03-14 R/Tracks.R

@@ -305,13 +305,20 @@ SimulatedTracks <- setRefClass(
       library(plyr)
       message("Randomizing observation days = ", days, " and filtering tracks...")
 
-      observationTracksDF <- ddply(tracks, .(year), function(x, days) {
+      observationTracksDF <- ddply(x, .(year, id), function(x, days) {
+        days <- days + 1 # Cos we want also the last vector of the day
         randomDays <- x$yday
         if (max(randomDays) + 1 - days < 0)
-          stop("Too many days to randomize, max days available = ", max(randomDays) + 1)
+          stop("Too many days to randomize, max days available = ", max(randomDays) + 1   -1) # So max avail days is -1
+        
         randomDays <- randomDays[randomDays <= max(randomDays) + 1 - days]
         randomDay <- base::sample(randomDays, 1)
-        return(subset(x, yday %in% randomDay:(randomDay + days - 1)))
+        
+        # From the last day, remove everything except the first relocation
+        y <- subset(x, yday %in% randomDay:(randomDay + days - 1))
+        lastDay <- max(y$yday)
+        y <- rbind(y[y$yday != lastDay,,drop=FALSE], y[y$yday == lastDay,,drop=FALSE][1,,drop=FALSE])
+        return(y)
       }, days=days)
       
       observationTracks <- copy(shallow=TRUE) # Avoid possible recursion

### 2015-03-14 inst/simulation/count_intersections.R

@@ -9,7 +9,7 @@
 countIntersections <- function(scenario, iteration, nSurveyRoutes=as.integer(500), countDays=as.integer(1), isTest=F) {
   mss <- getMSS(scenario=scenario, nSurveyRoutes=nSurveyRoutes, isTest=isTest)
   study <- mss$study
-  study$countIntersections(iteration=iteration, days=countDays)
+  study$countIntersections(iteration=iteration, days=countDays, save=TRUE)
 }
 
 library(parallel)

### 2015-03-16 inst/simulation/manuscript.Rmd
...

@@
 
-```{r figure-1-init}
-#scenario <- "A"
-#iterations <- as.integer(1)
-
+```{r figure-1}
 i <- 1
@@
 trackData$scenario <- factor(trackData$scenario)
-```
 
-```{r figure-1}
 #window <- c(3.375, 3.4, 6.9, 6.925)*1e6
@@
 
-```{r, eval=F}
-#x<-data.frame(scenario="C", iteration=as.integer(3), stringsAsFactors=F)
-modelName <- modelNames[2]
-x<-data.frame(scenario="A", iteration=as.integer(10), stringsAsFactors=F)
-study <- getStudy(scenario=x$scenario)
-estimates <- study$getModel(modelName=modelName, iteration=x$iteration)
-estimates$modelName <- modelName
-estimates <- study$loadEstimates(estimates)
-intersections <- study$loadIntersections(iteration=x$iteration)
-head(subset(intersections$intersections@data, intersections>0), 20)
-head(subset(estimates$data, intersections>0), 20)
-
-model <- SimulatedSmoothModelTemporal(study=study, iteration=as.integer(1))
-params <- list(family="nbinomial", offsetScale=1000^2, timeModel="ar1")
-intersections <- study$loadIntersections(iteration = model$iteration)
-model$setup(intersections = intersections, params = params)
-model$estimate()
-
-head(subset(model$data, intersections>0), 20)
-head(subset(intersections$intersections@data, intersections>0), 20)
-
-./count_intersections.sh ~/git/RParallelScreen/ A 1 1 1
-./estimate.sh ~/git/RParallelScreen/ A 1 1 SmoothModel-nbinomial-ar1
-
-./estimate.sh ~/git/RParallelScreen/ A 1 1 FMPModel
-
-```
-
-```{r figure-2-init}
+```{r figure-2}
 getSurveyRouteLocations = function(scenarios, modelName, iterations, years) {
@@
 
-getEstimatedDensity = function(scenarios, modelName, iterations, year, sd=FALSE, log=FALSE) {
+getEstimatedDensity = function(scenarios, modelName, iterations, year, sd=FALSE, weightHabitats=F) {
   ddply(data.frame(scenario=scenarios, year=rep(year, length(scenarios)), iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, sd) {
     message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
-    study <- getStudy(scenario=x$scenario)
+    withHabitatWeights <- if (weightHabitats && substr(x$scenario, 1, 1) %in% c("E","F")) T else F
+    study <- getStudy(scenario=x$scenario, withHabitatWeights=withHabitatWeights)
     estimates <- study$getModel(modelName=modelName, iteration=x$iteration)
@@
     estimates$collectEstimates()
+    estimates$data <- subset(estimates$data, year == x$year)
+    #studyArea <- FinlandStudyArea$new(context=context)$setup(tolerance=0.001)
+    #density <- estimates$getPopulationDensityInterpolate(getSD=sd, maskPolygon=studyArea$boundary)
 
-    studyArea <- FinlandStudyArea$new(context=context)$setup(tolerance=0.001)
+    #habitatWeights <- study$getHabitatWeights(iteration=x$iteration)
+    #density <- estimates$getPopulationDensity(habitatWeights=habitatWeights, .parallel=T)
+    density <- estimates$getPopulationDensity(.parallel=T)
+    density <- if (withHabitatWeights) {
+      habitatWeights <- study$getHabitatWeights(iteration=x$iteration)
+      habitatWeightsRaster <- habitatWeights$getWeightsRaster(save=F)
+      density$weight(habitatWeightsRaster)
+    } else wdensity <- density$copy()
 
-    estimates$data <- subset(estimates$data, year == x$year)
-    #if (log) estimates$data$fittedMean <- log(estimates$data$fittedMean)
-    density <- estimates$getPopulationDensityInterpolate(getSD=sd, maskPolygon=studyArea$boundary)
-    density <- if (sd) density$sd else density$mean
+    #density <- if (sd) density$sd else density$mean
     if (is.null(density) || maxValue(density$rasterStack[[1]]) > 1) stop("Estimation failed.")
@@
 }
-```
+#iterations <- as.integer(c(1,4,1,2,5,31))
+#densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, sd=F, log=F)
 
-```{r figure-2}
 study <- getStudy(scenario="A")
 limits <- extent(study$studyArea$boundary)
-iterations <- as.integer(c(1,6,3,2,5,31))
+iterations <- as.integer(c(1,4,1,2,5,31))
 year <- 2010
@@
 surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
-#surveyRoutes <- rbind(surveyRoutes[surveyRoutes$intersections == FALSE,], surveyRoutes[surveyRoutes$intersections == TRUE,])
 agents <- getAgentLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
-densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, sd=F, log=F)
-# TODO: do not use scaling??
-#densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer) { x[,yearLayer] <- scale(x[,yearLayer]); return(x) }, yearLayer=yearLayer)
+densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year)
+wdensities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, weightHabitats=T)
 
 # Scale density to 1/km^2
-densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer, weight, logTrans) {
+scaleDensity <- function(x, yearLayer, weight) {
   x[,yearLayer] <- x[,yearLayer] * weight
-  if (logTrans) x[,yearLayer] <- log(x[,yearLayer])
   return(x)
-}, yearLayer=yearLayer, weight=1 / (prod(res(study$getTemplateRaster())) / 1000^2), logTrans=F)
+}
+wscale <- 1 / (prod(res(study$getTemplateRaster())) / 1000^2)
+densitiesScaled <- ddply(densities, .(scenario), scaleDensity, yearLayer=yearLayer, weight=wscale)
+wdensitiesScaled <- ddply(wdensities, .(scenario), scaleDensity, yearLayer=yearLayer, weight=wscale)
 
@@
 
-p <- ggplot(densitiesScaled) +
+p <- ggplot(wdensitiesScaled) +
   geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
-  stat_contour(aes_string(x="x", y="y", z=yearLayer, colour=yearLayer), size=0.1)
+  stat_contour(data=densitiesScaled, aes_string(x="x", y="y", z=yearLayer, colour=yearLayer), size=0.1)
 #p <- direct.label(p)
@@
 ```{r figure-3-init}
+cutoffPopulationSizes <- function(populationSizes, cutoff=c(0,Inf), models) {
+  x <- subset(populationSizes, modelName %in% models & cutoff[1] <= Estimated & Estimated <= cutoff[2])
+  y <- subset(populationSizes, !(modelName %in% models))
+  rbind(x, y)
+}
+
 getIntersections <- function(scenarios) {
@@
 }
-
-cutoffPopulationSizes <- function(populationSizes, cutoff=c(0,Inf)) subset(populationSizes, cutoff[1] <= Estimated & Estimated <= cutoff[2])
 
@@
 
-meanPopulationSizes  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff))
-logPopulationSizes <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff))
+meanPopulationSizes  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff, models=modelNames[1]))
+logPopulationSizes <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff, models=modelNames[1]))
 logMeanPopulationSizes  <- getMeanPopulationSizes(logPopulationSizes)
 
-meanPopulationSizes2000  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff))
-logPopulationSizes2000 <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff))
+meanPopulationSizes2000  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[1]))
+logPopulationSizes2000 <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[1]))
 logMeanPopulationSizes2000  <- getMeanPopulationSizes(logPopulationSizes2000)
 
-meanPopulationSizes10Days  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff))
-logPopulationSizes10Days <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff))
+meanPopulationSizes10Days  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[1]))
+logPopulationSizes10Days <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[1]))
 logMeanPopulationSizes10Days  <- getMeanPopulationSizes(logPopulationSizes10Days)
+```
+
+```{r table-1abc}
+getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, overEstimate, models, file, ...) {
+  library(plyr)
+  x <- ddply(populationSizes, .(Scenario, modelName), function(x, overEstimate, models) {
+    xrows <- nrow(x)
+    old <- x
+    x <- cutoffPopulationSizes(x, overEstimate, models=models)
+    if (nrow(x) == 0) {
+      print(old)
+      stop("ERROR")
+    }
+
+    if (nrow(x) == 0) return(NULL)
+    failed <- (1 - nrow(x) / xrows) * 100
+
+    y <- lm(Estimated~Observed, x)
+    prop <- x$Estimated/x$Observed
+    prop <- prop[!is.infinite(prop) & prop > 0]
+    prop.mean <- mean(log(prop))
+    prop.sd <- sd(log(prop))
+    prop.min <- quantile(log(prop), .025)
+    prop.max <- quantile(log(prop), .975)
+    slope.sd <- coef(summary(y))[2,2]
+
+    #print(prop)
+    #message("prop = ", mean(prop), " log mean = ", prop.mean)
+
+    df <- data.frame(Intercept=coef(y)[1], Slope=coef(y)[2], SlopeSD=slope.sd, R2=summary(y)$r.squared, EstTrue=prop.mean, EstTrueSD=prop.sd, EstTrueMin=prop.min, EstTrueMax=prop.max, Failed=failed, N=nrow(x))
+    #print(df)
+    return(df)
+  }, overEstimate=overEstimate, models=models, .inform=T)
+
+  y <- unique(meanPopulationSizes)
+  y$Scenario <- y$scenario  
+  z <- plyr::join(x, y)
+  z$Model <- prettyModelName(z$modelName)
+  z$Scenario <- prettyScenario(z$Scenario)  
+  return(z)
+}
+
+table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff, models=modelNames[1])
+logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(cutoff+1), models=modelNames[1])
+logTable1$Failed <- table1$Failed
+
+table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff, models=modelNames[1])
+logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1), models=modelNames[1])
+logTable2$Failed <- table2$Failed
+
+table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff, models=modelNames[1])
+logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1), models=modelNames[1])
+logTable3$Failed <- table3$Failed
+
+subset(logTable1, Model != "VT")
+subset(logTable2, Model != "VT")
+subset(logTable3, Model != "VT")
+
+writeTable <- function(df, filename, context) {
+  library(ReporteRs)
+  doc <- docx()
+  flextable <- FlexTable(df)
+  addFlexTable(doc, flextable)
+  dir <- normalizePath(context$figuresDirectory)
+  writeDoc(doc, file=file.path(dir, filename))
+}
+
+plotPopulationSizesTable <- function(logTable) {
+  library(reshape2)
+  facet_labeller <- function(variable, value) {
+    measured <- list(Slope=expression(beta), R2=expression(R^2), EstTrue=expression(alpha), Failed="Failed (%)")
+    print(paste0(variable,"=",value))
+    if (variable=="variable") return(measured[value])
+    if (variable=="Case") return(paste("Case", value))
+  }
+  x <- melt(logTable, id.vars=c("Scenario", "Model", "Case"), measure.vars=c("Slope", "R2", "EstTrue", "Failed"))
+  bestFit <- data.frame(variable=unique(x$variable), y=c(1, 1, 0, 0))
+  errorBars <- x
+  errorBars$value <- NA
+  errorBars$value[errorBars$variable=="Slope"] <- logTable$Slope
+  errorBars$qmin[errorBars$variable=="Slope"] <- logTable$Slope-logTable$SlopeSD
+  errorBars$qmax[errorBars$variable=="Slope"] <- logTable$Slope+logTable$SlopeSD
+  errorBars$value[errorBars$variable=="EstTrue"] <- logTable$EstTrue
+  errorBars$qmin[errorBars$variable=="EstTrue"] <- logTable$EstTrueMin
+  errorBars$qmax[errorBars$variable=="EstTrue"] <- logTable$EstTrueMax
+
+  p <- ggplot(x, aes(Scenario, value, group=Model)) +
+    geom_hline(data=bestFit, aes(yintercept=y), color="darkgray") + 
+    geom_errorbar(data=errorBars, aes(Scenario, group=Model, ymax=qmax, ymin=qmin), position=position_dodge(width=.7), colour="darkgray") +
+    geom_point(aes(shape=Model), position=position_dodge(width=.7)) +
+    facet_grid(variable~Case, scales="free_y", labeller=facet_labeller) + ylab("") +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=10),
+                                        legend.position="bottom",
+                                        strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
+                                        plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
+                                        legend.margin=unit(0, units="line"))# , legend.key=element_rect(fill="transparent"))
+  return(p)
+}
+
+logTable1$Case <- "1"
+logTable2$Case <- "2"
+logTable3$Case <- "3"
+logTable <- rbind(subset(logTable1, Model != "VT"), subset(logTable2, Model != "VT"), subset(logTable3, Model != "VT"))
+p <- plotPopulationSizesTable(logTable)
+saveFigure(p, "manuscript-figure-3.png", width=17, height=17, dpi=200) # draft quality
+saveFigure(p, "manuscript-figure-3.pdf", width=17, height=17, dpi=300)
+
+logTablePretty <- logTable
+logTablePretty$Slope <- round(logTablePretty$Slope, 2)
+logTablePretty$R2 <- round(logTablePretty$R2, 2)
+logTablePretty$EstTrue <- round(logTablePretty$EstTrue, 2)
+logTablePretty$EstTrueSD <- round(logTablePretty$EstTrueSD, 2)
+logTablePretty$Failed <- round(logTablePretty$Failed, 0)
+logTablePretty <- logTablePretty[,c("Case","Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
+writeTable(subset(logTablePretty, Model != "VT"), "table-1.docx", context)
+
+#writeTable(subset(logTable1, Model != "VT"), "table-1a.docx", context)
+#writeTable(subset(logTable2, Model != "VT"), "table-1b.docx", context)
+#writeTable(subset(logTable3, Model != "VT"), "table-1c.docx", context)
 ```
@@
 
-```{r table-1abc}
-getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, overEstimate, full=F, file, ...) {
-  library(plyr)
-  x <- ddply(populationSizes, .(Scenario, modelName), function(x, overEstimate) {
-    failed <- (1 - nrow(cutoffPopulationSizes(x, overEstimate)) / nrow(x)) * 100
-      #(sum(x$Estimated < overEstimate[1] | x$Estimated > overEstimate[2]) / nrow(x)) * 100)
-    x <- cutoffPopulationSizes(x, overEstimate)
-    #subset(x, Estimated <= overEstimate)
-    #if (x$modelName[1] == "FMPModel") failed <- NA
-    #if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, SlopeSD=NA, R2=NA, Failed=failed, EstTrue=NA, EstTrueSD=NA, EstTrueMin=NA, EstTrueMax=NA, N=NA))
-    y <- lm(Estimated~Observed, x)
-    prop.mean <- mean(log(x$Estimated/x$Observed))
-    prop.sd <- sd(log(x$Estimated/x$Observed))
-    prop.min <- quantile(log(x$Estimated/x$Observed), .025)
-    prop.max <- quantile(log(x$Estimated/x$Observed), .975)
-    slope.sd <- coef(summary(y))[2,2]
-    return(data.frame(Intercept=coef(y)[1], Slope=coef(y)[2], SlopeSD=slope.sd, R2=summary(y)$r.squared, EstTrue=prop.mean, EstTrueSD=prop.sd, EstTrueMin=prop.min, EstTrueMax=prop.max, Failed=failed, N=nrow(x)))
-  }, overEstimate=overEstimate)
+```{r habitat-bias}
+ddply(logTable, .(modelName, Case, Scenario), function(x) {
+  data.frame(p=mean(exp(x$EstTrue)))
+})
+ddply(logTable, .(modelName, Case), function(x) {
+  data.frame(p=mean(exp(x$EstTrue)))
+})
+ddply(logTable, .(modelName, Scenario), function(x) {
+  data.frame(p=mean(exp(x$EstTrue)))
+})
 
-  y <- unique(meanPopulationSizes)
-  y$Scenario <- y$scenario
-  z <- plyr::join(x, y)
-  z$Model <- prettyModelName(z$modelName)
-  z$Scenario <- prettyScenario(z$Scenario)
-  if (full) return(z)
-
-  return(z)
-}
-
-table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff)
-logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(cutoff+1))
-logTable1$Failed <- table1$Failed
-
-table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff)
-logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1))
-logTable2$Failed <- table2$Failed
-
-table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff)
-logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1))
-logTable3$Failed <- table3$Failed
-
-subset(logTable1, Model != "VT")
-subset(logTable2, Model != "VT")
-subset(logTable3, Model != "VT")
-
-writeTable <- function(df, filename, context) {
-  library(ReporteRs)
-  doc <- docx()
-  flextable <- FlexTable(df)
-  addFlexTable(doc, flextable)
-  dir <- normalizePath(context$figuresDirectory)
-  writeDoc(doc, file=file.path(dir, filename))
-}
-
-plotPopulationSizesTable <- function(logTable) {
-  library(reshape2)
-  facet_labeller <- function(variable, value) {
-    measured <- list(Slope=expression(beta), R2=expression(R^2), EstTrue=expression(alpha), Failed="Failed (%)")
-    print(paste0(variable,"=",value))
-    if (variable=="variable") return(measured[value])
-    if (variable=="Case") return(paste("Case", value))
-  }
-  x <- melt(logTable, id.vars=c("Scenario", "Model", "Case"), measure.vars=c("Slope", "R2", "EstTrue", "Failed"))
-  bestFit <- data.frame(variable=unique(x$variable), y=c(1, 1, 0, 0))
-  errorBars <- x
-  errorBars$value <- NA
-  errorBars$value[errorBars$variable=="Slope"] <- logTable$Slope
-  errorBars$qmin[errorBars$variable=="Slope"] <- logTable$Slope-logTable$SlopeSD
-  errorBars$qmax[errorBars$variable=="Slope"] <- logTable$Slope+logTable$SlopeSD
-  errorBars$value[errorBars$variable=="EstTrue"] <- logTable$EstTrue
-  errorBars$qmin[errorBars$variable=="EstTrue"] <- logTable$EstTrueMin
-  errorBars$qmax[errorBars$variable=="EstTrue"] <- logTable$EstTrueMax
-
-  p <- ggplot(x, aes(Scenario, value, group=Model)) +
-    geom_hline(data=bestFit, aes(yintercept=y), color="darkgray") + 
-    geom_errorbar(data=errorBars, aes(Scenario, group=Model, ymax=qmax, ymin=qmin), position=position_dodge(width=.7), colour="darkgray") +
-    geom_point(aes(shape=Model), position=position_dodge(width=.7)) +
-    facet_grid(variable~Case, scales="free_y", labeller=facet_labeller) + ylab("") +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=10),
-                                        legend.position="bottom",
-                                        strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
-                                        plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
-                                        legend.margin=unit(0, units="line"))# , legend.key=element_rect(fill="transparent"))
-  return(p)
-}
-
-logTable1$Case <- "1"
-logTable2$Case <- "2"
-logTable3$Case <- "3"
-logTable <- rbind(subset(logTable1, Model != "VT"), subset(logTable2, Model != "VT"), subset(logTable3, Model != "VT"))
-p <- plotPopulationSizesTable(logTable)
-saveFigure(p, "manuscript-figure-3.png", width=17, height=17, dpi=200) # draft quality
-saveFigure(p, "manuscript-figure-3.pdf", width=17, height=17, dpi=300)
-
-logTablePretty <- logTable
-logTablePretty$Slope <- round(logTablePretty$Slope, 2)
-logTablePretty$R2 <- round(logTablePretty$R2, 2)
-logTablePretty$EstTrue <- round(logTablePretty$EstTrue, 2)
-logTablePretty$EstTrueSD <- round(logTablePretty$EstTrueSD, 2)
-logTablePretty$Failed <- round(logTablePretty$Failed, 0)
-logTablePretty <- logTablePretty[,c("Case","Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
-writeTable(subset(logTablePretty, Model != "VT"), "table-1.docx", context)
-
-#writeTable(subset(logTable1, Model != "VT"), "table-1a.docx", context)
-#writeTable(subset(logTable2, Model != "VT"), "table-1b.docx", context)
-#writeTable(subset(logTable3, Model != "VT"), "table-1c.docx", context)
+p <- ddply(subset(logTable, Scenario %in% c("A","E")), .(modelName, Case), function(x) {
+  p <- 1 - exp(subset(x, Scenario %in% "A")$EstTrue) / exp(subset(x, Scenario %in% "E")$EstTrue)
+  data.frame(p=p)
+})
+mean(p$p)
+p <- ddply(subset(logTable, Scenario %in% c("A","F")), .(modelName, Case), function(x) {
+  p <- 1 - exp(subset(x, Scenario %in% "A")$EstTrue) / exp(subset(x, Scenario %in% "F")$EstTrue)
+  data.frame(p=p)
+})
 ```
@@
 iall <- rbind(i, i2000, i10days)
+iall
 p <- plotIntersectionPopulationSizeTable(iall)
@@
 #writeTable(i10days, "table-2c.docx", context)
-```
-
-```{r figure-x-debug}
-getTrueAndEstimatedCounts <- function(scenarios, modelNames, cutoff=c(-Inf,Inf)) {
-  study <- getStudy(scenario="A")
-  validation <- Validation(study=study, populationSizeCutoff=cutoff)
-  counts <- validation$getMeanCounts(scenarios=scenarios, modelNames=modelNames)
-  return(counts)
-}
-
-c1 <- getTrueAndEstimatedCounts(scenarios=scenarios, modelNames=modelNames[c(3:5)]); c1$Case <- 1
-c2 <- getTrueAndEstimatedCounts(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(3:5)]); c1$Case <- 2
-c3 <- getTrueAndEstimatedCounts(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(3:5)]); c1$Case <- 3
-c.all <- rbind(c1, c2, c3)
-c.agg <- ddply(c.all, .(iteration), function(x) data.frame(Scenario=x$Scenario[1], Model=x$Model[1], Ratio=mean(x$True/x$Estimated), RatioSD=sd(x$True/x$Estimated)))
-
-#ggplot(c.all, aes())
 ```
@@
 ```{r figure-5}
-getCIProportions <- function(scenario, modelName, panel, scenarioNames, cutoff=Inf) {
+getCIProportions <- function(scenarios, modelNames, panel, scenarioNames, cutoff=Inf) {
   study <- getStudy(scenario="A")
   validation <- Validation(study=study, populationSizeCutoff=cutoff)
-  a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenario, modelNames=modelName)
-  b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenario, modelNames=modelName, probs=c(.25, .75), probsName="50%")
+  a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames)
+  b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames, probs=c(.25, .75), probsName="50%")
   credibilityIntervals <- rbind(a, b)
@@
 cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")
+#cip <- subset(rbind(cip1, cip2, cip3), !(modelName %in% c("SmoothModel-nbinomial-ar1", "SmoothModel-nbinomial-matern-ar1")))
 p <- plotCIProportions(cip)

### 2015-03-16 inst/simulation/estimate.R

@@ -21,6 +21,8 @@ estimateSpatioTemporal <- function(scenario, modelName, iteration, isTest=FALSE,
     #meshParams <- list(coordsScale=1e-6, maxEdge=c(.1e6, .2e6), cutOff=.05e6)
     #meshParams <- list(coordsScale=1e-6, maxEdge=c(.01e6, .2e6), cutOff=.01e6)
     meshParams <- list(coordsScale=1e-6, maxEdge=c(.05e6, .2e6), cutOff=.03e6)
+    
+    meshParams <- list(coordsScale=1e-6, maxEdge=c(.01e6, .02e6), cutOff=.007e6)
     modelParams <- list(family="nbinomial", offsetScale=1000^2, meshParams=meshParams, timeModel="ar1")
     model$setup(intersections=intersections, params=modelParams)
     plot(model$mesh)

### 2015-03-16 inst/simulation/manuscript.Rmd
...

@@
 
-meanPopulationSizes  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff, models=modelNames[1]))
-logPopulationSizes <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff, models=modelNames[1]))
+meanPopulationSizes  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff, models=modelNames[c(1,3,5)]))
+logPopulationSizes <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff, models=modelNames[c(1,3,5)]))
 logMeanPopulationSizes  <- getMeanPopulationSizes(logPopulationSizes)
 
-meanPopulationSizes2000  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[1]))
-logPopulationSizes2000 <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[1]))
+meanPopulationSizes2000  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
+logPopulationSizes2000 <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
 logMeanPopulationSizes2000  <- getMeanPopulationSizes(logPopulationSizes2000)
 
-meanPopulationSizes10Days  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[1]))
-logPopulationSizes10Days <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[1]))
+meanPopulationSizes10Days  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
+logPopulationSizes10Days <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
 logMeanPopulationSizes10Days  <- getMeanPopulationSizes(logPopulationSizes10Days)
@@
 
-table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff, models=modelNames[1])
+#x <- subset(populationSizes, modelName=="SmoothModelMean-nbinomial-ar1-priors1" & Scenario=="C")
+#y <- lm(Estimated~Observed, x)
+
+table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff, models=modelNames[c(1,3,5)])
 logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(cutoff+1), models=modelNames[1])
@@
 
-table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff, models=modelNames[1])
+table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff, models=modelNames[c(1,3,5)])
 logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1), models=modelNames[1])
@@
 
-table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff, models=modelNames[1])
+table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff, models=modelNames[c(1,3,5)])
 logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1), models=modelNames[1])
@@
   errorBars$value <- NA
-  errorBars$value[errorBars$variable=="Slope"] <- logTable$Slope
-  errorBars$qmin[errorBars$variable=="Slope"] <- logTable$Slope-logTable$SlopeSD
-  errorBars$qmax[errorBars$variable=="Slope"] <- logTable$Slope+logTable$SlopeSD
+  #errorBars$value[errorBars$variable=="Slope"] <- logTable$Slope
+  #errorBars$qmin[errorBars$variable=="Slope"] <- logTable$Slope-logTable$SlopeSD
+  #errorBars$qmax[errorBars$variable=="Slope"] <- logTable$Slope+logTable$SlopeSD
   errorBars$value[errorBars$variable=="EstTrue"] <- logTable$EstTrue
@@
 ```{r figure-3sample}
-xlim <- log10(c(10,800)); ylim <- xlim
+xlim <- log10(c(30,800)); ylim <- xlim
 logPopulationSizesSample <- subset(logPopulationSizes, scenario == "A")
@@
 
-pc1 <- getPopulationCorrelations(scenarios=scenarios, modelNames=modelNames[1], cutoff=cutoff)
-pc2a <- getPopulationCorrelations(scenarios="Acombined", modelNames=modelNames[1], cutoff=cutoff)
-pc2b <- getPopulationCorrelations(scenarios="Bcombined", modelNames=modelNames[1], cutoff=cutoff)
-pc2c <- getPopulationCorrelations(scenarios="Ccombined", modelNames=modelNames[1], cutoff=cutoff)
-pc2d <- getPopulationCorrelations(scenarios="Dcombined", modelNames=modelNames[1], cutoff=cutoff)
-pc2e <- getPopulationCorrelations(scenarios="Ecombined", modelNames=modelNames[1], cutoff=cutoff)
-pc2f <- getPopulationCorrelations(scenarios="Fcombined", modelNames=modelNames[1], cutoff=cutoff)
+pc1a <- getPopulationCorrelations(scenarios="A", modelNames=modelNames[1], cutoff=cutoff)
+pc1b <- getPopulationCorrelations(scenarios="B", modelNames=modelNames[1], cutoff=cutoff)
+pc1c <- getPopulationCorrelations(scenarios="C", modelNames=modelNames[1], cutoff=cutoff)
+pc1d <- getPopulationCorrelations(scenarios="D", modelNames=modelNames[1], cutoff=cutoff)
+pc1e <- getPopulationCorrelations(scenarios="E", modelNames=modelNames[1], cutoff=cutoff)
+pc1f <- getPopulationCorrelations(scenarios="F", modelNames=modelNames[1], cutoff=cutoff)
+pc1 <- rbind(pc1a,pc1b,pc1c,pc1d,pc1e,pc1f)
+pc2a <- getPopulationCorrelations(scenarios="Acombined", modelNames=modelNames[1], cutoff=10*cutoff)
+pc2b <- getPopulationCorrelations(scenarios="Bcombined", modelNames=modelNames[1], cutoff=10*cutoff)
+pc2c <- getPopulationCorrelations(scenarios="Ccombined", modelNames=modelNames[1], cutoff=10*cutoff)
+pc2d <- getPopulationCorrelations(scenarios="Dcombined", modelNames=modelNames[1], cutoff=10*cutoff)
+pc2e <- getPopulationCorrelations(scenarios="Ecombined", modelNames=modelNames[1], cutoff=10*cutoff)
+pc2f <- getPopulationCorrelations(scenarios="Fcombined", modelNames=modelNames[1], cutoff=10*cutoff)
 pc2 <- rbind(pc2a,pc2b,pc2c,pc2d,pc2e,pc2f)

### 2015-03-16 R/Validation.R

@@ -166,6 +166,8 @@ Validation <- setRefClass(
         estimates <- study$loadEstimates(estimates)
         estimates$collectEstimates()
         
+        if (any(!is.finite(estimates$data$fittedMean))) next
+        
         estimated <- estimates$getPopulationDensity(templateRaster=template)
         #estimated <- estimates$getPopulationDensityInterpolate(templateRaster=template, maskPolygon=NULL, getSD=F)
         #if (is.null(estimated$mean)) {
@@ -208,6 +210,8 @@ Validation <- setRefClass(
         spatialCorrelation <- rbind(spatialCorrelation, x)
       }
       
+      if (nrow(spatialCorrelation) == 0) return(NULL)
+      
       spatialCorrelation$Scenario <- study$response
       spatialCorrelation$modelName <- modelName
       
@@ -228,7 +232,7 @@ Validation <- setRefClass(
         s <- getStudy(scenario=x$scenario, isTest=F)
         validation <- Validation(study=s, populationSizeCutoff=populationSizeCutoff)
         z <- validation$validateSpatialPopulationSize(x$modelName, debugMaxIterations=debugMaxIterations)
-        if (nrow(z) == 0) next
+        if (is.null(z)) next
         spatialCorrelations <- rbind(spatialCorrelations, z)
       }
       return(spatialCorrelations)

### 2015-03-18 inst/simulation/manuscript.Rmd
...

@@
 pc2 <- rbind(pc2a,pc2b,pc2c,pc2d,pc2e,pc2f)
+pc3a <- getPopulationCorrelations(scenarios="A10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3b <- getPopulationCorrelations(scenarios="B10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3c <- getPopulationCorrelations(scenarios="C10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3d <- getPopulationCorrelations(scenarios="D10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3e <- getPopulationCorrelations(scenarios="E10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3f <- getPopulationCorrelations(scenarios="F10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3 <- rbind(pc3a,pc3b,pc3c,pc3d,pc3e,pc3f)
+
 pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1], cutoff=cutoff)

### 2015-03-18 inst/simulation/manuscript.Rmd
...

--- 2015-03-16 inst/simulation/manuscript.Rmd
@@
 pc2 <- rbind(pc2a,pc2b,pc2c,pc2d,pc2e,pc2f)
-pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1], cutoff=cutoff)
+pc3a <- getPopulationCorrelations(scenarios="A10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3b <- getPopulationCorrelations(scenarios="B10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3c <- getPopulationCorrelations(scenarios="C10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3d <- getPopulationCorrelations(scenarios="D10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3e <- getPopulationCorrelations(scenarios="E10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3f <- getPopulationCorrelations(scenarios="F10days", modelNames=modelNames[1], cutoff=cutoff)
+pc3 <- rbind(pc3a,pc3b,pc3c,pc3d,pc3e,pc3f)
+#pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1], cutoff=cutoff)
 

### 2015-04-15 inst/simulation/manuscript.Rmd
...

@@
 populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames[c(1,3,4,5)])
-populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,4,5)])
-populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4,5)])
+#populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,4,5)])
+#populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4,5)])
 
@@
 
-meanPopulationSizes2000  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
-logPopulationSizes2000 <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
-logMeanPopulationSizes2000  <- getMeanPopulationSizes(logPopulationSizes2000)
+#meanPopulationSizes2000  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
+#logPopulationSizes2000 <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
+#logMeanPopulationSizes2000  <- getMeanPopulationSizes(logPopulationSizes2000)
 
-meanPopulationSizes10Days  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
-logPopulationSizes10Days <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
-logMeanPopulationSizes10Days  <- getMeanPopulationSizes(logPopulationSizes10Days)
+#meanPopulationSizes10Days  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
+#logPopulationSizes10Days <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
+#logMeanPopulationSizes10Days  <- getMeanPopulationSizes(logPopulationSizes10Days)
 ```
@@
     y <- lm(Estimated~Observed, x)
-    prop <- x$Estimated/x$Observed
+    #prop <- x$Estimated/x$Observed
+    prop <- (x$Estimated+1)/(x$Observed+1)
     prop <- prop[!is.infinite(prop) & prop > 0]
@@
 table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff, models=modelNames[c(1,3,5)])
-logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(cutoff+1), models=modelNames[1])
+logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(cutoff+1), models=modelNames[c(1,3,5)])
 logTable1$Failed <- table1$Failed
 
-table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff, models=modelNames[c(1,3,5)])
-logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1), models=modelNames[1])
-logTable2$Failed <- table2$Failed
+#table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff, models=modelNames[c(1,3,5)])
+#logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1), models=modelNames[1])
+#logTable2$Failed <- table2$Failed
 
-table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff, models=modelNames[c(1,3,5)])
-logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1), models=modelNames[1])
-logTable3$Failed <- table3$Failed
+#table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff, models=modelNames[c(1,3,5)])
+#logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1), models=modelNames[1])
+#logTable3$Failed <- table3$Failed
 
@@
 
+facet_wrap_labeller <- function(gg.plot,labels=NULL) {
+  #works with R 3.0.1 and ggplot2 0.9.3.1
+  require(gridExtra)
+
+  g <- ggplotGrob(gg.plot)
+  gg <- g$grobs      
+  strips <- grep("strip_t", names(gg))
+
+  for(ii in seq_along(labels))  {
+    modgrob <- getGrob(gg[[strips[ii]]], "strip.text", 
+                       grep=TRUE, global=TRUE)
+    gg[[strips[ii]]]$children[[modgrob$name]] <- editGrob(modgrob,label=labels[ii])
+  }
+
+  g$grobs <- gg
+  class(g) = c("arrange", "ggplot",class(g)) 
+  g
+}
+
 plotPopulationSizesTable <- function(logTable) {
@@
   facet_labeller <- function(variable, value) {
-    measured <- list(Slope=expression(beta), R2=expression(R^2), EstTrue=expression(alpha), Failed="Failed (%)")
+    measured <- list(EstTrue=expression(alpha), Slope=expression(beta), R2=expression(R^2), Failed="Failed (%)")
     print(paste0(variable,"=",value))
@@
   }
-  x <- melt(logTable, id.vars=c("Scenario", "Model", "Case"), measure.vars=c("Slope", "R2", "EstTrue", "Failed"))
-  bestFit <- data.frame(variable=unique(x$variable), y=c(1, 1, 0, 0))
+  #x <- melt(logTable, id.vars=c("Scenario", "Model", "Case"), measure.vars=c("EstTrue", "Slope", "R2", "Failed"))
+  x <- melt(logTable, id.vars=c("Scenario", "Model"), measure.vars=c("EstTrue", "Slope", "R2", "Failed"))
+
+  bestFit <- data.frame(variable=unique(x$variable), y=c(0, 1, 1, 0))
   errorBars <- x
@@
     geom_errorbar(data=errorBars, aes(Scenario, group=Model, ymax=qmax, ymin=qmin), position=position_dodge(width=.7), colour="darkgray") +
-    geom_point(aes(shape=Model), position=position_dodge(width=.7)) +
-    facet_grid(variable~Case, scales="free_y", labeller=facet_labeller) + ylab("") +
+    geom_point(aes(shape=Model), position=position_dodge(width=.7)) + ylab("") +
+    #facet_grid(variable~., scales="free_y", labeller=facet_labeller) +
+    facet_wrap(~variable, scales="free_y", nrow=2) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=10),
@@
                                         strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
-                                        plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
-                                        legend.margin=unit(0, units="line"))# , legend.key=element_rect(fill="transparent"))
+                                        plot.margin=unit(c(0, 0.2, 0, -.5), units="line"),  #top, right, bottom, and left margins
+                                        legend.margin=unit(-1.5, units="line"))# , legend.key=element_rect(fill="transparent"))
+
+  # Expand y-axis 100 for Failed
+  x <- data.frame(Scenario="A", Model="FMP", variable="Failed", value=100)
+  p <- p + geom_blank(data=x, aes(x=Scenario, y=value, group=Model))
+
+  p <- facet_wrap_labeller(p, labels=c(expression(alpha), Slope=expression(beta), R2=expression(R^2), Failed="Failed (%)"))
+
   return(p)
@@
 
-logTable1$Case <- "1"
-logTable2$Case <- "2"
-logTable3$Case <- "3"
-logTable <- rbind(subset(logTable1, Model != "VT"), subset(logTable2, Model != "VT"), subset(logTable3, Model != "VT"))
+#logTable1$Case <- "1"
+#logTable2$Case <- "2"
+#logTable3$Case <- "3"
+#logTable <- rbind(subset(logTable1, Model != "VT"), subset(logTable2, Model != "VT"), subset(logTable3, Model != "VT"))
+logTable <- logTable1
 p <- plotPopulationSizesTable(logTable)
-saveFigure(p, "manuscript-figure-3.png", width=17, height=17, dpi=200) # draft quality
-saveFigure(p, "manuscript-figure-3.pdf", width=17, height=17, dpi=300)
+saveFigure(p, "manuscript-figure-3.png", width=17, dpi=200) # draft quality
+saveFigure(p, "manuscript-figure-3.pdf", height=17, dpi=300)
 
@@
 logTablePretty$Failed <- round(logTablePretty$Failed, 0)
-logTablePretty <- logTablePretty[,c("Case","Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
+#logTablePretty <- logTablePretty[,c("Case","Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
+logTablePretty <- logTablePretty[,c("Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
 writeTable(subset(logTablePretty, Model != "VT"), "table-1.docx", context)
@@
 #writeTable(subset(logTable3, Model != "VT"), "table-1c.docx", context)
+```
+
+```{r figure-3sample}
+xlim <- log10(c(30,800)); ylim <- xlim
+logPopulationSizesSample <- subset(logPopulationSizes, scenario == "A")
+logPopulationSizesSample <- cutoffPopulationSizes(logPopulationSizesSample, cutoff, models=modelNames[c(1,3,5)])
+logPopulationSizesSample$modelName <- prettyModelName(logPopulationSizesSample$modelName)
+logPopulationSizesSample <- rbind(
+  subset(logPopulationSizesSample, modelName=="ST"),
+  subset(logPopulationSizesSample, modelName=="T1"),
+  subset(logPopulationSizesSample, modelName=="T2"),
+  subset(logPopulationSizesSample, modelName=="FMP"))
+
+lm(Estimated~Observed, subset(logPopulationSizes, modelName=="FMPModel" & Scenario=="A"))
+lm(Estimated~Observed, subset(logPopulationSizesSample, modelName=="FMP" & Scenario=="A"))
+
+plotLogPopulationSizeSample <- function(size, modelName, case, xlim, ylim) {
+  library(ggplot2)
+  library(scales)  
+  size$Scenario <- prettyScenario(size$Scenario)
+  meanSize <- getMeanPopulationSizes(size)
+  meanSize$Scenario <- prettyScenario(meanSize$Scenario)
+
+  p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
+    geom_point(size=1, colour="darkgray", alpha=.5) +
+    geom_abline(colour="black", linetype="longdash") + 
+    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black")
+
+  #if (!any(abs(size$Estimated) == Inf)) p <- p + geom_smooth(method="lm", se=F, fullrange=T, colour="black")
+  ls <- ddply(size, .(modelName), function(x) if (!any(abs(x$Estimated) == Inf)) coef(lm(Estimated~Observed, x)) else c(NA,NA))
+  print(ls)
+  p <- p + geom_abline(data=ls, aes(intercept=`(Intercept)`, slope=Observed))
+
+  p <- p + facet_wrap(~modelName) +
+    scale_x_continuous(limits=xlim, labels=trans_format("identity", math_format(10^.x))) +
+    scale_y_continuous(limits=ylim, labels=trans_format("identity", math_format(10^.x))) +
+    xlab("True population size") + ylab("Estimated population size") +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+                                        plot.title=element_text(size=10),
+                                        plot.margin=unit(c(0, 0.1, 0, 0), units="line"), #top, right, bottom, and left margins
+                                        legend.margin=unit(0, units="line"),
+                                        panel.margin=unit(0, units="line"))
+
+  return(p)
+}
+
+p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName %in% c("ST","T1","T2","FMP")), modelName=c("ST","T1","T2","FMP"), "a", xlim=xlim, ylim=ylim)
+saveFigure(p, "manuscript-figure-3sample.png", height=7.87) # draft quality
+saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87, dpi=300)
 ```
@@
 saveFigure(p, "manuscript-figure-3c4.pdf", dpi=300)
-```
-
-```{r figure-3sample}
-xlim <- log10(c(30,800)); ylim <- xlim
-logPopulationSizesSample <- subset(logPopulationSizes, scenario == "A")
-logPopulationSizesSample$modelName <- prettyModelName(logPopulationSizesSample$modelName)
-logPopulationSizesSample <- rbind(
-  subset(logPopulationSizesSample, modelName=="ST"),
-  subset(logPopulationSizesSample, modelName=="T1"),
-  subset(logPopulationSizesSample, modelName=="T2"),
-  subset(logPopulationSizesSample, modelName=="FMP"))
-
-#p <- plotLogPopulationSize(subset(logPopulationSizesSample, modelName==modelNames[c(1,3,4)]), modelName=modelNames[c(1,3,4)], "a", xlim=xlim, ylim=ylim)
-
-plotLogPopulationSizeSample <- function(size, modelName, case, xlim, ylim) {
-  library(ggplot2)
-  library(scales)  
-  size$Scenario <- prettyScenario(size$Scenario)
-  meanSize <- getMeanPopulationSizes(size)
-  meanSize$Scenario <- prettyScenario(meanSize$Scenario)
-
-  p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
-    geom_point(size=1, colour="darkgray", alpha=.5) +
-    geom_abline(colour="black", linetype="longdash") + 
-    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black")
-
-  #if (!any(abs(size$Estimated) == Inf)) p <- p + geom_smooth(method="lm", se=F, fullrange=T, colour="black")
-  ls <- ddply(size, .(modelName), function(x) if (!any(abs(x$Estimated) == Inf)) coef(lm(Estimated~Observed, x)) else c(NA,NA))
-  p <- p + geom_abline(data=ls, aes(intercept=`(Intercept)`, slope=Observed))
-
-  p <- p + facet_wrap(~modelName) +
-    scale_x_continuous(limits=xlim, labels=trans_format("identity", math_format(10^.x))) +
-    scale_y_continuous(limits=ylim, labels=trans_format("identity", math_format(10^.x))) +
-    xlab("True population size") + ylab("Estimated population size") +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
-                                        plot.title=element_text(size=10),
-                                        plot.margin=unit(c(0, 0.1, 0, 0), units="line"), #top, right, bottom, and left margins
-                                        legend.margin=unit(0, units="line"),
-                                        panel.margin=unit(0, units="line"))
-
-  return(p)
-}
-
-p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName==c("ST","T1","T2","FMP")), modelName=c("ST","T1","T2","FMP"), "a", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3sample.png", height=7.87*2/3) # draft quality
-saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87*2/3, dpi=300)
 ```
@@
   scale_y_continuous(limits=c(-1,1)) +
-  ggtitle(paste("Case:", case)) +
+  #ggtitle(paste("Case:", case)) +
   theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
                                       plot.title=element_text(size=10),
-                                      plot.margin=unit(c(0.2, 0, 0, 0), units="line"),
+                                      plot.margin=unit(c(0, 0.2, 0, 0), units="line"),
                                       panel.margin=unit(0, units="line"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
@@
 
+cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,5)], panel="Case 1", scenarioNames=scenarios, cutoff=cutoff)
+#cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,5)], panel="Case 2", scenarioNames=scenarios, cutoff=10*cutoff)
+#cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,5)], panel="Case 3", scenarioNames=scenarios, cutoff=cutoff)
+
 plotCIProportions <- function(credibilityIntervals) {
-  modelLabels <- prettyModelName(unique(credibilityIntervals$modelName))
+  #modelLabels <- prettyModelName(unique(credibilityIntervals$modelName))
+  credibilityIntervals$Model <- prettyModelName(credibilityIntervals$modelName)
   credibilityIntervals$CI <- credibilityIntervals$Interval
-  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=CI, colour=modelName)) +
+  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Model, colour=CI)) +
     geom_hline(aes(yintercept=.5), color="grey90") +
     geom_hline(aes(yintercept=.95), color="grey90") +
-    geom_point() +
-    scale_shape_manual(values=c(16, 1)) +
-    scale_colour_manual("Model", values=c("black", "gray26", "gray80"), labels=modelLabels) +
+    geom_point(aes(shape=Model, group=Model), position=position_dodge(width=.7)) +
+    scale_shape_manual("Model", values=c(17,15,3)) +
+    scale_colour_manual("CI", values=c("gray", "black")) +
     #scale_colour_manual("Model", values=c("blue", "green", "red"), labels=modelLabels) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
-    xlab("Scenario") + ylab("Proportion") +
-    facet_wrap(~panel) + ylim(0, 1) +
+    xlab("Scenario") + ylab("Proportion") + ylim(0, 1) +
+    #facet_wrap(~panel) +
     theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
-          plot.margin=unit(c(0.2, 0, 0, 0), units="line"),
+          plot.margin=unit(c(0.2, 0.2, 0, 0), units="line"),
           #legend.background=element_rect(fill="transparent"),
@@
 }
-#p <- plotCIProportions(cip)
-#saveFigure(p, "manuscript-figure-5.png", height=7.87/2, dpi=200)
 
-cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,5)], panel="Case 1", scenarioNames=scenarios, cutoff=cutoff)
-cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,5)], panel="Case 2", scenarioNames=scenarios, cutoff=10*cutoff)
-cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,5)], panel="Case 3", scenarioNames=scenarios, cutoff=cutoff)
-
-cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")
+cip <- cip1
+#cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")
 #cip <- subset(rbind(cip1, cip2, cip3), !(modelName %in% c("SmoothModel-nbinomial-ar1", "SmoothModel-nbinomial-matern-ar1")))

### 2015-04-15 R/Covariates.R

@@ -1,3 +1,53 @@
+Covariates <- setRefClass(
+  Class = "Covariates",
+  fields = list(
+    study = "Study",
+    covariatesName = "character",
+    covariates = "data.frame"
+  ),
+  methods = list(
+    getCovariatesFileName = function() {
+      if (inherits(study, "uninitializedField"))
+        stop("Provide study parameters.")
+      study$context$getFileName(dir=study$context$scratchDirectory, name=covariatesName, region=study$studyArea$region)
+    },
+    
+    saveCovariates = function(fileName=getCovariatesFileName()) {
+      save(covariates, file=fileName)
+      return(invisible(.self))
+    },
+    
+    loadCovariates = function(fileName=getCovariatesFileName()) {
+      load(fileName, envir=as.environment(.self))  
+      return(invisible(.self))
+    },
+    
+    obtain = function(save=F) {
+      stop("Unimplemented abstract method.")
+    }  
+  )
+)
+
+SmoothHabitatCovariates <- setRefClass(
+  Class =" SmoothHabitatCovariates",
+  contains = "Covariates",
+  field = list(
+    scales = "numeric"
+  ),
+  methods = list(
+    obtain = function(save=F) {
+      transects <- study$loadSurveyRoutes(findLengths=F)
+      habitatWeights <- study$loadHabitatWeights()
+      habitatValues <- dlply(habitatWeights$weights[habitatWeights$weights$type != 0,], .(type), function(x) x$habitat)
+      edgeValues <- habitatWeights$weights$habitat[habitatWeights$weights$type == 0]
+      covariates <<- smoothSubsets(study$studyArea$habitat, coords=transects$centroids, scales=scales, processValues=habitatValues, edgeValues=edgeValues)
+      if (save) saveCovariates()
+    }
+  )
+)
+
+
+### TODO: fix
 FinlandCovariates <- setRefClass(
   Class = "FinlandCovariates",
   fields = list(
@@ -6,10 +56,10 @@ FinlandCovariates <- setRefClass(
     covariates = "data.frame"
   ),
   methods = list(
-    initialize = function(...) {
-      callSuper(...)
-      return(invisible(.self))
-    },
+    #initialize = function(study, covariatesName, ...) {
+    #  callSuper(study=study, covariatesName=covariatesName, ...)
+    #  return(invisible(.self))
+    #},
     
     # Note: Currently uses the same raster file for all inheriting classes.
     # Can be changed by adding covariatesName field in the file name.

### 2015-04-17 inst/simulation/manuscript.Rmd
...

@@
 limits <- extent(study$studyArea$boundary)
-iterations <- as.integer(c(1,4,1,2,5,31))
+iterations <- as.integer(c(1,4,1,2,5,34))
 year <- 2010
@@
 
-#library(directlabels)
+gridTemplate <- raster(extent(study$studyArea$boundary), nrows=12, ncols=6, crs=study$studyArea$proj4string)
+gridTemplate@extent@xmax <- gridTemplate@extent@xmin + dim(gridTemplate)[2] * 100 * 1e3
+gridTemplate@extent@ymax <- gridTemplate@extent@ymin + dim(gridTemplate)[1] * 100 * 1e3
 
-p <- ggplot(wdensitiesScaled) +
+p1 <- ggplot(wdensitiesScaled) +
   geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
-  stat_contour(data=densitiesScaled, aes_string(x="x", y="y", z=yearLayer, colour=yearLayer), size=0.1)
-#p <- direct.label(p)
-p <- p + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent",
+  stat_contour(data=densitiesScaled, aes_string(x="x", y="y", z=yearLayer, colour=yearLayer), size=0.1) +
+  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent",
                        guide=guide_legend(title=expression(paste("Population density ", (individuals/km^2)))), trans="log10",
@@
   facet_grid(~scenario) + coord_equal() + theme_raster(base_size=10) +
-  theme(legend.position="bottom", legend.margin=unit(-2, units="line"),
-        plot.margin=unit(c(-.5, 0, 0, -.5), units="line"))
-#p <- direct.label(p)
-saveFigure(p, "manuscript-figure-2a.png", width=17, height=7.2, dpi=200) # draft quality
-saveFigure(p, "manuscript-figure-2a.pdf", width=17, height=7.2, dpi=300)
+  theme(legend.position="bottom", legend.margin=unit(-2, units="line"), legend.title=element_text(size=10), legend.text=element_text(size=10),
+        plot.margin=unit(c(-3, 0, 0, -.5), units="line")) #top, right, bottom, and left margins
+g1 <- arrangeGrob(p1, main=textGrob("a)", x=0.01, just="left", gp=gpar(fontsize=10)))
+saveFigure(g1, "manuscript-figure-2a.png", width=17, height=8.8, dpi=200) # draft quality
+saveFigure(g1, "manuscript-figure-2a.pdf", width=17, height=8.8, dpi=300)
 
-###
-
-gridTemplate <- raster(extent(study$studyArea$boundary), nrows=12, ncols=6, crs=study$studyArea$proj4string)
-gridTemplate@extent@xmax <- gridTemplate@extent@xmin + dim(gridTemplate)[2] * 100 * 1e3
-gridTemplate@extent@ymax <- gridTemplate@extent@ymin + dim(gridTemplate)[1] * 100 * 1e3
-
-p <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey") +
+p2 <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey") +
   geom_point(data=surveyRoutes, aes(x, y, colour=intersections), size=1) + scale_color_manual(values=c("darkgrey", "black")) +
@@
   facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal() +
-  theme(panel.grid.major=element_line(colour="grey"), plot.margin=unit(c(0, 0, 0, -.5), units="line"))
-saveFigure(p, "manuscript-figure-2b.png", width=17, height=6.7, dpi=200) # draft quality
-saveFigure(p, "manuscript-figure-2b.pdf", width=17, height=6.7, dpi=300)
+  theme(panel.grid.major=element_line(colour="grey"),
+        plot.margin=unit(c(0, 0, 0, -.5), units="line")) #top, right, bottom, and left margins
+g2 <- arrangeGrob(p2, main=textGrob("b)", x=0.01, just="left", gp=gpar(fontsize=10)))
+saveFigure(g2, "manuscript-figure-2b.png", width=17, height=6.9, dpi=200) # draft quality
+saveFigure(g2, "manuscript-figure-2b.pdf", width=17, height=6.9, dpi=300)
+
+p <- arrangeGrob(g1, g2)
+saveFigure(p, "manuscript-figure-2.png", width=17, height=17, dpi=200) # draft quality
+saveFigure(p, "manuscript-figure-2.pdf", width=17, height=17, dpi=300)
+
 ```
@@
 
-plotPopulationSizesTable <- function(logTable) {
+plotPopulationSizesTable_DEPRECATED <- function(logTable) {
   library(reshape2)
@@
   }
+
   #x <- melt(logTable, id.vars=c("Scenario", "Model", "Case"), measure.vars=c("EstTrue", "Slope", "R2", "Failed"))
@@
   p <- facet_wrap_labeller(p, labels=c(expression(alpha), Slope=expression(beta), R2=expression(R^2), Failed="Failed (%)"))
+
+  return(p)
+}
+
+plotPopulationSizesTable <- function(logTable) {
+  library(reshape2)
+  library(dplyr)
+  library(gtable)
+  library(gridExtra)
+
+  x <- melt(logTable, id.vars=c("Scenario", "Model"), measure.vars=c("EstTrue", "Slope", "R2", "Failed"))
+
+  bestFit <- data.frame(variable=unique(x$variable), y=c(0, 1, 1, 0))
+  errorBars <- x
+  errorBars$value <- NA
+  errorBars$value[errorBars$variable=="EstTrue"] <- logTable$EstTrue
+  errorBars$qmin[errorBars$variable=="EstTrue"] <- logTable$EstTrueMin
+  errorBars$qmax[errorBars$variable=="EstTrue"] <- logTable$EstTrueMax
+
+  p1 <- x %>% filter(variable=="EstTrue") %>%
+    ggplot(aes(Scenario, value, group=Model)) +
+    geom_hline(data=data.frame(y=0), aes(yintercept=y), color="darkgray") + 
+    geom_errorbar(data=errorBars, aes(Scenario, group=Model, ymax=qmax, ymin=qmin), position=position_dodge(width=.7), colour="darkgray") +
+    geom_point(aes(shape=Model), position=position_dodge(width=.7)) +
+    xlab("") + ylab(expression(alpha)) +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+                                        legend.title=element_text(size=10), legend.position="none",
+                                        strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
+                                        plot.margin=unit(c(0, 0.2, -.75, 0), units="line")) #top, right, bottom, and left margins
+  g1 <- arrangeGrob(p1, main=textGrob("a)", x=0.01, just="left", gp=gpar(fontsize=10)))
+
+  p2 <- x %>% filter(variable=="Slope") %>%
+    ggplot(aes(Scenario, value, group=Model)) +
+    geom_hline(data=data.frame(y=1), aes(yintercept=y), color="darkgray") + 
+    geom_point(aes(shape=Model), position=position_dodge(width=.7)) +
+    xlab("") + ylab(expression(beta)) +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+                                        legend.title=element_text(size=10), legend.position="none",
+                                        strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
+                                        plot.margin=unit(c(0, 0.2, -.75, 0), units="line"))
+  g2 <- arrangeGrob(p2, main=textGrob("b)", x=0.01, just="left", gp=gpar(fontsize=10)))
+
+  p3 <- x %>% filter(variable=="R2") %>%
+    ggplot(aes(Scenario, value, group=Model)) +
+    geom_hline(data=data.frame(y=1), aes(yintercept=y), color="darkgray") + 
+    geom_point(aes(shape=Model), position=position_dodge(width=.7)) +
+    xlab("") + ylab(expression(R^2)) +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+                                        legend.title=element_text(size=10), legend.position="none",
+                                        strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
+                                        plot.margin=unit(c(0, 0.2, -.75, 0), units="line"))
+  g3 <- arrangeGrob(p3, main=textGrob("c)", x=0.01, just="left", gp=gpar(fontsize=10)))
+
+  p4 <- x %>% filter(variable=="Failed") %>%
+    ggplot(aes(Scenario, value, group=Model)) +
+    geom_hline(data=data.frame(y=0), aes(yintercept=y), color="darkgray") + 
+    geom_point(aes(shape=Model), position=position_dodge(width=.7)) +
+    xlab("") + ylab("Failed (%)") + ylim(0,100) +
+    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+                                        legend.title=element_text(size=10), legend.position="none",
+                                        strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
+                                        plot.margin=unit(c(0, 0.2, -.75, 0), units="line"))
+  g4 <- arrangeGrob(p4, main=textGrob("d)", x=0.01, just="left", gp=gpar(fontsize=10)))
+
+  legend <- gtable_filter(ggplot_gtable(ggplot_build(p1 + theme(legend.position="bottom", legend.margin=unit(-1, units="line")))), "guide-box")
+  lheight <- sum(legend$height)
+  p <- arrangeGrob(g1, g2, g3, g4, ncol=2)
+  theight <- unit(10, "points")
+  p <- arrangeGrob(p, textGrob("Scenario", gp=gpar(fontsize=10)), heights=unit.c(unit(1, "npc") - theight, theight))
+  p <- arrangeGrob(p, legend, heights=unit.c(unit(1, "npc") - lheight, lheight), ncol=1)
 

### 2015-04-17 R/Covariates.R

@@ -24,12 +24,12 @@ Covariates <- setRefClass(
     
     obtain = function(save=F) {
       stop("Unimplemented abstract method.")
-    }  
+    }
   )
 )
 
 SmoothHabitatCovariates <- setRefClass(
-  Class =" SmoothHabitatCovariates",
+  Class = "SmoothHabitatCovariates",
   contains = "Covariates",
   field = list(
     scales = "numeric"
@@ -45,6 +45,44 @@ SmoothHabitatCovariates <- setRefClass(
     }
   )
 )
+
+PopulationDensityCovariates <- setRefClass(
+  Class = "PopulationDensityCovariates",
+  contains = "Covariates",
+  field = list(
+    years = "numeric"
+  ),
+  methods = list(
+    obtain = function(save=F) {
+      request <- gisfin::GeoStatFiWFSRequest$new()$getPopulationLayers()
+      client <- gisfin::GeoStatFiWFSClient$new(request)
+      layers <- client$listLayers()
+      
+      availableLayers <- layers[grep("vaki\\d+_5km", layers)]
+      availableYears <- as.integer(substr(availableLayers, 17, 20))
+      nearestIndex <- sapply(years, function(x) which.min(abs(x-availableYears)))
+      nearestYears <- unique(availableYears[nearestIndex])
+
+      for (i in 1:length(availableYears)) {
+        nearestYear <- availableYears[i]
+        nearestLayer <- availableLayers[i]
+        focalYears <- years[availableYears[nearestIndex] == nearestYear]
+        
+        request$getPopulation(nearestLayer)
+        client <- gisfin::GeoStatFiWFSClient$new(request)
+        populationLayer <- client$getLayer(layerName)
+        x <- sp::SpatialPixelsDataFrame(coordinates(populationLayer), populationLayer@data[,"vaesto", drop=F], proj4string=populationLayer@proj4string)
+        population <- raster::stack(x)
+                
+        for (year in focalYears) {
+          # process movements
+        }
+      }
+    }
+  )
+)
+
+
 
 
 ### TODO: fix

### 2015-04-29 inst/simulation/manuscript.Rmd
...

@@
   populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
-  p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
+  p <- ggplot(populationCorrelations, aes(True, Correlation)) +
+    geom_hline(data=data.frame(y=0), aes(yintercept=y), color="darkgray") +
+    geom_point(size=1, colour="black") +
     facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
@@
   p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Model, colour=CI)) +
-    geom_hline(aes(yintercept=.5), color="grey90") +
-    geom_hline(aes(yintercept=.95), color="grey90") +
+    geom_hline(aes(yintercept=.5), color="darkgrey") +
+    geom_hline(aes(yintercept=.95), color="darkgrey") +
     geom_point(aes(shape=Model, group=Model), position=position_dodge(width=.7)) +

### 2015-04-30 inst/wtc-finland/manuscript2.Rmd
...

---
title: "Manuscript"
output: html_document
---

```{r initialize}
#library(devtools); install_github("statguy/Winter-Track-Counts")
library(parallel)
library(doMC)
registerDoMC(cores=detectCores())
library(WTC)
source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
library(ggplot2)
library(scales)
library(grid)
library(plyr)

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
boundaryDF <- ggplot2::fortify(FinlandWTCStudy(context=context, response="canis.lupus")$studyArea$boundary)
study <- FinlandWTCStudy$new(context=context, response="canis.lupus")
#study$studyArea$loadBoundary(thin=F, tolerance=0.001)
```

```{r figure-convolution-example}
r <- crop(study$studyArea$habitat, extent(3400000, 3410000, 7000000, 7010000))
plot(r)
forest <- calc(r, function(x) if (is.na(x) || x %in% c(0,45:255)) NA else x %in% 18:35)
plot(forest)
coords <- expand.grid(1:dim(r)[1], 1:dim(r)[2])
x <- smoothSubsets(r=forest, coords=coords, kernelFun=expKernel, scales=c(125, 1000), .parallel=T)
plot(flip(rasterFromXYZ(x[,c("x","y","value")]), direction="y"), asp=1)
```

### 2015-05-02 inst/wtc-finland/manuscript2.Rmd
---

@@
 library(plyr)
+library(gridExtra)
 
@@
 r <- crop(study$studyArea$habitat, extent(3400000, 3410000, 7000000, 7010000))
-plot(r)
+#plot(r)
 forest <- calc(r, function(x) if (is.na(x) || x %in% c(0,45:255)) NA else x %in% 18:35)
-plot(forest)
-coords <- expand.grid(1:dim(r)[1], 1:dim(r)[2])
+#plot(forest)
+coords <- xyFromCell(r, cell=1:ncell(r))
+#x0 <- smoothSubsets(r=forest, coords=coords[1:(400*10),], kernelFun=expKernel, scales=125, .parallel=F)
 x <- smoothSubsets(r=forest, coords=coords, kernelFun=expKernel, scales=c(125, 1000), .parallel=T)
-plot(flip(rasterFromXYZ(x[,c("x","y","value")]), direction="y"), asp=1)
+#plot(flip(rasterFromXYZ(x[,c("x","y","value")]), direction="y"), asp=1)
+fileName <- study$context$getFileName(dir=study$context$processedDataDirectory, name="figure-b-1", region=study$studyArea$region)
+save(x, file=fileName)
+load(fileName)
+
+r1 <- as.data.frame(rasterToPoints(r))
+colnames(r1) <- c("x","y","value")
+r1$value <- factor(r1$value, levels=0:255)
+r2 <- as.data.frame(rasterToPoints(forest))
+colnames(r2) <- c("x","y","value")
+r2$value <- factor(r2$value, levels=0:1)
+
+labelPanel <- function(p, text, x, y) {
+  g <- tableGrob(data.frame(a=text), rows=NULL, cols=NULL, gpar.corefill=gpar(fill="white", col=NA))
+  p + annotation_custom(grob=g, xmin=x, ymin=y, xmax=x, ymax=y)
+}
+
+gray <- gray.colors(2, start=0.2, end=0.8)
+p1 <- ggplot(r1) + geom_raster(aes(x, y, fill=value)) + scale_fill_grey() + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p2 <- ggplot(r2) + geom_raster(aes(x, y, fill=value)) + scale_fill_grey() + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p3 <- ggplot(subset(x, scale==125)) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p4 <- ggplot(subset(x, scale==1000)) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+
+p1 <- labelPanel(p1, "a)", min(r1$x), max(r1$y))
+p2 <- labelPanel(p2, "b)", min(r1$x), max(r1$y))
+p3 <- labelPanel(p3, "c)", min(r1$x), max(r1$y))
+p4 <- labelPanel(p4, "d)", min(r1$x), max(r1$y))
+
+p <- arrangeGrob(p1, p2, p3, p4, ncol=2)
+saveFigure(p, "figure-b-1.png", width=17, height=17, dpi=200)
+
 ```

### 2015-05-02 R/static_global_functions.R

@@ -91,7 +91,8 @@ theme_raster <- function(base_size=12, base_family="", ...) {
     axis.ticks=element_blank(),
     strip.background=element_blank(),
     #plot.margin=unit(c(0,0,-1,-1), "lines"),
-    plot.margin=unit(c(0,0,-.5,-.5), "lines"),
+    #plot.margin=unit(c(0,0,-.5,-.5), "lines"),
+    plot.margin=unit(c(0,0,0,0), "lines"),
     axis.ticks.length=unit(0,"lines"),axis.ticks.margin=unit(0,"lines"),
     #plot.margin=rep(unit(0,"null"),4),panel.margin=unit(0,"null"),axis.ticks.length=unit(0,"null"),axis.ticks.margin=unit(0,"null"),
     legend.position="none",
@@ -371,7 +372,9 @@ gaussKernel <- function(size, scale) {
 }
 
 smoothSubset <- function(r, x, y, kernelFun=expKernel, scale) {
-  if (is.na(r[y,x]))
+  col <- colFromX(r, x)
+  row <- rowFromY(r, y)  
+  if (is.na(r[row, col]))
     stop("The point is outside the effective area. The smoothing cannot be proceeded.")  
   
   # Construct the full kernel
@@ -380,20 +383,20 @@ smoothSubset <- function(r, x, y, kernelFun=expKernel, scale) {
   k <- kernelFun(kernelSize, resScale)
   fullArea <- prod(dim(k))
   kernelSize <- kernelSize + 1
-  
+    
   # Cut the kernel if partially outside the effective area
-  row <- max(0, y-kernelSize) + 1
-  col <- max(0, x-kernelSize) + 1
-  nrows <- min(dim(r)[1]+1, y+kernelSize) - row
-  ncols <- min(dim(r)[2]+1, x+kernelSize) - col
-  xmin <- row - (y-kernelSize) - 1
-  ymin <- col - (x-kernelSize) - 1
+  startRow <- max(0, row-kernelSize) + 1
+  startCol <- max(0, col-kernelSize) + 1
+  nrows <- min(dim(r)[1]+1, row+kernelSize) - startRow
+  ncols <- min(dim(r)[2]+1, col+kernelSize) - startCol
+  xmin <- startRow - (row-kernelSize) - 1
+  ymin <- startCol - (col-kernelSize) - 1
   k <- k[1:nrows+xmin, 1:ncols+ymin]
[…]

### 2015-05-06 inst/wtc-finland/manuscript2.Rmd
...

--- 2015-04-30 inst/wtc-finland/manuscript2.Rmd
@@
 #library(devtools); install_github("statguy/Winter-Track-Counts")
+#install_github("ropengov/gisfin")
 library(parallel)
@@
 #study$studyArea$loadBoundary(thin=F, tolerance=0.001)
+gray <- gray.colors(2, start=0.2, end=0.8)
+labelPanel <- function(p, text, x, y) {
+  g <- tableGrob(data.frame(a=text), rows=NULL, cols=NULL, gpar.corefill=gpar(fill="white", col=NA))
+  p + annotation_custom(grob=g, xmin=x, ymin=y, xmax=x, ymax=y)
+}
 ```
@@
 ```{r figure-convolution-example}
-r <- crop(study$studyArea$habitat, extent(3400000, 3410000, 7000000, 7010000))
-plot(r)
-forest <- calc(r, function(x) if (is.na(x) || x %in% c(0,45:255)) NA else x %in% 18:35)
-plot(forest)
-coords <- expand.grid(1:dim(r)[1], 1:dim(r)[2])
+# Habitat
+
+r <- raster::crop(study$studyArea$habitat, extent(3400000, 3410000, 7000000, 7010000))
+#plot(r)
+forest <- raster::calc(r, function(x) if (is.na(x) || x %in% c(0,45:255)) NA else x %in% 18:35)
+#plot(forest)
+coords <- raster::xyFromCell(r, cell=1:ncell(r))
+#x0 <- smoothSubsets(r=forest, coords=coords[1:(400*10),], kernelFun=expKernel, scales=125, .parallel=F)
 x <- smoothSubsets(r=forest, coords=coords, kernelFun=expKernel, scales=c(125, 1000), .parallel=T)
-plot(flip(rasterFromXYZ(x[,c("x","y","value")]), direction="y"), asp=1)
+#plot(flip(rasterFromXYZ(x[,c("x","y","value")]), direction="y"), asp=1)
+fileName <- study$context$getFileName(dir=study$context$processedDataDirectory, name="figure-b-1", region=study$studyArea$region)
+save(x, file=fileName)
+load(fileName)
+
+r1 <- as.data.frame(rasterToPoints(r))
+colnames(r1) <- c("x","y","value")
+r1$value <- factor(r1$value, levels=0:255)
+r2 <- as.data.frame(rasterToPoints(forest))
+colnames(r2) <- c("x","y","value")
+r2$value <- factor(r2$value, levels=0:1)
+
+p1 <- ggplot(r1) + geom_raster(aes(x, y, fill=value)) + scale_fill_grey() + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p2 <- ggplot(r2) + geom_raster(aes(x, y, fill=value)) + scale_fill_grey() + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p3 <- ggplot(subset(x, scale==125)) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p4 <- ggplot(subset(x, scale==1000)) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+
+p1 <- labelPanel(p1, "a)", min(r1$x), max(r1$y))
+p2 <- labelPanel(p2, "b)", min(r1$x), max(r1$y))
+p3 <- labelPanel(p3, "c)", min(r1$x), max(r1$y))
+p4 <- labelPanel(p4, "d)", min(r1$x), max(r1$y))
+
+p <- arrangeGrob(p1, p2, p3, p4, ncol=2)
+saveFigure(p, "figure-b-1a.png", width=17, height=17, dpi=200)
+
+# Human population density
+
+library(gisfin)
+request <- gisfin::GeoStatFiWFSRequest$new()$getPopulation("vaestoruutu:vaki2005_1km")
+client <- gisfin::GeoStatFiWFSClient$new(request)
+population <- client$getLayer("vaestoruutu:vaki2005_1km")
+population <- raster::stack(sp::SpatialPixelsDataFrame(coordinates(population), population@data, proj4string=population@proj4string))
+population <- projectRaster(population$vaesto, crs="+init=epsg:2393")
+r <- raster::crop(population$vaesto, extent(3400000, 3410000, 7000000, 7010000) * 10)
+r[is.na(r)] <- 0
+r1 <- as.data.frame(rasterToPoints(r))
+colnames(r1) <- c("x","y","value")
+coords <- raster::xyFromCell(r, cell=1:ncell(r))
+x <- smoothSubsets(r=r, coords=coords, kernelFun=expKernel, scales=c(2000), .parallel=T)
+
+p5 <- ggplot(r1) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p6 <- ggplot(x) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+
+# Elevation
+
+# Elevation model data: http://www.eea.europa.eu/data-and-maps/data/digital-elevation-model-of-europe
+elevation <- raster(file.path(context$rawDataDirectory, "elevation1x1_new.tif"))
+r <- raster::crop(elevation, extent(elevation, 200, 1500, 2600, 3500))
+raster::projection(r) <- "+proj=laea +lat_0=52 +lon_0=20 +x_0=5071000 +y_0=3210000 +a=6378137 +b=6378137 +units=m +no_defs"
+elevation <- raster::projectRaster(r, crs="+init=epsg:2393")
+r <- raster::crop(elevation, extent(3400000, 3410000, 7000000, 7010000) * 10)
+s <- raster(extent(3355003, 3455003, 6955246, 7055246), 100, 100, crs="+proj=laea +lat_0=52 +lon_0=20 +x_0=5071000 +y_0=3210000 +a=6378137 +b=6378137 +units=m +no_defs")
+r <- raster::resample(r, s) # Changing projection does not work perfectly. Has to resample to fix.
+r1 <- as.data.frame(rasterToPoints(r))
+colnames(r1) <- c("x","y","value")
+coords <- raster::xyFromCell(r, cell=1:ncell(r))
+x <- smoothSubsets(r=r, coords=coords, kernelFun=expKernel, scales=c(2000), .parallel=T)
+
+p7 <- ggplot(r1) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p8 <- ggplot(x) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+
+p5 <- labelPanel(p5, "e)", min(r1$x), max(r1$y))
+p6 <- labelPanel(p6, "f)", min(r1$x), max(r1$y))
+p7 <- labelPanel(p7, "g)", min(r1$x), max(r1$y))
+p8 <- labelPanel(p8, "h)", min(r1$x), max(r1$y))
+p <- arrangeGrob(p5, p6, p7, p8, ncol=2)
+saveFigure(p, "figure-b-1b.png", width=17, height=17, dpi=200)
 ```

### 2015-05-09 R/HabitatSelection.R
...

@@ -2,8 +2,6 @@ HabitatSelection <- setRefClass(
   Class = "HabitatSelection",
   fields = list(
     study = "Study",
-    iteration = "integer",
-    tracks = "Tracks",
     nullModelUsage = "data.frame",
     realizedUsage = "data.frame",
     relativeUsage = "data.frame",
@@ -73,8 +71,8 @@ HabitatSelection <- setRefClass(
       return(as.data.frame(x))
     },
 
-    getMovements = function(tracks) {
-      tracks$sample(nCollaredIndividuals)
+    DEPRECATED_getMovements = function(tracks) {
+      #tracks$sample(nCollaredIndividuals)
       intervals <- tracks$getSampleIntervals()
       maxIntervalH <- as.numeric(names(which.max(table(intervals$intervals$intervalH))))
       maxIntervals <- subset(intervals$intervals, intervalH == maxIntervalH)
@@ -91,7 +89,7 @@ HabitatSelection <- setRefClass(
       return(tracksDF)
     },
     
-    getHabitatPreferences = function(tracks, habitatWeightsTemplate, nSamples=10, save=FALSE) {
+    DEPRECATED_getHabitatPreferences = function(tracks, habitatWeightsTemplate, nSamples=10, save=FALSE) {
       message("Finding tracks with highest number of samples and constant frequency...")
       tracksDF <- getMovements(tracks)
       
@@ -107,11 +105,25 @@ HabitatSelection <- setRefClass(
       return(invisible(.self))
     },
 
+    getHabitatPreferences = function(intervals, habitatWeightsTemplate, nSamples=10, save=FALSE) {
+      if (!inherits(intervals, "MovementSampleIntervals"))
+        stop("Argument 'intervals' must be of class 'MovementSampleIntervals'.")
+      tracksDF <- intervals$getSampleIntervals()
+            
+      message("Estimating potential habitat usage...")
+      nullModelUsage <<- getNullModelMovementHabitatDistributions(movements=tracksDF, habitatWeightsTemplate=habitatWeightsTemplate, nSamples=nSamples)
+      message("Counting actual habitat usage...")
+      realizedUsage <<- getRealizedMovementHabitatDistributions(movements=tracksDF, habitatWeightsTemplate=habitatWeightsTemplate)
+      relativeUsage <<- as.data.frame(as.list(colSums(realizedUsage) / colSums(nullModelUsage)))
+      #relativeUsage95 <<- 
+      
[…]

### 2015-05-09 R/StudyArea.R

@@ -128,7 +128,7 @@ StudyArea <- setRefClass(
     thinGADM = function(gadm, tolerance) {
       library(maptools)
       p <- findLargestPolygon(gadm)
-      return(thinnedSpatialPoly(p, tolerance=tolerance))
+      return(thinnedSpatialPoly(p, tolerance=tolerance, avoidGEOS=T))
     },    
     
     loadBoundaryGADM = function(country, level=0, subregion, mainland=FALSE, thin=FALSE, coordinateScale=as.integer(1), tolerance=0.1) {  

### 2015-05-09 R/TracksCollection.R

@@ -1,42 +1,0 @@
-TracksCollection <- setRefClass(
-  Class = "TracksCollection",
-  fields = list(
-    study = "Study",
-    tracksList = "list"
-  ),
-  methods = list(
-    initialize = function(...) {
-      callSuper(...)
-      return(invisible(.self))
-    },
-
-    addTracks = function(tracks) {
-      tracksList[[length(tracksList) + 1]] <<- tracks
-      return(invisible(.self))
-    },
-    
-    getTracks = function(index) {
-      if (length(tracksList) < index) stop("Invalid index = ", index, ".")
-      return(tracksList[[index]])
-    },
-    
-    getTracksFileName = function() {
-      return(context$getFileName(context$resultDataDirectory, name="TracksCollection", response=study$response, region=study$studyArea$region))
-    },
-    
-    loadTracks = function(fileName=getTracksFileName()) {
-      load(file=fileName, envir=as.environment(.self))
-    },
-    
-    saveTracks = function(fileName=getTracksFileName()) {
-      save(tracksList, file=fileName)
-    }
-  )
-)
-
-SimulatedTracksCollection <- setRefClass(
-  Class = "SimulatedTracksCollection",
-  contains = "TracksCollection",
-  methods = list(
-  )
-)

### 2015-05-17 inst/wtc-finland/manuscript2.Rmd
...

@@
 s <- raster(extent(3355003, 3455003, 6955246, 7055246), 100, 100, crs="+proj=laea +lat_0=52 +lon_0=20 +x_0=5071000 +y_0=3210000 +a=6378137 +b=6378137 +units=m +no_defs")
-r <- raster::resample(r, s) # Changing projection does not work perfectly. Has to resample to fix.
+r <- raster::resample(r, s) # Changing projection does not work good enough. Have to resample to fix.
 r1 <- as.data.frame(rasterToPoints(r))

### 2015-05-17 R/MovementSampleIntervals.R

@@ -12,24 +12,6 @@ MovementSampleIntervals <- setRefClass(
     },
     
     getSampleIntervals = function() intervals
-    
-    #getSampleIntervalsFileName = function() {
-    #  if (inherits(study, "uninitializedField"))
-    #    stop("Parameter 'study' must be provided.")
-    #  if (length(study$response) == 0)
-    #    stop("Parameter 'response' for 'study' must be defined.")
-    #  return(context$getFileName(dir=study$context$processedDataDirectory, name="MovementSampleIntervals", response=study$response, region=study$studyArea$region))
-    #},
-    
-    #saveSampleIntervals = function(fileName=getSampleIntervalsFileName()) {
-    #  save(intervals, file=fileName)
-    #  return(invisible(.self))
-    #},
-    
-    #loadSampleIntervals = function(fileName=getSampleIntervalsFileName()) {
-    #  load(fileName, env=as.environment(.self))
-    #  return(invisible(.self))
-    #}
   )
 )
 

### 2015-05-17 R/SpatioTemporalRaster.R

@@ -30,7 +30,7 @@ SpatioTemporalRaster <- setRefClass(
     },
     
     getRasterFileName = function(name, layerName, ext=.self$ext) {
-      return(context$getFileName(dir=study$context$figuresDirectory, name=paste(name, layerName, sep="-"), response=study$response, region=study$studyArea$region, ext=paste(".", ext, sep="")))
+      return(study$context$getFileName(dir=study$context$figuresDirectory, name=paste(name, layerName, sep="-"), response=study$response, region=study$studyArea$region, ext=paste(".", ext, sep="")))
     },
     
     saveRasterFile = function(p, layer, name, layerName, ext=.self$ext, ...) {
@@ -165,7 +165,7 @@ SpatioTemporalRaster <- setRefClass(
       }
       
       message("Converting...")
-      outputFile <- context$getFileName(dir=study$context$figuresDirectory, name=name, response=study$response, region=study$studyArea$region, ext=".gif")
+      outputFile <- study$context$getFileName(dir=study$context$figuresDirectory, name=name, response=study$response, region=study$studyArea$region, ext=".gif")
       layerFileNameMask <- getRasterFileName(name=name, layerName="*", ext="png")
       cmd <- paste(convertCmd, " -loop 0 -delay ", delay, " ", layerFileNameMask, " ", outputFile, sep="")
       message(cmd)

### 2015-05-17 inst/wtc-finland/preprocess.R

@@ -9,8 +9,188 @@ registerDoMC(cores=round(detectCores()))
 source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
 library(WTC)
 
-context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
-study <- FinlandWTCStudy$new(context=context)
+
+preprocessIntersections <- function(response) {
+  context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
+  study <- FinlandWTCStudy$new(context=context)
+  study$response <- response
+
+  intersections <- study$loadIntersections(predictDistances=FALSE)
+  intersections$saveIntersections()
+}
+
+preprocessHabitatPreferences <- function(response) {
+  context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
+  study <- FinlandWTCStudy$new(context=context)
+  study$response <- response
+
+  tracks <- study$loadTracks()
+  habitatWeights <- CORINEHabitatWeights$new(study=study)
+  habitatSelection <- tracks$getHabitatPreferences(habitatWeightsTemplate=habitatWeights, nSamples=30, save=T)
+}
+
+preprocessSampleIntervals <- function(response) {
+  context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
+  study <- FinlandWTCStudy$new(context=context)
+  study$response <- response
+
+  human <- HumanPopulationDensityCovariates$new(study=study)
+  weather <- WeatherCovariates$new(study=study, apiKey=fmiApiKey)
+  sampleIntervals$setCovariatesId()$addCovariates(human, weather)
+  sampleIntervals$saveCovariates()
+  
+  human <- HumanPopulationDensityCovariates$new(study=study)
+  weather <- WeatherCovariates$new(study=study, apiKey=fmiApiKey)
[…]

### 2015-05-29 inst/simulation/manuscript.Rmd
...

@@
   validation <- Validation(study=study, populationSizeCutoff=cutoff)
+  credibilityIntervals <- list()
+
+  if ("FMPModel" %in% modelNames) {
+    modelNames <- modelNames[!modelNames %in% "FMPModel"]
+
+    for (scenario in scenarios) {
+      study <- getStudy(scenario=scenario)
+      validation <- Validation(study=study, populationSizeCutoff=cutoff)
+      iterations <- validation$getCredibilityIntervalsValidationIterations("FMPModel")
+      p95 <- c()
+      p50 <- c()
+      for (i in iterations) {
+        load(file=validation$getCredibilityIntervalsValidationFileName(modelName=modelName, iteration=i))
+        p95 <- c(p95, bootCI$p95)
+        p50 <- c(p50, bootCI$p50)
+      }
+      a <- data.frame(scenario=scenario, Proportion=mean(p95, na.rm=T), n=NA, True=NA, Estimated=NA, Estimated.q1=NA, Estimated.q2=NA, modelName="FMPModel", Interval="95%")
+      b <- data.frame(scenario=scenario, Proportion=mean(p50, na.rm=T), n=NA, True=NA, Estimated=NA, Estimated.q1=NA, Estimated.q2=NA, modelName="FMPModel", Interval="50%")
+      credibilityIntervals <- rbind(credibilityIntervals, rbind(a, b))
+    }
+  }
+
   a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames)
   b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames, probs=c(.25, .75), probsName="50%")
-  credibilityIntervals <- rbind(a, b)
+  #credibilityIntervals <- rbind(a, b)
+  credibilityIntervals <- rbind(credibilityIntervals, rbind(a, b))
   credibilityIntervals$panel <- panel

### 2015-05-29 inst/simulation/validate_boot_ci.R

@@ -13,8 +13,8 @@ modelName <- extraArgs[1]
 iteration <- as.integer(task_id)
 if (F) {
 modelName <- "FMPModel"
-scenario <- "A"
-iteration <- as.integer(1)
+scenario <- "E"
+iteration <- as.integer(10)
 }
 mss <- getMSS(scenario=scenario)
 study <- mss$study
@@ -41,6 +41,7 @@ getPopulationSize <- function(fitted, index, model, readHabitatIntoMemory) {
 findCI <- function(fitted, iteration, model, modelName) {
   populationSize <- study$loadPopulationSize(iteration, modelName)
   
+  # x <- subset(fitted, year=2001)
   b <- ddply(fitted, .(year), function(x, iteration, model, populationSize) {
     message("Resampling year ", x$year[1], "...")
     year <- as.integer(x$year[1])
@@ -57,7 +58,7 @@ findCI <- function(fitted, iteration, model, modelName) {
       return(data.frame(year=year, p95=p95, p50=p50))
     }
     else return(data.frame(year=year, p95=NA, p50=NA))
-  }, iteration=iteration, model=model, populationSize)
+  }, iteration=iteration, model=model, populationSize=populationSize)
   return(b)
 }
 
@@ -70,7 +71,14 @@ bootCI <- do.call(rbind, bootCI)
 save(bootCI, file=validation$getCredibilityIntervalsValidationFileName(modelName=modelName, iteration=iteration))
 
 if (F) {
-load(file=validation$getCredibilityIntervalsValidationFileName(modelName=modelName, iteration=iteration))
-mean(bootCI$p95)
-mean(bootCI$p50)
+  iterations <- validation$getCredibilityIntervalsValidationIterations(modelName)
+  p95 <- c()
+  p50 <- c()
+  for (i in iterations) {
+    load(file=validation$getCredibilityIntervalsValidationFileName(modelName=modelName, iteration=i))
+    p95 <- c(p95, bootCI$p95)
+    p50 <- c(p50, bootCI$p50)
+  }
+  mean(p95, na.rm=T)
+  mean(p50, na.rm=T)
 }

### 2015-06-15 inst/simulation/manuscript.Rmd
...

--- 2015-04-29 inst/simulation/manuscript.Rmd
@@
 boundaryDF <- ggplot2::fortify(FinlandWTCStudy(context=context, response="canis.lupus")$studyArea$boundary)
-
+cutoff <- c(1,1000)
 ```
@@
 #populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4,5)])
-
-cutoff <- c(1,1000)
 
@@
   validation <- Validation(study=study, populationSizeCutoff=cutoff)
-  a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames)
-  b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames, probs=c(.25, .75), probsName="50%")
-  credibilityIntervals <- rbind(a, b)
+  credibilityIntervals <- list()
+
+  if ("FMPModel" %in% modelNames) {
+    for (scenario in scenarios) {
+      study <- getStudy(scenario=scenario)
+      validation <- Validation(study=study, populationSizeCutoff=cutoff)
+      iterations <- validation$getCredibilityIntervalsValidationIterations("FMPModel")
+      p95 <- c()
+      p50 <- c()
+      for (i in iterations) {
+        load(file=validation$getCredibilityIntervalsValidationFileName(modelName="FMPModel", iteration=i))
+        p95 <- c(p95, bootCI$p95)
+        p50 <- c(p50, bootCI$p50)
+      }
+      a <- data.frame(scenario=scenario, Proportion=mean(p95, na.rm=T), n=NA, True=NA, Estimated=NA, Estimated.q1=NA, Estimated.q2=NA, modelName="FMPModel", Interval="95%")
+      b <- data.frame(scenario=scenario, Proportion=mean(p50, na.rm=T), n=NA, True=NA, Estimated=NA, Estimated.q1=NA, Estimated.q2=NA, modelName="FMPModel", Interval="50%")
+      credibilityIntervals <- rbind(credibilityIntervals, rbind(a, b))
+    }
+  }
+
+  a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames[!modelNames %in% "FMPModel"])
+  b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames[!modelNames %in% "FMPModel"], probs=c(.25, .75), probsName="50%")
+  credibilityIntervals <- rbind(credibilityIntervals, rbind(a, b))
   credibilityIntervals$panel <- panel
-  credibilityIntervals$scenario <- rep(scenarioNames, times=2, each=length(modelNames))
+  #credibilityIntervals$scenario <- rep(scenarioNames, times=2, each=length(modelNames))
   return(credibilityIntervals)
-}
-
-cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,5)], panel="Case 1", scenarioNames=scenarios, cutoff=cutoff)
+
+  #scenario=rep(scenarioNames, times=2, each=length(modelNames[!modelNames %in% "FMPModel"])))
+}
+
+cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,4,5)], panel="Case 1", scenarioNames=scenarios, cutoff=cutoff)
 #cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,5)], panel="Case 2", scenarioNames=scenarios, cutoff=10*cutoff)
@@
     geom_point(aes(shape=Model, group=Model), position=position_dodge(width=.7)) +
-    scale_shape_manual("Model", values=c(17,15,3)) +
-    scale_colour_manual("CI", values=c("gray", "black")) +
-    #scale_colour_manual("Model", values=c("blue", "green", "red"), labels=modelLabels) +
+    scale_colour_manual("CI", values=c("black", "grey")) +
     theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
@@
           plot.margin=unit(c(0.2, 0.2, 0, 0), units="line"),
-          #legend.background=element_rect(fill="transparent"),
-          legend.margin=unit(-2, units="line"),
           panel.margin=unit(0, units="line"),
-          legend.position="bottom",
+          legend.margin=unit(-1, units="line"),
+          #legend.position="bottom",
+          legend.direction="horizontal",
+          #legend.direction="vertical",
           legend.box="horizontal",
-          legend.key.width=unit(.7, "line"))
+          legend.key.width=unit(.7, "line"),
+          legend.key.height=unit(0, "line"))
+
+  library(gridExtra)
+  g_legend <- function(a.gplot) {
+    tmp <- ggplot_gtable(ggplot_build(a.gplot))
+    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+    legend <- tmp$grobs[[leg]]
+    legend
+  }
+
+  legend <- g_legend(p)
+  p <- arrangeGrob(p + theme(legend.position="none"), legend, heights=c(9/10,1/10))
   return(p)
@@
 saveFigure(p, "manuscript-figure-5.pdf", height=7.87/2, dpi=300)
-```
+
+```

### 2015-06-29 inst/simulation/manuscript.Rmd
...

@@
 
+#studyArea <- FinlandStudyArea$new(context=context)$setup(tolerance=0.001)
 boundaryDF <- ggplot2::fortify(FinlandWTCStudy(context=context, response="canis.lupus")$studyArea$boundary)
@@
   tracks <- study$loadTracks(iteration=iterations[i])
-  intersections <- tracks$countIntersections(surveyRoutes=mss$surveyRoutes, save=F)
+  intersections <- tracks$countIntersections(surveyRoutes=study$surveyRoutes, save=F)
 
@@
 
-  #trackData$type <- "track"
-  #surveyRoutesData$type <- "transect"
-  #x <- rbind(trackData[,c("long","lat","order","piece","group","observation","type")], surveyRoutesData[,c("long","lat","order","piece","group","observation","type")])
-  #x$colour <- factor(paste(x$type, x$observation))
-  #ggplot(x, aes(long, lat, group=group, colour=colour)) + geom_path()
-
     i <- i + 1
@@
 habitatData <- as.data.frame(rasterToPoints(z))
-#habitatData <- as.data.frame(rasterToPoints(habitatRaster))
 colnames(habitatData) <- c("x","y","z")
@@
 
+habitatData$x <- habitatData$x / 1000
+habitatData$y <- habitatData$y / 1000
+trackData$long <- trackData$long / 1000
+trackData$lat <- trackData$lat / 1000
+surveyRoutesData$long <- surveyRoutesData$long / 1000
+surveyRoutesData$lat <- surveyRoutesData$lat / 1000
+window <- window / 1000
+
 p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z), alpha=.4) + scale_fill_manual(values=col) +
@@
   facet_wrap(~scenario) + #coord_equal()
-  theme_raster(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
-#print(p)
-#saveFigure(p, "manuscript-figure-1.png", width=8, height=5)
+  theme_ms() + theme(panel.border=element_rect(fill="transparent")) +
+  xlab("X-coordinate (km)") + ylab("Y-coordinate (km)") +
+  theme(legend.position="none", panel.margin=unit(0, "lines"), panel.grid.minor=element_blank(), panel.grid.major=element_blank(), plot.margin=unit(c(0,0.1,0,0), "lines")) # top, right, bottom, and left
 saveFigure(p, "manuscript-figure-1.png", width=17, height=15, dpi=200) # draft quality
+saveFigure(p, "manuscript-figure-1.pdf", width=17, height=15, dpi=300)
 
@@
 c((window[2]-window[1]), (window[4]-window[3]))/1000
-```
-
-```{r figure-1-separate, eval=FALSE}
-getRaster <- function(habitatRaster, xlim, ylim, size=10000) {
-  x <- crop(habitatRaster, extent(c(xlim, ylim)))
-  y <- sampleRegular(x, size=size, asRaster=TRUE)
-  habitatData <- as.data.frame(rasterToPoints(y))
-  colnames(habitatData) <- c("x","y","z")
-  habitatData$z <- factor(habitatData$z, levels=0:255)
-  col <- habitatRaster@legend@colortable
-  names(col) <- 0:255
-  return(habitatData)
-}
-
-getPanel <- function(scenario, xlim=c(3.375,3.4)*1e6, ylim=c(6.9,6.925)*1e6, habitatRaster, trackData, surveyRoutesData) {
-  s <- scenario
-  p <- if (!missing(habitatRaster)) {
-    habitatData <- getRaster(habitatRaster=habitatRaster, xlim=xlim, ylim=ylim)
-    ggplot(habitatData) + geom_raster(aes(x, y, fill=z)) + scale_fill_manual(values=col) +
-      geom_path(data=subset(trackData, scenario==s), aes(long, lat, group=group, colour=observation))
-  }
-  else {
-     ggplot(subset(trackData, scenario==s)) +
-       geom_path(aes(long, lat, group=group, colour=observation))
-  }
-
-  p <- p + scale_colour_manual(values=c("gray40","black")) +
-    geom_path(data=subset(surveyRoutesData, scenario==s), aes(long, lat, group=group)) +
-    coord_fixed(ratio=1, xlim=xlim, ylim=ylim) +
-    theme_raster(base_size=10, panel.border=element_rect(fill="transparent"), plot.title=element_text(size=12)) +
-    ggtitle(s)
-
-  return(p)
-}
-
-library(gridExtra)
-
-p1 <- getPanel(scenario="A", trackData=trackData, surveyRoutesData=surveyRoutesData)
-p2 <- getPanel(scenario="B", trackData=trackData, surveyRoutesData=surveyRoutesData)
-p3 <- getPanel(scenario="C", trackData=trackData, surveyRoutesData=surveyRoutesData)
-p4 <- getPanel(scenario="D", trackData=trackData, surveyRoutesData=surveyRoutesData)
-p5 <- getPanel(scenario="E", habitatRaster=study$studyArea$habitat, trackData=trackData, surveyRoutesData=surveyRoutesData)
-p6 <- getPanel(scenario="F", habitatRaster=study$studyArea$habitat, trackData=trackData, surveyRoutesData=surveyRoutesData)
-p <- arrangeGrob(p1, p2, p3, p4, p5, p6, ncol=3, nrow=2)
-saveFigure(p, "manuscript-figure-1.png")
 ```
@@
   ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, years) {
-
-    #study <- getStudy(scenario=x$scenario)
-    #intersections <- study$loadIntersections(iteration=x$iteration)
-    #observationTracks <- SimulatedTracks$new(study=study, iteration=x$iteration)
-    #observationTracks$loadTracks(fileName=observationTracks$getObservationTracksFileName(), addColumns=TRUE)
-    # TODO: use observation tracks ^^^^
-
     study <- getStudy(scenario=x$scenario)
@@
     observationTracks <- tracks
-
-    #d_ply(observationTracks$tracks, .(id, year, burst), function(x) {
-    #  with(x[1,], message(id, " ", year, " ", burst))
-    #})
 
@@
     estimates$data <- subset(estimates$data, year == x$year)
-    #studyArea <- FinlandStudyArea$new(context=context)$setup(tolerance=0.001)
-    #density <- estimates$getPopulationDensityInterpolate(getSD=sd, maskPolygon=studyArea$boundary)
 
-    #habitatWeights <- study$getHabitatWeights(iteration=x$iteration)
-    #density <- estimates$getPopulationDensity(habitatWeights=habitatWeights, .parallel=T)
     density <- estimates$getPopulationDensity(.parallel=T)
@@
 
-    #density <- if (sd) density$sd else density$mean
     if (is.null(density) || maxValue(density$rasterStack[[1]]) > 1) stop("Estimation failed.")
@@
 }
-#iterations <- as.integer(c(1,4,1,2,5,31))
-#densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, sd=F, log=F)
 
@@
 
+wdensitiesScaled$x <- wdensitiesScaled$x / 1000 - 3000
+wdensitiesScaled$y <- wdensitiesScaled$y / 1000 - 6000
+densitiesScaled$x <- densitiesScaled$x / 1000 - 3000
+densitiesScaled$y <- densitiesScaled$y / 1000 - 6000
+boundaryDF$lat <- boundaryDF$lat / 1000 - 6000
+boundaryDF$long <- boundaryDF$long / 1000 - 3000
+agents$x <- agents$x / 1000 - 3000
+agents$y <- agents$y / 1000 - 6000
+limits@xmin <- limits@xmin / 1000 - 3000
+limits@ymin <- limits@ymin / 1000 - 6000
+limits@xmax <- limits@xmax / 1000 - 3000
+limits@ymax <- limits@ymax / 1000 - 6000
+surveyRoutes$x <- surveyRoutes$x / 1000 - 3000
+surveyRoutes$y <- surveyRoutes$y / 1000 - 6000
+gridTemplate@extent@xmin <- round(gridTemplate@extent@xmin / 1000) - 3000
+gridTemplate@extent@ymin <- round(gridTemplate@extent@ymin / 1000) - 6000
+gridTemplate@extent@xmax <- round(gridTemplate@extent@xmax / 1000) - 3000
+gridTemplate@extent@ymax <- round(gridTemplate@extent@ymax / 1000) - 6000
+
 p1 <- ggplot(wdensitiesScaled) +
@@
                        guide=guide_legend(title=expression(paste("Population density ", (individuals/km^2)))), trans="log10",
-                       #breaks=c(10^-4.5, 10^-4, 10^-3.5, 10^-3, 10^-2.5),
                        breaks=trans_breaks("log10", function(x) 10^x),
-                       #labels=c(expression(10^-4.5), expression(10^-4.0), expression(10^-3.5), expression(10^-3.0), expression(10^-2.5)), limits=c(10^-4.5, 10^-2.5)
                        labels=trans_format("log10", math_format(10^.x))
                        ) +
-  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
+  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="black", size=0.5) +
   geom_point(data=agents, aes(x, y, colour=Observed), size=1.3, colour="white") +
   geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black"), guide=FALSE) +
-  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0.01,0.01)) +
+  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0.01,0.01), breaks=pretty_breaks(n=3)) +
   scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0.01,0.01)) +
-  facet_grid(~scenario) + coord_equal() + theme_raster(base_size=10) +
-  theme(legend.position="bottom", legend.margin=unit(-2, units="line"), legend.title=element_text(size=10), legend.text=element_text(size=10),
-        plot.margin=unit(c(-3, 0, 0, -.5), units="line")) #top, right, bottom, and left margins
-g1 <- arrangeGrob(p1, main=textGrob("a)", x=0.01, just="left", gp=gpar(fontsize=10)))
-saveFigure(g1, "manuscript-figure-2a.png", width=17, height=8.8, dpi=200) # draft quality
-saveFigure(g1, "manuscript-figure-2a.pdf", width=17, height=8.8, dpi=300)
+  facet_grid(~scenario) + coord_equal() + theme_ms() +
+  theme(legend.position="bottom", legend.margin=unit(-1, units="line"), legend.title=element_text(size=10), legend.text=element_text(size=10)) + #top, right, bottom, and left margins
+  theme(strip.background=element_blank(), panel.background=element_rect(fill="transparent", colour=NA), panel.margin=unit(0, "lines"), panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
+  xlab("X-coordinate (km)") + ylab("Y-coordinate (km)")
+g1 <- arrangeGrob(p1, main=textGrob("a)", x=0.01, y=0.01, just="left", gp=gpar(fontsize=10)))
+saveFigure(g1, "manuscript-figure-2a.png", width=17, height=8.07, dpi=200) # draft quality
+saveFigure(g1, "manuscript-figure-2a.pdf", width=17, height=8.07, dpi=300)
 
-p2 <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey") +
+xlabels <- seq(gridTemplate@extent@xmin, gridTemplate@extent@xmax, 1e5/1000)
+xlabels[c(2,4,6,7)] <- ""
+p2 <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey", size=0.5) +
   geom_point(data=surveyRoutes, aes(x, y, colour=intersections), size=1) + scale_color_manual(values=c("darkgrey", "black")) +
-  scale_x_continuous(breaks=seq(gridTemplate@extent@xmin, gridTemplate@extent@xmax, 1e5), expand=c(0.01,0.01)) +
-  scale_y_continuous(breaks=seq(gridTemplate@extent@ymin, gridTemplate@extent@ymax, 1e5), expand=c(0.01,0.01)) +
-  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal() +
-  theme(panel.grid.major=element_line(colour="grey"),
-        plot.margin=unit(c(0, 0, 0, -.5), units="line")) #top, right, bottom, and left margins
-g2 <- arrangeGrob(p2, main=textGrob("b)", x=0.01, just="left", gp=gpar(fontsize=10)))
-saveFigure(g2, "manuscript-figure-2b.png", width=17, height=6.9, dpi=200) # draft quality
-saveFigure(g2, "manuscript-figure-2b.pdf", width=17, height=6.9, dpi=300)
+  scale_x_continuous(breaks=seq(gridTemplate@extent@xmin, gridTemplate@extent@xmax, 1e5/1000), expand=c(0.01,0.01), labels=xlabels) +
+  scale_y_continuous(breaks=seq(gridTemplate@extent@ymin, gridTemplate@extent@ymax, 1e5/1000), expand=c(0.01,0.01)) +
+  facet_grid(~scenario) + coord_equal() + theme_ms() +
+  theme(panel.grid.major=element_line(colour="grey")) + #top, right, bottom, and left margins
+  theme(strip.background=element_blank(), panel.background=element_rect(fill="transparent", colour=NA), panel.margin=unit(0, "lines"), panel.grid.minor=element_blank(), legend.position="none", axis.ticks=element_line(colour="grey", size=0.2)) +
+  xlab("X-coordinate (km)") + ylab("Y-coordinate (km)")
+g2 <- arrangeGrob(p2, main=textGrob("b)", x=0.01, y=0.01, just="left", gp=gpar(fontsize=10)))
+saveFigure(g2, "manuscript-figure-2b.png", width=17, height=7.2, dpi=200) # draft quality
+saveFigure(g2, "manuscript-figure-2b.pdf", width=17, height=7.2, dpi=300)
 
-p <- arrangeGrob(g1, g2)
-saveFigure(p, "manuscript-figure-2.png", width=17, height=17, dpi=200) # draft quality
-saveFigure(p, "manuscript-figure-2.pdf", width=17, height=17, dpi=300)
-
+p <- arrangeGrob(g1, g2, heights=c(8.07, 7.2)/(8.07+7.2))
+saveFigure(p, "manuscript-figure-2.png", width=17, height=8.07+7.2, dpi=200) # draft quality
+saveFigure(p, "manuscript-figure-2.pdf", width=17, height=8.07+7.2, dpi=300)
 ```
@@
   populationSizes <- validation$getValidationTemporalPopulationSizes(scenarios=scenarios, modelNames=modelNames)
-  #validation$summarizePopulationSize(populationSizes)  
 }
@@
 populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames[c(1,3,4,5)])
-#populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,4,5)])
-#populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4,5)])
-
 meanPopulationSizes  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes, cutoff, models=modelNames[c(1,3,5)]))
@@
 
-#meanPopulationSizes2000  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
-#logPopulationSizes2000 <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
-#logMeanPopulationSizes2000  <- getMeanPopulationSizes(logPopulationSizes2000)
-
-#meanPopulationSizes10Days  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
-#logPopulationSizes10Days <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
-#logMeanPopulationSizes10Days  <- getMeanPopulationSizes(logPopulationSizes10Days)
-```
-
-```{r table-1abc}
 getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, overEstimate, models, file, ...) {
@@
 
-#x <- subset(populationSizes, modelName=="SmoothModelMean-nbinomial-ar1-priors1" & Scenario=="C")
-#y <- lm(Estimated~Observed, x)
-
 table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff, models=modelNames[c(1,3,5)])
@@
 
-#table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff, models=modelNames[c(1,3,5)])
-#logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1), models=modelNames[1])
-#logTable2$Failed <- table2$Failed
-
-#table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff, models=modelNames[c(1,3,5)])
-#logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1), models=modelNames[1])
-#logTable3$Failed <- table3$Failed
-
-subset(logTable1, Model != "VT")
-subset(logTable2, Model != "VT")
-subset(logTable3, Model != "VT")
+#subset(populationSizes, Scenario=="C" & modelName=="FMPModel")
+#subset(populationSizes, Scenario=="B" & modelName=="FMPModel")
 
@@
 
-plotPopulationSizesTable_DEPRECATED <- function(logTable) {
-  library(reshape2)
-  facet_labeller <- function(variable, value) {
-    measured <- list(EstTrue=expression(alpha), Slope=expression(beta), R2=expression(R^2), Failed="Failed (%)")
-    print(paste0(variable,"=",value))
-    if (variable=="variable") return(measured[value])
-    if (variable=="Case") return(paste("Case", value))
-  }
-
-  #x <- melt(logTable, id.vars=c("Scenario", "Model", "Case"), measure.vars=c("EstTrue", "Slope", "R2", "Failed"))
-  x <- melt(logTable, id.vars=c("Scenario", "Model"), measure.vars=c("EstTrue", "Slope", "R2", "Failed"))
-
-  bestFit <- data.frame(variable=unique(x$variable), y=c(0, 1, 1, 0))
-  errorBars <- x
-  errorBars$value <- NA
-  #errorBars$value[errorBars$variable=="Slope"] <- logTable$Slope
-  #errorBars$qmin[errorBars$variable=="Slope"] <- logTable$Slope-logTable$SlopeSD
-  #errorBars$qmax[errorBars$variable=="Slope"] <- logTable$Slope+logTable$SlopeSD
-  errorBars$value[errorBars$variable=="EstTrue"] <- logTable$EstTrue
-  errorBars$qmin[errorBars$variable=="EstTrue"] <- logTable$EstTrueMin
-  errorBars$qmax[errorBars$variable=="EstTrue"] <- logTable$EstTrueMax
-
-  p <- ggplot(x, aes(Scenario, value, group=Model)) +
-    geom_hline(data=bestFit, aes(yintercept=y), color="darkgray") + 
-    geom_errorbar(data=errorBars, aes(Scenario, group=Model, ymax=qmax, ymin=qmin), position=position_dodge(width=.7), colour="darkgray") +
-    geom_point(aes(shape=Model), position=position_dodge(width=.7)) + ylab("") +
-    #facet_grid(variable~., scales="free_y", labeller=facet_labeller) +
-    facet_wrap(~variable, scales="free_y", nrow=2) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=10),
-                                        legend.position="bottom",
-                                        strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
-                                        plot.margin=unit(c(0, 0.2, 0, -.5), units="line"),  #top, right, bottom, and left margins
-                                        legend.margin=unit(-1.5, units="line"))# , legend.key=element_rect(fill="transparent"))
-
-  # Expand y-axis 100 for Failed
-  x <- data.frame(Scenario="A", Model="FMP", variable="Failed", value=100)
-  p <- p + geom_blank(data=x, aes(x=Scenario, y=value, group=Model))
-
-  p <- facet_wrap_labeller(p, labels=c(expression(alpha), Slope=expression(beta), R2=expression(R^2), Failed="Failed (%)"))
-
-  return(p)
-}
-
 plotPopulationSizesTable <- function(logTable) {
@@
 
+  margin <- unit(c(0.1, 0.1, -.75, 0), units="line")
+
   p1 <- x %>% filter(variable=="EstTrue") %>%
@@
     xlab("") + ylab(expression(alpha)) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+    theme_ms() + theme(panel.border=element_rect(fill="transparent"),
                                         legend.title=element_text(size=10), legend.position="none",
                                         strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
-                                        plot.margin=unit(c(0, 0.2, -.75, 0), units="line")) #top, right, bottom, and left margins
-  g1 <- arrangeGrob(p1, main=textGrob("a)", x=0.01, just="left", gp=gpar(fontsize=10)))
+                                        plot.margin=margin) #top, right, bottom, and left margins
+  g1 <- arrangeGrob(p1, main=textGrob("a)", x=0.01, y=0.01, just="left", gp=gpar(fontsize=10)))
 
@@
     xlab("") + ylab(expression(beta)) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+    theme_ms() + theme(panel.border=element_rect(fill="transparent"),
                                         legend.title=element_text(size=10), legend.position="none",
                                         strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
-                                        plot.margin=unit(c(0, 0.2, -.75, 0), units="line"))
-  g2 <- arrangeGrob(p2, main=textGrob("b)", x=0.01, just="left", gp=gpar(fontsize=10)))
+                                        plot.margin=margin)
+  g2 <- arrangeGrob(p2, main=textGrob("b)", x=0.01, y=0.01, just="left", gp=gpar(fontsize=10)))
 
@@
     xlab("") + ylab(expression(R^2)) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+    theme_ms() + theme(panel.border=element_rect(fill="transparent"),
                                         legend.title=element_text(size=10), legend.position="none",
                                         strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
-                                        plot.margin=unit(c(0, 0.2, -.75, 0), units="line"))
-  g3 <- arrangeGrob(p3, main=textGrob("c)", x=0.01, just="left", gp=gpar(fontsize=10)))
+                                        plot.margin=margin)
+  g3 <- arrangeGrob(p3, main=textGrob("c)", x=0.01, y=0.01, just="left", gp=gpar(fontsize=10)))
 
@@
     xlab("") + ylab("Failed (%)") + ylim(0,100) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+    theme_ms() + theme(panel.border=element_rect(fill="transparent"),
                                         legend.title=element_text(size=10), legend.position="none",
                                         strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
-                                        plot.margin=unit(c(0, 0.2, -.75, 0), units="line"))
-  g4 <- arrangeGrob(p4, main=textGrob("d)", x=0.01, just="left", gp=gpar(fontsize=10)))
+                                        plot.margin=margin)
+  g4 <- arrangeGrob(p4, main=textGrob("d)", x=0.01, y=0.01, just="left", gp=gpar(fontsize=10)))
 
@@
   lheight <- sum(legend$height)
-  p <- arrangeGrob(g1, g2, g3, g4, ncol=2)
   theight <- unit(10, "points")
-  p <- arrangeGrob(p, textGrob("Scenario", gp=gpar(fontsize=10)), heights=unit.c(unit(1, "npc") - theight, theight))
+  #p <- arrangeGrob(g1, g2, g3, g4, ncol=2)
+
+  p1 <- arrangeGrob(g1, g3, nrow=2)
+  p2 <- arrangeGrob(g2, g4, nrow=2)
+  p1 <- arrangeGrob(p1, textGrob("Scenario", x=0.58, gp=gpar(fontsize=10)), heights=unit.c(unit(1, "npc") - theight, theight))
+  p2 <- arrangeGrob(p2, textGrob("Scenario", x=0.57, gp=gpar(fontsize=10)), heights=unit.c(unit(1, "npc") - theight, theight))
+  p <- arrangeGrob(p1, p2, ncol=2)
+
+  #p <- arrangeGrob(p, textGrob("Scenario", gp=gpar(fontsize=10)), heights=unit.c(unit(1, "npc") - theight, theight))
   p <- arrangeGrob(p, legend, heights=unit.c(unit(1, "npc") - lheight, lheight), ncol=1)
@@
 
-#logTable1$Case <- "1"
-#logTable2$Case <- "2"
-#logTable3$Case <- "3"
-#logTable <- rbind(subset(logTable1, Model != "VT"), subset(logTable2, Model != "VT"), subset(logTable3, Model != "VT"))
 logTable <- logTable1
@@
 saveFigure(p, "manuscript-figure-3.png", width=17, dpi=200) # draft quality
-saveFigure(p, "manuscript-figure-3.pdf", height=17, dpi=300)
+saveFigure(p, "manuscript-figure-3.pdf", width=17, dpi=300)
 
+logTable <- logTable1
 logTablePretty <- logTable
@@
 logTablePretty$Failed <- round(logTablePretty$Failed, 0)
-#logTablePretty <- logTablePretty[,c("Case","Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
 logTablePretty <- logTablePretty[,c("Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
 writeTable(subset(logTablePretty, Model != "VT"), "table-1.docx", context)
-
-#writeTable(subset(logTable1, Model != "VT"), "table-1a.docx", context)
-#writeTable(subset(logTable2, Model != "VT"), "table-1b.docx", context)
-#writeTable(subset(logTable3, Model != "VT"), "table-1c.docx", context)
 ```
@@
 ```{r figure-3sample}
-xlim <- log10(c(30,800)); ylim <- xlim
+#xlim <- log10(c(30,800)); ylim <- xlim
 logPopulationSizesSample <- subset(logPopulationSizes, scenario == "A")
@@
     xlab("True population size") + ylab("Estimated population size") +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
+    theme_ms() + theme(panel.border=element_rect(fill="transparent"),
                                         plot.title=element_text(size=10),
-                                        plot.margin=unit(c(0, 0.1, 0, 0), units="line"), #top, right, bottom, and left margins
+                                        #plot.margin=unit(c(0, 0.1, 0, 0), units="line"), #top, right, bottom, and left margins
                                         legend.margin=unit(0, units="line"),
@@
 p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName %in% c("ST","T1","T2","FMP")), modelName=c("ST","T1","T2","FMP"), "a", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3sample.png", height=7.87) # draft quality
-saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87, dpi=300)
+saveFigure(p, "manuscript-figure-3sample.png", height=7.87, width=17) # draft quality
+saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87, width=17, dpi=300)
 ```
@@
 
-  #if (!any(abs(size$Estimated) == Inf)) p <- p + geom_smooth(method="lm", se=F, fullrange=T, colour="black")
   ls <- ddply(size, .(Scenario), function(x) if (!any(abs(x$Estimated) == Inf)) coef(lm(Estimated~Observed, x)) else c(NA,NA))
@@
   p <- p + facet_wrap(~Scenario) +
-    #scale_x_continuous(limits=xlim, labels=trans_format("identity", math_format(10^.x))) +
-    #scale_y_continuous(limits=ylim, labels=trans_format("identity", math_format(10^.x))) +
     scale_x_continuous(limits=xlim, labels=trans_format("identity", math_format(.x))) +
     scale_y_continuous(limits=ylim, labels=trans_format("identity", math_format(.x))) +
-    xlab(expression(paste(Log[10], " true population size"))) + ylab(expression(paste(Log[10], " estimated population size"))) + ggtitle(paste("Case:", case, " Model:", prettyModelName(modelName[1]))) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
-                                        plot.title=element_text(size=10),
-                                        plot.margin=unit(c(0, 0.2, 0, 0), units="line"), #top, right, bottom, and left margins
+    xlab(expression(paste(Log[10], " true population size"))) + ylab(expression(paste(Log[10], " estimated population size"))) + ggtitle(paste("Model:", prettyModelName(modelName[1]))) +
+    theme_ms() + theme(panel.border=element_rect(fill="transparent"),
                                         legend.margin=unit(0, units="line"),
@@
 
-xlim <- log10(c(50,1200)); ylim <- xlim
+#xlim <- log10(c(1,1200)); ylim <- xlim
+xlim <- log10(c(1,2000)); ylim <- xlim
 p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], "1", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3a1.png")
-saveFigure(p, "manuscript-figure-3a1.pdf", dpi=300)
+saveFigure(p, "manuscript-figure-3a1.png", width=17, dpi=200)
+saveFigure(p, "manuscript-figure-3a1.pdf", width=17, dpi=300)
 p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], "1", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3a2.png")
-saveFigure(p, "manuscript-figure-3a2.pdf", dpi=300)
+saveFigure(p, "manuscript-figure-3a2.png", width=17, dpi=200)
+saveFigure(p, "manuscript-figure-3a2.pdf", width=17, dpi=300)
 p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], "1", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3a3.png")
-saveFigure(p, "manuscript-figure-3a3.pdf", dpi=300)
+saveFigure(p, "manuscript-figure-3a3.png", width=17, dpi=200)
+saveFigure(p, "manuscript-figure-3a3.pdf", width=17, dpi=300)
 p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[5]), modelName=modelNames[5], "1", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3a4.png")
-saveFigure(p, "manuscript-figure-3a4.pdf", dpi=300)
-
-xlim <- log10(c(500,4000)); ylim <- xlim
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], "2", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3b1.png")
-saveFigure(p, "manuscript-figure-3b1.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], "2", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3b2.png")
-saveFigure(p, "manuscript-figure-3b2.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], "2", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3b3.png")
-saveFigure(p, "manuscript-figure-3b3.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[5]), modelName=modelNames[5], "2", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3b4.png")
-saveFigure(p, "manuscript-figure-3b4.pdf", dpi=300)
-
-xlim <- log10(c(5,1200)); ylim <- xlim
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], "3", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3c1.png")
-saveFigure(p, "manuscript-figure-3c1.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], "3", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3c2.png")
-saveFigure(p, "manuscript-figure-3c2.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], "3", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3c3.png")
-saveFigure(p, "manuscript-figure-3c3.pdf", dpi=300)
-p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[5]), modelName=modelNames[5], "3", xlim=xlim, ylim=ylim)
-saveFigure(p, "manuscript-figure-3c4.png")
-saveFigure(p, "manuscript-figure-3c4.pdf", dpi=300)
+saveFigure(p, "manuscript-figure-3a4.png", width=17, dpi=200)
+saveFigure(p, "manuscript-figure-3a4.pdf", width=17, dpi=300)
 ```
@@
     #ggtitle(paste("Case:", case)) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
-                                        plot.title=element_text(size=10),
-                                        plot.margin=unit(c(0, 0.2, 0, 0), units="line"),
+    theme_ms() + theme(panel.border=element_rect(fill="transparent"),
                                         panel.margin=unit(0, units="line"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
@@
 pc1 <- rbind(pc1a,pc1b,pc1c,pc1d,pc1e,pc1f)
-pc2a <- getPopulationCorrelations(scenarios="Acombined", modelNames=modelNames[1], cutoff=10*cutoff)
-pc2b <- getPopulationCorrelations(scenarios="Bcombined", modelNames=modelNames[1], cutoff=10*cutoff)
-pc2c <- getPopulationCorrelations(scenarios="Ccombined", modelNames=modelNames[1], cutoff=10*cutoff)
-pc2d <- getPopulationCorrelations(scenarios="Dcombined", modelNames=modelNames[1], cutoff=10*cutoff)
-pc2e <- getPopulationCorrelations(scenarios="Ecombined", modelNames=modelNames[1], cutoff=10*cutoff)
-pc2f <- getPopulationCorrelations(scenarios="Fcombined", modelNames=modelNames[1], cutoff=10*cutoff)
-pc2 <- rbind(pc2a,pc2b,pc2c,pc2d,pc2e,pc2f)
-pc3a <- getPopulationCorrelations(scenarios="A10days", modelNames=modelNames[1], cutoff=cutoff)
-pc3b <- getPopulationCorrelations(scenarios="B10days", modelNames=modelNames[1], cutoff=cutoff)
-pc3c <- getPopulationCorrelations(scenarios="C10days", modelNames=modelNames[1], cutoff=cutoff)
-pc3d <- getPopulationCorrelations(scenarios="D10days", modelNames=modelNames[1], cutoff=cutoff)
-pc3e <- getPopulationCorrelations(scenarios="E10days", modelNames=modelNames[1], cutoff=cutoff)
-pc3f <- getPopulationCorrelations(scenarios="F10days", modelNames=modelNames[1], cutoff=cutoff)
-pc3 <- rbind(pc3a,pc3b,pc3c,pc3d,pc3e,pc3f)
-#pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1], cutoff=cutoff)
 
@@
 saveFigure(p1, "manuscript-figure-4a.pdf", dpi=300)
-p2 <- plotPopulationCorrelations(pc2, "Case 2")
-saveFigure(p2, "manuscript-figure-4b.png")
-saveFigure(p2, "manuscript-figure-4b.pdf", dpi=300)
-p3 <- plotPopulationCorrelations(pc3, "Case 3")
-saveFigure(p3, "manuscript-figure-4c.png")
-saveFigure(p3, "manuscript-figure-4c.pdf", dpi=300)
 ```
@@
   credibilityIntervals$panel <- panel
-  #credibilityIntervals$scenario <- rep(scenarioNames, times=2, each=length(modelNames))
   return(credibilityIntervals)
-
-  #scenario=rep(scenarioNames, times=2, each=length(modelNames[!modelNames %in% "FMPModel"])))
 }
@@
 cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,4,5)], panel="Case 1", scenarioNames=scenarios, cutoff=cutoff)
-#cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,5)], panel="Case 2", scenarioNames=scenarios, cutoff=10*cutoff)
-#cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,5)], panel="Case 3", scenarioNames=scenarios, cutoff=cutoff)
 
 plotCIProportions <- function(credibilityIntervals) {
-  #modelLabels <- prettyModelName(unique(credibilityIntervals$modelName))
   credibilityIntervals$Model <- prettyModelName(credibilityIntervals$modelName)
@@
     scale_colour_manual("CI", values=c("black", "grey")) +
-    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
+    theme_ms() + theme(panel.border=element_rect(fill="transparent")) +
     xlab("Scenario") + ylab("Proportion") + ylim(0, 1) +
-    #facet_wrap(~panel) +
     theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
-          plot.margin=unit(c(0.2, 0.2, 0, 0), units="line"),
           panel.margin=unit(0, units="line"),
-          legend.margin=unit(-1, units="line"),
+          #legend.margin=unit(-1, units="line"),
           #legend.position="bottom",
-          legend.direction="horizontal",
-          #legend.direction="vertical",
+          legend.direction="vertical",
+          #legend.direction="horizontal",
           legend.box="horizontal",
-          legend.key.width=unit(.7, "line"),
-          legend.key.height=unit(0, "line"))
+          legend.key.height=unit(0.7, "line"))
+          #legend.key.width=unit(.7, "line"),
+          #legend.key.height=unit(0, "line"))
 
@@
   legend <- g_legend(p)
-  p <- arrangeGrob(p + theme(legend.position="none"), legend, heights=c(9/10,1/10))
+  p <- arrangeGrob(p + theme(legend.position="none"), legend, heights=c(6/10,4/10))
   return(p)
@@
 cip <- cip1
-#cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")
-#cip <- subset(rbind(cip1, cip2, cip3), !(modelName %in% c("SmoothModel-nbinomial-ar1", "SmoothModel-nbinomial-matern-ar1")))
 p <- plotCIProportions(cip)
-saveFigure(p, "manuscript-figure-5.png", height=7.87/2)
-saveFigure(p, "manuscript-figure-5.pdf", height=7.87/2, dpi=300)
-
+saveFigure(p, "manuscript-figure-5.png", height=7.87*2/3)
+saveFigure(p, "manuscript-figure-5.pdf", height=7.87*2/3, dpi=300)
 ```

### 2015-07-15 inst/wtc-finland/manuscript2.Rmd
...

@@
 ```{r initialize}
-#library(devtools); install_github("statguy/Winter-Track-Counts")
+#library(devtools)
+#install_github("statguy/WTC")
 #install_github("ropengov/gisfin")
+#install_github("statguy/SpaceTimeModels")
 library(parallel)
@@
 library(gridExtra)
+library(SpaceTimeModels)
 
@@
 # Habitat
-
 r <- raster::crop(study$studyArea$habitat, extent(3400000, 3410000, 7000000, 7010000))
@@
 coords <- raster::xyFromCell(r, cell=1:ncell(r))
-#x0 <- smoothSubsets(r=forest, coords=coords[1:(400*10),], kernelFun=expKernel, scales=125, .parallel=F)
-x <- smoothSubsets(r=forest, coords=coords, kernelFun=expKernel, scales=c(125, 1000), .parallel=T)
+x <- SpaceTimeModels::smoothContinuousSubsets(r=forest, coords=coords, kernel=SpaceTimeModels::ExponentialKernel$new(), scales=c(125, 1000), .parallel=T)
 #plot(flip(rasterFromXYZ(x[,c("x","y","value")]), direction="y"), asp=1)
@@
 load(fileName)
+
+# Tests:
+if (F) {
+r <- crop(forest, extent(forest, 10, 20, 10, 20))
+x <- smoothContinuousSubsets(r=r, coords=xyFromCell(r, cell=1:ncell(r)), kernel=ExponentialKernel$new(), scales=125, .parallel=F)
+p <- ggplot(x) + geom_raster(aes(x, y, fill=`125`)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+plot(r)
+print(p)
+xy <- xyFromCell(r, cell=1:ncell(r)); x=xy[1,1]; y=xy[1,2]
+kernel <- ExponentialKernel$new()$constructKernel(r, 125)
+smoothContinuousSubset(r=r, x=x, y=y, scale=125, kernel=kernel)
+}
 
@@
 p2 <- ggplot(r2) + geom_raster(aes(x, y, fill=value)) + scale_fill_grey() + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
-p3 <- ggplot(subset(x, scale==125)) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
-p4 <- ggplot(subset(x, scale==1000)) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p3 <- ggplot(x) + geom_raster(aes(x, y, fill=`125`)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p4 <- ggplot(x) + geom_raster(aes(x, y, fill=`1000`)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
 
@@
 p4 <- labelPanel(p4, "d)", min(r1$x), max(r1$y))
-
 p <- arrangeGrob(p1, p2, p3, p4, ncol=2)
@@
 coords <- raster::xyFromCell(r, cell=1:ncell(r))
-x <- smoothSubsets(r=r, coords=coords, kernelFun=expKernel, scales=c(2000), .parallel=T)
+x <- SpaceTimeModels::smoothContinuousSubsets(r=r, coords=coords, kernel=SpaceTimeModels::ExponentialKernel$new(), scales=c(2000), .parallel=T)
 
 p5 <- ggplot(r1) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
-p6 <- ggplot(x) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p6 <- ggplot(x) + geom_raster(aes(x, y, fill=`2000`)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
 
@@
 coords <- raster::xyFromCell(r, cell=1:ncell(r))
-x <- smoothSubsets(r=r, coords=coords, kernelFun=expKernel, scales=c(2000), .parallel=T)
+x <- SpaceTimeModels::smoothContinuousSubsets(r=r, coords=coords, kernel=SpaceTimeModels::ExponentialKernel$new(), scales=c(2000), .parallel=T)
 
 p7 <- ggplot(r1) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
-p8 <- ggplot(x) + geom_raster(aes(x, y, fill=value)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
+p8 <- ggplot(x) + geom_raster(aes(x, y, fill=`2000`)) + scale_fill_gradient(low=gray[1], high=gray[2]) + theme_raster() + labs(x=NULL, y=NULL) + scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0))
 

### 2015-07-15 inst/wtc-finland/manuscript2.Rmd
...

@@
 #install_github("ropengov/gisfin")
-#install_github("statguy/SpaceTimeModels")
+#install_github("statguy/Blur")
 library(parallel)
@@
 library(gridExtra)
-library(SpaceTimeModels)
+library(Blur)
 
@@
 coords <- raster::xyFromCell(r, cell=1:ncell(r))
-x <- SpaceTimeModels::smoothContinuousSubsets(r=forest, coords=coords, kernel=SpaceTimeModels::ExponentialKernel$new(), scales=c(125, 1000), .parallel=T)
+x <- Blur::smoothContinuousSubsets(r=forest, coords=coords, kernel=Blur::ExponentialKernel$new(), scales=c(125, 1000), .parallel=T)
 #plot(flip(rasterFromXYZ(x[,c("x","y","value")]), direction="y"), asp=1)
@@
 coords <- raster::xyFromCell(r, cell=1:ncell(r))
-x <- SpaceTimeModels::smoothContinuousSubsets(r=r, coords=coords, kernel=SpaceTimeModels::ExponentialKernel$new(), scales=c(2000), .parallel=T)
+x <- Blur::smoothContinuousSubsets(r=r, coords=coords, kernel=Blur::ExponentialKernel$new(), scales=c(2000), .parallel=T)
 

### 2016-02-14 inst/wtc-finland/manuscript.Rmd
Change package name

@@
 registerDoMC(cores=detectCores())
-library(WTC)
-source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
+library(STREM)
+source("~/git/STREM/setup/WTC-Boot.R")
 library(sp)

### 2016-02-14 inst/wtc-finland/manuscript2.Rmd
Change package name

@@
 #library(devtools)
-#install_github("statguy/WTC")
+#install_github("statguy/STREM")
 #install_github("ropengov/gisfin")
@@
 registerDoMC(cores=detectCores())
-library(WTC)
-source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
+library(STREM)
+source("~/git/STREM/setup/WTC-Boot.R")
 library(ggplot2)
@@
 ```
+
+```{r table-1}
+writeTable <- function(df, filename, context) {
+  library(ReporteRs)
+  doc <- docx()
+  flextable <- FlexTable(df)
+  addFlexTable(doc, flextable)
+  dir <- normalizePath(context$figuresDirectory)
+  writeDoc(doc, file=file.path(dir, filename))
+}
+
+models <- data.frame(Species=c("Canis Lupus"), Model=c("Intercept-only"), WAIC=c("1074.72"))
+models <- rbind(models, data.frame(Species=c("Canis Lupus", "Canis Lupus"), Model=c("All PCs", "PC1 habitat 250 + PC2 all 64000"), WAIC=c("1074.78", "1072.57")))
+models <- rbind(models, data.frame(Species=c("Canis Lupus", "Canis Lupus"), Model=c("All non-PCs 32000", "All non-PCs 16000"), WAIC=c("1061.45", "1067.00")))
+models <- rbind(models, data.frame(Species=c("Lynx Lynx"), Model=c("Intercept-only"), WAIC=c("3922.51")))
+models <- rbind(models, data.frame(Species=c("Lynx Lynx", "Lynx Lynx"), Model=c("All PCs", "PC3 all 8000"), WAIC=c("3930.95", "3914.81")))
+models <- rbind(models, data.frame(Species=c("Lynx Lynx", "Lynx Lynx"), Model=c("All non-PCs 2000", "All non-PCs 8000"), WAIC=c("3917.34", "3917.35")))
+writeTable(models, "table-b-1.docx", context)
+```

### 2016-02-14 R/RcppExports.R

@@ -2,6 +2,6 @@
 # Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
 internal_countIntersections <- function(track, surveyRoute) {
-    .Call('WTC_internal_countIntersections', PACKAGE = 'WTC', track, surveyRoute)
+    .Call('STREM_internal_countIntersections', PACKAGE = 'STREM', track, surveyRoute)
 }
 

### 2016-02-14 inst/simulation/combine_tracks.R

@@ -3,8 +3,8 @@
 library(parallel)
 library(doMC)
 registerDoMC(cores=detectCores())
-library(WTC)
-source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
+library(STREM)
+source("~/git/STREM/setup/WTC-Boot.R")
 
 combineTracks <- function(scenario, combineFactor) {
   mss <- getMSS(scenario=scenario)

### 2016-02-14 inst/simulation/count_intersections.R

@@ -15,8 +15,8 @@ countIntersections <- function(scenario, iteration, nSurveyRoutes=as.integer(500
 library(parallel)
 library(doMC)
 registerDoMC(cores=detectCores())
-library(WTC)
-source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
+library(STREM)
+source("~/git/STREM/setup/WTC-Boot.R")
 
 parseArguments()
 countDays <- if (!is.null(extraArgs)) as.integer(extraArgs[1]) else as.integer(1)

### 2016-02-14 inst/simulation/summarize.R

@@ -1,8 +1,8 @@
 library(parallel)
 library(doMC)
 registerDoMC(cores=detectCores())
-library(WTC)
-source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
+library(STREM)
+source("~/git/STREM/setup/WTC-Boot.R")
 
 context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
 mss <- if (scenario == "A") MovementSimulationScenarioA$new()$setup(context=context) else stop("Unknown scenario ", scenario)

### 2016-02-14 inst/simulation/test_template.R

@@ -3,8 +3,8 @@ library(devtools); install_github("statguy/Winter-Track-Counts")
 library(parallel)
 library(doMC)
 registerDoMC(cores=detectCores())
-library(WTC)
-source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
+library(STREM)
+source("~/git/STREM/setup/WTC-Boot.R")
 
 scenarios <- c("A","B","C","D","E","F")
 modelNames <- c("FMPModel","SmoothModel-nbinomial-ar1","SmoothModel-nbinomial-matern-ar1")

### 2016-02-14 inst/simulation/validate_boot_ci.R

@@ -2,8 +2,8 @@
 library(parallel)
 library(doMC)
 registerDoMC(cores=round(detectCores()))
-library(WTC)
-source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
+library(STREM)
+source("~/git/STREM/setup/WTC-Boot.R")
 library(plyr)
 #library(reshape2)
 library(boot)

### 2016-02-14 inst/wtc-finland/estimate.R

@@ -1,7 +1,7 @@
 # ./parallel_r.py -t 1:3 -n 6 -l 10.0 -b ~/tmp/blacklist.txt -v ~/git/Winter-Track-Counts/inst/wtc/estimate.R notest FMPModel
 # ./parallel_r.py -t 3 -n 2 -l 10.0 -b ~/tmp/blacklist.txt -v ~/git/Winter-Track-Counts/inst/wtc/estimate.R notest FMPModel
 
-# library(devtools); install_github("statguy/Winter-Track-Counts")
+# library(devtools); install_github("statguy/WTC")
 
 args <- commandArgs(trailingOnly=TRUE)
 if (length(args) != 3) stop("Invalid arguments.")
@@ -14,12 +14,82 @@ print(args)
 library(parallel)
 library(doMC)
 registerDoMC(cores=detectCores())
-library(WTC)
-source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
+library(STREM)
+source("~/git/STREM/setup/WTC-Boot.R")
 
 if (test == "test") {
+  response <- "canis.lupus"
+  
   context <- Context(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
   study <- FinlandWTCStudy(context=context, response=response)
+  
+  intersections <- FinlandWTCIntersections$new(study=study)$setCovariatesId(tag="density")$loadCovariates()$removeMissingCovariates()
+  #plot(study$studyArea$boundary)
+  #lines(matrix(c(3186635,7300000,3732935,7300000), byrow=T, ncol=2))
+  index <- coordinates(intersections$intersections)[,2] < 7.3e6 # Cut off Lapland since all wolves are killed there
+  index <- index & intersections$intersections$duration == 1 # Use durations of 1 only
+  intersections$intersections <- intersections$intersections[index,]
+  colnames(intersections$intersections@data) <- stringr::str_replace_all(colnames(intersections$intersections@data), "[\\[\\]_]", ".") # there is a problem with column names in INLA or somewhere
+  intersections$covariateNames <- stringr::str_replace_all(intersections$covariateNames, "[\\[\\]_]", ".")
+
+  #ddply(intersections$intersections@data, .(year), nrow)
+  
+  dataNames <- colnames(intersections$intersections@data)
+  for (scale in 2^(0:10) * 62.5) {
[…]

### 2016-02-14 inst/wtc-finland/movement_habitat_distribution.R

@@ -1,8 +1,8 @@
 library(parallel)
 library(doMC)
 registerDoMC(cores=round(detectCores()))
-source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
-library(WTC)
+library(STREM)
+source("~/git/STREM/setup/WTC-Boot.R")
 library(ggplot2)
 
 plotMovementHabitatDistribution <- function(response) {

### 2016-02-14 inst/wtc-finland/visualize.R

@@ -1,5 +1,5 @@
-library(WTC)
-source("~/git/Winter-Track-Counts/setup/WTC-Boot.R")
+library(STREM)
+source("~/git/STREM/setup/WTC-Boot.R")
 library(sp)
 library(ggplot2)
 library(rasterVis)

### 2016-03-06 inst/simulation/manuscript.Rmd
...

@@
 library(plyr)
+library(gridExtra)
 
-context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
+context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures, figuresHost=wd.figuresHost)
 scenarios <- c("A","B","C","D","E","F")
@@
   facet_wrap(~scenario) + #coord_equal()
-  theme_ms() + theme(panel.border=element_rect(fill="transparent")) +
+  theme_raster() + theme(panel.border=element_rect(fill="transparent")) +
   xlab("X-coordinate (km)") + ylab("Y-coordinate (km)") +
@@
 study <- getStudy(scenario="A")
-limits <- extent(study$studyArea$boundary)
-iterations <- as.integer(c(1,4,1,2,5,34))
 year <- 2010
 yearLayer <- paste0("X", year)
+#iterations <- as.integer(c(1,4,1,4,5,38))
+#iterations <- as.integer(c(1,4,1,4,5,16))
+iterations <- as.integer(c(1,4,1,4,5,21))
+densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year)
+wdensities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, weightHabitats=T)
 surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
 agents <- getAgentLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
-densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year)
-wdensities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, weightHabitats=T)
+limits <- extent(study$studyArea$boundary)
 
@@
 
+boundaryDF <- ggplot2::fortify(FinlandWTCStudy(context=context, response="canis.lupus")$studyArea$boundary)
 wdensitiesScaled$x <- wdensitiesScaled$x / 1000 - 3000
@@
   stat_contour(data=densitiesScaled, aes_string(x="x", y="y", z=yearLayer, colour=yearLayer), size=0.1) +
-  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent",
+  scale_fill_gradientn(colours=c(rep("#00A600FF", 60), terrain.colors(40)), na.value="transparent",
                        guide=guide_legend(title=expression(paste("Population density ", (individuals/km^2)))), trans="log10",
                        breaks=trans_breaks("log10", function(x) 10^x),
-                       labels=trans_format("log10", math_format(10^.x))
-                       ) +
+                       labels=trans_format("log10", math_format(10^.x))) +
   geom_path(data=boundaryDF, aes(long, lat, group=group), colour="black", size=0.5) +
@@
   scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0.01,0.01)) +
-  facet_grid(~scenario) + coord_equal() + theme_ms() +
+  facet_grid(~scenario) + coord_equal() + theme_raster() +
   theme(legend.position="bottom", legend.margin=unit(-1, units="line"), legend.title=element_text(size=10), legend.text=element_text(size=10)) + #top, right, bottom, and left margins
@@
   scale_y_continuous(breaks=seq(gridTemplate@extent@ymin, gridTemplate@extent@ymax, 1e5/1000), expand=c(0.01,0.01)) +
-  facet_grid(~scenario) + coord_equal() + theme_ms() +
+  facet_grid(~scenario) + coord_equal() + theme_raster() +
   theme(panel.grid.major=element_line(colour="grey")) + #top, right, bottom, and left margins

### 2016-03-06 R/static_global_functions.R

@@ -111,7 +111,7 @@ theme_presentation <- function(base_size=20, base_family="", ...) {
 }
 
 # full width = 2*7.87+1.25=17
-saveFigure <- function(p, filename, width=7.87, height=7.87, units="cm", dpi=200, dimensions, ...) {
+saveFigure <- function(p, filename, width=7.87, height=7.87, units="cm", dpi=200, dimensions, copy=T, ...) {
   library(ggplot2)
   if (missing(p) | missing(filename))
     stop("Argument p or filename missing.")
@@ -122,6 +122,10 @@ saveFigure <- function(p, filename, width=7.87, height=7.87, units="cm", dpi=200
     height <- width * aspectRatio
   }
   ggsave(p, filename=filename, width=width, height=height, units=units, ...)
+  if (copy) {
+    system(paste("scp", filename, paste0(context$figuresHost, ":", filename)))
+    system(paste("ssh", context$figuresHost, "\"chmod 644", filename, "\""))
+  }
 }
 
 addDtDist <- function(tracksDF) {

### 2016-03-06 inst/simulation/manuscript.Rmd
...

@@
 ```{r initialize}
-#library(devtools); install_github("statguy/Winter-Track-Counts")
+#library(devtools); install_github("statguy/STREM")
 library(parallel)
@@
                                         plot.title=element_text(size=10),
-                                        #plot.margin=unit(c(0, 0.1, 0, 0), units="line"), #top, right, bottom, and left margins
                                         legend.margin=unit(0, units="line"),
@@
   credibilityIntervals$Model <- prettyModelName(credibilityIntervals$modelName)
-  credibilityIntervals$CI <- credibilityIntervals$Interval
-  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Model, colour=CI)) +
+  credibilityIntervals$CrI <- credibilityIntervals$Interval
+  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Model, colour=CrI)) +
     geom_hline(aes(yintercept=.5), color="darkgrey") +
@@
     geom_point(aes(shape=Model, group=Model), position=position_dodge(width=.7)) +
-    scale_colour_manual("CI", values=c("black", "grey")) +
-    theme_ms() + theme(panel.border=element_rect(fill="transparent")) +
+    scale_colour_manual("CrI", values=c("black", "grey"), guide=guide_legend(override.aes=list(size=3.5, shape=15))) +
+    theme_ms() +
+    theme(panel.border=element_rect(fill="transparent")) +
     xlab("Scenario") + ylab("Proportion") + ylim(0, 1) +
@@
           panel.margin=unit(0, units="line"),
-          #legend.margin=unit(-1, units="line"),
-          #legend.position="bottom",
           legend.direction="vertical",
-          #legend.direction="horizontal",
           legend.box="horizontal",
           legend.key.height=unit(0.7, "line"))
-          #legend.key.width=unit(.7, "line"),
-          #legend.key.height=unit(0, "line"))
+    #guides(scale_colour_manual=guide_colourbar())
 

### 2016-03-06 R/static_global_functions.R

@@ -70,34 +70,50 @@ inverseDistanceWeightningImpute <- function(data, varname, formula=as.formula(pa
   return(rbind(data.full, data.na))
 }
 
+theme_ms <- theme_raster <- function(base_size=12, base_family="", ...) {
+  library(grid)
+  theme_minimal(base_size=base_size, base_family=base_family) %+replace%
+    theme(
+      panel.background=element_rect(fill="transparent", colour=NA),
+      panel.grid.minor=element_blank(), 
+      panel.grid.major=element_blank(),
+      plot.background=element_rect(fill="transparent", colour=NA),
+      panel.border=element_blank(),
+      panel.margin=unit(0, "lines"),
+      strip.background=element_blank(),
+      plot.margin=unit(c(0.1,0,0,0), "lines"), # top, right, bottom, and left margins
+      ...
+    )
+}
+
 theme_raster <- function(base_size=12, base_family="", ...) {
   library(grid)
   theme_minimal(base_size=base_size, base_family=base_family) %+replace%
-  theme(
-    panel.background=element_rect(fill="transparent", colour=NA),
-    panel.grid.minor=element_blank(), 
-    panel.grid.major=element_blank(),
-    plot.background=element_rect(fill="transparent", colour=NA),
-    #panel.background=element_blank(),
-    panel.border=element_blank(),
-    #panel.grid.major=element_blank(),
-    #panel.grid.minor=element_blank(),
-    panel.margin=unit(0, "lines"),
-    axis.line=element_blank(),
-    axis.text.x=element_blank(),
-    axis.text.y=element_blank(),
-    axis.title.x=element_blank(),
-    axis.title.y=element_blank(),
-    axis.ticks=element_blank(),
-    strip.background=element_blank(),
-    #plot.margin=unit(c(0,0,-1,-1), "lines"),
-    #plot.margin=unit(c(0,0,-.5,-.5), "lines"),
-    plot.margin=unit(c(0,0,0,0), "lines"),
-    axis.ticks.length=unit(0,"lines"),axis.ticks.margin=unit(0,"lines"),
-    #plot.margin=rep(unit(0,"null"),4),panel.margin=unit(0,"null"),axis.ticks.length=unit(0,"null"),axis.ticks.margin=unit(0,"null"),
-    legend.position="none",
-    ...
-  )
+    theme(
+      panel.background=element_rect(fill="transparent", colour=NA),
[…]

### 2016-03-06 inst/simulation/manuscript.Rmd
...

@@
 ```{r figure-3sample}
-#xlim <- log10(c(30,800)); ylim <- xlim
 logPopulationSizesSample <- subset(logPopulationSizes, scenario == "A")
@@
   p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
-    geom_point(size=1, colour="darkgray", alpha=.5) +
+    geom_point(size=1, colour="gray40", alpha=.5) +
     geom_abline(colour="black", linetype="longdash") + 
@@
     xlab("True population size") + ylab("Estimated population size") +
-    theme_ms() + theme(panel.border=element_rect(fill="transparent"),
+    theme_ms2() + theme(panel.border=element_rect(fill="transparent"),
                                         plot.title=element_text(size=10),
@@
 
-p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName %in% c("ST","T1","T2","FMP")), modelName=c("ST","T1","T2","FMP"), "a", xlim=xlim, ylim=ylim)
+p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName %in% c("ST","T1","T2","FMP")), modelName=c("ST","T1","T2","FMP"), "a", xlim=log10(c(1, 10^3)), ylim=log10(c(1,10^3)))
 saveFigure(p, "manuscript-figure-3sample.png", height=7.87, width=17) # draft quality
@@
     xlab(expression(paste(Log[10], " true population size"))) + ylab(expression(paste(Log[10], " estimated population size"))) + ggtitle(paste("Model:", prettyModelName(modelName[1]))) +
-    theme_ms() + theme(panel.border=element_rect(fill="transparent"),
+    theme_ms2() + theme(panel.border=element_rect(fill="transparent"),
                                         legend.margin=unit(0, units="line"),

### 2016-03-06 R/static_global_functions.R

@@ -86,6 +86,20 @@ theme_ms <- theme_raster <- function(base_size=12, base_family="", ...) {
     )
 }
 
+theme_ms2 <- theme_raster <- function(base_size=12, base_family="", ...) {
+  library(grid)
+  theme_minimal(base_size=base_size, base_family=base_family) %+replace%
+    theme(
+      panel.background=element_rect(fill="transparent", colour=NA),
+      plot.background=element_rect(fill="transparent", colour=NA),
+      panel.border=element_blank(),
+      panel.margin=unit(0, "lines"),
+      strip.background=element_blank(),
+      plot.margin=unit(c(0.1,0,0,0), "lines"), # top, right, bottom, and left margins
+      ...
+    )
+}
+
 theme_raster <- function(base_size=12, base_family="", ...) {
   library(grid)
   theme_minimal(base_size=base_size, base_family=base_family) %+replace%

### 2016-03-06 bin/manuscript2_pdf2tiff.sh

@@ -1,0 +1,39 @@
+#!/bin/bash
+# Run in figures output directory
+
+function pdf2eps {
+  if [ ! -z "$2" ]; then
+    pdfcrop --bbox "$2" $1.pdf "$1-crop.pdf"
+  else
+    pdfcrop $1.pdf "$1-crop.pdf"
+  fi
+
+  convert -density 600 -units PixelsPerInch "$1-crop.pdf" -background white \
+    -flatten -alpha off -auto-level -depth 8 -compress lzw $1.tif
+  chmod a+r $1.tif
+}
+
+pdf2eps manuscript-figure-1
+pdf2eps manuscript-figure-2
+pdf2eps manuscript-figure-3
+pdf2eps manuscript-figure-3sample
+pdf2eps manuscript-figure-4a
+pdf2eps manuscript-figure-5
+
+# 1 2 481 219
+pdf2eps manuscript-figure-3a1 "1 2 481 205"
+pdf2eps manuscript-figure-3a2 "1 2 481 205"
+pdf2eps manuscript-figure-3a3 "1 2 481 205"
+pdf2eps manuscript-figure-3a4 "1 2 481 205"
+
+mv manuscript-figure-1.tif Fig1.tif
+mv manuscript-figure-2.tif Fig2.tif
+mv manuscript-figure-3.tif Fig3.tif
+mv manuscript-figure-3sample.tif Fig4.tif
+mv manuscript-figure-4a.tif Fig6.tif
+mv manuscript-figure-5.tif Fig5.tif
+
+mv manuscript-figure-3a1.tif S1_Fig.tif
+mv manuscript-figure-3a2.tif S2_Fig.tif
+mv manuscript-figure-3a3.tif S3_Fig.tif
+mv manuscript-figure-3a4.tif S4_Fig.tif

### 2016-03-14 inst/simulation/manuscript.Rmd
...

@@
     geom_point(aes(shape=Model, group=Model), position=position_dodge(width=.7)) +
-    scale_colour_manual("CrI", values=c("black", "grey"), guide=guide_legend(override.aes=list(size=3.5, shape=15))) +
+    scale_colour_manual("CrI / CI", values=c("black", "grey"), guide=guide_legend(override.aes=list(size=3.5, shape=15))) +
     theme_ms() +

### 2016-07-09 inst/simulation/manuscript.Rmd
...

@@
                                         plot.title=element_text(size=10),
-                                        legend.margin=unit(0, units="line"),
-                                        panel.margin=unit(0, units="line"))
+                                        legend.margin=unit(0, units="line"))
 

### 2016-07-09 R/Intersections.R

@@ -313,14 +313,20 @@ FinlandWTCIntersections <- setRefClass(
       
       wtc <- read.xls(file.path(study$context$rawDataDirectory, "Kolmiot_1989_2011.xls"))
       names(wtc)[1:9] <- c("id","y","x","length","forest","field","year","date","duration")
-      names(wtc)[16] <- "canis.lupus"
-      names(wtc)[29] <- "lynx.lynx"
-      names(wtc)[33] <- "rangifer.tarandus.fennicus"
+      names(wtc)[names(wtc) == "CANLUP"] <- "canis.lupus"
+      names(wtc)[names(wtc) == "LYNLYN"] <- "lynx.lynx"
+      names(wtc)[names(wtc) == "RANTAR"] <- "rangifer.tarandus.fennicus"
+      names(wtc)[names(wtc) == "ODOVIR"] <- "odocoileus.virginianus"
+      names(wtc)[names(wtc) == "CAPCAP"] <- "capreolus.capreolus"
+      names(wtc)[names(wtc) == "ALCALC"] <- "alces.alces"
       
       responseIndex <- switch (study$response,
-              canis.lupus=16,
-              lynx.lynx=29,
-              rangifer.tarandus.fennicus=33)
+              canis.lupus=names(wtc) == "canis.lupus",
+              lynx.lynx=names(wtc) == "lynx.lynx",
+              rangifer.tarandus.fennicus=names(wtc) == "rangifer.tarandus.fennicus",
+              odocoileus.virginianus=names(wtc) == "odocoileus.virginianus",
+              capreolus.capreolus=names(wtc) == "capreolus.capreolus",
+              alces.alces=names(wtc) == "alces.alces")
       wtc <- wtc[,c(1:9,responseIndex)]
       colnames(wtc)[ncol(wtc)] <- "intersections"
       wtc$response <- study$response
@@ -338,6 +344,11 @@ FinlandWTCIntersections <- setRefClass(
       wtc <- subset(wtc, duration<4 & duration>0 & length>0)  
       # Transects 168 and 1496 are at the same location but separate, take average, ignore or something...
       wtc <- subset(wtc, id!=1496)
+      
+      #plot(study$studyArea$boundary)
+      #x <- subset(wtc, response > 0)
+      #points(x$x, x$y)
+      
       xy <- cbind(wtc$x, wtc$y)
       wtc$x <- NULL
       wtc$y <- NULL

### 2016-07-09 R/static_global_functions.R

@@ -95,7 +95,7 @@ theme_ms2 <- theme_raster <- function(base_size=12, base_family="", ...) {
       panel.border=element_blank(),
       panel.margin=unit(0, "lines"),
       strip.background=element_blank(),
-      plot.margin=unit(c(0.1,0,0,0), "lines"), # top, right, bottom, and left margins
+      plot.margin=unit(c(.1,.1,0,0), "lines"), # top, right, bottom, and left margins
       ...
     )
 }
